### 2009-09-15 doc/paper.tex
Add documentation.
Ignore-this: d508e0b173d47625e5e05ebe1741da9c

darcs-hash:20090915081658-bab43-6e5d01ef90a159c728802195b952eb62065eb2bc.gz

\documentclass[runningheads]{llncs}

\newcommand{\eelis}{Eelis van der Weegen}
\newcommand{\herman}{Herman Geuvers}
\newcommand{\dan}{Dan Synek}
\newcommand{\adam}{Adam Koprowski}

\newcommand{\thetitle}{Automated Machine-Checked Hybrid System Safety Proofs}

\usepackage{color}
\usepackage{amsmath}
\usepackage[colorlinks,pdftitle=\thetitle,pdfauthor=]{hyperref}

\spnewtheorem*{propo}{Proposition}{\bfseries}{\rmfamily}

%include polycode.fmt

%format pipe = "\mathopen{|}"

%format times = "\times"

%format . = "."
%format forall = "\forall"
%format Fixpoint = "\mathbf{Fixpoint}"
%format Definition = "\mathbf{Definition}"
%format Eval = "\mathbf{Eval}"
%format Notation = "\mathbf{Notation}"
%format Lemma = "\mathbf{Lemma}"
%format Proof = "\mathbf{Proof}"
%format Qed = "\mathbf{Qed}"
%format Defined = "\mathbf{Defined}"
%format Function = "\mathbf{Function}"
%format Proposition = "\mathbf{Proposition}"
%format Record = "\mathbf{Record}"
%format Program = "\mathbf{Program}"
%format Inductive = "\mathbf{Inductive}"
%format Theorem = "\mathbf{Theorem}"
%format Let = "\mathbf{Let}"
%format Instance = "\mathbf{Instance}"
%format Variable = "\mathbf{Variable}"
%format Variables = "\mathbf{Variables}"
%format measure = "\mathbf{measure}"

%format prod a b = a "\times " b
%format sp = "\ "
%format frac a b = "\frac{" a "}{" b "}"

%format == = "= "
%format </ = "\nless "
%format /= = "\ne "
  % Without these, lhs2tex uses three-line equal signs.

%format fcmp = "\circ"
%format match = "\mathbf{match}"
%format with = "\mathbf{with}"
%format end = "\mathbf{end}"
%format fun = "\lambda\hspace{-1mm} "
%format nat = "\mathbb{N}"
%format CRplain = "\texttt{CR}"
%format CR = "\mathbb{R}"
%format CRasCSetoid = CR
%format exists = "\exists "
%format log2ceil = "\log_{2}\hspace{-0.5mm}"
%format over = "\mathbf{over}"
%format INR (a) = a
%format `union` = "\cup "
%format /\ = "\wedge "
%format ~ = "\neg "
%format \/ = "\vee"
%format <-> = "\leftrightarrow "
%format ==> = "\Longrightarrow "
%format `subsetof` = "\subseteq "
%format phi = "\phi "
%format False = "\bot "

%format sp1mm = "\hspace{1mm} "

%%format { = "\hspace{0.9mm}\{"
  % so that parameters declared with {...} are spaced like those declared with (...)
    % disabled because it makes record defs ugly

%format (pow (a) (b)) = a"^"b

%format In (a) (b) = a "\in " b
%format `NotIn` = "\notin"

\def\commentbegin{\quad\{\ }
\def\commentend{\}}

\def\typeset{}

\usepackage{macros}

\arrayhs % i.e. no page breaks in code blocks

% todo: mention deadlock

% meta:
%  - shoudl this be readable without having read alur?
% 
% presentation simplifications:
% - don't show concrete./abstract.State system. just use "State", "Point", "Location", etc.
% - use "a <= x <= b"
% - don't show backtic coercions or notation scope tags
% - some implicit arguments like plain eq instead of (@eq _)
% - leave out all the NoDup stuff.
% - instead of saying ProdCSetoid and CRasCSetoid, just say A*B and CR
% - "exists x, x `in` c /\ P" -> "exists x `in` c, P"
% - sigma projections and other coercions
% - use plain "=" for setoid equality.
% - patterns in fun's

\title{\thetitle}
\subtitle{An Implementation of the Abstraction Method In Coq}

\author{}
\institute{Institute for Computing and Information Sciences\\ Nijmegen University}

\begin{document}

\maketitle

%\authorrunning{\em}
%\titlerunning{\em}

\begin{abstract}
This technical report documents our development of a hybrid system safety prover, implemented in Coq using the abstraction method introduced by Alur in \cite{alur}. The development includes: a formalization of the structure of hybrid systems; a systematic approach and generic set of support utilities for the construction of an abstract system (consisting of decidable ``overestimators'' of abstract transitions and initiality) faithfully representing a (concrete) hybrid system; a translation of abstract systems to graphs enabling decision of abstract state reachability using a certified graph reachability algorithm; a proof of an example hybrid system (taken from \cite{alur}) generated using this tool stack. The development critically relies on the computable real number implementation part of the C-CoRN library of formalized constructive mathematics. \todo{Hm, anything else worth mentioning?}

\end{abstract}

\setcounter{tocdepth}{2}

\tableofcontents
\todo{How do I get rid of that silly author index at the top of the TOC?!}

\nocite{*}

\section{Introduction}

In \cite{alur}, Alur presents a method for automated hybrid system safety verification based on the construction of an \emph{abstract} hybrid system (essentially a finite automaton) corresponding to the hybrid system of interest. The abstract system is constructed such that traces in the original system are represented in the abstract system. Consequently, one can draw conclusions about reachability of states in the concrete system from analysis of state reachability in the abstract system. Since the abstract system is an entirely finite discrete object (unlike the concrete system), reachability can simply be computed (using any graph reachability algorithm). Thus, the abstraction method brings the safety verification problem from a continuous and infinite domain into a discrete and finite domain, where it is amenable to what amounts to brute force.

The prototype implementation described in \cite{alur} was developed in \todo{some ordinary programming language}, using \todo{some ordinary programming libraries}. The potential for bugs and floating point artifacts inherent in this approach limit the confidence one can justifiably have in safety judgements made by such an implementation. In this report, we describe a reimplementation of the basic technique in Coq, a proof assistant based on a rich type theory that also functions as a programming language, letting one develop ``certified'' programs: programs whose correctness is proved formally within the system. In our case, the program will be one that produces machine-checked proofs of hybrid system safety, obtained using the abstraction method.

Our development relies crucially on computation using real numbers, for which we use the computable real number implementation developed by Russell O'Connor \cite{something}, and part of the C-CoRN library of formalized constructive mathematics \cite{something}. Indeed, this development showcases its use in a concrete and practical application, and subtleties regarding the nature of these computable reals are reflected in this development in several ways, which we will discuss in some detail.

% todo: document that fallback-to-making transitions
% todo: document separability, and why we don't use alur's technique
% read my mail about DN to see if all its contents is now part of section bla

 % In our case, the specification of the program we write is this: given a description of a concrete system, the program  . The whole implementation, including the computational part, serves only to prove a single theorem. Hence, one essentially only needs to trust Coq to be convinced of the safety proofs created by our development.

%A key ingredient in our development

%Wherever we say ``|CR|'', we are really referring to the type |CRplain| of computable reals, part of C-CoRN.

%The computational steps become part of the complete proof of the safety theorem.

%We formally prove safety of the thermostat example used by Alur.

\paragraph{Organization}

Throughout this report, we present in parallel our general formalized framework (with particular emphasis on differences with Alur's presentation), and the way it is used to obtain a safety proof for an example hybrid system, taken from \cite{alur}, modeling the operation of a thermostat. In section \ref{concrete}, we begin by defining the structure of normal (concrete) hybrid systems, their semantics, reachability, and safety. We also define the concrete thermostat and the safety condition we wish to prove. In section \ref{dn}, we take a moment to discuss some of the limits on decidability of properties of computable real numbers, as these complicate matters in the remainder of the development. Next, in section \ref{estimation} we describe notions of underestimation and overestimation that are applied throughout the development. Then, in section \ref{abstraction}, we begin our presentation of the abstraction method by discussing abstract regions, states, and spaces. We build on these in section \ref{abs.reach}, where we first develop the ``obvious'' notions of abstract system and reachability, discuss why they fail, and then rework them to embrace reachability sharing. In section \ref{underestimatingsafety} we show how, assuming abstract reachability is decidable, we may prove hybrid system safety by computation. Then, in section \ref{graphs}, we show how an abstract system consisting of transition and initiality ``overestimators'' can be used to decide abstract reachability by construction of a directed graph corresponding to the system, and the use of an ordinary but verified graph reachability algorithm. Finally, in sections \ref{over_cont_trans} and \ref{over_disc_trans} we describe the implementation of the aforementioned overestimators. We end with conclusions in section \ref{conclusions}.

% proceed to define abstraction bla. Following that, we show that a simple (certified) graph reachability can be employed to compute reachability in the graph corresponding to the abstract system, and that the resulting reachability judgement can be translated back through the abstract system all the way to the concrete system. We end with conclusions in section \ref{conclusions}.

\section{(Concrete) Hybrid Systems}
\label{concrete}

We begin by showing our definition of a concrete system, the different parts of which we discuss in the remainder of this section.

\begin{code}
Record System: Type :=
  { Point: CSetoid
  ; Location: Set
  ; Location_eq_dec: EqDec Location eq
  ; locations: ExhaustiveList Location
  ; State := prod Location Point
  ; initial: State -> Prop
  ; invariant: State -> Prop
  ; invariant_initial: initial `subsetof` invariant
  ; invariant_mor: Morphism (eq ==> cs_eq ==> iff) (curry invariant)
  ; invariant_stable: forall s, Stable (invariant s)
  ; flow: Location -> Flow Point
  ; guard: State -> Location -> Prop
  ; reset: Location -> Location -> Point -> Point
  }.
\end{code}

%  (* this separation of guard and reset seems to cause a problem:
%    the paper allows having different transitions from a given (l, x) to
%    some (l', x'), because you can have different transitions. we only allow one! *)

\subsection{States}

A hybrid system is a model of how a software system, described as a finite set of \emph{locations} \footnote{We reserve the term ``state'' for something else.} with (discrete) transitions between them, acts on and responds to a set of continuous variables (called the \emph{continuous state space}), typically representing physical properties of some environment (such as temperature and pressure).

In \cite{alur}, Alur requires that the continuous state space be a convex polyhedron in |pow CR n|. In our definition of a hybrid system, we initially let the continuous state space be an arbitrary (constructive) setoid, called |Point|. We also explicitly require decidability of location equality.

Our running example (taken from \cite{alur}) is a thermostat whose discrete component consists of three locations (called |Heat|, |Cool|, and |Check|), and whose continuous state space is |pow CR 2|, where |CR| are the \emph{computable} reals. The |Heat| and |Cool| locations represent states in which the thermostat heats and cools the environment it operates in, respectively. The |Check| state is a self-diagnostic state in which the thermostat does not heat or cool. \todo{bah, so awkward!}  The first |CR| component in the continuous state space represents an internally resettable clock, while the second represents the temperature in the environment in which the thermostat operates.

A \emph{State} in a hybrid system is a pair |(l, p)| consisting of a location |l| and a point |p| in the continuous state space (representing an assignment of the continuous variables).

The |invariant| predicate defines, for each location, a set of permitted values for the continuous variables. We will use this in the definitions of transition relations in a moment. The morphism field expresses that this invariant respects |Point|'s setoid equality. We further require that it is \emph{stable}, where |stable P| is defined as |~ ~P -> P|. We discuss the reason for this stability requirement in section \ref{dn}.

For the thermostat, the invariant is as follows:
\begin{code}
Definition invariant (s: State): Prop :=
  0 <= clock s /\
  match location s with
  | Heat => temp s <= 10 /\ clock s <= 3
  | Cool => 5 <= temp s
  | Check => clock s <= 1
  end.
\end{code}

Associated with the hybrid system is a set of \emph{initial states}. For our thermostat example, the initial states are
\begin{code}
  {s: State pipe location s = Heat /\ 5 <= temp s <= 10 /\ clock s == 0}
\end{code}
where |location|, |temp|, and |clock| are the obvious projections. For the thermostat, the requirement that the invariant holds at each initial state is easily proved.

The remaining parts of a hybrid system (i.e. |flow|, |guard| and |reset|) describe transitions between states, which, together with the set of initial states, determine the set of \emph{reachable} states, representing the possible behaviors exhibited by a hypothetical real-world implementation of the hybrid system (as software running on a device with sensors and actuators).

\subsection{Flow and Continuous Transitions}
\label{concreteflow}

Each location in a concrete system has an accompanying \emph{flow function} which describes how the continuous variables change over time while the system is in that location. The idea is that the different locations corresponds to different uses of actuators available to the software system, the effects of which are described by the flow function. For instance, in our thermostat, the flow function corresponding to the |Heat| location will have the temperature increase with time, modeling the effect of the heater component in our imagined thermostat deviced.

%format Duration = CR"_{\ge 0}"
%format CRge0 = CR"_{\ge 0}"

In the canonical definition of hybrid systems, flow functions are specified as solutions to differential equations describing the dynamics of the continuous variables. We follow Alur's example in abstracting from these, taking instead functions |phi| of type |Point -> Duration -> Point| which satisfy:
\begin{align*}
  |phi p 0| &= |p| \\
  |phi p (t + t')| &= |phi (phi t p) t'|
\end{align*}
We further require that |phi| is a morphism respecting |Point|'s setoid equality. We bundle functions with these properties as a record type called |Flow| (not shown), which is parameterized over the continuous state space type, and is equipped with an implicit coercion to the |phi| function it contains.

\todo{Unlike Alur, we don't have ``input'' (why?).}

We now say that there is a (concrete) \emph{continuous transition} from a state |(l, p)| to a state |(l', p')| if |l = l'| and there is a nonnegative duration |d| such that |p' = flow l p d| with the invariant for |l| holding at every point along the way:

\begin{code}
  Definition can_flow (l: Location): relation Point
    := fun p p' => exists d: Duration, flow l p d = p' /\
      forall t, 0 <= t <= d -> invariant (l, flow l p t).

  Definition cont_trans: relation State
    := fun (l, p) (l', p') => l = l' /\ can_flow l p p'.
\end{code}

In our thermostat example, we express the flow function as the product of two flow functions on |CR|:

\begin{code}
  Definition thermo_flow (l: Location): Flow (pow CR 2)
    := product_flow (clock_flow l) (temp_flow l).
\end{code}
Here, |product_flow| has type |forall X Y, Flow X -> Flow Y -> Flow (prod X Y)|, and includes the trivial proofs showing that flow in a product space can be formed by combining flows in the respective components.

Intuitively, |clock_flow| and |temp_flow| are the functions
\begin{code}
Definition clock_flow (l: Location) (c: CR) (d: Duration): CR :=
  c + d.

Definition temp_flow (l: Location) (t: CR) (d: Duration): CR :=
  match l with
  | Heat => t + 2 * d
  | Cool => t * exp (-d)
  | Check => t * exp(- frac 1 2 * d)
  end.
\end{code}
However, to define the system, we actually require |Flow| records carrying additional proofs that these functions are proper flow functions. While these proofs could be given in an ad-hoc fashion for this thermostat system, we have instead developed a modest library of reusable flow functions and adaptors that let one compose flow functions that are proper by construction. Using these, |clock_flow| and |temp_flow| are defined as:
\begin{code}
Definition clock_flow (l: Location): Flow CRasCSetoid :=
  flow.positive_linear

Definition temp_flow (l: Location): Flow CRasCSetoid :=
  match l with
  | Heat => flow.scale 2 flow.positive_linear
  | Cool => flow.decreasing_exponential
  | Check => flow.scale (frac 1 2) flow.decreasing_exponential
  end.
\end{code}

\subsection{Discrete Transitions}
\label{conc.disc_trans}

Where continuous transitions describe the flow of continuous variables, \emph{discrete} transitions between locations describe the logic of the software system. Each such transition is comprised of two components: a \emph{guard} predicate, and a \emph{reset} function. The former defines a subset of the continous state space in which the transition is enabled (permitted), while the latter describes an instantaneous change applied as a side effect of the transition, as seen in the following definition of the discrete transition relation:

\begin{code}
  Definition disc_trans: relation State := fun (l, p) (l', p') =>
    guard (l, p) l' /\ reset l l' p = p' /\
    invariant (l, p) /\ invariant (l', p').
\end{code}

\todo{we don't support multiple discrete transitions between a pair of locations}

The thermostat we are modeling has four transitions, as expressed by the following guard:

\begin{code}
Definition thermo_guard (s: State) (l: Location): Prop :=
  match location s, l with
  | Heat, Cool => 9 <= temp s
  | Cool, Heat => temp s <= 6
  | Heat, Check => 2 <= clock s
  | Check, Heat => frac 1 2 <= clock s
  | _, _ => False
  end.
\end{code}

Our reset function resets the clock for all but one of these transitions, and leaves the temperature variable as is:

\begin{code}
Definition thermo_reset (l l': Location) (p: Point): Point :=
  ( match l, l' with
    | Cool, Heat | Heat, Check | Check, Heat => 0
    | _, _ => fst p
    end
  , snd p).
\end{code}
Here we can discern a conceptual distinction between continuous variables directly controlled by the system (such as the clock in our thermostat), and variables that model a physical phenomenon (such as the temperature in our thermostat). This distinction is not made explicit in the definition of a hybrid system; in principle, nothing is stopping the thermostat from treating temperature as a variable of the former kind and resetting it to whatever value it pleases. However, this would simply make the system unimplementable.

\subsection{Traces and Safety}

A transition is either continuous or discrete:
\begin{code}
  Definition trans: relation State := disc_trans `union` cont_trans.
\end{code}
% we can't actually do this in the code, but it's plenty clear what this means

We now say that a state |s| is \emph{reachable} if there is an initial state |i| from which one can, by a finite number of transitions, end up in |s|:
%format trans_refl_closure (x) = x "^{*}"
\begin{code}
  Definition reachable (s: State): Prop :=
    exists i: State, initial i /\ trans_refl_closure trans i s.
\end{code}
Here, |trans_refl_closure trans| is the transitive reflexive closure of |trans|.

As mentioned before, the set of reachable states represent the possible behaviors exhibited by a hypothetical real-world implementation of the hybrid system (as software running on a device with sensor and actuators).

The idea now is that the purpose of a hybrid system is typically to keep the continuous variables within certain limits. In other words, to limit the set of reachable states to some ``safe'' subset of the complete state space. For our thermostat example, the intent is to keep the temperature above 4.5 degrees at all times, and so the safe states are defined to be those in which the temperature component is |> 4.5|:

\begin{code}
  Definition unsafe_thermo_state: Ensemble State :=
    fun s => temp s <= 4.5.
\end{code}
(An |Ensemble T| is just a |T -> Prop|.)

The goal, then, is to verify that the reachable states are a subset of the safe states. Hence, for the thermostat, our main theorem is the following:
\begin{code}
  Theorem ThermoSafe: unsafe_thermo_state `subsetof` unreachable.
\end{code}
``|A `subsetof` B|'' is just notation for ``|forall x, A x -> B x|'', and |unreachable| is merely the complement of |reachable|.

Since we required that the invariant held at each initial state, and further defined the continous and discrete transitions such that the invariant had to hold everywhere along the path, a simple induction proof shows that
\begin{code}
  reachable `subsetof` invariant.
\end{code}
\todo{Who cares? Where do we use this?}

Before we continue with our presentation of the abstraction method, we first take a moment to discuss some of the limits on decidability of properties of computable real numbers, as their consequences can be seen in many places in the development. Indeed, we already saw one such instance in the stability requirement for invariants.

% The check location seems pretty weird, as if it's only purpose is to let the system potentially fail.

% Alur starts by defining linear hybrid systems as tuples consisting of
%   * a convex polyhedron subset X of R^n representing the continuous statespace
%   * a finite set of locations
%   * a set X_0 subsetof L * X of initial states (where states are pairs containing a location and a continuous state ("point"))
%   * a function I: L -> C_n giving a linear invariant predicate for each location
%   * a function f: L -> (X * blabla flow function.
%   * a relation T subsetof L*L*C_n*yada.

\section{Double Negation and Stability}
\label{dn}

%format le_lt_dec_overlap = le_lt_dec"_{"overlap"}"
%format le_lt_dec_DN = le_lt_dec"_{DN}"

One obvious and useful property we can derive is transitivity of the continuous transition relation, the proof of which reveals the need for invariant stability. Suppose we have |cont_trans (l, p) (l', p')| and |cont_trans (l', p') (l'', p'')| for locations |l, l', l''| and points |p, p', p''|. To show |cont_trans (l, p) (l'', p'')|, we must show two things. The first, |l = l''|, follows immediately from transitivity of |=|. For the second, |can_flow l p p''|, we simply take the flow duration to be the sum of the durations from |p| to |p'| and from |p'| to |p''| (call these |d| and |d'|, respectively), and observe:
\begin{align}
  |flow l p (d + d')| & = |flow l (flow l p d) d'| \nonumber \\
    & = |flow l p' d'| \nonumber \\
    & = |p''| \nonumber
\end{align}
What remains is to show that the invariant holds at each point along the way. That is,
\begin{code}forall t, 0 <= t <= d + d' -> invariant (l, flow l p t).\end{code}
From |can_flow p p'| we know this is true for |0 <= t <= d|, and from |can_flow p' p''| we know that this is true for |d <= t <= d + d'|. Classically, then, the proof is a done deal, for one can simply distinguish cases |t <= d| and |d < t|.

Unfortunately, such case distinction is a luxury we do not have, because for computable reals, the proposition
\begin{code}
  le_lt_dec: forall x y: CR, x <= y \/ y < x.
\end{code}
is not provable. After all, to have a constructive proof of a disjunction |A\/B| (where A and B are arbitrary propositions/types) is to have either a proof of |A| or a proof of |B|. Similarly, to have a proof of |X -> A \/ B| is to have a function that, given an |X|, either returns a proof of |A|, or a proof of |B|. With this in mind, suppose we try to implement |le_lt_dec|. Then given |x| and |y| in |CR|, we are to produce a proof either of |x <= y| or of |y < x|. Unfortunately, the nature of computable reals only lets us observe arbitrarily close approximations of |x| and |y|. Now suppose |x = y|. Then no matter how closely we approximate |x| and |y|, the error margins (however small) will always leave open the possibility that |y| is really smaller than |x|. Consequently, we will never be able to definitively conclude that |x <= y|.

Computable reals do admit two variations on the proposition:
\begin{enumerate}
\item |le_lt_dec_overlap: forall x y: R, x < y -> forall z, z <= y \/ x <= z|
\item |le_lt_dec_DN: forall x y: R, ~ ~ (x <= y \/ y < x)|
\end{enumerate}

Both are weaker than the original, and are less straightforward to use. Nevertheless, this is the path we will take in our development, partly out of necessity (because in some parts of the development, we really need to ``run'' these lemmas to obtain |<=|/|<| proofs), and partly because just taking |le_lt_dec| as an axiom amounts to cheating.

For our transitivity proof, we will use the variant expressed using double negation. Two questions immediately arise when considering this variant. First, why does the double negation make it provable? And second, how does one actually use this doubly negated variant in proofs?

For the first, we need only observe that |x <= y| is equivalent to |y </ x|, and that the law of the excluded middle holds under double negation. That is,
\begin{code}
  forall P, ~ ~ (P \/ ~ P)
\end{code}
is a trivial tautology considering that |~ P| is taken to mean |P -> False|.

One practical way to answer the question of how such a doubly negated proposition might be used to prove things is to observe that double negation, as a function on types/propositions, is a monad \cite{bla}. Writing |DN P| for |~ ~ P|, we have the following two key operations that make |DN| a monad:

%format return_DN
%format bind_DN

\begin{code}
  return_DN: forall A, A -> DN A
  bind_DN: forall A B, DN A -> (A -> DN B) -> DN B
\end{code}
The first expresses that any previously obtained result can always be inserted ``into'' the monad. The second expresses that results inside the monad may be used freely in proofs of additional properties in the monad. For instance, one may |bind_DN| a proof of |DN (x <= y \/ y < x)| (obtained from |le_lt_dec_DN| above) with a proof of |(x <= y \/ y < x) -> DN P|, yielding a proof of |DN P|.

Thus, |DN| establishes a ``proving context'' in which one may make use of lemmas yielding results inside |DN| that may not hold outside of it (such as |le_lt_dec_DN|), as well as lemmas yielding results not in |DN|, which can always be injected into |DN| using |return_DN|. The catch is that such proofs always end up with results in |DN|, which begs the question: what good is any of this? In particular, can |le_lt_dec_DN| be used to prove anything not doubly negated?

As it happens, there is a class of \emph{stable} propositions that are equivalent to their own double negation. Examples include negations, non-strict inequalities on real numbers, and any decidable proposition.

We now see why we required invariant stability in section \ref{concrete}: in the transitivity proof for |cont_trans|, it allows us to employ |le_lt_dec_DN| to do case distinction on the |t| variable when showing that the invariant holds at each point along the composite path. That is, we simply bind |le_lt_dec_DN t d| of type |DN (t <= d \/ d < t)| with the straightforward proof of |(t <= d \/ d < t) -> DN (invariant (l, flow system l p t))|, and then pull the latter out of |DN| on account of its stability.

Invariants are typically conjunctions of inequalities, which are stable only if the inequalities are non-strict. Hence, the limits on observability of computable real numbers ultimately mean that our development cannot cope with hybrid systems whose location invariants use strict inequalities. We feel that this is not a terrible loss. In section \ref{abstraction} we will see analogous limitations in the choice of one's abstraction parameters.

% Since in our development we do not want to use any axioms,
%  Since in development we actually need to \emph{run} these functions, we have no choice but to 
% In some cases, the former is actually a perfectly usable substitu
% In the first alternative, we change the problem from "deciding whether one number is less than another" to "deciding whether, given two numbers 
% because the error margins of our approximations will always
% we will never be able to decisively conclude |x <= y|, because the possibility that
%  can never be excluded from these approximations.
% but with a difference smaller .
% Writing x <= y as ~ y < x

\section{Underestimation and Overestimation}
\label{estimation}

% It \emph{will} always terminate, so we are not writing a semidecision procedure either.
Ultimately, in our development we are writing a program that \emph{attempts} to produce hybrid system safety proofs. Importantly, we are \emph{not} writing a complete hybrid system safety decision procedure: if the concrete system is unsafe or the abstraction method fails, our program will simply not produce a safety proof. It might seem, then, that we are basically writing a \emph{tactic} for a particular problem domain. However, tactics in Coq are normally written in a language called Ltac, and typically rely on things like pattern matching on syntax. Our development, on the other hand, is very much written in regular Gallina, with hardly any significant use of Ltac. This was never a conscious design decision though---it is just the way the development let itself be written. \todo{While I know too little about complex use of Ltac to know whether heavier use of it might have been useful, there should definitely be more to say about this.} In any case, to characterize tactic-like functions in regular Gallina, we define |underestimation P| to be either a proof of |P|, or not. In Coq, either of the following will do:
\begin{code}
Definition underestimation (P: Prop): Set := option P.
Definition underestimation (P: Prop): Set := { b: bool | b = true -> P }.
\end{code}
The latter tends to work better with the |Program| family of commands \cite{bla} which have special support for dependent pairs. Using |Program|, an underestimation of the second variety may be provisionally defined strictly as a bool, and then separately proved to be a valid underestimation in a proof obligation generated by |Program|. The second form also nicely illustrates why we call this an underestimation: it may be |false| even when |P| holds. We can now describe the functionality of our program by saying that it underestimates hybrid system safety, yielding a term of type |underestimation Safe|, where |Safe| is a proposition expressing safety of a hybrid system.

Considered as theorems, underestimations are not very interesting, because they can be trivially ``proved'' by taking the |false|/|None| estimation. Hence, the value of our program is not witnessed by the mere fact that it manages to produce terms of type |underestimation Safe|, but rather by the fact that when run, it actually manages to returns |true|/|Just P| for the hybrid system we are interested in (i.e. the thermostat). Again, this is typical tactic stuff. It is for this reason that we primarily think of the development as a program rather than a proof, even though the program's purpose is to produce proofs.

\

\noindent The opposite of an underestimation is an overestimation:
\begin{code}
Definition overestimation (P: Prop): Set := { b: bool | b = false -> ~ P }.
\end{code}
(Here, too, one could use an option type: |option (~ P)|.)

Since hybrid system safety is defined as unreachability of unsafe states, we may equivalently express the functionality of our development by saying that it overestimates unsafe state reachability. Indeed, most subroutines in our programs will be overestimators rather than underestimators. Notions of overestimation and underestimation trickle down through all layers of our development, down to basic arithmetic. For instance, we employ functions such as:

%format CRle_unapplied = <= "_{" CR "}"
%format CRle a b = a <= "_{" CR "}" b
%format Qpos = "\mathbb{Q}^{+}"
%format overestimate_CRle = overestimate "_{" CRle_unapplied "}"

%format epsilon = "\epsilon "

\begin{code}
overestimate_CRle (epsilon: Qpos): forall x y: CR, overestimation (CRle x y)
\end{code}
As discussed in the last section, |CRle_unapplied| is not decidable. |overestimate_CRle| merely makes a ``best effort'' to prove |~ (CRle x y)| using |epsilon|-approximations. A smaller |epsilon| will result in fewer spurious |true| results.

%\todo{|CRle_unapplied| can also be underestimated, right?}

%format A_0
%format A_n
%format a_0
%format a_n

The types of underestimators and overerestimators often merely repeat a certain predicate's parameters, as in the |overestimate_CRle| example above. Using a bit of type class magic to achieve variadicity, we can rid us of this repetition. Given a predicate |P: A_0 -> ... -> A_n -> Prop|, we define |underestimator P| as the type \begin{code}forall (a_0: A_0) ... (a_n: A_n), underestimation (P a_0 ... a_n).\end{code} (And similar for |overestimator|.) With these, the type of |overestimate_CRle| may be written as |Qpos -> overestimator (CRle_unapplied)|.

Underestimators and overestimators can be combined to form underestimators (resp. overestimators) for things like conjunctions and quantifications over finite domains. Applying these combinators by hand is somewhat tedious, and for this reason we have made |overestimation| and |underestimation| type classes. We will not go into the details, but the net effect is that whenever we need an overestimation of some proposition |P|, we may say |overestimate P|, and Coq's type class resolution will try to construct an overestimation for |P| based on a collection of type class instances that amount to a database of registered overestimator combinators and primitives.

Finally, while we're on the subject of decision-procedure-like things, we note that we use the same techniques for real decision procedures (in the few places we use those): we define |decision P| as |{P}+{~P}|, and given a predicate |P: A_0 -> ... -> A_n -> Prop|, we define |decider P| as the type |forall (a_0: A_0) ... (a_n: A_n), decision (P a_0 ... a_n)|. And here, too, we make |decision| a type class, so that we may simply say |decide P| and have Coq use type class resolution to try and construct a decider deciding |P|.

%\todo{Things like the Qpos above throw dirt in the food though. We could make a type class for ``epsilon-parameterized overestimators'', but meh.}

\section{Abstraction: Regions, States, and Spaces}
\label{abstraction}

The abstraction method for verification of (concrete) hybrid system safety as we implement it can be summarized as follows:
\begin{enumerate}
\item build an abstract hybrid system corresponding to the concrete hybrid system;
\item \label{steptwo} show that each trace in the concrete system corresponds to some trace in the abstract system (and as a corollary, that reachability in the concrete system implies reachability in the abstract system, and most importantly, that unreachability in the abstract system implies unreachability in the concrete system);
\item \label{stepthree} run a certified graph reachability algorithm on the (finite, discrete) abstract system to verify that no unsafe \emph{abstract} states are reachable;
\item conclude from \ref{steptwo} and \ref{stepthree} that no unsafe \emph{concrete} states are reachable either.
\end{enumerate}

An abstract system corresponding to some concrete hybrid system is ``like'' the concrete system, but with the continuous state space replaced with a finite set of \emph{regions}, each corresponding to a subset of the continuous state space. We begin by describing these regions, after which we will describe abstract transitions and reachability.

\subsection{Regions and States}
\label{regionsandstates}

Whereas in a concrete hybrid system states consist of a location paired with a point in the continuous state space, in an abstract hybrid system states consist of a location paired with the ``name'' of a region corresponding to a subset of the continuous state space:

\begin{code}
  Definition abstract.State := prod Location Region.
\end{code}
From now on we will use a ``|concrete.|'' prefix for names like |State| defined in section \ref{concrete}, which now have abstract counterparts. |Region| is a field from a record type |Space| bundling region sets with related requisites:
\begin{code}
  Record Space: Type :=
    { Region: Set
    ; Region_eq_dec: EqDec Region eq
    ; regions: ExhaustiveList Region
    ; NoDup_regions: NoDup regions
    ; in_region: Container Point Region
    ; in_region_mor: Morphism (cs_eq ==> eq ==> iff) in_region
    ; regions_cover: forall (l: Location) (p: Point),
        invariant (l, p) -> DN { r: Region | p `elem` r }
    }.
\end{code}
The |Container Point Region| type specified for |in_region| reduces to |Point -> Region -> Prop|. |Container| is a type class that provides the notation ``|x `elem` y|'', prettier than ``|in_region x y|''. |in_region_mor| states that |in_region| respects |Point|'s setoid equality.

|regions_cover| expresses that each concrete point belonging to a valid state must be represented by a region---a crucial ingredient when arguing that unreachability in the abstract system implies unreachability in the concrete system. The double negation in its result type is both necessary and sufficient:

It is \emph{necessary} because |regions_cover| boils down to a (partial) function that, given a concrete point, must select an abstract region containing that point. This means that it must be able to decide on which side of a border between two regions the given point lies. As we saw in section \ref{dn}, that kind of decidability is only available inside |DN| unless all region borders have nontrivial overlap, which as we will see later is undesireable.

Fortunately, the double negation is also \emph{sufficient}, because we will ultimately only use |regions_cover| in a proof of of |... -> ~ concrete.reachable s| (for some universally quantified variable s), which, due to its head type being a negation, is stable, and can therefore be proved in and then extracted from |DN|. Hence, we only need |regions_cover|'s result in |DN|.

\

We can lift the containment relation between points and regions to a containment relation between concrete states and abstract states:
\begin{code}
  Instance abs: Container concrete.State abstract.State
    := fun (l, p) (l', r) => l = l' /\ p `elem` r.
\end{code}
Again, the |Container| application reduces to |concrete.State -> abstract.State -> Prop|, but the use of |Container| lets us say |s `elem` s'| instead of |abs s s'|. \todo{Coq trunk has nameless instances, I believe, which would let us dispense with the pointless |abs| name altogether.}

%format inf = "\infty "
%format closesquare = "\hspace{-1mm}] "

% Suppose one has a very simple concrete system with the trivial invariant |True| and only a single continuous variable in |CR|. And suppose one wishes to abstract this system using two regions, corresponding to the ranges |(inf, 2 closesquare| and |(2, inf)|, respectively. Then implementing |regions_cover| amounts to deciding, for any |x|, whether |x <= 2| or |2 < x|. As discussed in section \ref{bla}, bla. And as before, the solution involved the double negation monad |DN|.

% Unfortunately, the constructive real number representation we use does not feature decidability of this particular query, because they fundamentally only let one observe arbitrarily close approximations, and if |x| happens to be 2, no approximation of it can decisely determine whether |x <= 2| or |2 < x| (because 2 will always lie within the error range of the approximation).

% One solution to this problem is to make sure one's abstract regions have nontrivial overlap. For example, one could instead have regions corresponding to |(inf, 2.01)| and |(2, inf)|. An approximation of |x| with error margin 0.05 would then let one find at least one range that definitely contains |x|. Unfortunately, as we will see later, region overlap can result in undesireable transitions (and therefore traces), which can adversely affect the abstract system's utility. We therefore use a different solution, which is where |DN| comes in.

% As it turns out, there are propositions that are not provable themselves, but whose double negation \emph{is} provable. The proposition |x <= 2 \/ 2 < x| is an example of such a proposition. Defining |DN P| as |~ ~P|, |DN (x <= 2 \/ 2 < x)| is provable (for any |x|). This |DN| is a monad, which means that one can reason ``inside'' of it, deriving ever more doubly negated propositions, possibly making use of lemmas/functions with result types also in |DN|. Such lemmas/functions are ``weak'' in the sense that their result can only be used in proofs inside the |DN| monad.

% Now, there exist propositions |P| for which |DN P| is equivalent to |P| (these are called ``stable'' propositions). One important example are propositions of the form |~ P|. To prove such a proposition, one may do so inside the |DN| monad, making use of propositions which are only true inside |DN|, until one has derived |DN P|, at which point |P| may be extracted.

% In our hybrid system development, we willl ultimately only use |regions_cover| in a proof of |... -> ~ concrete_reachable s| (for some universally quantified variable s), which, due to its head type being a negation, is stable. Hence, this proof can be performed inside the DN monad, which means it is sufficient for |regions_cover| to have the weaker |DN| result, which in turn means it may make use of DN-only lemmas such as the trichotomy property for |CR|, which in turn means we need not resort to non-trivial region overlap.

% The result type of |regions_cover|, |DN (sig (in_region p))|
% On the other hand, |regions_cover|

\subsection{Abstract Space Construction}
\label{abs.space.constr}

When building an abstract system, one is in principle free to divide the continuous state space up whichever way one likes. However:
\begin{itemize}
\item if the regions are too fine-grained, there will have to be very many of them to cover the continuous state space of the concrete system, resulting in poor performance;
\item if the regions are too coarse, they will fail to capture the subtleties of the hybrid system that actually make it safe (if indeed it is safe at all);
\item careless use of region overlap can result in undesireable abstract transitions (and therefore traces), adversely affecting the abstract system's utility (as we will discuss in detail in section \ref{trouble}).
\end{itemize}

Like |Alur|, we use regions formed by multiplying intervals on individual continuous variables. That is, for the thermostat, we first define a |Space| partitioning the continous state space into regions corresponding to clock intervals (which in the development amounts to enumerating the interval bounds, thanks to some generic space definition utilities we wrote), then define a |Space| partitioning the continuous state space into regions corresponding to temperature intervals, and then take the product of these to obtain a |Space| where regions correspond to ``squares'' in the continuous state space.

\todo{Mention optional unbounded-ness? Of no theoretical interest, but makes implementation a bit more work.}

In \cite{alur}, Alur describes a heuristic for interval bound selection, where the bounds are taken from the constants that occur in the invariant, guard, and safety predicates. For the thermostat, we initially attempted to follow this heuristic and use the same bounds Alur uses, but found that due to our use of computable reals, we had to tweak the bounds somewhat to let the system successfully produce a safety proof. We give one example of why such tweaking was required.

Following the heuristic, Alur derives the following two regions:
%format r_0
%format r_1
%format p_0
%format p_1
\begin{code}
  r_0 := { (c, t): Point | c <= 0 /\ 5 <= t <=6 }
  r_1 := { (c, t): Point | 2 <= c <= 3 /\ 6 < t < 9 }
\end{code}

In the |Heat| location, where both the clock and the temperature increase linearly, the latter twice as fast as the former, there is a concrete continuous transition from |p_0 := (0, 5) `elem` r_0| to |p_1 := (2, 9)|. |p_1| is not an element of |r_1|, and since the floating point representation Alur uses presumably easily lets one conclude that |5+2*2 </ 9|, Alur's procedure manages to determine that there is no continuous transition from any point in |r_0| to a point in |r_1|, justifying suppression of any abstract continuous transition from |(Heat, r_0)| to |(Heat, r_1)| (we will discuss abstract transitions in detail in the next section).

Unfortunately, our computable reals only let us compute arbitrarily close approximations of |5+2*2|. 
Consequently, in the process of determining whether there ought to be an abstract continuous transition from |(Heat, r_0)| to |(Heat, r_1)|, our procedure (which we will describe in \ref{over_cont_trans}) will fail to conclude with certainty that this value does not lie below 9, will thus be unable to definitively establish the absence of concrete continuous transitions from points in |r_0| to points in |r_1|, and will be forced to include the transition, to ensure that the abstract transition relation respects its concrete counterpart (in a way discussed in detail in the next section).

As it happens, the thermostat's safety indirectly depends on unreachability of |(Heat, r_1)|, and since |(Heat, r_0)| is a reachable state, the addition of the transition mentioned above makes the abstract system unsafe, obviously preventing us from using it to conclude safety of the concrete system.

% Hence, by the fallback mechanism we will describe in \ref{bla}, the procedure \emph{will} default to emitting an abstract continuous transition from |(Heat, r_0)| to |(Heat, r_1)|.

To correct this kind of problem, we tweaked many of the bounds, nudging them slightly toward one side or the other. For this particular example, we changed the 9 bound to 8.9. That way, the comparison reduces to |5+2*2 </ 8.9|, which a sufficiently close approximation \emph{can} automatically establish.

\todo{Mention that the epsilon we use in computation to determine whether there's a transition from some point to another (and in the side conditions and box comparisons and such) is a parameter of the abstraction procecudure. for the actual safety proof, we take this epsilon to be 0.001, which makes the 0.1 nudge in things like 8.9 large enough.}

\todo{Note that since the abstract system is only an overestimation of the concrete system, its unsafety does not imply unsafety of the concrete system.}

\todo{Emphasize that changing the abstraction interval bounds does not mean our result isn't as "pure" as Alur's, because the whole abstraction business only serves to prove concrete system safety, and we do prove safety of the same concrete system as Alur.}

\todo{Having to do this "nudging" manually is obviously not idea. Can we think of heuristics for this nudging? Could be that it's just that this thermostat example is rotten though..}

\

% The main idea of abstraction is to come up with a finite set of "regions", each corresponding to a subset of the continuous state space, and to analyze traces through these regions instead of traces through points.

Another way in which our thermostat regions differ from |Alur|'s lies in the fact that our bounds are always inclusive, which means adjacent regions overlap in lines. We will discuss this in more detail later.

 %(* todo: hide the nodup part? *)
% todo: embed AState := ?

\todo{mention that the reason the space definition could be mostly automated is that |regions_cover| is essentially a whole bunch of if-then-else's (which are only constructive inside |DN|)}

\section{Abstract Transitions and Reachability}
\label{abs.reach}

% hm, where shoudl we introduce alternating stuff?

Once we have a satisfactory abstract |Space|, our goal is to construct an overestimatable notion of abstract reachability implied by concrete reachability, so that concrete unreachability results may be obtained simply by executing the abstract reachability overestimator. It seems reasonable, then, to look for an overestimatable predicate |abstract.reachable| such that |reachable_respect| holds:
\begin{code}
Definition reachable_respect: Prop :=
  forall (s: concrete.State), concrete.reachable s ->
    forall (s': abstract.State), s `elem` s' -> abstract.reachable s'.
\end{code}
After all, this would imply
\begin{code}
forall (s: concrete.State) (s': abstract.State),
   s `elem` s' -> ~ abstract.reachable s' -> ~ concrete.reachable s,
\end{code}
expressing that to conclude unreachability of a concrete state, one need only establish unreachability of an abstract state that contains it. As we will see shortly, the above definition of |reachable_respect| leads to problems down the line, so this is not actually the definition we use in the development. However, in order to see the motivation for the actual definition we use, let us proceed as if without foresight for a few more moments.

\subsection{The Straightforward (but Inadequate) Approach}
\label{straightforward}

The definition of |reachable_respect| above suggests that the property we would wish to overestimate is simply
\begin{code}
Definition abstract.reachable (s: abstract.State): Prop
  := exists s': concrete.State, s' `elem` s /\ concrete.reachable s'.
\end{code}
In order to overestimate it, we could introduce the following obvious abstract versions of initiality and continuous and discrete transitions:
\begin{code}
  Definition abstract.initial (s: abstract.State): Prop
    := exists c `elem` s, concrete.initial c.
  Definition abstract.cont_trans: relation abstract.State
    := fun s s' => exists (c `elem` s) (c' `elem` s'), concrete.cont_trans c c'.
  Definition abstract.disc_trans: relation abstract.State
    := fun s s' => exists (c `elem` s) (c' `elem` s'), concrete.disc_trans c c'.
\end{code}
We could then prove (by induction over traces) that |abstract.reachable| is included in the transitive closure of |abstract.cont_trans `union` abstract.disc_trans| starting at |abstract.initial|:\footnote{One would actually take the \emph{alternating} transitive closure, but we will get to that later.}

\begin{code}
  Definition abstract.trans: relation abstract.State :=
    abstract.cont_trans `union` abstract.disc_trans.

  Definition reachable_by_abstract_trace (s: abstract.State): Prop
    := exists i, abstract.initial i /\ trans_refl_closure (abstract.trans) i s.

  Lemma: abstract.reachable `subsetof` reachable_by_abstract_trace.
\end{code}
This would show that |abstract.reachable| could be overestimated by overestimating |reachable_by_abstract_trace|. For the latter, we would first define an \emph{abstract system} as a triple containing overestimators for |abstract.initial|, |abstract.cont_trans|, and |abstract.disc_trans|:
\begin{code}
Record abstract.System: Type :=
  { over_initial: overestimator abstract.initial
  ; over_cont_trans: overestimator abstract.cont_trans
  ; over_disc_trans: overestimator abstract.disc_trans
  }.
\end{code}
Given an instance of this record, we would then overestimate |reachable_by_abstract_trace| by constructing a graph with vertices representing abstract states and edges representing overestimated abstract transitions, and running an ordinary (but verified) graph reachability algorithm on this graph.

Thus, the task would be reduced to construction of the three overestimators.

\subsection{Clouds on the Horizon: Drift}
\label{trouble}

All of the above seems perfectly reasonable, but having gotten to the level of abstract transition overestimators, we can now see where things would (and did, in early versions of our development) go wrong. Unfolding the |overestimator| type specified for the |over_cont_trans| member in |abstract.System|, we get
\begin{code}
  forall (s s': abstract.State), { b: bool | b = false -> ~ abstract.cont_trans s s' }.
\end{code}
Hence, |over_cont_trans| would need to satisfy:
\begin{code}
  forall (s s': abstract.State), abstract.cont_trans s s' ->
    abstract.cont_trans_over s s' = true
\end{code}
Which rewrites to:
\begin{code}
  forall (c c': concrete.State), concrete.cont_trans c c' ->
  forall (s s': abstract.State), c `elem` s) -> c' `elem` s' ->
    over_cont_trans s s' = true
\end{code}
In other words, the existence of a (continuous) transition from one concrete state to another would force inclusion of transitions between any two abstract states that contain the respective concrete states. When regions do not overlap, this is perfectly appropriate. However, consider the implications in the following example involving overlapping regions.

Suppose our continuous state space is |CRge0|, and our abstract regions are intervals of the form |[n, n+1]| with |n `elem` nat|. Further suppose that in some location |l|, the flow function monotonically increases. Then by the above, |over_cont_trans| must yield |true| given |(l, [1, 2])| and |(l, [0, 1])| (in that order), because in |l|, there actually is a concrete continuous transition from a point in the former to a point in the latter, namely from 1 to 1 (since the continuous transition relation is reflexive). Hence, we would get an abstract transition from |(l, [1, 2])| to |(l, [0, 1])|, essentially introducing abstract flow in the opposite direction of the concrete flow. This can clearly have disastrous consequences for unreachability of abstract states containing unsafe states, which can render the abstraction useless.

Intuitively, the transition from |(l, [1, 2])| and |(l, [0, 1])| seems \emph{redundant}, because the only point in |[0, 1]| that can be flowed to from |[1, 2]| is a point that is actually still in |[1, 2]|. Making |[0, 1]| reachable on account of this redundant transition seems wasteful. Let us formalize this notion of redundance:
\begin{code}
  Definition redundant_cont_trans (l: Location): relation Region
    := fun r r' => forall p, p `elem` r -> forall t, 0 <= t -> flow l p t `elem` r' -> flow l p t `elem` r.
\end{code}
This relation can be lifted to abstract states. An analogous definition can be given for abstract discrete transitions.

\todo{Discuss strong redundancy which holds over space composition.}

We call the phenomenon of redundant transitions being generated by abstract transition overestimators: ``drift''. With the specifications for abstract transition overestimators given in the previous section, drift is an inescapable consequence when regions overlap, occurring not only for continuous abstract transitions (as shown above), but also for discrete abstract transitions (as will be discussed in section \ref{over_disc_trans}).

It is important to note that drift is not just a consequence of ill-chosen specifications or regions. Suppose we use strictly non-overlapping regions. While this would solve the problem at the specification level (in that the specification would no longer force the overestimators to produce redundant transitions), the overestimator \emph{implementations} would still be bound to yield redundant transitions when no specific countermeasures are employed. As a trivial example, consider non-overlapping regions [0,1) and [1,2), again with monotonically increasing flow. To justify omission of an abstract continous transition from [1,2) to [0,1), an overestimator implementation would need to prove that |1 </ 1|, which is not decidable or meaningfully underestimatable given only arbitrarily close approximations of 1.

Hence, our drift countermeasures will be twofold. First, we will first ensure that at the \emph{specification} level, transition overestimators are no longer forced to emit redundant transitions. Then, in sections \ref{over_cont_trans} and \ref{over_disc_trans}, we show how our overestimator \emph{implementations} exploit this freedom and use a redundancy hint mechanism to detect and omit redundant transitions.

\todo{Describe somewhere how things go \emph{completely} wrong when regions have surface overlap.}

\subsection{Specification Drift Avoidance: Sharing Overestimators}

To avoid drift at the specification level, we must either alter the specifications of abstract transition overestimators to account for redundancy (which will necessitate analogous changes in the specifications of abstract reachability and respect, as we will see in a moment), or avoid region overlap.

Practically speaking, using non-overlapping regions implies the use of non-strict equalities, as in the example above. As it happens, non-strict and strict constructive inequalities on real numbers live in different universes: the former live in |Prop|, while the latter live in |Type| \todo{explain why}. Since derived propositions (such as |p `elem` r| for a point |p| and a region |r|) all inherit this trait, we would essentially have to give up on |Prop| for the entire development. Unfortunately, several components in the Coq system, including the highly useful subset coercion functionality of the |Program| family of commands, currently only support |Prop|. While these are strictly engineering concerns that may or may not have theoretical substance, they do matter when developing a working system.

On the other hand, changing the abstract transition overestimator specification, abstract reachability, and respect specifications, turns out to be a relatively local change, affecting only a few hundred lines of code, most of them in the one module that introduces these concepts and proves some key lemmas about them.

From an engineering perspective, then, the choice is easy. We will now describe the changes in abstract transition overestimator specification in detail.

% todo: The continuous hint mechanism only works with border-overlap.
% todo: how about making abstraction bounds rationals?
% Drift at the specification level (as discussed in the previous section) \emph{can} be avoided by restricting oneself to abstract spaces whose regions do not overlap, but drift at the implementation level cannot. 
% blabla evil---even though we're using alternating transitions.
% Discrete abstract transitions suffer the same fate, but in a slightly

\

\noindent In the discussion of the first example in the previous section, we judged the redundant abstract transition from |(l, [1, 2]|) to |(l, [0, 1])| to be ``wasteful'' because the destination of the sole concrete continuous transition from 1 to 1 that spawned it was already covered by |[1, 2]|, and so if |[1, 2]| is reachable, there should be no reason to make |[0, 1]| reachable as well. Implicit in this intuition is the idea that regions ought to \emph{share} the burden of being reachable on behalf of the points they represent: if 1 is a reachable concrete point, then reachability of |[1, 2]| should remove the need for reachability of |[0, 1]| (to represent reachability of 1), and vice versa.

Unfortunately, as shown in the previous section, the overestimator specifications used in the definition of |abstract.System| given in section \ref{straightforward} do not permit this sharing. Let us consider how they would need to be changed to allow for it. We can easily make the unfolded specification of |over_cont_trans| shown in the previous section permit sharing by substituting one universal quantifier with an existential one:
\begin{code}
  forall c c', concrete.cont_trans c c' ->
  forall (s: abstract.State), c `elem` s ->
  exists (s': abstract.State), c' `elem` s' /\
    over_cont_trans s s' = true
\end{code}
Applied to the example above, the existence of a concrete continuous transition from 1 to 1 now only results in the requirement that
\begin{code}
  forall (s: abstract.State), (l, 1) `elem` s ->
  exists (s': abstract.State), (l, 1) `elem` s' /\
    over_cont_trans s s' = true
\end{code}
This means |over_cont_trans| now has the freedom to return |false| given |(l, [1, 2])| and |(l, [0, 1])| (thereby suppressing creation of that abstract transition) if it returns |true| given |(l, [0, 1])| and |(l, [0, 1])| (which it will have to anyway).

To give |over_cont_trans| and |over_disc_trans| new types that embody this new specification, to replace the old |overestimator| types, we first introduce |shared_cover|:
\begin{code}
  Definition shared_cover
    `{Container concrete.State C} `{Container abstract.State D}
      (cs: C) (ss: D): Prop :=
        forall s: concrete.State, In s cs -> DN (exists r: abstract.State, In s r /\ In r ss).
\end{code}
The details of the |Container| type class are of no interest to us right now. What matters is that a ``container'' of concrete states is said to be sharedly-covered by a ``container'' of abstract states if for each of the concrete states in the former there is an abstract state in the latter that contains it.\footnote{The double negation is here for much the same reasons as were given in section \ref{regionsandstates} for the double negation in the type of |regions_cover|.} We state this in terms of |Container|s so that the definition applies to any types which have a reasonable notion of containment.

Stated in terms of |shared_cover|, a reasonably straightforward type for |over_cont_trans| would be:
\begin{code}
over_cont_trans: forall s: abstract.State,
  { p: abstract.State -> bool |
    shared_cover (fun c' => exists c `elem` s, concrete.cont_trans c c') p }
\end{code}
That is, |over_cont_trans| would be a function that, given an abstract state, returns a boolean predicate on abstract states that, when interpreted as a container in the obvious way, covers all concrete states reachable by a single concrete continuous transition from a concrete state contained in the original abstract state. We make a few last adjustments to arrive at the actual type used in the development.

First, the abstract transition only actually needs to cover points at which the invariant holds. Integrating that fact in the type of |over_cont_trans| allows us to use |regions_cover| in its implementation:
\begin{code}
over_cont_trans: forall s: abstract.State,
  { p: abstract.State -> bool | shared_cover
    (fun c' => concrete.invariant c' /\ exists c `elem` s, concrete.cont_trans c c') p }
\end{code}
Using some |Container| utilities, we may write this in a more point-free fashion:
%format `inter` = "\cap "
%format `compose` = "\circ "
%format utilflip (a) = a "^{-1}"
%format prettysig = "\Sigma"
\begin{code}
over_cont_trans: forall s: abstract.State,
  { p: abstract.State -> bool | shared_cover
    (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p }
\end{code}
The use of a boolean predicate is less than ideal because it means that to enumerate all abstract states directly reachable from some given abstract state (which we will have to do eventually when computing reachability in a graph whose edges are built from these overestiamtors), one basically needs to filter an exhaustive list of all abstract regions using the predicate, while one could easily imagine that clever overestimator implementations could exploit locality to exclude certain classes of potential transition destinations a priori from consideration. To facilitate such overestimators (which we do not develop in this report), we replace the boolean predicate with a \emph{list} of abstract states. Since lists are also |Container|s, the rest of the definition remains unchanged:
\begin{code}
over_cont_trans: forall s: abstract.State,
  { p: list abstract.State | shared_cover
    (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p }
\end{code}
Lists may contain duplicates. The graph reachability algorithm which we will use in section \ref{graphs} relies on absence of duplicates in the edge relation (expressed as a function returning lists) for its termination proof. While we could filter out duplicates at a much later stage to satisfy that requirement, we choose to let the no-duplicates requirement propagate through to our overestimator specification, avoiding premature pessimization:
\begin{code}
over_cont_trans: forall s: abstract.State,
  { p: list abstract.State | NoDup l /\ shared_cover
    (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p }
\end{code}
The |vm_compute| tactic we will ultimately use to run our program tends to reduce terms in |Prop| a bit too eagerly, adversely affecting performance. To work around this, we introduce a bit of laziness in the form of a little unit abstraction:
\begin{code}
over_cont_trans: forall s: abstract.State,
  { p: list abstract.State | () -> (NoDup l /\ shared_cover
    (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p) }
\end{code}
This will still let us use the property in proofs, but will stop |vm_compute| from unnecessarily evaluating its proof term.

Finally, we actually need the same definition for abstract discrete transitions, so we factor out the common part:
\begin{code}
Definition sharing_transition_overestimator
  (R: relation concrete.State): Set := forall s: abstract.State,
    { p: list abstract.State | () -> (NoDup l /\ shared_cover
      (concrete.invariant `inter` (overlap s `compose` utilflip R)) p) }.
\end{code}
We can now show the definition of |abstract.System| as it appears in the development:
\begin{code}
  Record abstract.System: Type :=
    { over_initial: overestimator (overlap concrete.initial)
    ; over_disc_trans: sharing_transition_overestimator concrete.disc_trans
    ; over_cont_trans: sharing_transition_overestimator concrete.cont_trans
    }.
\end{code}

\subsection{Propagating Sharing Upwards}

Consider again the definition of |abstract.reachable| given in section \ref{straightforward}:
\begin{code}
Definition abstract.reachable (s: abstract.State): Prop
  := exists s': concrete.State, s' `elem` s /\ concrete.reachable s'.
\end{code}
Clearly, this definition does not let abstract states share the burden of reachability, and can therefore not be overestimated using our retyped abstract transition overestimators. But recall that it followed directly from the following more basic specification we had in mind:
\begin{code}
Definition reachable_respect: Prop :=
  forall (s: concrete.State), concrete.reachable s ->
    forall (s': abstract.State), s `elem` s' -> abstract.reachable s'.
\end{code}
Which we chose because it implied
\begin{code}
forall (s: concrete.State)
  (exists s': abstract.State, s `elem` s' /\ ~ abstract.reachable s') ->
    ~ concrete.reachable s,
\end{code}
expressing that to conclude unreachability of a concrete state, one need only establish unreachability of \emph{any} abstract state that contains it. However, we now see that this definition, too, neglects to facilitate sharing: when abstract states may share the burden of reachability, one should establish unreachability of \emph{all} abstract states containing the concrete state. That is, what we really want is an |abstract.reachable| satisfying:
\begin{code}
forall s: concrete.State,
  (forall s': abstract.State, s `elem` s' -> ~ abstract.reachable s') ->
  ~ concrete.reachable s.
\end{code}
\todo{Nice symmetry here, on this side of the table an existential quantifier changes to a universal one.}
This property follows from the following new definition of |reachable_respect| we will use:
\begin{code}
Definition reachable_respect: Prop :=
  shared_cover concrete.reachable abstract.reachable.
\end{code}
% This is only natural--in the example, to conclude unreachability of the concrete state |(l, 1)|, it is no longer sufficient to establish unreachability of the abstract state |(l, [0, 1])|.
This new definition does not give rise to a nice simple definition of |abstract.reachable| distinct from the simple transitive closure of initiality and transition judgements made by the overestimators (analogous to |reachable_by_abstract_trace|), so we just take the latter and prove |reachable_respect| by induction over traces. There is, however, one last subtlety involved.

%  Let us see what happens when we redefine it directly in terms of our abstract system overestimators, much like how |reachable_by_abstract_trace| was defined earlier:
% \begin{code}
% Definition trans (b: bool): relation abstract.State :=
%   fun s1 s2 => s2 `elem` if b
%     then cont_trans_over ahs s1
%     else disc_trans_over ahs s1.
% 
% Definition abstract.reachable (s: abstrat.State): Prop :=
%   exists is: State ap,
%     overestimation_bool (initial_dec ahs is) = true /\ alternate trans is s.
% \end{code}
% By induction over traces, we can prove
% \begin{code}
% shared_cover concrete.reachable abstract.reachable,
% \end{code}
% which rewrites to
% Having changed the abstract transition overestimator specifications to permit sharing, the definition of |abstract.reachable| given in \ref{inadequate} is no longer appropriate, 
% we can no longer use these to overestimate |abstract.reachable| as defined in \ref{inadequate}. After all, when the burden of reachability is shared among abstract states, some abstract states may actually be unreachable even though they contain reachable concrete states.
% 
% In section \ref{inadequate}, we defined |reachable_by_abstract_trace| in terms of |abstract.cont_trans| and |abstract.desc_trans|, which were the simplistic ``is there a concrete transition between the regions?'' relations that our abstract transition overestimators overestimated, but which had no notion of sharing. Having changed the abstract transition overestimator specifications to permit sharing, we must mimic this change in |abstract.cont_trans| and |abstract.desc_trans| for |reachable_by_abstract_trace| to remain representative. Hence, we redefine these relations to simply reflect overestimator judgements:
% \begin{code}
% Definition abstract.cont_trans: relation abstract.State
%   := fun s1 s2 => s2 ∈ cont_trans_over ahs s1.
% 
% Definition abstract.disc_trans: relation abstract.State
%   := fun s1 s2 => s2 ∈ disc_trans_over ahs s1.
% \end{code}
% Unfortunately, after these changes, |reachable_by_abstract_trace `subsetof` abstract.reachable| is no longer provable, because after all, when the burden of reachability is shared among abstract states, some abstract states may actually be unreachable even though they contain reachable concrete states. However, we \emph{can} prove (again, by induction over traces) a variation of that bypasses |abstract.reachable| altogether, and embraces sharing:
% \begin{code}
% (* SharedCover concrete.reachable reachable *)
% 
% Definition reachable_respect: Prop :=
%   forall (s: concrete.State),
%     (forall s': abstract.State, s `elem` s' -> ~ reachable_by_abstract_trace s') ->
%     ~ concrete.reachable s.
% \end{code}
% Where the previous definition of |unreachable_respect| stated that to conclude unreachability of a concrete state one needed only to prove unreachability of \emph{any} abstract state that contained it, this revised definition requires one to prove unreachability of \emph{all} abstract states containing the concrete state. We can see this in the example: to conclude unreachability of the concrete state |(l, 1)|, it is no longer sufficient to establish unreachability of the abstract state |(l, [0, 1])|.
% 
% At this point, what's left is to implement the three overestimators that make up an abstract system, and to use them to overestimate |reachable_by_abstract_trace|. However, before we do that, we make one more change.

\subsection{Alternating Traces}
\label{alter}

In the definition of |reachable_by_abstract_trace|, we used the simple transitive closure of the union of |abstract.cont_trans| and |abstract.desc_trans|. This formulation implied that abstract traces could contain successive continuous transitions. Unfortunately, as Alur observes, such repetition results in pathological abstract traces that do not represent any concrete trace. \todo{Example needed?}

The solution, as described by Alur, lies in the fact that concrete reachability by arbitrary traces is equivalent to concrete reachability by traces that alternate between continuous and discrete transitions:
\begin{code}
  Definition concrete.trans_kind (b: bool): relation concrete.State :=
    if b then concrete.disc_trans else concrete.cont_trans.

  Definition concrete.reachable_alternating (s: concrete.State): Prop :=
    exists i: concrete.State, concrete.initial i /\ alternate concrete.trans_kind i s.

  Lemma concrete.alternating_reachable_equiv:
    forall s, concrete.reachable s <-> concrete.reachable_alternating s.
\end{code}
(Here, |alternate| forms the alternating transitive reflexive closure of a pair of relations. Indexing the latter by a |bool| simplifies the definition of |alternate|.) \todo{Say something about the proof of this lemma?} As an immediate corollary, we have:
\begin{code}
  ~ concrete.reachable_alternating s -> ~ concrete.reachable s
\end{code}
Hence, in order to show that unsafe concrete states are not reachable, we need only show that they are not reachable by an alternating trace. Consequently, we may define |abstract.reachable| in terms of alternating traces as well, and still prove |reachable_respect|:
\begin{code}
Definition abstract.trans_kind (b: bool): relation abstract.State :=
  fun s s' => s `elem` (if b then over_cont_trans else over_disc_trans) s'.

Definition abstract.reachable (s: abstract.State): Prop :=
  exists i: abstract.State, over_initial ahs i = true /\ alternate abstract.trans_kind i s.
\end{code}
\todo{Say something about the induction proof over alternating traces?}
% todo: the use of this kind of alternation instead of (F o G) o F is that it's simpler, thought it may not look it. but we don't need reflexivity of F. TODO: is this true?

% We now turn our attention to overestimation of |reachable_by_abstract_trace|, after which we will consider the implementation of the overstimators that comprise an abstract system, in sections \ref{bla} and \ref{bla}.

% repeat somewhere that an overestimator could just return "true" all teh time, but that would make everything reachable, makign the thing useless for unreachability (and hence, safety) determination.
% and repeat somewhere that overestimation is all about inclusion of predicates, heuristics. basically, this paragraph:

% The idea now is to define |abstract.reachable| by decidable heuristics that make a ``best effort'' to establish absence of reachable concrete states represented by a given abstract state, marking the latter unreachable if successful, and reachable otherwise. Better heuristics will result in fewer spuriously reachable abstract states, decreasing the risk that abstract states containing unsafe concrete states become reachable (these have to remain unreachable for the abstraction to be of use).

% in terms of new computable transition relations expressing abstract discrete and continuous transitions, as well as abstract initiality. All of these will be ``decidable overestimators'', occasionally emitting spurious transitions or initial states when heuristics fail to establish their absence, to ensure that their concrete counterparts are respected. |unreachable_respect| will then follow by induction on abstract traces. \todo{right?}

% We emphasize that weakening the specification in this way only \emph{permits} a computable implementation of the continuous abstract transition relation to omit |[0, 1]| in the list of continuous abstract transitions originating from |[1,2]| in the example. Actually constructing a ``smart'' abstract continuous transition function that exploits this freedom and takes care to omit silly regions from its result list is a different matter, discussed in section \ref{drift}.

% However, these definitions do make good \emph{minimal} specifications: any relations (decidable or otherwise) that include them are sufficient to prove |unreachable_respect|. Hence, we may work with \emph{decidable overestimations}: decidable transition relations that include \emph{at least} all transitions required to ensure that the concrete transition relation is represented, but \emph{may} include additional transitions. In essence, this means that in the implementation of these relations, we are free to employ heuristics that \emph{may} definitively establish absence of concrete transitions between concrete states represented by a pair of abstract states, justifying omission of an abstract transition between the two, but may also ``give up''. By simply including the transition in the latter case, we ensure that the concrete relation is respected. This does mean that abstract states may become reachable that do not represent any reachable concrete states. Thus, if the heuristic gives up too often, abstract states representing unsafe concrete states may become reachable, rendering the abstraction useless.

% show that an abstract system of this type admits a decidable notion of reachability relates to concrete reachability in the right way.

\section{Underestimating Safety}
\label{underestimatingsafety}

In the next section we show that thanks to decidability of our transition and initiality overstimators, |abstract.reachable| is decidable. But first, we end the current section by showing how a decision procedure for |abstract.reachable| lets us underestimate hybrid system safety, and in particular, lets us obtain a proof of thermostat safety. So suppose we have
\begin{code}
  reachable_dec: decider abstract.reachable
\end{code}
And suppose we are given the following specification of unsafe concrete states, covered by a finite list of abstract states:
\begin{code}
  Variables
    (unsafe: concrete.State -> Prop)
    (astates: list abstract.State)
    (astates_cover_unsafe: forall s, unsafe s -> forall r, s `elem` r -> r `elem` astates).
\end{code}
\todo{Mention why |shared_cover| isn't appropriate here.}
Then, using |reachable_dec| and |unreachable_respect|, we can easily define
\begin{code}
  Definition over_unsafe_reachable: overestimation (overlap unsafe concrete.reachable).
\end{code}
Taking |unsafe := thermo_unsafe| and a suitable abstract cover, we obtain
\begin{code}
  Definition over_thermo_unsafe_reachable:
    overestimation (overlap thermo_unsafe concrete.reachable).
\end{code}
Recall that |ThermoSafe| was defined as |thermo_unsafe `subsetof` concrete.unreachable| in section \ref{concrete}. Since we trivially have |~ overlap unsafe concrete.reachable -> ThermoSafe|, we also have:
\begin{code}
Definition under_thermo_unsafe_unreachable: underestimation ThermoSafe.
\end{code}
Finally, we \emph{run} the underestimation:
\begin{code}
Theorem: ThermoSafe.
Proof.
  apply (underestimation_true under_unsafe_unreachable).
  vm_compute. sp1mm reflexivity.
Qed.
\end{code}
|underestimation_true| is a tiny utility of type |forall P (o: underestimation P), o = true -> P|, whose application in the proof reduces the goal to
\begin{code}under_thermo_unsafe_unreachable = true.\end{code} The |vm_compute| tactic invocation then forces evaluation of the left hand side, which will in turn evaluate |over_thermo_unsafe_reachable|, which will evaluate |reachable_dec|, which will (as we will see in the next section) evaluate the |abstract.System| overestimators (which we will build in sections \ref{over_cont_trans} and \ref{over_disc_trans}). This process, which takes about 35 seconds on a modern desktop machine, eventually reduces |under_thermo_unsafe_unreachable| to |true|, leaving |true = true|, proved by |reflexivity|.

We can now also clearly see what happens when the abstraction method ``fails'' due to poor region selection, overly simplistic transition/initiality overestimators, or plain old unsafety of the system. In all these cases, |vm_compute| reduces |under_thermo_unsafe_unreachable| to |false|, and the subsequent |reflexivity| invocation will fail.

\todo{Say something about automatic generation of offending traces?}

\

This concludes the high level story of our development. What remains are the implementation of |reachable_dec| in terms of the decidable overestimators for abstract initiality and transitions bundled in the |abstract.System| record, and the implementation of those overestimators themselves. The former is treated in section \ref{graphs}, the latter in sections \ref{over_cont_trans} and \ref{over_disc_trans}.

 % using  the particular heuristics and techniques we used to implement the decidable overestimators for abstract initiality and transitions, the specifications (types) of which were given in section \ref{abs.trans}.

% next up:
% - defining cont_trans
% - defining disc_trans
% - definign initial
% - square_abstraction

\section{Graphs for Reachability Decision}
\label{graphs}

Given an |abstract.System| containing overestimators for abstract transitions and initiality, we can decide |abstract.reachable| (as defined in the last section) by constructing a directed graph in which vertices and edges correspond to abstract states and transitions, respectively, and then using an ordinary graph reachability algorithm. Because the hybrid system safety proofs we ultimately wish to produce obviously depend on the correctness of the reachability determinations obtained this way, we have implemented a certified graph reachability algorithm inside Coq. \todo{say something about algorithm? does it have a name?} For efficiency reasons, it computes a list of all vertices reachable from a given list of initial vertices in one go:
\begin{code}
graph_reachables: forall (g : DiGraph) (start: list Vertex), NoDup start ->
  {l : list Vertex | forall w : Vertex, In w l <-> digraph.reachable start w}
\end{code}
Here, |DiGraph| and |digraph.reachable| are defined as
\begin{code}
Record DiGraph: Type := Build_DiGraph
  { Vertex: Set
  ; Vertex_eq_dec: EqDec Vertex eq
  ; vertices: ExhaustiveList Vertex
  ; edges: Vertex -> list Vertex
  ; edges_NoDup: forall v, NoDup (edges v)
  }.

Definition edge: relation Vertex := fun v w => In w (edges v).

Definition digraph.reachable (start: list Vertex) (v: Vertex): Prop :=
  exists s `elem` start, trans_refl_closure edge s v.
\end{code}

Naively equating vertices with abstract states and edges with abstract transitions (either continuous or discrete) would make vertex reachability in the graph equivalent to potentially-non-alternating abstract state reachability, which as we saw in section \ref{alter} would result in many needlessly reachable abstract states, potentially rendering the abstraction useless.

To make vertex reachability in the graph equivalent to \emph{alternating} abstract state reachability, we first add a |bool| component to |Vertex|:
\begin{code}
Definition Vertex: Set := prod bool abstract.State.
\end{code}
Next, we only add edges corresponding to abstract \emph{continuous} transitions from a vertex |v| with |fst v = true| to a vertex |w| with |fst w = false|, and only add edges corresponding to abstract \emph{discrete} transitions from a vertex |v| with |fst v = false| to a vertex |w| with |fst w = true|:

\begin{code}
  Definition nexts (v: Vertex): list abstract.State :=
    let (k, s) := v in
      if k then over_cont_trans s
      else over_disc_trans s.

  Definition edges (v: Vertex): list Vertex := map (pair (negb (fst v))) (nexts v).

  Definition graph: DiGraph := Build_DiGraph Vertex edges.
    (*{-"\text{ Remaining record fields omitted for the sake of exposition. }"-}*)
\end{code}
This ensures that paths through the graph alternate between vertices |v| with |fst v = true| and vertices with |fst v = false|, and that the corresponding abstract traces alternate between continuous and discrete transitions. A few simple inductions now show that if we take as |start| those vertices selected by our initiality overestimator, |abstract.reachable| is equivalent to |digraph.reachable|:
\begin{code}
  Definition start: list Vertex := { v: Vertex | abstract.over_initial (snd v) }.

  Theorem graph_respect: forall s: abstract.State ap,
    abstract.reachable s <-> exists b: bool, digraph.reachable start (b, s).
\end{code}

Combining |graph_respect|, |abstract.unreachable_respect|, and |reachables|, it is now straightforward to construct a |decider abstract.reachable|.
\todo{Show some more detail?}

Through the definitions of |edges|, |nexts|, and |start|, this decider builds on the three overestimators in |abstract.System|. In the next two sections, we discuss how to define an |abstract.System| instance for the discretization (described in section \ref{abs.space.constr}) of the concrete thermostat's continuous state space.

\section{Overestimating Continuous Abstract Transitions}
\label{over_cont_trans}

We now discuss the implementation of the second component of an |abstract.System|, |over_cont_trans|:
\begin{code}
  over_cont_trans: sharing_transition_overestimator concrete.cont_trans
\end{code}
where |sharing_transition_overestimator| is defined as
\begin{code}
Definition sharing_transition_overestimator
  (R: relation concrete.State): Set := forall s: abstract.State,
    { l: list abstract.State | () -> (NoDup l /\ shared_cover
      (concrete.invariant `inter` (overlap s `compose` utilflip R)) l) }.
\end{code}
where |shared_cover| is defined as
\begin{code}
  Definition shared_cover
    `{Container concrete.State C} `{Container abstract.State D}
      (cs: C) (ss: D): Prop :=
        forall s: concrete.State, In s cs -> DN (exists r: abstract.State, In s r /\ In r ss).
\end{code}

\subsection{Avoiding Drift}

A first observation is that since the concrete continuous transition relation is reflexive, |over_cont_trans| must be reflexive as well (meaning its result list must include the abstract state it received as an argument). By the sharing principle, this means that redundantly reachable regions (whose only reachable points lie within the source region) are accounted for, which in turn means the problem can be reduced to the overestimation of non-redundant transitions, characterized by the following relation:
\begin{code}
Definition substantial_cont_trans (l: Location): relation Region
  := fun r r' => exists p p', p `elem` r /\ p' `elem` r' /\ ~ p' `elem` r /\ can_flow l p q.
\end{code}
Hence, we seek to build an |overestimator substantial_cont_trans|. Note that even though redundantly reachable regions are ``accounted for'' specification-wise, we still need a mechanism enabling us to notice their redundancy, so that our overestimator may return |false|. Unfortunately, detecting whether a particular abstract continuous transition would be redundant is undecidable in general. For now, let us simply assume existence of a redundance underestimator, and worry about its implementation later:

% 
% 
% not notice their redundancy.
% 
% In section \ref{trouble}, we illustrated how drift would cause problems if we gave the |over_cont_trans| member in |abstract.System| the type
% \begin{code}
% overestimator abstract.cont_trans.
% \end{code}
% While appealing in its simplicity---such overestimators only ever consider two regions at a time---this specification would force inclusion of redundant transitions like the one from |[1,2]| to |[0,1]| in the example. To counteract this outcome, we reformulated the specification for abstract continuous transitions as a predicate |cont_respect| on functions that, given an abstract state, yield a list of regions representing abstract continuous transitions. This permitted such functions to exploit the sharing principle and omit redundant transitions.

\begin{code}
Variable under_redundant: underestimator redundant_cont_trans.
\end{code}
The idea now is to build a drift-oblivious |overestimator| for the simplistic |abstract.cont_trans| relation, which \emph{will} emit redundant transitions, but to only run it if the redundancy underestimator does not indicate redundance. That is, supposing we have
\begin{code}
Variable simple_over_cont_trans: overestimator abstract.cont_trans,
\end{code}
we can define

%format r_src
%format r_dst
%format p_src
%format p_dst
%format negb (x) = ~ x
\begin{code}
  Program Definition over_subst_cont_trans (l: Location) (r_src r_dst: Region)
    : overestimation (substantial_cont_trans l r_src r_dst)
    := negb (under_redundant l r r') && simple_over_cont_trans l r_src r_dst.
\end{code}

Thus, the problem is reduced to construction of |under_redundant| and |simple_over_cont_trans|. We discuss the latter in the next section. For the former, we observe that the redundant transitions we worry about appear as transitions from a region to a region adjacent to it in the opposite direction of the flow corresponding to the current location. This means that if we know which regions are adjacent, and further know when (e.g. in which locations) individual flow components are monotonic (and in which direction), we can underestimate redundancy quite well. Since in our development abstract regions are formed by explicitly enumerating interval bounds, we have a-priori knowledge of which regions will be adjacent. Hence, the only additional work we have to do is to provide as many proofs of flow monotonicity (in the different locations and for the different flow components) as possible, with which some generic utilities can construct a proper redundancy underestimator.

Having taken care of drift, all we need now (to complete the definition of |over_cont_trans|) is an implementation of |simple_over_cont_trans|, which we present in the next section.

\subsection{Simple Transition Overestimation}

Recall that |abstract.cont_trans| is defined as
\begin{code}
Definition abstract.cont_trans: relation abstract.State
  := fun s s' => exists (c `elem` s) (c' `elem` s'), concrete.cont_trans c c'.
\end{code}
Since location does not change in continuous transitions, and location equality is decidable (since locations are just names), we can easily reduce the problem down to overestimation of
\begin{code}
Definition region_cont_trans (l: Location): relation Region
  := fun r_src r_dst => exists (p_src `elem` r_src) (p_dst `elem` r_dst'), can_flow l p_src p_dst,
\end{code}
where |can_flow| is still
\begin{code}
Definition can_flow (l: Location): relation Point
  := fun p_src p_dst => exists d: Duration, flow l p_src d = p_dst /\
    forall t, 0 <= t <= d -> invariant (l, flow l p_src t).
\end{code}
So given a location |l| and regions |r_src| and |r_dst|, we are looking to establish absence of concrete transitions from points in |r_src| to points in |r_dst|.

\todo{story about how Alur does it, and why we don't just do the same}

%\subsection{Ruling out Transitions by Invariant Absence}

\

\noindent Let us first consider whether we can rule out transitions by looking at the invariant, which |can_flow| demands must hold at every point along the way. Clearly, if we are able to determine that there do not exists paths from points in |r_src| to points in |r_dst| with the invariant holding at each point along the way, then there need be no continuous transition between |r_src| and |r_dst|. Unfortunately, \emph{deciding} this property is too hard. However, suppose we can at least overestimate whether the invariant holds somewhere in an abstract region:
\begin{code}
  Variable invariant_overestimator: overestimator abstract.invariant.
\end{code}
With |invariant_overestimator|, we can start defining |simple_over_cont_trans| as follows:
\begin{code}
  Definition simple_over_cont_trans: overestimator abstract.cont_trans
    := fun r_src r_dst =>
      invariant_overestimator r_src &&
      invariant_overestimator r_dst && (*{-"\text{ .. further tests, to be discussed. }"-}*).
\end{code}
That is, if either |~ invariant_overestimator r_src| or |~ invariant_overestimator r_dst|, then we conclude that there can be no concrete transition from a point in |r_src| to a point in |r_dst|. To keep matters simple, we will not bother to check the invariant anywhere else. This does mean that we might generate a spurious abstract transition from |r_src| to |r_dst| if the only reason for lack of concrete transitions between |r_src| and |r_dst| is that the invariant be broken somewhere in the middle of the flow path. Fortunately, this turns out not to cause problems for the thermostat hybrid system. \todo{Can anything be said about to what extent this simplification is justifiable or even necessary for hybrid systems in general?} In our development, the definition of |invariant_overestimator| is almost automatic, requiring only that the thermostat's invariant be reformulated in terms of (possibly unbounded) squares in the continuous state space, whose intersection with abstract regions can be automatically overestimated.

\

Next, let us consider how we might rule out absence of concrete continuous transitions from points in |r_src| to points |r_dst| by looking at the actual flow function. Clearly, if we are able to determine that there are no points in |r_src| which the flow function maps to points in |r_dst|, then there need be no continuous transition between |r_src| and |r_dst|. Equally clearly, this is utterly impossible to meaningfully overestimate for an general flow function and general regions. However, the thermostat's posesses three key properties that we can exploit:
\begin{enumerate}
\item its continuous space is of the form |pow CR n|;
\item abstract regions correspond to multiplied |CR| intervals;
\item its flow functions are both separable and range invertible.
\end{enumerate}

% Finally, we get to the real ``meat'' of our overestimation: overestimating whether there is a point in |r_src| that the flow function maps to a point in |r_dst| in the first place.

A flow function on a product space is \emph{separable} if it can be written as the product of flow function on the respective component spaces. For instance, in |pow CR 2|, a flow function is separable if future values of the first component only depend on past values of that same component, and the same is true for the second component. Being defined as the product of two flow functions on |CR| (|temp_flow| and |temp_flow|), the thermostat's flow function (in any location) has this property by construction.

A flow function |f| on |CR| is range invertible if
\begin{code}
  exists (range_inverse: OpenRange -> OpenRange -> OpenRange),
    forall (a: OpenRange) (p: CR), p `elem` a ->
    forall (b: OpenRange) (d: Duration), f p d `elem` b -> d `elem` range_inverse a b
\end{code}
Here, |OpenRange| represents potentially unbounded intervals in |CR| (with bounds closed if present \todo{meh, need a better name for OpenRange..}). Range invertibility is a less demanding alternative to point invertibility:
\begin{code}
  exists (point_inverse: CR -> CR -> CR), forall p q: CR, f p (point_inverse p q) = q
\end{code}
We need range invertibility because the exponential flow functions are not point invertible.

As mentioned in section \ref{concreteflow}, we used a modest library of flow functions when defining the thermostat's flow. Included in that library are range-inverses, which consequently automatically apply to the thermostat's flow. Hence, no ad-hoc work is needed to show that the thermostat's flow functions are range-invertible.

%format r_src_temp = r_src _temp
%format r_dst_temp = r_dst _temp
%format r_src_clock = r_src _clock
%format r_dst_clock = r_dst _clock

Having defined the class of separable range-invertible flow functions, and having argued that the thermostat's flow is in this class, we now show how to proceed with our overestimation of existence of points in |r_src| which the flow function map to points in |r_dst|. As discussed in section \ref{abs.space.constr}, the abstract space for our thermostat was constructed as the product of two abstract spaces based on temperature and clock intervals, respectively. Regions in such a product space are pairs of regions in the composite spaces, so |r_src| and |r_dst| can be written as |(r_src_temp, r_src_clock)| and |(r_dst_temp, r_dst_clock)|, respectively.

We now simply use an |OpenRange| overlap overestimator of type
\begin{code}
Qpos -> forall a b: OpenRange, overestimation (overlap a b)
\end{code}
(defined in terms of things like |overestimate_CRle| shown in section \ref{estimation}) to overestimate whether the following three ranges overlap:
\begin{enumerate}
\item |[0, inf]|
\item |range_inverse temp_flow r_src_temp r_dst_temp|
\item |range_inverse clock_flow r_src_clock r_dst_clock|
\end{enumerate}
Overlap of 2 and 3 is equivalent to existence of a point in |r_src| from which one can flow to a point in |r_dst|. After all, if these two range inverses overlap, then there is a duration |d| that takes a certain temperature value in |r_src_temp| to a value in |r_dst_temp| and also takes a certain clock value in |r_src_clock| to a value in |r_dst_clock|. If 2 and 3 do \emph{not} overlap, then either it takes so long for the temperature to flow from |r_src_temp| to |r_dst_temp| that any clock value in |r_src_clock| would ``overshoot'' |r_dst_clock|, or vice versa. Finally, if 1 does not overlap with 2 and 3, then apparently one could only flow backward in time, which is not permitted. Hence, overlap of these three ranges is a necessary condition for existence of concrete flow from points in |r_src| to points in |r_dst|, and so our |abstract.cont_trans| overestimator may justifiably return ``false'' when the overlap overestimator manages to prove absence of overlap.

% todo: these paragraphs make it seem as though this algorithm was hard coded for temperature/clock, while it just works because the thermostat has product flow/space.
%format halfrange (a) (b) = [a, b)

Having shown at last the ``real'' flow test condition used in the overestimation, we can show more precisely how drift would arise if we avoided region overlap (thereby eliminating drift problems at the specification level) but did not bother to employ a redundancy underestimator. Suppose there were clock intervals [0, 1) and [1, 2), and a temperature interval [5, 6). Let us consider what our simple |abstract.cont_trans| overestimator would return given the abstract states |(Heat, (half_range 5 6, halfrange 1 2)| and |(Heat, (halfrange 5 6, halfrange 0 1)|. After determining that the locations are equal as required, and that the invariant holds in both regions, our overestimator would proceed to overestimate overlap of the ranges mentioned above. In this particular case, with |clock_flow| simply being linear, it would compute that
\begin{code}range_inverse clock_flow r_src_clock r_dst_clock = halfrange (-2) 0\end{code} Unfortunately, the overestimator would not be able to determine that [0, inf) does not overlap with [-2, 0), because given only two approximations of 0, it cannot determine that |0 <= 0|. Thus, it would have no choice but to bless the transition.

\section{Overestimating Discrete Abstract Transitions}
\label{over_disc_trans}

We now turn our attention to the implementation of the |over_disc_trans| component in |abstract.System|. Recall (from section \ref{conc.disc_trans}) the definition of concrete discrete transitions:
\begin{code}
  Definition concrete.disc_trans: relation concrete.State := fun s s' =>
    guard s (location s') /\
     reset (location s) (location s') (point s) = point s' /\
      invariant s /\ invariant s'.
\end{code}

To overestimate whether there exists a concrete discrete transition from a concrete state in one abstract state to a concrete state in another abstract state, we first have |over_disc_trans| simply check whether the invariant holds at both ends (exactly as in the continuous case), and (in the same way) whether the guard holds at the origin. The interesting part is the overestimation of whether there is a point in the source region that the reset function maps to a point in the destination region. This problem sounds rather similar to that of flow overestimation discussed in section \ref{over_cont_trans}, but the similarity is superficial for two reasons: reset functions are much simpler as they don't have a time parameter, and discrete transitions suffer from a substantially different variety of drift.

Suppose one's continuous state space is |CRge0|, and one's abstract regions are intervals of the form |[n, n+1]| with |n `elem` nat|. Further suppose that there is a discrete transition from location |l| to location |l'| with guard |g = const True| and reset function |r = id|. And suppose we are overestimating whether there needs to be an abstract discrete transition from |(l, [0, 1])| to |(l', [1, 2])|. Including this transition seems wasteful, because the only point in |[1, 2]| that the reset function maps a point in |[0, 1]| to is 1, but that point is still in |[0, 1]|.

For continuous transitions, we avoided drift by explicitly filtering out transitions found to be redundant by a redundance underestimator that relied on reflexivity of the continuous transition relation, knowledge of adjacent regions, and knowledge of where the flow function was monotonic. However, the discrete transition relation is not necessarily reflexive, so we use a slightly different mechanism. Basically, we simply take as an argument an underestimation of whether a particular reset function is the identity function, and if the underestimator says that it is, then we only make abstract discrete transition to abstract states with the same region as the source state. If on the other hand the underestimator does \emph{not} indicate that the reset function in question is the identity function, then we simply take the bounds of the source region, map them with the reset function, and overestimate whether the resulting region overlaps with the candidate destination (for this to be valid, the reset function must be nondecreasing\footnote{When we say that a \emph{flow} function is ``nondecreasing'', we mean that it does not decrease over time, and so this is really a statement about flow direction. But when we say that a \emph{reset} function (which does not take a time argument) is ``nondecreasing'', we simply mean that it does not flip ranges around, which is a rather mundane property.}

\todo{In the code, this is actually done in a more roundabout way, including an extra case for constant reset functions, but I really think this should not be necessary. Preliminary experiments with removing the third case mysteriously made the safety proof fail though, so deeper investigation is needed.}

For all this to work for a system with a continuous state space in |pow CR n|, the reset function must additionally be separable. Our thermostat's reset functions are trivially separable.

% todo: mention that we don't discuss boring things like over_initial and the abstract safety list, even though they're all treated uniformly: the user must specify these as lists of opensquares when setting up the abstraction.

\section{Conclusions}
\label{conclusions}

\begin{itemize}
\item nice showcase of proof-by-computation-on-computable-reals
\item systematic use of estimators to make tactic-like optional-deciders, at each level in the stack
\item computable reals do complicate things
\item heuristic for bound selection doesn't work out of the box. manual tweaking obviously not ideal. more experimentation required
\item region overlap
\item drift and redundance
\item code stats(?)
\item requirements on system: separability of flow and reset, invariant stability
\item lots of room for more clever heuristics with less restrictive preconditions (e.g. separability)
\end{itemize}

\bibliographystyle{splncs}
\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{paper}

\end{document}

### 2009-09-15 doc/clean.sh

@@ -1,0 +1,2 @@
+#!/bin/sh
+rm -f paper.log processed.tex paper.out paper.aux

### 2009-09-15 doc/compile.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+
+lhs2TeX < paper.tex > processed.tex
+cp processed.tex tmp.tex
+
+set -e
+pdflatex tmp.tex
+bibtex tmp
+pdflatex tmp.tex
+bibtex tmp
+pdflatex tmp.tex
+cat tmp.pdf > paper.pdf
+rm tmp.*

### 2010-01-20 doc/paper.tex
Elaborate on automatic estimator/decider construction.
Ignore-this: 8ea91feb3eb23c3b24df3704e298141a

darcs-hash:20100120155613-bab43-1c866906a07fd3c566b29a229c41a724acff6759.gz

@@
 %format forall = "\forall"
+%format Next = "\mathbf{Next}"
+%format Obligation = "\mathbf{Obligation}"
+%format Class = "\mathbf{Class}"
 %format Fixpoint = "\mathbf{Fixpoint}"
@@
 The idea is that the different locations corresponds to different uses of actuators available to the software system, the effects of which are described by the flow function.
-For instance, in our thermostat, the flow function corresponding to the |Heat| location will have the temperature increase with time, modeling the effect of the heater component in our imagined thermostat deviced.
+For instance, in our thermostat, the flow function corresponding to the |Heat| location will have the temperature increase with time, modeling the effect of the heater component in our imagined thermostat device.
 
@@
 
-Underestimators and overestimators can be combined to form underestimators (resp.
-overestimators) for things like conjunctions and quantifications over finite domains.
-Applying these combinators by hand is somewhat tedious, and for this reason we have made |overestimation| and |underestimation| type classes.
-We will not go into the details, but the net effect is that whenever we need an overestimation of some proposition |P|, we may say |overestimate P|, and Coq's type class resolution will try to construct an overestimation for |P| based on a collection of type class instances that amount to a database of registered overestimator combinators and primitives.
-
-Finally, while we're on the subject of decision-procedure-like things, we note that we use the same techniques for real decision procedures (in the few places we use those): we define |decision P| as |{P}+{~P}|, and given a predicate |P: A_0 -> ...
+We use the same techniques for real decision procedures (in the few places we use those): we define |decision P| as |{P}+{~P}|, and given a predicate |P: A_0 -> ...
 -> A_n -> Prop|, we define |decider P| as the type |forall (a_0: A_0) ...
@@
 a_n)|.
-And here, too, we make |decision| a type class, so that we may simply say |decide P| and have Coq use type class resolution to try and construct a decider deciding |P|.
-
-%\todo{Things like the Qpos above throw dirt in the food though. We could make a type class for ``epsilon-parameterized overestimators'', but meh.}
+
+\subsection{Automating estimator/decider composition}
+
+Underestimators, overestimators and deciders can be combined to form underestimators (resp.
+overestimators, deciders) for things like conjunctions and quantifications over finite domains.
+For instance, here is a combinator that forms conjunction overestimators:
+
+\begin{code}
+Program Definition overestimate_conj {P Q: Prop}
+  (x: overestimation P) (y: overestimation Q):
+    overestimation (P /\ Q) := x && y.
+Next Obligation.
+  intros [A B].
+  destruct x. destruct y.
+  simpl in H.
+  destruct (andb_false_elim _ _ H); intuition.
+Qed.
+\end{code}
+\todo{clean up proof}
+Applying these combinators by hand to form estimators/deciders for composite propositions is somewhat tedious, but fortunately this process can be automated if we make |overestimation|, |underestimation|, and |decision| type classes:
+
+\begin{code}
+Class overestimation (P: Prop): Set
+  := overestimate: { b: bool | b = false -> ~ P }.
+\end{code}
+If we now say |overestimate P|, the type class instance resolution mechanism will try to find (or build) a term of type |overestimation P| by recursively applying declared type class instances.
+For instance, if we declare |overestimate_conj| above as a type class instance (instead of a plain definition) and then say |overestimate (A /\ B)| where |A| and |B| are already known to be overestimatable (meaning the resolution mechanism is able to build terms of types |overestimation A| and |overestimation B|), then the resolution mechanism will be able to build a term of type |overestimation (A /\ B)| by applying |overestimate_conj|.
+
+We show one nontrivial example to illustrate the convenience of this mechanism: at one point in the development we have Coq automatically construct an overestimator of the property
+\begin{code}
+  exists u: abstract.State, In u astates /\ reachable u,
+\end{code}
+simply by saying
+\begin{code}
+  overestimate (exists u: abstract.State, In u astates /\ reachable u).
+\end{code}
+To construct this overestimator, Coq recursively applies a variety of instances that show that (1) decidable properties are overestimatable; (2) existentially quantified decidable properties over finite domains are decidable; (3) |abstract.State| is a finite domain; (4) conjunctions of decidable properties are decidable; (5) list membership is decidable if equality is decidable for the element type; (6) equality is decidable for products (such as |abstract.State|) if equality is decidable for the components; (7) equality is decidable for |Location|s; (8) equality is decidable for abstract (symbolic) regions; (9) reachability in the abstract system is decidable.
+
+In \cite{hintsinunification}, Asperti et al.
+use a similar technique (but implemented using unification hints instead of type classes) to automate reflection of algebraic expressions in concrete syntax into a an inductively defined abstract syntax.
 
@@
 Fortunately, the double negation is also \emph{sufficient}, because we will ultimately only use |regions_cover| in a proof of of |...
--> ~ concrete.reachable s| (for some universally quantified variable s), which, due to its head type being a negation, is stable, and can therefore be proved in and then extracted from |DN|.
+-> ~ concrete.reachable s| (for some universally quantified variable |s|), which, due to its head type being a negation, is stable, and can therefore be proved in and then extracted from |DN|.
 Hence, we only need |regions_cover|'s result in |DN|.

### 2010-01-22 doc/paper.tex
doc: Make grammar work with hidden todo's; add note about DN rationale.
Ignore-this: e77bee7f267bcef4ae01532a943f2b54

darcs-hash:20100122114551-bab43-549fcb4983f7aaedfa14528684d666cebd99ee0c.gz

@@
 
-The prototype implementation described in \cite{alur} was developed in \todo{some ordinary programming language}, using \todo{some ordinary programming libraries}.
+The prototype implementation described in \cite{alur} was developed in a conventional programming language using libraries that use ordinary floating point arithmetic \todo{which language/libraries exactly?}.
 The potential for bugs and floating point artifacts inherent in this approach limit the confidence one can justifiably have in safety judgements made by such an implementation.
@@
 Nevertheless, this is the path we will take in our development, partly out of necessity (because in some parts of the development, we really need to ``run'' these lemmas to obtain |<=|/|<| proofs), and partly because just taking |le_lt_dec| as an axiom amounts to cheating.
+\todo{the former is a lie.
+paste the new perspective as written in mail to Herman, and/or stuff from our paper}
 

### 2010-01-22 doc/paper.tex
doc: Fix some spelling/grammar bugs Adam noticed.
Ignore-this: 14d924069b65821274815be53efa212b

darcs-hash:20100122125228-bab43-44cba8fb50f03ad6e0c33cbb8ec5cb84b970263f.gz

@@
 The prototype implementation described in \cite{alur} was developed in a conventional programming language using libraries that use ordinary floating point arithmetic \todo{which language/libraries exactly?}.
-The potential for bugs and floating point artifacts inherent in this approach limit the confidence one can justifiably have in safety judgements made by such an implementation.
+The potential for bugs and floating point artifacts inherent in this approach limits the confidence one can justifiably have in safety judgements made by such an implementation.
 In this report, we describe a reimplementation of the basic technique in Coq, a proof assistant based on a rich type theory that also functions as a programming language, letting one develop ``certified'' programs: programs whose correctness is proved formally within the system.
@@
 
-We now say that there is a (concrete) \emph{continuous transition} from a state |(l, p)| to a state |(l', p')| if |l = l'| and there is a nonnegative duration |d| such that |p' = flow l p d| with the invariant for |l| holding at every point along the way:
+We now say that there is a (concrete) \emph{continuous transition} from a state |(l, p)| to a state |(l', p')| if |l = l'| and there is a non-negative duration |d| such that |p' = flow l p d| with the invariant for |l| holding at every point along the way:
 
@@
 Each such transition is comprised of two components: a \emph{guard} predicate, and a \emph{reset} function.
-The former defines a subset of the continous state space in which the transition is enabled (permitted), while the latter describes an instantaneous change applied as a side effect of the transition, as seen in the following definition of the discrete transition relation:
+The former defines a subset of the continuous state space in which the transition is enabled (permitted), while the latter describes an instantaneous change applied as a side effect of the transition, as seen in the following definition of the discrete transition relation:
 
@@
 
-As mentioned before, the set of reachable states represent the possible behaviors exhibited by a hypothetical real-world implementation of the hybrid system (as software running on a device with sensor and actuators).
+As mentioned before, the set of reachable states represents the possible behaviors exhibited by a hypothetical real-world implementation of the hybrid system (as software running on a device with sensor and actuators).
 
@@
 
-% repeat somewhere that an overestimator could just return "true" all teh time, but that would make everything reachable, makign the thing useless for unreachability (and hence, safety) determination.
+% repeat somewhere that an overestimator could just return "true" all the time, but that would make everything reachable, makign the thing useless for unreachability (and hence, safety) determination.
 % and repeat somewhere that overestimation is all about inclusion of predicates, heuristics. basically, this paragraph:

### 2010-01-22 doc/paper.tex
doc: Fix monad citation.
Ignore-this: 53ce628e2cccea337ff4760129dcafed

darcs-hash:20100122150926-bab43-89092f3f26e5c45021656a68a7077b6cbe8f856d.gz

@@
 
-One practical way to answer the question of how such a doubly negated proposition might be used to prove things is to observe that double negation, as a function on types/propositions, is a monad \cite{bla}.
+One practical way to answer the question of how such a doubly negated proposition might be used to prove things is to observe that double negation, as a function on types/propositions, is a monad \cite{wadler95monads}.
 Writing |DN P| for |~ ~ P|, we have the following two key operations that make |DN| a monad:

### 2010-02-01 doc/paper.tex
doc: Fix university name and credit BRICKS/FOCUS project.
Ignore-this: 5627502dd551b7c29e383bd26bb96246

darcs-hash:20100201111628-bab43-f2099273d43230bcaee67fe50c8af341708d75b8.gz

@@
 
-\author{}
-\institute{Institute for Computing and Information Sciences\\ Nijmegen University}
+\author{\eelis}
+\institute{Institute for Computing and Information Sciences\\ Raboud University Nijmegen}
 
@@
 The development critically relies on the computable real number implementation part of the C-CoRN library of formalized constructive mathematics.
-\todo{Hm, anything else worth mentioning?}
+\todo{Hm, anything else worth mentioning?}\footnote{This research was supported by the BRICKS/FOCUS project 642.000.501 ``Advancing the Real use of Proof Assistants''.}
 

### 2010-03-22 ITP10/hybrid-12oclock.tex
Add ITP10 paper sources.
Ignore-this: 49de24983449dd9b9daef53ef1da6000

darcs-hash:20100322174102-bab43-dcc3cc0af50eb50c30ef52b815fff99223eb4669.gz

--- 2010-01-22 doc/paper.tex
@@
 \documentclass[runningheads]{llncs}
-
-\newcommand{\eelis}{Eelis van der Weegen}
-\newcommand{\herman}{Herman Geuvers}
-\newcommand{\dan}{Dan Synek}
-\newcommand{\adam}{Adam Koprowski}
-
+\ifx\pdftexversion\undefined
+  \usepackage[dvips]{graphics}
+\else
+  \usepackage[pdftex]{graphics}
+\fi
+\usepackage{amsmath}
+%include format.tex
+%\usepackage{rotating}
+\usepackage{graphicx}
+\input{diagrams}
 \newcommand{\thetitle}{Automated Machine-Checked Hybrid System Safety Proofs}
-
+\newcommand{\eelis}{Eelis van der Weegen$^{\dagger}$}
+\newcommand{\herman}{Herman Geuvers$^{\dagger\ddagger}$}
+\newcommand{\dan}{Dan Synek$^{\dagger}$}
+\newcommand{\adam}{Adam Koprowski$^{\P}$}
+
+% For the final version:
+%\renewcommand{\todo}[1]{}
+\newcommand{\ADAM}[1] {\todo{\textsc{Adam says}: #1}}
+\newcommand{\HERMAN}[1] {\todo{\textsc{Herman says}: #1}}
+\newcommand{\DAN}[1] {\todo{\textsc{Dan says}: #1}}
+\newcommand{\EELIS}[1] {\todo{\textsc{Eelis says}: #1}}
+
+\newcommand{\Exists}[2]{\exists_{#1} \, . \, #2}
+\newcommand{\Forall}[2]{\forall_{#1} \, . \, #2}
+
+\def\phi{\varphi}
+\newcommand{\weg}[1]{}
+
+\newcommand{\DN}{{\sf DN}}
+\newcommand{\IR}{{\mathbf R}}
+\newcommand{\Heat}{{\sf Heat}}
+\newcommand{\Cool}{{\sf Cool}}
+\newcommand{\Chec}{{\sf Check}}
+\newcommand{\Off}{{\rm Off}} 
+\newcommand{\Loc}{{\rm Loc}}
+\newcommand{\AState}{{\rm AState}}
+\newcommand{\State}{{\rm State}}
+\newcommand{\Inv}{{\rm Inv}}
+\newcommand{\Init}{{\rm Init}}
+\newcommand{\Region}{{\rm Region}}
+\newcommand{\List}{{\rm List}}
+\newcommand{\Unsafe}{{\rm Unsafe}}
+\newcommand{\Safe}{{\rm Safe}}
+\newcommand{\AUnsafe}{{\rm AUnsafe}}
+\newcommand{\ASafe}{{\rm ASafe}}
+\newcommand{\Reach}{{\rm Reach}}
+\newcommand{\AReach}{{\rm AReach}}
+\newcommand{\guard}{{\rm guard}}
+\newcommand{\reset}{{\rm reset}}
+\newcommand{\trans}{\rightarrow_{CD}}
+\newcommand{\doubleheadrightarrow}{\mathrel{\rightarrow\!\!\!\!\rightarrow}}
+\newcommand{\ttrans}{\doubleheadrightarrow_{CD}}
+\newcommand{\attrans}{\stackrel{A}{\doubleheadrightarrow_{CD}}}
+\newcommand{\contrans}{\rightarrow_C}
+\newcommand{\distrans}{\rightarrow_D}
+\newcommand{\acontrans}{\stackrel{A}{\rightarrow_C}}
+\newcommand{\adistrans}{\stackrel{A}{\rightarrow_D}}
+\newcommand{\Dur}{{\IR_{\geq 0}}}
+\newcommand{\leltdec}{{\tt leltdec}}
+\newcommand{\leltdecoverlap}{{\tt leltdecoverlap}}
+\newcommand{\leltdecDN}{{\tt leltdecDN}}
+
+\newcommand{\overcontrans}{{\rm over\_cont\_trans}}
+\newcommand{\overdistrans}{{\rm over\_disc\_trans}}
 \usepackage{color}
 \usepackage{amsmath}
-\usepackage[colorlinks,pdftitle=\thetitle,pdfauthor=]{hyperref}
 
 \spnewtheorem*{propo}{Proposition}{\bfseries}{\rmfamily}
-
-%include polycode.fmt
-
-%format pipe = "\mathopen{|}"
-
-%format times = "\times"
-
-%format . = "."
-%format forall = "\forall"
-%format Next = "\mathbf{Next}"
-%format Obligation = "\mathbf{Obligation}"
-%format Class = "\mathbf{Class}"
-%format Fixpoint = "\mathbf{Fixpoint}"
-%format Definition = "\mathbf{Definition}"
-%format Eval = "\mathbf{Eval}"
-%format Notation = "\mathbf{Notation}"
-%format Lemma = "\mathbf{Lemma}"
-%format Proof = "\mathbf{Proof}"
-%format Qed = "\mathbf{Qed}"
-%format Defined = "\mathbf{Defined}"
-%format Function = "\mathbf{Function}"
-%format Proposition = "\mathbf{Proposition}"
-%format Record = "\mathbf{Record}"
-%format Program = "\mathbf{Program}"
-%format Inductive = "\mathbf{Inductive}"
-%format Theorem = "\mathbf{Theorem}"
-%format Let = "\mathbf{Let}"
-%format Instance = "\mathbf{Instance}"
-%format Variable = "\mathbf{Variable}"
-%format Variables = "\mathbf{Variables}"
-%format measure = "\mathbf{measure}"
-
-%format prod a b = a "\times " b
-%format sp = "\ "
-%format frac a b = "\frac{" a "}{" b "}"
-
-%format == = "= "
-%format </ = "\nless "
-%format /= = "\ne "
-  % Without these, lhs2tex uses three-line equal signs.
-
-%format fcmp = "\circ"
-%format match = "\mathbf{match}"
-%format with = "\mathbf{with}"
-%format end = "\mathbf{end}"
-%format fun = "\lambda\hspace{-1mm} "
-%format nat = "\mathbb{N}"
-%format CRplain = "\texttt{CR}"
-%format CR = "\mathbb{R}"
-%format CRasCSetoid = CR
-%format exists = "\exists "
-%format log2ceil = "\log_{2}\hspace{-0.5mm}"
-%format over = "\mathbf{over}"
-%format INR (a) = a
-%format `union` = "\cup "
-%format /\ = "\wedge "
-%format ~ = "\neg "
-%format \/ = "\vee"
-%format <-> = "\leftrightarrow "
-%format ==> = "\Longrightarrow "
-%format `subsetof` = "\subseteq "
-%format phi = "\phi "
-%format False = "\bot "
-
-%format sp1mm = "\hspace{1mm} "
-
-%%format { = "\hspace{0.9mm}\{"
-  % so that parameters declared with {...} are spaced like those declared with (...)
-    % disabled because it makes record defs ugly
-
-%format (pow (a) (b)) = a"^"b
-
-%format In (a) (b) = a "\in " b
-%format `NotIn` = "\notin"
 
@@
 
-\arrayhs % i.e. no page breaks in code blocks
+%\arrayhs % i.e. no page breaks in code blocks
 
@@
 
-\author{}
-\institute{Institute for Computing and Information Sciences\\ Nijmegen University}
+\author{Herman Geuvers\inst{1,2}, Adam Koprowski\inst{3}, Dan Synek\inst{1}, \\ Eelis van der Weegen\inst{1}}
+\authorrunning{H. Geuvers, A. Koprowski, D. Synek, E. van der Weegen}
+
+\institute{
+Institute for Computing and Information Sciences\\ 
+Radboud University Nijmegen 
+\and
+Faculty of Mathematics and Computer Science\\ 
+Technical University Eindhoven
+\and
+MLState, Paris
+}
+
+\usepackage[colorlinks,pdftitle=\thetitle,pdfauthor=]{hyperref}
 
@@
 \begin{abstract}
-This technical report documents our development of a hybrid system safety prover, implemented in Coq using the abstraction method introduced by Alur in \cite{alur}.
-The development includes: a formalization of the structure of hybrid systems; a systematic approach and generic set of support utilities for the construction of an abstract system (consisting of decidable ``overestimators'' of abstract transitions and initiality) faithfully representing a (concrete) hybrid system; a translation of abstract systems to graphs enabling decision of abstract state reachability using a certified graph reachability algorithm; a proof of an example hybrid system (taken from \cite{alur}) generated using this tool stack.
-The development critically relies on the computable real number implementation part of the C-CoRN library of formalized constructive mathematics.
-\todo{Hm, anything else worth mentioning?}
+We have developed a hybrid system safety prover, implemented in Coq
+using the abstraction method introduced by Alur, Dang and Ivan\v ci\'c
+(2006).  The development includes: a formalization of the structure of
+hybrid systems; a set of utilities for the construction of an abstract
+system (consisting of decidable ``overestimators'' of abstract
+transitions and initiality) faithfully representing a (concrete)
+hybrid system; a translation of abstract systems to graphs enabling
+the decision of abstract state reachability using a certified graph
+reachability algorithm; a proof of the safety of an example hybrid
+system generated using this tool stack.  The development critically
+relies on the computable real number implementation part of the CoRN
+library of formalized constructive mathematics in Coq.  The obtained
+safety proofs are fully certified and do not use floating point real
+numbers, but arbitrary precise approximations of computable real
+numbers. It also uses a nice interplay between constructive and
+classical logic via the double negation monad.
+
+\weg{ This technical report documents our development of a hybrid
+  system safety prover, implemented in Coq using the abstraction
+  method introduced by Alur in \cite{alur}. The development includes:
+  a formalization of the structure of hybrid systems; a systematic
+  approach and generic set of support utilities for the construction
+  of an abstract system (consisting of decidable ``overestimators'' of
+  abstract transitions and initiality) faithfully representing a
+  (concrete) hybrid system; a translation of abstract systems to
+  graphs enabling decision of abstract state reachability using a
+  certified graph reachability algorithm; a proof of an example hybrid
+  system (taken from \cite{alur}) generated using this tool stack. The
+  development critically relies on the computable real number
+  implementation part of the CoRN library of formalized constructive
+  mathematics. \todo{Hm, anything else worth mentioning?}  }
 
@@
 
-\setcounter{tocdepth}{2}
-
-\tableofcontents
-\todo{How do I get rid of that silly author index at the top of the TOC?!}
+%\setcounter{tocdepth}{2}
+
+%\tableofcontents
+%\todo{How do I get rid of that silly author index at the top of the TOC?!}
 
@@
 
-In \cite{alur}, Alur presents a method for automated hybrid system safety verification based on the construction of an \emph{abstract} hybrid system (essentially a finite automaton) corresponding to the hybrid system of interest.
-The abstract system is constructed such that traces in the original system are represented in the abstract system.
-Consequently, one can draw conclusions about reachability of states in the concrete system from analysis of state reachability in the abstract system.
-Since the abstract system is an entirely finite discrete object (unlike the concrete system), reachability can simply be computed (using any graph reachability algorithm).
-Thus, the abstraction method brings the safety verification problem from a continuous and infinite domain into a discrete and finite domain, where it is amenable to what amounts to brute force.
-
-The prototype implementation described in \cite{alur} was developed in a conventional programming language using libraries that use ordinary floating point arithmetic \todo{which language/libraries exactly?}.
-The potential for bugs and floating point artifacts inherent in this approach limits the confidence one can justifiably have in safety judgements made by such an implementation.
-In this report, we describe a reimplementation of the basic technique in Coq, a proof assistant based on a rich type theory that also functions as a programming language, letting one develop ``certified'' programs: programs whose correctness is proved formally within the system.
-In our case, the program will be one that produces machine-checked proofs of hybrid system safety, obtained using the abstraction method.
-
-Our development relies crucially on computation using real numbers, for which we use the computable real number implementation developed by Russell O'Connor \cite{something}, and part of the C-CoRN library of formalized constructive mathematics \cite{something}.
-Indeed, this development showcases its use in a concrete and practical application, and subtleties regarding the nature of these computable reals are reflected in this development in several ways, which we will discuss in some detail.
+In \cite{alur}, Alur et al.\ present an automated method for hybrid system
+safety verification based on the construction of an \emph{abstract}
+hybrid system (essentially a finite automaton) corresponding to the
+hybrid system of interest. The abstract system is constructed such
+that traces in the original system are represented in the abstract
+system. Consequently, one can draw conclusions about reachability of
+states in the concrete system from analysis of state reachability in
+the abstract system. Since the abstract system is an entirely finite
+discrete object (unlike the concrete system), reachability can simply
+be computed using any graph reachability algorithm. Thus, the
+abstraction method brings the safety verification problem from a
+continuous and infinite domain into a discrete and finite domain,
+where the problem can be dealt with using standard graph algorithms.
+%where it is amenable to what amounts to brute force.
+
+The prototype implementation described in \cite{alur} was developed in
+a conventional programming language using libraries that use ordinary
+floating point arithmetic. The potential for bugs and floating point
+artifacts inherent in this approach limits the confidence one can
+justifiably have in safety judgments made by such an
+implementation. In this paper, we describe a reimplementation of the
+basic technique of Alur et al.\ in Coq, a proof assistant based on a
+rich type theory that also functions as a programming language,
+letting one develop ``certified'' programs: programs whose correctness
+is proved formally within the system. In our case, the program will be
+one that produces machine-checked proofs of hybrid system safety,
+obtained using the abstraction method.
+
+Our development relies crucially on computations involving real
+numbers, for which we use the computable real number implementation
+developed by Russell O'Connor \cite{oconnor}, and part of the CoRN
+library of formalized constructive mathematics \cite{corn}. Indeed,
+this development showcases its use in a concrete and practical
+application. Subtleties regarding the nature of these computable reals
+are reflected in this development in several ways, which we will
+discuss in some detail.
+
+\section{Motivation}
+The motivation for this work is threefold.
+
+First, we want to avoid the use of floating point number computation,
+because rounding errors may affect the correctness of the safety
+verification that has been performed. Instead, we want to use exact
+real arithmetic, which deals with infinitely precise real numbers and
+more importantly, infinitely precise algorithms for computing with
+real numbers. This means that, e.g.\ an algorithm for the function
+$f(x) = e^x$ can produce arbitrary many output precision (``digits'')
+by asking for more input precision. So, using exact real number
+algorithms, we can decide $\cos(\sin(e^2))<\cos(\sin(\frac{1}{e^2}))$
+by computing better and better approximations of the value of both
+expressions and observing at a certain point that the inequality
+holds. (Note that with this method we will never be able to determine
+in finite time that two such expressions denote the same number; for
+that we need a symbolic proof.) In Nijmegen we have developed a
+verified library of exact real arithmetic, formalized in the proof
+assistant Coq, as part of CoRN \cite{corn,oconnor}, our Constructice
+Coq Repository at Nijmegen. The algorithms (for $e^x$, $\cos$, $\sin$
+etc.) have been written in the programming language of Coq and have
+been proved correct using its logic. Coq is sufficiently fast that the
+algorithms can also be executed in reasonable time within the system,
+as can be witnessed by our online certified calculator
+\url{http://prover.cs.ru.nl/calc.html}. So, from this perspective,
+this work can also be seen as a case study in the use of this formal
+repository of exact real arithmetic: Can we use its computational
+capacities to verify safety of hybrid systems? Can we use exact real
+arithmetic to model the environment and verify the correctness of the
+controller?
+
+A second related point is that hybrid systems include continuous
+values. Some of these are controlled by the system itself, like a
+clock that can be reset, a valve that can be opened or a gas handle
+that can be opened. Others represent characteristics of the
+environment and can only be observed, like speed, pressure and
+temperature. Due to the fact that we deal with physical systems,
+measurements are always imprecise, and setting a valve in a certain
+position %may on a continuous scale 
+is also imprecise. 
+%Of course, we don't know exactly how imprecise, and we may assume
+%that better measuring instruments may arise
+This implies that when a controller makes a case distinction on the
+basis of $T>20$ or $T\leq 20$ -- e.g.\ if the temperature is larger
+that $20$, turn the heater off, if it is $20$ or below, keep it on --
+is always only ``approximately'' effective: The actual heater may
+remain on until just above $20$ degrees, because of a measuring
+imprecision. In a verification, one should actually take these things
+into account and we believe that constructive analysis provides the
+right way to model and verify hybrid systems. In constructive
+analysis, we do not have (for $x,y\in\IR$) $x>y \vee x\leq y$, so we
+cannot define an algorithm on the reals by making such a case
+distinction. On the other hand, we {\em do\/} have, for all
+$\epsilon>0$, $x+\epsilon>y \vee x-\epsilon< y$, which means that we
+can get an $\epsilon$-close enough approximation to a decision between
+$x>y$ and $x\leq y$.  So, the present work can also be seen as a
+confirmation of our belief that constructive analysis is the
+proper setting to model and verify hybrid systems.
+
+As a third point, the method to verify safety of hybrid systems of
+\cite{alur} uses so called ``predicate abstraction''. We will detail
+this approach in the next section; for now it suffices to say that it
+involves mapping the hybrid system to another, finite transition
+system, which is then verified. However, this abstraction itself is
+not verified. In this formalization we want to also verify the
+abstraction itself: show that the abstract system models the concrete
+one.
 
@@
 
- % In our case, the specification of the program we write is this: given a description of a concrete system, the program  .
-The whole implementation, including the computational part, serves only to prove a single theorem.
-Hence, one essentially only needs to trust Coq to be convinced of the safety proofs created by our development.
+ % In our case, the specification of the program we write is this:
+ % given a description of a concrete system, the program . The whole
+ % implementation, including the computational part, serves only to
+ % prove a single theorem. Hence, one essentially only needs to trust
+ % Coq to be convinced of the safety proofs created by our
+ % development.
 
@@
 
-%Wherever we say ``|CR|'', we are really referring to the type |CRplain| of computable reals, part of C-CoRN.
-
-%The computational steps become part of the complete proof of the safety theorem.
+%Wherever we say ``|CR|'', we are really referring to the type
+%|CRplain| of computable reals, part of C-CoRN.
+
+%The computational steps become part of the complete proof of the
+%safety theorem.
 
@@
 
+\weg{
 \paragraph{Organization}
 
-Throughout this report, we present in parallel our general formalized framework (with particular emphasis on differences with Alur's presentation), and the way it is used to obtain a safety proof for an example hybrid system, taken from \cite{alur}, modeling the operation of a thermostat.
-In section \ref{concrete}, we begin by defining the structure of normal (concrete) hybrid systems, their semantics, reachability, and safety.
-We also define the concrete thermostat and the safety condition we wish to prove.
-In section \ref{dn}, we take a moment to discuss some of the limits on decidability of properties of computable real numbers, as these complicate matters in the remainder of the development.
-Next, in section \ref{estimation} we describe notions of underestimation and overestimation that are applied throughout the development.
-Then, in section \ref{abstraction}, we begin our presentation of the abstraction method by discussing abstract regions, states, and spaces.
-We build on these in section \ref{abs.reach}, where we first develop the ``obvious'' notions of abstract system and reachability, discuss why they fail, and then rework them to embrace reachability sharing.
-In section \ref{underestimatingsafety} we show how, assuming abstract reachability is decidable, we may prove hybrid system safety by computation.
-Then, in section \ref{graphs}, we show how an abstract system consisting of transition and initiality ``overestimators'' can be used to decide abstract reachability by construction of a directed graph corresponding to the system, and the use of an ordinary but verified graph reachability algorithm.
-Finally, in sections \ref{over_cont_trans} and \ref{over_disc_trans} we describe the implementation of the aforementioned overestimators.
-We end with conclusions in section \ref{conclusions}.
-
-% proceed to define abstraction bla.
-Following that, we show that a simple (certified) graph reachability can be employed to compute reachability in the graph corresponding to the abstract system, and that the resulting reachability judgement can be translated back through the abstract system all the way to the concrete system.
-We end with conclusions in section \ref{conclusions}.
-
-\section{(Concrete) Hybrid Systems}
-\label{concrete}
-
-We begin by showing our definition of a concrete system, the different parts of which we discuss in the remainder of this section.
-
-\begin{code}
-Record System: Type :=
-  { Point: CSetoid
-  ; Location: Set
-  ; Location_eq_dec: EqDec Location eq
-  ; locations: ExhaustiveList Location
-  ; State := prod Location Point
-  ; initial: State -> Prop
-  ; invariant: State -> Prop
-  ; invariant_initial: initial `subsetof` invariant
-  ; invariant_mor: Morphism (eq ==> cs_eq ==> iff) (curry invariant)
-  ; invariant_stable: forall s, Stable (invariant s)
-  ; flow: Location -> Flow Point
-  ; guard: State -> Location -> Prop
-  ; reset: Location -> Location -> Point -> Point
-  }.
-\end{code}
-
-%  (* this separation of guard and reset seems to cause a problem:
-%    the paper allows having different transitions from a given (l, x) to
-%    some (l', x'), because you can have different transitions. we only allow one! *)
-
-\subsection{States}
-
-A hybrid system is a model of how a software system, described as a finite set of \emph{locations} \footnote{We reserve the term ``state'' for something else.} with (discrete) transitions between them, acts on and responds to a set of continuous variables (called the \emph{continuous state space}), typically representing physical properties of some environment (such as temperature and pressure).
-
-In \cite{alur}, Alur requires that the continuous state space be a convex polyhedron in |pow CR n|.
-In our definition of a hybrid system, we initially let the continuous state space be an arbitrary (constructive) setoid, called |Point|.
-We also explicitly require decidability of location equality.
-
-Our running example (taken from \cite{alur}) is a thermostat whose discrete component consists of three locations (called |Heat|, |Cool|, and |Check|), and whose continuous state space is |pow CR 2|, where |CR| are the \emph{computable} reals.
-The |Heat| and |Cool| locations represent states in which the thermostat heats and cools the environment it operates in, respectively.
-The |Check| state is a self-diagnostic state in which the thermostat does not heat or cool.
-\todo{bah, so awkward!}  The first |CR| component in the continuous state space represents an internally resettable clock, while the second represents the temperature in the environment in which the thermostat operates.
-
-A \emph{State} in a hybrid system is a pair |(l, p)| consisting of a location |l| and a point |p| in the continuous state space (representing an assignment of the continuous variables).
-
-The |invariant| predicate defines, for each location, a set of permitted values for the continuous variables.
-We will use this in the definitions of transition relations in a moment.
-The morphism field expresses that this invariant respects |Point|'s setoid equality.
-We further require that it is \emph{stable}, where |stable P| is defined as |~ ~P -> P|.
-We discuss the reason for this stability requirement in section \ref{dn}.
-
-For the thermostat, the invariant is as follows:
-\begin{code}
-Definition invariant (s: State): Prop :=
-  0 <= clock s /\
-  match location s with
-  | Heat => temp s <= 10 /\ clock s <= 3
-  | Cool => 5 <= temp s
-  | Check => clock s <= 1
-  end.
-\end{code}
-
-Associated with the hybrid system is a set of \emph{initial states}. For our thermostat example, the initial states are
+Throughout this report, we present in parallel our general formalized
+framework (with particular emphasis on differences with Alur's
+presentation), and the way it is used to obtain a safety proof for an
+example hybrid system, taken from \cite{alur}, modeling the operation
+of a thermostat. In section \ref{concrete}, we begin by defining the
+structure of normal (concrete) hybrid systems, their semantics,
+reachability, and safety. We also define the concrete thermostat and
+the safety condition we wish to prove. In section \ref{dn}, we take a
+moment to discuss some of the limits on decidability of properties of
+computable real numbers, as these complicate matters in the remainder
+of the development. Next, in section \ref{estimation} we describe
+notions of underestimation and overestimation that are applied
+throughout the development. Then, in section \ref{abstraction}, we
+begin our presentation of the abstraction method by discussing
+abstract regions, states, and spaces. We build on these in section
+\ref{abs.reach}, where we first develop the ``obvious'' notions of
+abstract system and reachability, discuss why they fail, and then
+rework them to embrace reachability sharing. In section
+\ref{underestimatingsafety} we show how, assuming abstract
+reachability is decidable, we may prove hybrid system safety by
+computation. Then, in section \ref{graphs}, we show how an abstract
+system consisting of transition and initiality ``overestimators'' can
+be used to decide abstract reachability by construction of a directed
+graph corresponding to the system, and the use of an ordinary but
+verified graph reachability algorithm. Finally, in sections
+\ref{over_cont_trans} and \ref{over_disc_trans} we describe the
+implementation of the aforementioned overestimators. We end with
+conclusions in section \ref{conclusions}.
+}
+
+% proceed to define abstraction bla. Following that, we show that a
+% simple (certified) graph reachability can be employed to compute
+% reachability in the graph corresponding to the abstract system, and
+% that the resulting reachability judgement can be translated back
+% through the abstract system all the way to the concrete system. We
+% end with conclusions in section \ref{conclusions}.
+
+\section{Hybrid Systems and the Abstraction method}
+\label{sec:hybsys}
+We briefly introduce the notion of a {\em linear hybrid system}, as
+used by \cite{alur}, and we show how the abstraction method is used in
+that paper in order to verify safety. There are various notions
+(formal definitions) of hybrid systems
+\cite{henziger,lynchvaandrager}. To illustrate the definition, we give
+the example of a thermostat, which is the same as \cite{alur} uses as
+a running example to illustrate the abstraction method. We will use
+the same example to illustrate our formal verification in Coq.
+
+\begin{center}
+\begin{figure}[htb!]
+%\begin{turn}{180}
+\begin{center}
+%    \includegraphics[scale=.6, viewport=0 500 700 710, clip]{FIGS/traject}
+    \includegraphics[scale=.5]{FIGS/ThermoFig}
+\end{center}
+%\end{turn}
+\caption{The Thermostat as an example of a Hybrid Systems \label{fig:thermostat}}
+\end{figure}
+\end{center}
+
+A hybrid system is a model of how a software system, described as a
+finite set of \emph{locations}, $\Loc$ 
+%\footnote{We reserve the term ``state''
+%  for something else.} 
+with (discrete) transitions between them, acts
+on and responds to a set of continuous variables (called the
+\emph{continuous state space}), typically representing physical
+properties of some environment (such as temperature and pressure).
+
+In \cite{alur}, Alur requires that the continuous state space be a
+convex polyhedron in $\IR^ n$. 
+%In our definition of a hybrid system,
+%we initially let the continuous state space be an arbitrary
+%(constructive) setoid, called |Point|. We also explicitly require
+%decidability of location equality.
+In the running example of Figure \ref{fig:thermostat},
+the discrete component consists of three locations (called $\Heat$, $\Cool$,
+and $\Chec$), whose continuous state space is $\IR^2$,
+where $\IR$ are the \emph{computable} reals as defined in CoRN. 
+The $\Heat$ and $\Cool$ locations represent states in which the
+thermostat heats and cools the environment it operates in,
+respectively. The $\Chec$ state is a self-diagnostic state in which
+the thermostat does not heat or cool. The first $\IR$ component in the
+continuous state space represents an internally resettable clock,
+while the second represents the temperature in the environment in
+which the thermostat operates.
+
+A \emph{State} in a hybrid system is a pair $(l, p)$ consisting of a
+location $l$ and a point $p$ in the continuous state space
+(representing the value of the continuous variables). We use $\State$
+for the set of states of a hybrid system.
+
+The {\em invariant\/} predicate defines, for each location, a set of
+permitted values for the continuous variables. 
+\weg{
+We will use this in the
+definitions of transition relations in a moment. The morphism field
+expresses that this invariant respects |Point|'s setoid equality. We
+further require that it is \emph{stable}, where |stable P| is defined
+as |~ ~P -> P|. We discuss the reason for this stability requirement
+in section \ref{dn}.
+}
+For the thermostat, the invariant is as follows: $$\Inv_{\Heat}(c,T) := T\leq 10\wedge c\leq 3,\quad \Inv_{\Cool}(C,t) := 5 \leq T,\quad \Inv_{\Chec} := c\leq 1.$$
+
+A hybrid system also has a set of \emph{initial states}, for which we
+will use the predicate $\Init(s)$.  For the thermostat example,
+$\Init(l,c,T) := l= \Heat \wedge c= 0 \wedge 5 \leq T \leq 10.$
+\weg{
 \begin{code}
@@
 \end{code}
-where |location|, |temp|, and |clock| are the obvious projections. For the thermostat, the requirement that the invariant holds at each initial state is easily proved.
-
-The remaining parts of a hybrid system (i.e.
-|flow|, |guard| and |reset|) describe transitions between states, which, together with the set of initial states, determine the set of \emph{reachable} states, representing the possible behaviors exhibited by a hypothetical real-world implementation of the hybrid system (as software running on a device with sensors and actuators).
-
-\subsection{Flow and Continuous Transitions}
-\label{concreteflow}
-
-Each location in a concrete system has an accompanying \emph{flow function} which describes how the continuous variables change over time while the system is in that location.
-The idea is that the different locations corresponds to different uses of actuators available to the software system, the effects of which are described by the flow function.
-For instance, in our thermostat, the flow function corresponding to the |Heat| location will have the temperature increase with time, modeling the effect of the heater component in our imagined thermostat device.
+where |location|, |temp|, and |clock| are the obvious projections.
+} 
+There is the obvious requirement that the invariant holds at each
+initial state.
+
+The remaining parts of a hybrid system (i.e.\ {\em flow}, {\em guard}
+and {\em reset}) describe transitions between states, which, together
+with the set of initial states, determine the set of \emph{reachable}
+states, representing the possible behaviors exhibited by a
+hypothetical real-world implementation of the hybrid system (as
+software running on a device with sensors and actuators).
+
+%\subsection{Flow and Continuous Transitions}
+%\label{concreteflow}
+
+Each location in a concrete system has an accompanying \emph{flow
+  function} which describes how the continuous variables change over
+time while the system is in that location. The idea is that then
+different locations correspond to different uses of actuators
+available to the software system, the effects of which are described
+by the flow function. For instance, in the thermostat, the flow
+function corresponding to the $\Heat$ location will have the
+temperature increase with time, modeling the effect of the heater
+component in the imagined thermostat device.
 
@@
 
-In the canonical definition of hybrid systems, flow functions are specified as solutions to differential equations describing the dynamics of the continuous variables.
-We follow Alur's example in abstracting from these, taking instead functions |phi| of type |Point -> Duration -> Point| which satisfy:
-\begin{align*}
-  |phi p 0| &= |p| \\
-  |phi p (t + t')| &= |phi (phi t p) t'|
-\end{align*}
-We further require that |phi| is a morphism respecting |Point|'s setoid equality.
-We bundle functions with these properties as a record type called |Flow| (not shown), which is parameterized over the continuous state space type, and is equipped with an implicit coercion to the |phi| function it contains.
-
+In the canonical definition of hybrid systems, flow functions are
+specified as solutions to differential equations describing the
+dynamics of the continuous variables. We follow \cite{alur} in
+abstracting from these, taking instead the solutions of these
+differential equations, which are flow functions $\Phi$ which satisfy:
+\begin{eqnarray*}
+  \Phi( p, 0) &=& p \\
+  \Phi (p, t + t') &=& \Phi (\Phi( p, t), t')
+\end{eqnarray*}
+\weg{
+We further require that $\Phi$ is a morphism respecting |Point|'s
+setoid equality. We bundle functions with these properties as a record
+type called |Flow| (not shown), which is parametrized over the
+continuous state space type, and is equipped with an implicit coercion
+to the |phi| function it contains.
 \todo{Unlike Alur, we don't have ``input'' (why?).}
-
-We now say that there is a (concrete) \emph{continuous transition} from a state |(l, p)| to a state |(l', p')| if |l = l'| and there is a non-negative duration |d| such that |p' = flow l p d| with the invariant for |l| holding at every point along the way:
-
-\begin{code}
-  Definition can_flow (l: Location): relation Point
-    := fun p p' => exists d: Duration, flow l p d = p' /\
-      forall t, 0 <= t <= d -> invariant (l, flow l p t).
-
-  Definition cont_trans: relation State
-    := fun (l, p) (l', p') => l = l' /\ can_flow l p p'.
-\end{code}
-
-In our thermostat example, we express the flow function as the product of two flow functions on |CR|:
-
-\begin{code}
-  Definition thermo_flow (l: Location): Flow (pow CR 2)
-    := product_flow (clock_flow l) (temp_flow l).
-\end{code}
-Here, |product_flow| has type |forall X Y, Flow X -> Flow Y -> Flow (prod X Y)|, and includes the trivial proofs showing that flow in a product space can be formed by combining flows in the respective components.
-
-Intuitively, |clock_flow| and |temp_flow| are the functions
-\begin{code}
-Definition clock_flow (l: Location) (c: CR) (d: Duration): CR :=
-  c + d.
-
-Definition temp_flow (l: Location) (t: CR) (d: Duration): CR :=
-  match l with
-  | Heat => t + 2 * d
-  | Cool => t * exp (-d)
-  | Check => t * exp(- frac 1 2 * d)
-  end.
-\end{code}
-However, to define the system, we actually require |Flow| records carrying additional proofs that these functions are proper flow functions.
-While these proofs could be given in an ad-hoc fashion for this thermostat system, we have instead developed a modest library of reusable flow functions and adaptors that let one compose flow functions that are proper by construction.
-Using these, |clock_flow| and |temp_flow| are defined as:
-\begin{code}
-Definition clock_flow (l: Location): Flow CRasCSetoid :=
-  flow.positive_linear
-
-Definition temp_flow (l: Location): Flow CRasCSetoid :=
-  match l with
-  | Heat => flow.scale 2 flow.positive_linear
-  | Cool => flow.decreasing_exponential
-  | Check => flow.scale (frac 1 2) flow.decreasing_exponential
-  end.
-\end{code}
-
-\subsection{Discrete Transitions}
-\label{conc.disc_trans}
-
-Where continuous transitions describe the flow of continuous variables, \emph{discrete} transitions between locations describe the logic of the software system.
-Each such transition is comprised of two components: a \emph{guard} predicate, and a \emph{reset} function.
-The former defines a subset of the continuous state space in which the transition is enabled (permitted), while the latter describes an instantaneous change applied as a side effect of the transition, as seen in the following definition of the discrete transition relation:
-
+}
+
+We now say that there is a (concrete) \emph{continuous transition}
+from a state $(l, p)$ to a state $(l', p')$ if $l = l'$ and there is a
+non-negative duration $d$ such that $p' = \Phi_l( p,d)$ with the
+invariant for $l$ holding at every point along the way:
+$$ (l,p) \contrans (l',p') := l=l' \wedge \Exists{d\in
+\Dur}{\Phi_l(p, d) = p' \wedge \Forall{t}{0 \leq t\leq d \rightarrow
+\Inv_l(\Phi_l (p, t))}}.$$
+
+In case we are in $\IR ^2$, the flow function can be expressed as the
+product of two flow functions: $\Phi_l(c,T,t) = (\phi_{l,1}(c,T,t),
+\phi_{l,2}(c,T,t))$. In the thermostat example, as in many other
+examples of hybrid systems, $\phi_{l,1}(c,T,t)$ does not depend on $T$
+and $\phi_{l,2}(c,T)$ does not depend on $c$. We call this feature
+{\em separability\/} of the flow function; our development currently
+relies heavily on this property. Separability makes the form of the
+flow functions simpler:
+$$\Phi_l(c,T,t) = (\phi_{l,1}(c,t), \phi_{l,2}(T,t))$$
+
+In the thermostat, for all locations, $\phi_{l,1}(c,d) = c + d$ (for
+the clock) and $\phi_{\Heat,2}(T,d)= T + 2 * d$,
+$\phi_{\\Cool,2}(T,d)= T * e^{-d}$, $\phi_{\Chec,2}(T,d)= T * e^{-
+  \frac{1}{2}d}$.
+
+Where continuous transitions describe the flow of continuous
+variables, \emph{discrete} transitions between locations describe the
+logic of the software system. Each such transition is comprised of two
+components: a \emph{guard} predicate, and a \emph{reset} function. The
+former defines a subset of the continuous state space in which the
+transition is enabled (permitted), while the latter describes an
+instantaneous change applied as a side effect of the transition, as
+seen in the following definition of the discrete transition relation:
+
+$$ (l,p) \distrans (l',p') := \guard_{l,l'}(p) \wedge \reset_{l,l'}(p)= p' \wedge \Inv_l(p) \wedge \Inv_{l'}(p')$$
+
+\weg{
 \begin{code}
@@
 \end{code}
-
-\todo{we don't support multiple discrete transitions between a pair of locations}
-
-The thermostat we are modeling has four transitions, as expressed by the following guard:
-
-\begin{code}
-Definition thermo_guard (s: State) (l: Location): Prop :=
-  match location s, l with
-  | Heat, Cool => 9 <= temp s
-  | Cool, Heat => temp s <= 6
-  | Heat, Check => 2 <= clock s
-  | Check, Heat => frac 1 2 <= clock s
-  | _, _ => False
-  end.
-\end{code}
-
-Our reset function resets the clock for all but one of these transitions, and leaves the temperature variable as is:
-
+}
+
+It will be obvious from Figure \ref{fig:thermostat} what the guards
+and reset functions are. The reset function resets the clock for all
+but one of these transitions, and leaves the temperature variable as
+is. Here we can discern a conceptual distinction between continuous
+variables directly controlled by the system (such as the clock in the
+thermostat), and variables that model a physical phenomenon and that
+can only be observed (such as the temperature in the thermostat). This
+distinction is not made explicit in the definition of a hybrid system;
+in principle, nothing is stopping the thermostat from treating
+temperature as a variable of the former kind and resetting its
+value. However, this would simply make the system unimplementable.
+
+\weg{
 \begin{code}
@@
 \end{code}
-Here we can discern a conceptual distinction between continuous variables directly controlled by the system (such as the clock in our thermostat), and variables that model a physical phenomenon (such as the temperature in our thermostat).
-This distinction is not made explicit in the definition of a hybrid system; in principle, nothing is stopping the thermostat from treating temperature as a variable of the former kind and resetting it to whatever value it pleases.
-However, this would simply make the system unimplementable.
-
-\subsection{Traces and Safety}
-
-A transition is either continuous or discrete:
-\begin{code}
-  Definition trans: relation State := disc_trans `union` cont_trans.
-\end{code}
-% we can't actually do this in the code, but it's plenty clear what this means
-
-We now say that a state |s| is \emph{reachable} if there is an initial state |i| from which one can, by a finite number of transitions, end up in |s|:
+}
+
+%\subsection{Traces and Safety}
+
+A transition is either continuous or discrete $\trans:= \distrans \cup \contrans$.
+We now say that a state $s$ is \emph{reachable} if there is an initial state $i$ from which one can, by a finite number of transitions, end up in $s$:
 %format trans_refl_closure (x) = x "^{*}"
-\begin{code}
-  Definition reachable (s: State): Prop :=
-    exists i: State, initial i /\ trans_refl_closure trans i s.
-\end{code}
-Here, |trans_refl_closure trans| is the transitive reflexive closure of |trans|.
-
-As mentioned before, the set of reachable states represents the possible behaviors exhibited by a hypothetical real-world implementation of the hybrid system (as software running on a device with sensor and actuators).
-
-The idea now is that the purpose of a hybrid system is typically to keep the continuous variables within certain limits.
-In other words, to limit the set of reachable states to some ``safe'' subset of the complete state space.
-For our thermostat example, the intent is to keep the temperature above 4.5 degrees at all times, and so the safe states are defined to be those in which the temperature component is |> 4.5|:
-
-\begin{code}
-  Definition unsafe_thermo_state: Ensemble State :=
-    fun s => temp s <= 4.5.
-\end{code}
-(An |Ensemble T| is just a |T -> Prop|.)
-
-The goal, then, is to verify that the reachable states are a subset of the safe states. Hence, for the thermostat, our main theorem is the following:
-\begin{code}
-  Theorem ThermoSafe: unsafe_thermo_state `subsetof` unreachable.
-\end{code}
-``|A `subsetof` B|'' is just notation for ``|forall x, A x -> B x|'', and |unreachable| is merely the complement of |reachable|.
-
-Since we required that the invariant held at each initial state, and further defined the continous and discrete transitions such that the invariant had to hold everywhere along the path, a simple induction proof shows that
-\begin{code}
-  reachable `subsetof` invariant.
-\end{code}
-\todo{Who cares? Where do we use this?}
-
-Before we continue with our presentation of the abstraction method, we first take a moment to discuss some of the limits on decidability of properties of computable real numbers, as their consequences can be seen in many places in the development.
-Indeed, we already saw one such instance in the stability requirement for invariants.
-
-% The check location seems pretty weird, as if it's only purpose is to let the system potentially fail.
-
-% Alur starts by defining linear hybrid systems as tuples consisting of
-%   * a convex polyhedron subset X of R^n representing the continuous statespace
-%   * a finite set of locations
-%   * a set X_0 subsetof L * X of initial states (where states are pairs containing a location and a continuous state ("point"))
-%   * a function I: L -> C_n giving a linear invariant predicate for each location
-%   * a function f: L -> (X * blabla flow function.
-%   * a relation T subsetof L*L*C_n*yada.
-
-\section{Double Negation and Stability}
+$$\Reach(s):=
+    \Exists{i\in \State}{\Init(i) \wedge i \ttrans s}.$$
+Here, $\ttrans$ is the transitive reflexive closure of $\trans$.
+
+As mentioned before, the set of reachable states represents the
+possible behaviors exhibited by a hypothetical real-world
+implementation of the hybrid system (as software running on a device
+with sensors and actuators).
+
+The idea now is that the purpose of a hybrid system is typically to
+keep the continuous variables within certain limits. In other words,
+to limit the set of reachable states to some ``safe'' subset of the
+complete state space. For the thermostat example, the intent is to
+keep the temperature above $4.5$ degrees at all times, and so the safe
+states are defined to be those in which we have $T > 4.5$: $$\Safe(l,c,T) := T >4.5 \quad and \quad \Unsafe(l,c,T) := T\leq 4.5.$$
+The goal, then, is to verify that the reachable states are all safe.
+
+\subsection{The Abstraction Method}
+There are uncountably many traces in a hybrid system, so safety is
+undecidable in general. In concrete situations, safety may be (easily)
+provable, however, for example, if one finds the proper {\em proof
+  invariant}. As proof-invariants are often hard to find, one would
+like to have more automated heuristic methods for establishing
+safety. This is what the {\em predicate abstraction method\/} of
+\cite{alur} does. In general one divides the continuous state space
+into a finite number of convex subsets (polygons), $A_1, \ldots, A_n$,
+which yields a finite {\em abstract state space}, $\AState := \{ (l,
+A_i) \mid l \in \Loc, 1\leq i \leq n \}$, with an obvious embedding
+$A: \State \to \AState$ of concrete states into abstract states. On this
+abstract state space, one immediately defines {\em abstract continuous
+  transitions\/} and {\em abstract discrete transitions\/} as follows.
+\begin{eqnarray*}
+(l,P) \acontrans (l',Q) &:=& l=l' \wedge \Exists{p\in P, q\in Q}{(l,p) \contrans (l,q)}\\
+(l,P) \adistrans (l',Q) &:=& \Exists{p\in P, q\in Q}{(l,p) \distrans (l,q)}.
+\end{eqnarray*}
+Now {\em abstract reachability\/} can be defined in the obvious way as
+follows $\AReach(a) := \Exists{s_0\in\State}{\Init(s_0) \wedge A(s_0)
+\attrans a}$, and also the predicates $\ASafe$ and $\AUnsafe$, stating
+when abstract states are safe / unsafe are defined as expected.
+
+This yields a finite transition system (where it may be undecidable if
+there is a transition between two states!) which has the
+obvious property indicated in Figure \ref{diag:abstraction}.
+\begin{figure}[htb!]
+$$\begin{diagram}
+ s & \rArr_{CD} & s'&& s_0& \rArr_{CD} & s_1& \rArr_{CD} & s_2& \rArr_{CD} & \ldots\\
+\dTo &         &\dTo &&\dTo & &\dTo &&\dTo & &\\
+A(s)& \rArr^A_{CD} & A(s') && A(s_0)& \rArr^A_{CD} & A(s_1)& \rArr^A_{CD} & A(s_2)& \rArr^A_{CD} & \ldots\\
+\end{diagram}$$
+\caption{The abstraction function \label{diag:abstraction}}
+\end{figure}
+With this property, one can conclude safety of the concrete system
+from safety of the abstract system: $$\text{if}\ \Forall{a\in \AState}{\AReach(a) \rightarrow \ASafe(a)}, \text{then}\ \Forall{s\in \State}{
+\Reach(s) \rightarrow \Safe(s)}.$$
+
+The interest and power of the abstraction method lies in two
+facts. First, we do not need the {\em exact\/} definitions of
+$\acontrans$ and $\adistrans$ to conclude safety of the concrete
+system from safety of the abstract system. We only need the property
+of diagram \ref{diag:abstraction}, so we can {\em overestimate\/}
+$\acontrans$ and $\adistrans$ (that is: replace it with a transition
+relation that allows more transitions). Second, there is a fairly good
+heuristic for making the abstraction (how to divide the continuous
+state space into regions) and how to decide whether there should be an
+abstract transition from one abstract state to another.
+This is indicated in figure \ref{fig:abstraction}.
+
+\begin{center}
+\begin{figure}[htb!]
+%\begin{turn}{180}
+%    \includegraphics[scale=.6, viewport=0 500 700 710, clip]{FIGS/traject}
+    \includegraphics[scale=.5]{FIGS/traject}
+    \includegraphics[scale=.4]{FIGS/AThermo-new3}
+%\end{turn}
+\caption{The abstraction function computed
+\label{fig:abstraction}}
+\end{figure}
+\end{center}
+
+On the left side we schematically indicate how one can try to
+``follow'' the path that region $A$ takes when time passes and verify
+whether this path intersects with region $B$. The idea of
+overestimation is that one should only {\em not\/} put an arrow $A
+\acontrans B$ if we positively know that the path of $A$ will avoid
+$B$, which is the case for example for paths that follow the dashed lines.
+
+On the right side we see the abstract state space indicated for the
+location $\Heat$. The abstract state space consists of rectangles,
+possibly degenerated (extending to $-\infty$ or $+\infty$). For the
+bounds of these rectangles one takes the values that occur in the
+specification of the hybrid system (see Figure
+\ref{fig:thermostat}). According to \cite{alur}, this is a good
+choice. (In case one cannot prove safety in this, there is of course
+the opportunity for the user to refine it.)  The gray area indicates
+that from these states also abstract discrete transitions are
+possible. The dashed area is unreachable, because of the invariant for
+the $\Heat$ location %prevents these states from being reachable.  
+All the abstract
+transitions from the rectangle $[0.5,1) \times[5,6)$ are shown: as the
+    temperature flow function for $\Heat$ is $\phi_{\Heat,2}(T,d)= T +
+    2 * d$, and the clock flow function is $\phi_{\Heat,1}(c,d) = c +
+    d$, these are all the possible abstract transitions.
+
+There is a slight subtlety with the abstract transitions that we want
+to consider. If we just take $\attrans$ to be the reflexive transitive
+closure of $\acontrans \cup \adistrans$, we get far too many abstract
+traces. (There is no problem with the correctness of the method --
+Figure \ref{diag:abstraction} still applies -- but we get way too many
+traces to be able to prove any system safe.) The problem lies in the
+fact that there is no transition from any point in $[0.5,1)
+  \times[5,6)$ to $[0.5,1) \times[9,10)$, but the transitive closure
+        of $\adistrans$ gives us such a transition anyway. 
+% \ADAM{I will try to add some more explanation}
+Using this method, \cite{alur} proves the correctness of the thermostat.
+
+\weg{
+\section{Motivation}
+
+1. Verify the correctness of the abstraction itself
+
+2. Avoid Floating Point arithmetic
+
+3. Avoid correctness proofs of systems that are ``almost wrong''.
+   (Add one arrow to Thermostat!)
+
+- One cannot decide x>0 of a measured value
+- One can't define a function on the basis of distinguishing x<=y \/ x >y
+}
+
+\section{Formalization}
+We now describe the formalization in Coq and the design choices
+made. Some of the design choices were due to specifics in Coq or its
+type theory CiC, while other are due to our choice to work with exact
+real arithmetic and a constructive logic. We will mostly discuss the
+latter design choices, because we feel they are most relevant.
+
+\subsection{(Concrete) Hybrid Systems}
+\label{concrete}
+
+We begin by showing our definition of a concrete system, the different
+parts of which we discuss in the remainder of this section.
+
+\begin{code}
+Record System: Type :=
+  { Point: CSetoid
+  ; Location: Set
+  ; Location_eq_dec: EqDec Location eq
+  ; locations: ExhaustiveList Location
+  ; State :=  Location x Point
+  ; initial: State -> Prop
+  ; invariant: State -> Prop
+  ; invariant_initial: initial subsetof invariant
+  ; invariant_stable: forall s, Stable (invariant s)
+  ; flow: Location -> Flow Point
+  ; guard: State -> Location -> Prop
+  ; reset: Location -> Location -> Point -> Point
+  }.
+\end{code}
+
+This Coq definition defines the record type of ``Systems'', which
+contains a field ``Point'', representing the continuous state space
+and a field ``Location'', representing the set of locations. Here, we
+take Point to be an arbitrary Constructive Setoid, which is basically
+just a type with an equality on it. For Location, we assume a
+decidable equality and we assume a finite enumeration (``exhaustive
+list'') of locations. The other parts are as expected (``Flow Point''
+is the type of flow functions on the type ``Point''), except for the
+requirement that the invariant $\Inv_l$ is ``Stable'' for every
+location $l$.
+
+\subsection{Stability and Double Negation}
 \label{dn}
 
-%format le_lt_dec_overlap = le_lt_dec"_{"overlap"}"
-%format le_lt_dec_DN = le_lt_dec"_{DN}"
-
-One obvious and useful property we can derive is transitivity of the continuous transition relation, the proof of which reveals the need for invariant stability.
-Suppose we have |cont_trans (l, p) (l', p')| and |cont_trans (l', p') (l'', p'')| for locations |l, l', l''| and points |p, p', p''|.
-To show |cont_trans (l, p) (l'', p'')|, we must show two things.
-The first, |l = l''|, follows immediately from transitivity of |=|.
-For the second, |can_flow l p p''|, we simply take the flow duration to be the sum of the durations from |p| to |p'| and from |p'| to |p''| (call these |d| and |d'|, respectively), and observe:
-\begin{align}
-  |flow l p (d + d')| & = |flow l (flow l p d) d'| \nonumber \\
-    & = |flow l p' d'| \nonumber \\
-    & = |p''| \nonumber
-\end{align}
-What remains is to show that the invariant holds at each point along the way. That is,
-\begin{code}forall t, 0 <= t <= d + d' -> invariant (l, flow l p t).\end{code}
-From |can_flow p p'| we know this is true for |0 <= t <= d|, and from |can_flow p' p''| we know that this is true for |d <= t <= d + d'|.
-Classically, then, the proof is a done deal, for one can simply distinguish cases |t <= d| and |d < t|.
-
-Unfortunately, such case distinction is a luxury we do not have, because for computable reals, the proposition
-\begin{code}
-  le_lt_dec: forall x y: CR, x <= y \/ y < x.
-\end{code}
-is not provable.
-After all, to have a constructive proof of a disjunction |A\/B| (where A and B are arbitrary propositions/types) is to have either a proof of |A| or a proof of |B|.
-Similarly, to have a proof of |X -> A \/ B| is to have a function that, given an |X|, either returns a proof of |A|, or a proof of |B|.
-With this in mind, suppose we try to implement |le_lt_dec|.
-Then given |x| and |y| in |CR|, we are to produce a proof either of |x <= y| or of |y < x|.
-Unfortunately, the nature of computable reals only lets us observe arbitrarily close approximations of |x| and |y|.
-Now suppose |x = y|.
-Then no matter how closely we approximate |x| and |y|, the error margins (however small) will always leave open the possibility that |y| is really smaller than |x|.
-Consequently, we will never be able to definitively conclude that |x <= y|.
-
-Computable reals do admit two variations on the proposition:
+A property $P$ ({\tt P: Prop}) is {\em stable\/} if $\neg \neg P \rightarrow P$ holds. 
+Classically, this is trivially true, but constructively
+not. 
+%Especially when we work with real numbers, we don't have $\leltdec: \forall
+%x, y \in \IR (x\leq y \vee y<x)$. This conforms with the view of real
+%numbers as an infinite sequence of better approximations: if $x$
+%converges to $0$ we will not be able to decide $x\leq 0 \vee 0<x$ in a
+%finite amount of time.
+
+In constructive logic, a proof of a
+disjunction $A\vee B$ (where $A$ and $B$ are arbitrary propositions) is
+either a proof of $A$ or a proof of $B$. Similarly, to have a
+proof of $X \rightarrow A \vee B$ is to have a function that, given an $X$,
+either returns a proof of $A$, or a proof of $B$. With this in mind,
+suppose we try to implement $\leltdec: \forall
+%x, y \in \IR (x\leq y \vee y<x)$. Then given $x$ and $y$ in
+$\IR$, we are to produce a proof either of $x \leq y$ or of $y <
+x$. Unfortunately, the nature of computable reals only lets us observe
+arbitrarily close approximations of $x$ and $y$. Now suppose $x =
+y$. Then no matter how closely we approximate $x$ and $y$, the error
+margins (however small) will always leave open the possibility that
+$y$ is really smaller than $x$. Consequently, we will never be able to
+definitively conclude that $x \leq y$.
+
+Computable reals do admit two variations of the proposition:
 \begin{enumerate}
-\item |le_lt_dec_overlap: forall x y: R, x < y -> forall z, z <= y \/ x <= z|
-\item |le_lt_dec_DN: forall x y: R, ~ ~ (x <= y \/ y < x)|
+\item $\leltdecoverlap: \forall x y: \IR. ( x < y \rightarrow \forall z. (z \leq y \vee x \leq z))$
+\item $\leltdecDN: \forall x y: \IR. \neg  \neg  (x \leq y \vee y < x)$
 \end{enumerate}
 
-Both are weaker than the original, and are less straightforward to use.
-Nevertheless, this is the path we will take in our development, partly out of necessity (because in some parts of the development, we really need to ``run'' these lemmas to obtain |<=|/|<| proofs), and partly because just taking |le_lt_dec| as an axiom amounts to cheating.
-\todo{the former is a lie.
-paste the new perspective as written in mail to Herman, and/or stuff from our paper}
-
-For our transitivity proof, we will use the variant expressed using double negation.
-Two questions immediately arise when considering this variant.
-First, why does the double negation make it provable?
-And second, how does one actually use this doubly negated variant in proofs?
-
-For the first, we need only observe that |x <= y| is equivalent to |y </ x|, and that the law of the excluded middle holds under double negation. That is,
-\begin{code}
-  forall P, ~ ~ (P \/ ~ P)
-\end{code}
-is a trivial tautology considering that |~ P| is taken to mean |P -> False|.
-
-One practical way to answer the question of how such a doubly negated proposition might be used to prove things is to observe that double negation, as a function on types/propositions, is a monad \cite{bla}.
-Writing |DN P| for |~ ~ P|, we have the following two key operations that make |DN| a monad:
+Both are weaker than the original, and are less straightforward to
+use. Nevertheless, this is the path we will take in our development,
+because just taking $\leltdec$ as an axiom amounts to cheating.
+%and partly because in some parts of the development, we may need to
+%really ``run'' these lemmas to obtain $\leq$/$<$ proofs or to do a
+%case distinction in an algorithm. Having an axiom, disables such
+%algorithms to complete.
+
+A place where we definitely need the stability of the invariant is in
+proving the obvious and useful property of {\em transitivity of the
+  continuous transition relation}: Suppose we have $(l, p) \contrans
+(l, p')$ and $(l, p')\contrans (l, p'')$ for location $l$ and points
+$p, p', p''$. 
+To show $(l, p)\contrans (l, p'')$, we must show (among other things)
+%that
+%$\Exists{d''\in \Dur}{\Phi_l(p, d'') = p''}$ -- for which we
+%simply take $d''$ to be the sum of the durations from $p$
+%to $p'$ and from $p'$ to $p''$ (call these $d$ and $d'$,
+%respectively), and observe:
+%\begin{eqnarray*}
+%  \Phi_l(p, d + d') & =& \Phi_l( \Phi_l (p, d), d')\\
+%    & =& \Phi_l(p', d')\\
+%    & =& p'' \nonumber
+%\end{eqnarray*}
+%What remains is to show 
+that the invariant holds at each point along the way. That is,
+$$\forall t, 0 \leq t \leq d + d' \rightarrow \Inv_l( \Phi_L(p, t)).$$
+We know this is true for $0 \leq t \leq d$, and we know that this is
+true for $d \leq t \leq d + d'$. Classically, then, the proof is done,
+for one can simply distinguish cases $t \leq d$ or $d < t$.  In our
+transitivity proof, we will use the variant expressed using double
+negation. Now we need only observe that $x \leq y$ is equivalent to $\neg(y < x)$, and that the law of the excluded middle holds under double negation. That is,
+$\forall P( \neg  \neg  (P \vee \neg  P))$
+is a trivial tautology.
+
+A question that immediately arises is: How does one actually use this
+doubly negated variant in proofs?  One practical way is to observe
+that double negation, as a function on propositions, is a monad
+\cite{wadler95monads}. Writing $\DN P$ for $\neg  \neg  P$, we have the following two key
+operations that make $\DN$ a monad:
 
@@
 %format bind_DN
-
 \begin{code}
@@
 \end{code}
-The first expresses that any previously obtained result can always be inserted ``into'' the monad.
-The second expresses that results inside the monad may be used freely in proofs of additional properties in the monad.
-For instance, one may |bind_DN| a proof of |DN (x <= y \/ y < x)| (obtained from |le_lt_dec_DN| above) with a proof of |(x <= y \/ y < x) -> DN P|, yielding a proof of |DN P|.
-
-Thus, |DN| establishes a ``proving context'' in which one may make use of lemmas yielding results inside |DN| that may not hold outside of it (such as |le_lt_dec_DN|), as well as lemmas yielding results not in |DN|, which can always be injected into |DN| using |return_DN|.
-The catch is that such proofs always end up with results in |DN|, which begs the question: what good is any of this?
-In particular, can |le_lt_dec_DN| be used to prove anything not doubly negated?
-
-As it happens, there is a class of \emph{stable} propositions that are equivalent to their own double negation.
-Examples include negations, non-strict inequalities on real numbers, and any decidable proposition.
-
-We now see why we required invariant stability in section \ref{concrete}: in the transitivity proof for |cont_trans|, it allows us to employ |le_lt_dec_DN| to do case distinction on the |t| variable when showing that the invariant holds at each point along the composite path.
-That is, we simply bind |le_lt_dec_DN t d| of type |DN (t <= d \/ d < t)| with the straightforward proof of |(t <= d \/ d < t) -> DN (invariant (l, flow system l p t))|, and then pull the latter out of |DN| on account of its stability.
-
-Invariants are typically conjunctions of inequalities, which are stable only if the inequalities are non-strict.
-Hence, the limits on observability of computable real numbers ultimately mean that our development cannot cope with hybrid systems whose location invariants use strict inequalities.
-We feel that this is not a terrible loss.
-In section \ref{abstraction} we will see analogous limitations in the choice of one's abstraction parameters.
-
-% Since in our development we do not want to use any axioms,
-%  Since in development we actually need to \emph{run} these functions, we have no choice but to 
-% In some cases, the former is actually a perfectly usable substitu
-% In the first alternative, we change the problem from "deciding whether one number is less than another" to "deciding whether, given two numbers 
-% because the error margins of our approximations will always
-% we will never be able to decisively conclude |x <= y|, because the possibility that
-%  can never be excluded from these approximations.
-% but with a difference smaller .
-% Writing x <= y as ~ y < x
-
-\section{Underestimation and Overestimation}
+
+The first expresses that any previously obtained result can always be
+inserted ``into'' the monad. The second expresses that results inside
+the monad may be used freely in proofs of additional properties in the
+monad. For instance, one may $bind_{\DN}$ a proof of $\DN (x \leq y \vee y < x)$ 
+(obtained from $\leltdecDN$ above) with a proof of $(x \leq y \vee y < x) \rightarrow \DN P$, yielding a proof of $\DN P$.
+
+Thus, $\DN$ establishes a ``proving context'' in which one may make
+use of lemmas yielding results inside $\DN$ that may not hold outside
+of it (such as $\leltdecDN$), as well as lemmas yielding results
+not in $\DN$, which can always be injected into $\DN$ using
+$return_{\DN}$. The catch is that such proofs always end up with results
+in $\DN$, which begs the question: what good is any of this? In
+particular, can $\leltdecDN$ be used to prove anything not doubly
+negated?
+
+As it happens, there is the class of \emph{stable} propositions (that
+are equivalent to their own double negation). Examples include
+negations, non-strict inequalities on real numbers, and any decidable
+proposition.
+
+We have seen why we required invariant stability:
+%in section \ref{concrete}: 
+in the transitivity proof for continuous transitions
+it allows
+us to employ $\leltdecDN$ to do case distinction on the $t$ variable
+when showing that the invariant holds at each point along the
+composite path. That is, we simply bind $\leltdecDN t d$ of type
+$\DN (t \leq d \vee d < t)$ with the straightforward proof of $(t \leq d
+\vee d < t) \rightarrow \DN (\Inv_l(\Phi_l (p, t)))$, and
+then pull the latter out of $\DN$ on account of its stability.
+
+Invariants are typically conjunctions of inequalities, which are
+stable only if the inequalities are non-strict. Hence, the limits on
+observability of computable real numbers ultimately mean that our
+development cannot cope with hybrid systems whose location invariants
+use strict inequalities. We feel that this is not a terrible loss. In
+section \ref{abstraction} we will see analogous limitations in the
+choice of one's abstraction parameters.
+
+%An important point to make here is that the $\DN$ version is also
+%\emph{sufficient\/} for proving safety, because the ultimate goal to
+%prove is $\forall s\in \State( \Unsafe(s) \rightarrow \neg\Reach(s))$.
+%This is a stable proposition, because the conclusion is a negated
+%proposition, so is we have prove it ``under $\DN$'' we can extract it
+%and conclude safety.
+
+\subsection{Abstract Hybrid Systems}
+\label{abshybsys}
+\label{abstraction}
+We now want to define an abstract system and an abstraction function
+satisfying the properties indicated in diagram
+\ref{diag:abstraction}. However, this is not possible, because we
+cannot make a case distinction like $x\leq 0 \vee 0< x$ and therefore
+we can't define a function that sends a point $(c,T)$ to the rectangle $R$
+it is in. We can define a function that approximates a point $(c,T)$
+up to, say $\epsilon$ ($\epsilon >0$) and then decides to send that
+point to the rectangle $R$ the approximation is in. This implies that,
+when one is close to the edge of a rectangle,
+\begin{itemize}
+\item 
+different {\em representations\/} of the same point $(c,T)$ may be
+sent to different rectangles,
+\item
+a point that is $<\epsilon$ outside the rectangle $R$ may still be sent to $R$.
+\end{itemize}
+The second is very problematic, because the property for the
+abstraction function $A$, as depicted in diagram
+\ref{diag:abstraction} is now just false.
+
+These may seem just anoying problems,
+caused by the fact that our logic is constructive, but we claim it is
+not. It is is a fact of physical reality that one can only get a
+finite approximation of a measured value, so case distinctions like
+$x\leq 0 \vee 0< x$ are unrealistic. Moreover, we claim that the
+classical abstraction method allows one to prove the safety of systems
+that are unreliable in practice. We will not expand on this here, but
+suppose we add a fourth location $\Off$ to the thermostat of Figure
+\ref{fig:thermostat}, with $\cdot{T} = -1$, $\cdot{c} =1$ and an arrow
+from $\Heat$ to $\Off$ with guard $c\geq 2 \wedge T<9$. Clearly, if
+the system can end up in location $\Off$, its is unsafe. Now, using
+the classical abstraction method, there is no transition to any state
+involving location $\Off$ from the initial state, because as soon as
+$c\geq 2$, $T\geq 9$. However, when we get close to $c=2$, any small
+mistake in the measurement of $T$ may send the system to $\Off$,
+making the whole hybrid system very unreliable.
+
+The good thing is that we don't really need the property in Diagram
+\ref{diag:abstraction}, because we can do something better. 
+\begin{itemize}
+\item
+We let regions in the abstract hybrid systems overlap (ideally as
+little as possible, e.g.\ only at the edges).
+
+\item
+We replace the {\em abstract continuous transition\/} relation
+$\acontrans$ by a function $\overcontrans$ that takes a region $R_0$
+as input and outputs a {\em list of regions\/} including $R_0$: $(R_o,
+R_1, \ldots, R_n)$ in such a way that $\cup_{0\leq i\leq n}R_i$ is an
+overapproximation of the set of states reachable by a continuous step
+from a state in $R_0$.\\
+We do something similar for the {\em abstract discrete transition}, and obtain $\overdistrans :\Region \rightarrow List \Region$.
+\item
+We loosen the requirement on the abstraction function $A$; for $s \in \State$, we only require $\DN {\exists r\in \Region}{(s\in r)}$.
+\end{itemize}
+
+So, to summarize, if $s\contrans s'$, then we don't require $s'$ to
+$A(s')$ to be in the list $\overcontrans(A(s))$, but we just require
+$s'$ to be in the $\bigcup \overcontrans(A(s))$. (Basically it should
+be in one of the regions in the list $\overcontrans(A(s))$.) This
+simple change relieves us from having to determine the exact regions
+that points are in: they just should be covered. The functions
+$\overcontrans$ and $\overdistrans$ yield a notion of {\em trace\/} in the abstract
+hybrid system in the obvious way: starting from $R_0$, take an $R_1$ in $\overcontrans( R_0)$, then an $R_2$ in $\overdistrans(R_1)$, and so forth.
+
+\weg{
+The abstraction method for verification of (concrete) hybrid system safety as we implement it can be summarized as follows:
+\begin{enumerate}
+\item build an abstract hybrid system corresponding to the concrete hybrid system;
+\item 
+\label{steptwo} 
+show that each trace in the concrete system corresponds to some trace
+in the abstract system (and as a corollary, that reachability in the
+concrete system implies reachability in the abstract system, and most
+importantly, that unreachability in the abstract system implies
+unreachability in the concrete system);
+\item 
+\label{stepthree} 
+run a certified graph reachability algorithm on the (finite, discrete)
+abstract system to verify that no unsafe \emph{abstract} states are
+reachable;
+\item 
+conclude from \ref{steptwo} and \ref{stepthree} that no unsafe
+\emph{concrete} states are reachable either.
+\end{enumerate}
+}
+
+Whereas in a concrete hybrid system states consist of a location
+paired with a point in the continuous state space, in an abstract
+hybrid system states consist of a location paired with the ``name'' of
+a region corresponding to a subset of the continuous state space.
+
+From now on we will use a ``|concrete.|'' prefix for names like
+|State| defined in section \ref{concrete}, which now have abstract
+counterparts. |Region| is a field from a record type |Space| bundling
+region sets with related requirements:
+
+\begin{code}
+  Record Space: Type :=
+    { Region: Set
+    ; Region_eq_dec: EqDec Region eq
+    ; regions: ExhaustiveList Region
+    ; NoDup_regions: NoDup regions
+    ; in_region: Container Point Region
+    ; regions_cover: forall (l: Location) (p: Point),
+        invariant (l, p) -> DN { r: Region | p `elem` r }
+    }.
+\end{code}
+
+The |Container Point Region| type specified for |in_region| reduces to
+|Point -> Region -> Prop|. |Container| is a type class that provides
+the notation ``|x `elem` y|'', prettier than ``|in_region x
+y|''. 
+%|in_region_mor| states that |in_region| respects |Point|'s
+%setoid equality.
+
+|regions_cover| expresses that each concrete point belonging to a
+valid state must be represented by a region---a crucial ingredient
+when arguing that unreachability in the abstract system implies
+unreachability in the concrete system. The double negation in its
+result type is both necessary and sufficient:
+
+It is \emph{necessary} because |regions_cover| boils down to a
+(partial) function that, given a concrete point, must select an
+abstract region containing that point. This means that it must be able
+to decide on which side of a border between two regions the given
+point lies. As we saw in section \ref{dn}, that kind of decidability
+is only available inside |DN| unless all region borders have
+nontrivial overlap, which as we will see later is undesireable.
+
+Fortunately, the double negation is also \emph{sufficient}, because we
+will ultimately only use |regions_cover| in a proof of of |... -> ~
+concrete.reachable s| (for some universally quantified variable s),
+which, due to its head type being a negation, is stable, and can
+therefore be proved in and then extracted from |DN|. Hence, we only
+need |regions_cover|'s result in |DN|.
+
+\subsection{Underestimation and Overestimation}
 \label{estimation}
@@
 % It \emph{will} always terminate, so we are not writing a semidecision procedure either.
-Ultimately, in our development we are writing a program that \emph{attempts} to produce hybrid system safety proofs.
-Importantly, we are \emph{not} writing a complete hybrid system safety decision procedure: if the concrete system is unsafe or the abstraction method fails, our program will simply not produce a safety proof.
-It might seem, then, that we are basically writing a \emph{tactic} for a particular problem domain.
-However, tactics in Coq are normally written in a language called Ltac, and typically rely on things like pattern matching on syntax.
-Our development, on the other hand, is very much written in regular Gallina, with hardly any significant use of Ltac.
-This was never a conscious design decision though---it is just the way the development let itself be written.
-\todo{While I know too little about complex use of Ltac to know whether heavier use of it might have been useful, there should definitely be more to say about this.} In any case, to characterize tactic-like functions in regular Gallina, we define |underestimation P| to be either a proof of |P|, or not.
-In Coq, either of the following will do:
+Ultimately, in our development we are writing a program that
+\emph{attempts} to produce hybrid system safety proofs. Importantly,
+we are \emph{not} writing a complete hybrid system safety decision
+procedure: if the concrete system is unsafe or the abstraction method
+fails, our program will simply not produce a safety proof. It might
+seem, then, that we are basically writing a \emph{tactic} for a
+particular problem domain. However, tactics in Coq are normally
+written in a language called Ltac, and typically rely on things like
+pattern matching on syntax. Our development, on the other hand, is
+very much written in regular Gallina, with hardly any significant use
+of Ltac.
+
+We define |underestimation P| to be either a proof of
+|P|, or not. In Coq, either of the following will do:
 \begin{code}
@@
 \end{code}
-The latter tends to work better with the |Program| family of commands \cite{bla} which have special support for dependent pairs.
-Using |Program|, an underestimation of the second variety may be provisionally defined strictly as a bool, and then separately proved to be a valid underestimation in a proof obligation generated by |Program|.
-The second form also nicely illustrates why we call this an underestimation: it may be |false| even when |P| holds.
-We can now describe the functionality of our program by saying that it underestimates hybrid system safety, yielding a term of type |underestimation Safe|, where |Safe| is a proposition expressing safety of a hybrid system.
-
-Considered as theorems, underestimations are not very interesting, because they can be trivially ``proved'' by taking the |false|/|None| estimation.
-Hence, the value of our program is not witnessed by the mere fact that it manages to produce terms of type |underestimation Safe|, but rather by the fact that when run, it actually manages to returns |true|/|Just P| for the hybrid system we are interested in (i.e.
-the thermostat).
-Again, this is typical tactic stuff.
-It is for this reason that we primarily think of the development as a program rather than a proof, even though the program's purpose is to produce proofs.
-
-\
-
-\noindent The opposite of an underestimation is an overestimation:
+The latter tends to work better with the |Program| family of commands
+\cite{sozeau} which have special support for dependent pairs. Using
+|Program|, an underestimation of the second variety may be
+provisionally defined strictly as a bool, and then separately proved
+to be a valid underestimation in a proof obligation generated by
+|Program|. The second form also nicely illustrates why we call this an
+underestimation: it may be |false| even when |P| holds. We can now
+describe the functionality of our program by saying that it
+underestimates hybrid system safety, yielding a term of type
+|underestimation Safe|, where |Safe| is a proposition expressing
+safety of a hybrid system.
+
+Considered as theorems, underestimations are not very interesting,
+because they can be trivially ``proved'' by taking
+|false|/|None|. Hence, the value of our program is not witnessed by
+the mere fact that it manages to produce terms of type
+|underestimation Safe|, but rather by the fact that when run, it
+actually manages to returns |true|/|Just P| for the hybrid system we
+are interested in (e.g.\ the thermostat). It is for this reason that
+we primarily think of the development as a program rather than a
+proof, even though the program's purpose is to produce proofs.
+The opposite of an underestimation is an overestimation:
 \begin{code}
@@
 \end{code}
-(Here, too, one could use an option type: |option (~ P)|.)
-
-Since hybrid system safety is defined as unreachability of unsafe states, we may equivalently express the functionality of our development by saying that it overestimates unsafe state reachability.
-Indeed, most subroutines in our programs will be overestimators rather than underestimators.
-Notions of overestimation and underestimation trickle down through all layers of our development, down to basic arithmetic.
-For instance, we employ functions such as:
+
+Since hybrid system safety is defined as unreachability of unsafe
+states, we may equivalently express the functionality of our
+development by saying that it overestimates unsafe state
+reachability. Indeed, most subroutines in our programs will be
+overestimators rather than underestimators. Notions of overestimation
+and underestimation trickle down through all layers of our
+development, down to basic arithmetic. For instance, we employ
+functions such as:
 
@@
 \end{code}
-As discussed in the last section, |CRle_unapplied| is not decidable.
-|overestimate_CRle| merely makes a ``best effort'' to prove |~ (CRle x y)| using |epsilon|-approximations.
-A smaller |epsilon| will result in fewer spurious |true| results.
+As discussed earlier, |CRle_unapplied| is not
+decidable. |overestimate_CRle| merely makes a ``best effort'' to prove
+|~ (CRle x y)| using |epsilon|-approximations. A smaller |epsilon|
+will result in fewer spurious |true| results.
 
@@
 
-The types of underestimators and overerestimators often merely repeat a certain predicate's parameters, as in the |overestimate_CRle| example above.
-Using a bit of type class magic to achieve variadicity, we can rid us of this repetition.
-Given a predicate |P: A_0 -> ...
--> A_n -> Prop|, we define |underestimator P| as the type \begin{code}forall (a_0: A_0) ...
-(a_n: A_n), underestimation (P a_0 ...
-a_n).\end{code} (And similar for |overestimator|.) With these, the type of |overestimate_CRle| may be written as |Qpos -> overestimator (CRle_unapplied)|.
-
+\weg{
+The types of underestimators and overerestimators often merely repeat
+a certain predicate's parameters, as in the |overestimate_CRle|
+example above. Using a bit of type class magic to achieve variadicity,
+we can rid us of this repetition. Given a predicate |P: A_0 -> ... ->
+A_n -> Prop|, we define |underestimator P| as the
+type 
+\begin{code}forall (a_0: A_0) ... (a_n: A_n), underestimation (P a_0 ... a_n).\end{code} 
+(And similar for |overestimator|.) With
+these, the type of |overestimate_CRle| may be written as |Qpos ->
+overestimator (CRle_unapplied)|.
 We use the same techniques for real decision procedures (in the few places we use those): we define |decision P| as |{P}+{~P}|, and given a predicate |P: A_0 -> ...
@@
 
-\subsection{Automating estimator/decider composition}
-
-Underestimators, overestimators and deciders can be combined to form underestimators (resp.
-overestimators, deciders) for things like conjunctions and quantifications over finite domains.
-For instance, here is a combinator that forms conjunction overestimators:
+%\subsection{Automating estimator/decider composition}
+
+Underestimators, overestimators and deciders can be combined to form
+underestimators (resp. overestimators, deciders) for things like
+conjunctions and quantifications over finite domains. For instance,
+here is a combinator that forms conjunction overestimators:
 
@@
 \end{code}
-\todo{clean up proof}
-Applying these combinators by hand to form estimators/deciders for composite propositions is somewhat tedious, but fortunately this process can be automated if we make |overestimation|, |underestimation|, and |decision| type classes:
+%\todo{clean up proof}
+Applying these combinators by hand to form estimators/deciders for
+composite propositions is somewhat tedious, but fortunately this
+process can be automated if we make |overestimation|,
+|underestimation|, and |decision| type classes:
 
@@
 \end{code}
-If we now say |overestimate P|, the type class instance resolution mechanism will try to find (or build) a term of type |overestimation P| by recursively applying declared type class instances.
-For instance, if we declare |overestimate_conj| above as a type class instance (instead of a plain definition) and then say |overestimate (A /\ B)| where |A| and |B| are already known to be overestimatable (meaning the resolution mechanism is able to build terms of types |overestimation A| and |overestimation B|), then the resolution mechanism will be able to build a term of type |overestimation (A /\ B)| by applying |overestimate_conj|.
+
+If we now say |overestimate P|, the type class instance resolution
+mechanism will try to find (or build) a term of type |overestimation
+P| by recursively applying declared type class instances. For
+instance, if we declare |overestimate_conj| above as a type class
+instance (instead of a plain definition) and then say |overestimate (A
+/\ B)| where |A| and |B| are already known to be overestimatable
+(meaning the resolution mechanism is able to build terms of types
+|overestimation A| and |overestimation B|), then the resolution
+mechanism will be able to build a term of type |overestimation (A
+/\ B)| by applying |overestimate_conj|.
 
@@
 \end{code}
-To construct this overestimator, Coq recursively applies a variety of instances that show that (1) decidable properties are overestimatable; (2) existentially quantified decidable properties over finite domains are decidable; (3) |abstract.State| is a finite domain; (4) conjunctions of decidable properties are decidable; (5) list membership is decidable if equality is decidable for the element type; (6) equality is decidable for products (such as |abstract.State|) if equality is decidable for the components; (7) equality is decidable for |Location|s; (8) equality is decidable for abstract (symbolic) regions; (9) reachability in the abstract system is decidable.
-
-In \cite{hintsinunification}, Asperti et al.
-use a similar technique (but implemented using unification hints instead of type classes) to automate reflection of algebraic expressions in concrete syntax into a an inductively defined abstract syntax.
-
+To construct this overestimator, Coq recursively applies a variety of
+instances that show that (1) decidable properties are overestimatable;
+(2) existentially quantified decidable properties over finite domains
+are decidable; (3) |abstract.State| is a finite domain; (4)
+conjunctions of decidable properties are decidable; (5) list
+membership is decidable if equality is decidable for the element type;
+(6) equality is decidable for products (such as |abstract.State|) if
+equality is decidable for the components; (7) equality is decidable
+for |Location|s; (8) equality is decidable for abstract (symbolic)
+regions; (9) reachability in the abstract system is decidable.
+
+In \cite{hintsinunification}, Asperti et al. use a similar technique
+(but implemented using unification hints instead of type classes) to
+automate reflection of algebraic expressions in concrete syntax into a
+an inductively defined abstract syntax.
+}
+
+\weg{
 \section{Abstraction: Regions, States, and Spaces}
-\label{abstraction}
-
-The abstraction method for verification of (concrete) hybrid system safety as we implement it can be summarized as follows:
-\begin{enumerate}
-\item build an abstract hybrid system corresponding to the concrete hybrid system;
-\item \label{steptwo} show that each trace in the concrete system corresponds to some trace in the abstract system (and as a corollary, that reachability in the concrete system implies reachability in the abstract system, and most importantly, that unreachability in the abstract system implies unreachability in the concrete system);
-\item \label{stepthree} run a certified graph reachability algorithm on the (finite, discrete) abstract system to verify that no unsafe \emph{abstract} states are reachable;
-\item conclude from \ref{steptwo} and \ref{stepthree} that no unsafe \emph{concrete} states are reachable either.
-\end{enumerate}
-
-An abstract system corresponding to some concrete hybrid system is ``like'' the concrete system, but with the continuous state space replaced with a finite set of \emph{regions}, each corresponding to a subset of the continuous state space.
-We begin by describing these regions, after which we will describe abstract transitions and reachability.
-
-\subsection{Regions and States}
-\label{regionsandstates}
-
-Whereas in a concrete hybrid system states consist of a location paired with a point in the continuous state space, in an abstract hybrid system states consist of a location paired with the ``name'' of a region corresponding to a subset of the continuous state space:
-
-\begin{code}
-  Definition abstract.State := prod Location Region.
-\end{code}
-From now on we will use a ``|concrete.|'' prefix for names like |State| defined in section \ref{concrete}, which now have abstract counterparts.
-|Region| is a field from a record type |Space| bundling region sets with related requisites:
-\begin{code}
-  Record Space: Type :=
-    { Region: Set
-    ; Region_eq_dec: EqDec Region eq
-    ; regions: ExhaustiveList Region
-    ; NoDup_regions: NoDup regions
-    ; in_region: Container Point Region
-    ; in_region_mor: Morphism (cs_eq ==> eq ==> iff) in_region
-    ; regions_cover: forall (l: Location) (p: Point),
-        invariant (l, p) -> DN { r: Region | p `elem` r }
-    }.
-\end{code}
-The |Container Point Region| type specified for |in_region| reduces to |Point -> Region -> Prop|.
-|Container| is a type class that provides the notation ``|x `elem` y|'', prettier than ``|in_region x y|''.
-|in_region_mor| states that |in_region| respects |Point|'s setoid equality.
-
-|regions_cover| expresses that each concrete point belonging to a valid state must be represented by a region---a crucial ingredient when arguing that unreachability in the abstract system implies unreachability in the concrete system.
-The double negation in its result type is both necessary and sufficient:
-
-It is \emph{necessary} because |regions_cover| boils down to a (partial) function that, given a concrete point, must select an abstract region containing that point.
-This means that it must be able to decide on which side of a border between two regions the given point lies.
-As we saw in section \ref{dn}, that kind of decidability is only available inside |DN| unless all region borders have nontrivial overlap, which as we will see later is undesireable.
-
-Fortunately, the double negation is also \emph{sufficient}, because we will ultimately only use |regions_cover| in a proof of of |...
--> ~ concrete.reachable s| (for some universally quantified variable |s|), which, due to its head type being a negation, is stable, and can therefore be proved in and then extracted from |DN|.
-Hence, we only need |regions_cover|'s result in |DN|.
-
-\
 
@@
 % On the other hand, |regions_cover| 
+}
 
@@
 \item if the regions are too coarse, they will fail to capture the subtleties of the hybrid system that actually make it safe (if indeed it is safe at all);
-\item careless use of region overlap can result in undesireable abstract transitions (and therefore traces), adversely affecting the abstract system's utility (as we will discuss in detail in section \ref{trouble}).
+\item careless use of region overlap can result in undesireable abstract transitions (and therefore traces), adversely affecting the abstract system's utility 
+%(as we will discuss in detail in section \ref{trouble}).
 \end{itemize}
 
-Like |Alur|, we use regions formed by multiplying intervals on individual continuous variables.
-That is, for the thermostat, we first define a |Space| partitioning the continous state space into regions corresponding to clock intervals (which in the development amounts to enumerating the interval bounds, thanks to some generic space definition utilities we wrote), then define a |Space| partitioning the continuous state space into regions corresponding to temperature intervals, and then take the product of these to obtain a |Space| where regions correspond to ``squares'' in the continuous state space.
+\weg{ Like |Alur|, we use regions formed by multiplying intervals on
+  individual continuous variables. That is, for the thermostat, we
+  first define a |Space| partitioning the continous state space into
+  regions corresponding to clock intervals (which in the development
+  amounts to enumerating the interval bounds, thanks to some generic
+  space definition utilities we wrote), then define a |Space|
+  partitioning the continuous state space into regions corresponding
+  to temperature intervals, and then take the product of these to
+  obtain a |Space| where regions correspond to ``squares'' in the
+  continuous state space.
 
 \todo{Mention optional unbounded-ness? Of no theoretical interest, but makes implementation a bit more work.}
-
-In \cite{alur}, Alur describes a heuristic for interval bound selection, where the bounds are taken from the constants that occur in the invariant, guard, and safety predicates.
-For the thermostat, we initially attempted to follow this heuristic and use the same bounds Alur uses, but found that due to our use of computable reals, we had to tweak the bounds somewhat to let the system successfully produce a safety proof.
+}
+
+In \cite{alur}, a heuristic for interval bound selection is described,
+where the bounds are taken from the constants that occur in the
+invariant, guard, and safety predicates. For the thermostat, we
+initially attempted to follow this heuristic and use the same bounds,
+but found that due to our use of computable reals, we had to tweak the
+bounds somewhat to let the system successfully produce a safety
+proof. 
+\weg{
 We give one example of why such tweaking was required.
@@
 \todo{Emphasize that changing the abstraction interval bounds does not mean our result isn't as "pure" as Alur's, because the whole abstraction business only serves to prove concrete system safety, and we do prove safety of the same concrete system as Alur.}
-
-\todo{Having to do this "nudging" manually is obviously not idea. Can we think of heuristics for this nudging? Could be that it's just that this thermostat example is rotten though..}
-
+}
+Having to do this "tweaking" manually is obviously not idea. One may want to develop heuristics for this. 
 \
@@
 
-Another way in which our thermostat regions differ from |Alur|'s lies in the fact that our bounds are always inclusive, which means adjacent regions overlap in lines.
-We will discuss this in more detail later.
+Another way in which our thermostat regions differ from \cite{alur}
+lies in the fact that our bounds are always inclusive, which means
+adjacent regions overlap in lines. We will discuss this in more detail
+later.
 
@@
 
-\todo{mention that the reason the space definition could be mostly automated is that |regions_cover| is essentially a whole bunch of if-then-else's (which are only constructive inside |DN|)}
-
-\section{Abstract Transitions and Reachability}
+\weg{ \todo{mention that the reason the space definition could be
+    mostly automated is that |regions_cover| is essentially a whole
+    bunch of if-then-else's (which are only constructive inside |DN|)}
+}
+
+\subsection{Abstract Transitions and Reachability}
 \label{abs.reach}
@@
 
-Once we have a satisfactory abstract |Space|, our goal is to construct an overestimatable notion of abstract reachability implied by concrete reachability, so that concrete unreachability results may be obtained simply by executing the abstract reachability overestimator.
-It seems reasonable, then, to look for an overestimatable predicate |abstract.reachable| such that |reachable_respect| holds:
+Once we have a satisfactory abstract |Space|, our goal is to construct
+an overestimatable notion of abstract reachability implied by concrete
+reachability, so that concrete unreachability results may be obtained
+simply by executing the abstract reachability overestimator. We first overestimate the continuous transitions; we need the following definition for that.
+
+\weg{
+It seems
+reasonable, then, to look for an overestimatable predicate
+|abstract.reachable| such that |reachable_respect| holds:
 \begin{code}
@@
 \end{code}
-expressing that to conclude unreachability of a concrete state, one need only establish unreachability of an abstract state that contains it.
-As we will see shortly, the above definition of |reachable_respect| leads to problems down the line, so this is not actually the definition we use in the development.
-However, in order to see the motivation for the actual definition we use, let us proceed as if without foresight for a few more moments.
+expressing that to conclude unreachability of a concrete state, one
+need only establish unreachability of an abstract state that contains
+it. However, as abstract states may overlap, the above definition of
+|reachable_respect| leads to problems, so this is not
+actually the definition we use in the development. We first introduce |shared_cover|, expressing that a collection of abstract states together covers a collection of concrete states.
+}
+
+\weg{
+However, in order
+to see the motivation for the actual definition we use, let us proceed
+as if without foresight for a few more moments.
 
@@
 
-\
-
 \noindent In the discussion of the first example in the previous section, we judged the redundant abstract transition from |(l, [1, 2]|) to |(l, [0, 1])| to be ``wasteful'' because the destination of the sole concrete continuous transition from 1 to 1 that spawned it was already covered by |[1, 2]|, and so if |[1, 2]| is reachable, there should be no reason to make |[0, 1]| reachable as well.
@@
 To give |over_cont_trans| and |over_disc_trans| new types that embody this new specification, to replace the old |overestimator| types, we first introduce |shared_cover|:
+}
+
 \begin{code}
@@
 \end{code}
-The details of the |Container| type class are of no interest to us right now.
-What matters is that a ``container'' of concrete states is said to be sharedly-covered by a ``container'' of abstract states if for each of the concrete states in the former there is an abstract state in the latter that contains it.\footnote{The double negation is here for much the same reasons as were given in section \ref{regionsandstates} for the double negation in the type of |regions_cover|.} We state this in terms of |Container|s so that the definition applies to any types which have a reasonable notion of containment.
-
+
+The details of the |Container| type class are of no interest to us
+right now; one may think of |C| and |D| as the types of lists over
+concrete states, respectively over abstract states.  What matters is
+that a ``container'' of concrete states is said to be sharedly-covered
+by a ``container'' of abstract states if for each of the concrete
+states in the former there is an abstract state in the latter that
+contains it.
+%\footnote{The double negation is here for much the same
+%  reasons as were given in section \ref{regionsandstates} for the
+%  double negation in the type of |regions_cover|.} 
+We state this in
+terms of |Container|s so that the definition applies to any types
+which have a reasonable notion of containment.
+
+\weg{
 Stated in terms of |shared_cover|, a reasonably straightforward type for |over_cont_trans| would be:
@@
 \end{code}
+
 Lists may contain duplicates.
@@
 While we could filter out duplicates at a much later stage to satisfy that requirement, we choose to let the no-duplicates requirement propagate through to our overestimator specification, avoiding premature pessimization:
+}
+
+We now specify what the type of $\overcontrans$ should be.
 \begin{code}
 over_cont_trans: forall s: abstract.State,
-  { p: list abstract.State | NoDup l /\ shared_cover
+  { p: list abstract.State | NoDup p /\ shared_cover
     (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p }
 \end{code}
-The |vm_compute| tactic we will ultimately use to run our program tends to reduce terms in |Prop| a bit too eagerly, adversely affecting performance.
-To work around this, we introduce a bit of laziness in the form of a little unit abstraction:
+So, |over_cont_trans s| should produce a list of abstract states |p|
+without duplicates (to enhance the graph reachability algorithm that
+we don't detail here, but that will do the actual computation of the
+reachability set), such that |p| is a shared cover of the collection
+of concrete states |c| that satisfy the invariant and whose set of
+origins under |concrete.cont_trans| have an overlap with |s|. In more
+mathematical terms: |p| should form a shared cover of $\{ c \in \State
+\mid \Inv(c) \wedge s\cap \{c' \mid c' \contrans c\} \neq \emptyset\}$.
+
+\weg{
+The |vm_compute| tactic we will ultimately use to run our program
+tends to reduce terms in |Prop| a bit too eagerly, adversely affecting
+performance. So to work around this, we have introduced a bit of laziness in
+the form of a little ``unit abstraction'' that we will not detail here.
+}
+
+We similarly specify |over_disc_trans| as an overestimator for
+|concrete_disc_trans| and |over_initial| as an overestimator of
+|concrete.initial|.
+
+ \weg{
 \begin{code}
@@
 \end{code}
-Clearly, this definition does not let abstract states share the burden of reachability, and can therefore not be overestimated using our retyped abstract transition overestimators.
-But recall that it followed directly from the following more basic specification we had in mind:
+Clearly, this definition does not let abstract states share the burden
+of reachability, and can therefore not be overestimated using our
+retyped abstract transition overestimators. But recall that it
+followed directly from
+}
+
+We now consider the properties we require for |abstract.reachable|. The following basic specification we had in mind:
+
 \begin{code}
@@
 \end{code}
-Which we chose because it implied
+Which we chose because it implies
 \begin{code}
@@
 \end{code}
-expressing that to conclude unreachability of a concrete state, one need only establish unreachability of \emph{any} abstract state that contains it.
-However, we now see that this definition, too, neglects to facilitate sharing: when abstract states may share the burden of reachability, one should establish unreachability of \emph{all} abstract states containing the concrete state.
-That is, what we really want is an |abstract.reachable| satisfying:
+This expresses that to conclude unreachability of a concrete state,
+one need only establish unreachability of \emph{any} abstract state
+that contains it. However, this definition neglects to facilitate {\em
+  sharing}: remember that a concrete state may be in more then one
+abstract state. When abstract states ``share'' the burden of
+reachability, one should establish unreachability of \emph{all}
+abstract states containing the concrete state. Hence, what we really
+want is an |abstract.reachable| satisfying:
 \begin{code}
@@
 \end{code}
-\todo{Nice symmetry here, on this side of the table an existential quantifier changes to a universal one.}
 This property follows from the following new definition of |reachable_respect| we will use:
@@
 % This is only natural--in the example, to conclude unreachability of the concrete state |(l, 1)|, it is no longer sufficient to establish unreachability of the abstract state |(l, [0, 1])|.
-This new definition does not give rise to a nice simple definition of |abstract.reachable| distinct from the simple transitive closure of initiality and transition judgements made by the overestimators (analogous to |reachable_by_abstract_trace|), so we just take the latter and prove |reachable_respect| by induction over traces.
-There is, however, one last subtlety involved.
+\weg{
+This new definition does not give rise to a nice simple definition of
+|abstract.reachable| distinct from the simple transitive closure of
+initiality and transition judgements made by the overestimators
+(analogous to |reachable_by_abstract_trace|), so we just take the
+latter and prove |reachable_respect| by induction over traces. There
+is, however, one last subtlety involved.
 
@@
 However, before we do that, we make one more change.
-
-\subsection{Alternating Traces}
-\label{alter}
-
-In the definition of |reachable_by_abstract_trace|, we used the simple transitive closure of the union of |abstract.cont_trans| and |abstract.desc_trans|.
-This formulation implied that abstract traces could contain successive continuous transitions.
-Unfortunately, as Alur observes, such repetition results in pathological abstract traces that do not represent any concrete trace.
-\todo{Example needed?}
-
-The solution, as described by Alur, lies in the fact that concrete reachability by arbitrary traces is equivalent to concrete reachability by traces that alternate between continuous and discrete transitions:
-\begin{code}
-  Definition concrete.trans_kind (b: bool): relation concrete.State :=
-    if b then concrete.disc_trans else concrete.cont_trans.
-
-  Definition concrete.reachable_alternating (s: concrete.State): Prop :=
-    exists i: concrete.State, concrete.initial i /\ alternate concrete.trans_kind i s.
-
-  Lemma concrete.alternating_reachable_equiv:
-    forall s, concrete.reachable s <-> concrete.reachable_alternating s.
-\end{code}
-(Here, |alternate| forms the alternating transitive reflexive closure of a pair of relations.
-Indexing the latter by a |bool| simplifies the definition of |alternate|.) \todo{Say something about the proof of this lemma?} As an immediate corollary, we have:
-\begin{code}
-  ~ concrete.reachable_alternating s -> ~ concrete.reachable s
-\end{code}
-Hence, in order to show that unsafe concrete states are not reachable, we need only show that they are not reachable by an alternating trace.
-Consequently, we may define |abstract.reachable| in terms of alternating traces as well, and still prove |reachable_respect|:
-\begin{code}
-Definition abstract.trans_kind (b: bool): relation abstract.State :=
-  fun s s' => s `elem` (if b then over_cont_trans else over_disc_trans) s'.
-
-Definition abstract.reachable (s: abstract.State): Prop :=
-  exists i: abstract.State, over_initial ahs i = true /\ alternate abstract.trans_kind i s.
-\end{code}
-\todo{Say something about the induction proof over alternating traces?}
-% todo: the use of this kind of alternation instead of (F o G) o F is that it's simpler, thought it may not look it. but we don't need reflexivity of F. TODO: is this true?
-
-% We now turn our attention to overestimation of |reachable_by_abstract_trace|, after which we will consider the implementation of the overstimators that comprise an abstract system, in sections \ref{bla} and \ref{bla}.
-
-% repeat somewhere that an overestimator could just return "true" all the time, but that would make everything reachable, makign the thing useless for unreachability (and hence, safety) determination.
-% and repeat somewhere that overestimation is all about inclusion of predicates, heuristics. basically, this paragraph:
-
-% The idea now is to define |abstract.reachable| by decidable heuristics that make a ``best effort'' to establish absence of reachable concrete states represented by a given abstract state, marking the latter unreachable if successful, and reachable otherwise.
-Better heuristics will result in fewer spuriously reachable abstract states, decreasing the risk that abstract states containing unsafe concrete states become reachable (these have to remain unreachable for the abstraction to be of use).
-
-% in terms of new computable transition relations expressing abstract discrete and continuous transitions, as well as abstract initiality.
-All of these will be ``decidable overestimators'', occasionally emitting spurious transitions or initial states when heuristics fail to establish their absence, to ensure that their concrete counterparts are respected.
-|unreachable_respect| will then follow by induction on abstract traces.
-\todo{right?}
-
-% We emphasize that weakening the specification in this way only \emph{permits} a computable implementation of the continuous abstract transition relation to omit |[0, 1]| in the list of continuous abstract transitions originating from |[1,2]| in the example.
-Actually constructing a ``smart'' abstract continuous transition function that exploits this freedom and takes care to omit silly regions from its result list is a different matter, discussed in section \ref{drift}.
-
-% However, these definitions do make good \emph{minimal} specifications: any relations (decidable or otherwise) that include them are sufficient to prove |unreachable_respect|.
-Hence, we may work with \emph{decidable overestimations}: decidable transition relations that include \emph{at least} all transitions required to ensure that the concrete transition relation is represented, but \emph{may} include additional transitions.
-In essence, this means that in the implementation of these relations, we are free to employ heuristics that \emph{may} definitively establish absence of concrete transitions between concrete states represented by a pair of abstract states, justifying omission of an abstract transition between the two, but may also ``give up''.
-By simply including the transition in the latter case, we ensure that the concrete relation is respected.
-This does mean that abstract states may become reachable that do not represent any reachable concrete states.
-Thus, if the heuristic gives up too often, abstract states representing unsafe concrete states may become reachable, rendering the abstraction useless.
-
-% show that an abstract system of this type admits a decidable notion of reachability relates to concrete reachability in the right way.
-
-\section{Underestimating Safety}
+}
+
+\subsection{Underestimating Safety}
 \label{underestimatingsafety}
 
-In the next section we show that thanks to decidability of our transition and initiality overstimators, |abstract.reachable| is decidable.
-But first, we end the current section by showing how a decision procedure for |abstract.reachable| lets us underestimate hybrid system safety, and in particular, lets us obtain a proof of thermostat safety.
-So suppose we have
+In the next section we show that thanks to decidability of our
+transition and initiality overstimators, |abstract.reachable| is
+decidable. But first we show how a
+decision procedure for |abstract.reachable| lets us underestimate
+hybrid system safety, and in particular, lets us obtain a proof of
+thermostat safety. So suppose we have
 \begin{code}
@@
 \end{code}
-\todo{Mention why |shared_cover| isn't appropriate here.}
+
 Then, using |reachable_dec| and |unreachable_respect|, we can easily define
@@
 
-\todo{Say something about automatic generation of offending traces?}
-
-\
-
-This concludes the high level story of our development.
-What remains are the implementation of |reachable_dec| in terms of the decidable overestimators for abstract initiality and transitions bundled in the |abstract.System| record, and the implementation of those overestimators themselves.
-The former is treated in section \ref{graphs}, the latter in sections \ref{over_cont_trans} and \ref{over_disc_trans}.
+This concludes the high level story of our development. What remains
+are the implementation of |reachable_dec| in terms of the decidable
+overestimators for abstract initiality and continuous and discrete
+transitions, and the implementation of those overestimators
+themselves. The former is a formally verified graph reachability
+algorithm, that we don't detail here. The overestimator for continuous
+transitions, |over_cont_trans| will be detailed in the next section
+for the thermostat case.
+%latter
+%in sections \ref{over_cont_trans} and \ref{over_disc_trans}.
 
@@
 
+\weg{
 \section{Graphs for Reachability Decision}
@@
 In the next two sections, we discuss how to define an |abstract.System| instance for the discretization (described in section \ref{abs.space.constr}) of the concrete thermostat's continuous state space.
-
-\section{Overestimating Continuous Abstract Transitions}
+}
+
+\subsection{Overestimating Continuous Abstract Transitions}
 \label{over_cont_trans}
 
-We now discuss the implementation of the second component of an |abstract.System|, |over_cont_trans|:
+We now discuss the implementation of  |over_cont_trans|:
+\weg{
 \begin{code}
@@
 
-\
-
 \noindent Let us first consider whether we can rule out transitions by looking at the invariant, which |can_flow| demands must hold at every point along the way.
@@
 
-\
-
-Next, let us consider how we might rule out absence of concrete continuous transitions from points in |r_src| to points |r_dst| by looking at the actual flow function.
-Clearly, if we are able to determine that there are no points in |r_src| which the flow function maps to points in |r_dst|, then there need be no continuous transition between |r_src| and |r_dst|.
-Equally clearly, this is utterly impossible to meaningfully overestimate for an general flow function and general regions.
-However, the thermostat's posesses three key properties that we can exploit:
+Next, let us consider }
+
+Given two regions |r_src| and |r_dst|, we have to think how we can
+rule out the existence of concrete continuous transitions from points
+in |r_src| to points |r_dst|. Clearly, if we can determine that there
+are no points in |r_src| which the flow function maps to points in
+|r_dst|, then there need be no continuous transition between |r_src|
+and |r_dst|. Equally clearly, this is utterly impossible to
+meaningfully overestimate for a general flow function and general
+regions. However, the thermostat posesses three key properties that
+we can exploit:
 \begin{enumerate}
-\item its continuous space is of the form |pow CR n|;
-\item abstract regions correspond to multiplied |CR| intervals;
-\item its flow functions are both separable and range invertible.
+\item its continuous space is of the form $\IR^n$;
+\item abstract regions correspond to multiplied $\IR$ intervals;
+\item its flow functions are both {\em separable\/} and {\em range invertible}.
 \end{enumerate}
@@
 
-A flow function on a product space is \emph{separable} if it can be written as the product of flow function on the respective component spaces.
-For instance, in |pow CR 2|, a flow function is separable if future values of the first component only depend on past values of that same component, and the same is true for the second component.
-Being defined as the product of two flow functions on |CR| (|temp_flow| and |temp_flow|), the thermostat's flow function (in any location) has this property by construction.
-
-A flow function |f| on |CR| is range invertible if
+The notion of {\em separability\/} has already been discussed in
+Section \ref{sec:hybsys}.
+%: a flow function on a product space is
+%\emph{separable} if it can be written as the product of flow function
+%on the respective component spaces. In $\IR^2$, a flow function is
+%separable if future values of the first component only depend on past
+%values of that same component, and the similarly for the second
+%component. Being defined as the product of two flow functions on |CR|
+%(|temp_flow| and |temp_flow|), 
+The thermostat's flow function (in any
+location) is separable by construction.
+
+A flow function |f| on |CR| is {\em range invertible\/} if
 \begin{code}
@@
 \end{code}
-Here, |OpenRange| represents potentially unbounded intervals in |CR| (with bounds closed if present \todo{meh, need a better name for OpenRange..}).
-Range invertibility is a less demanding alternative to point invertibility:
+Here, |OpenRange| represents potentially unbounded intervals in $\IR$
+(with bounds closed if present. In other words, if $\phi : \IR^2
+\rightarrow \IR$ is a flow function with range inverse $F$ and $a,b$
+are intervals in $\IR$, then $F(a,b)$ is an interval that contains all
+$t$ for which $\phi(x,t)\in b$ for some $x\in a$. So $F(a,b)$ should
+contain all durations $t$ that take some elements of $a$ to some
+element of $b$ via flow $\phi$.
+
+Range invertibility is a less demanding alternative to point invertibility: $\phi^{-1}$ is the {\em point inverse\/} of $\phi$ if $\forall x,y\in \IR (\phi (x, \phi^-1(x,y)) =y)$.
+So a point inverse $\phi^{-1}(x,y)$ computes the exact time it takes to go from $x$ to $y$ via flow $\phi$.
+\weg{
 \begin{code}
@@
 \end{code}
-We need range invertibility because the exponential flow functions are not point invertible.
-
-As mentioned in section \ref{concreteflow}, we used a modest library of flow functions when defining the thermostat's flow.
-Included in that library are range-inverses, which consequently automatically apply to the thermostat's flow.
-Hence, no ad-hoc work is needed to show that the thermostat's flow functions are range-invertible.
+}
+We need range invertibility because the exponential flow functions are
+not point invertible.
+
+In the formalization we use a modest library
+of flow functions when defining the thermostat's flow. Included in
+that library are range-inverses, which consequently automatically
+apply to the thermostat's flow. Hence, no ad-hoc work is needed to
+show that the thermostat's flow functions are range-invertible.
 
@@
 
-Having defined the class of separable range-invertible flow functions, and having argued that the thermostat's flow is in this class, we now show how to proceed with our overestimation of existence of points in |r_src| which the flow function map to points in |r_dst|.
-As discussed in section \ref{abs.space.constr}, the abstract space for our thermostat was constructed as the product of two abstract spaces based on temperature and clock intervals, respectively.
-Regions in such a product space are pairs of regions in the composite spaces, so |r_src| and |r_dst| can be written as |(r_src_temp, r_src_clock)| and |(r_dst_temp, r_dst_clock)|, respectively.
+Having defined the class of separable range-invertible flow functions,
+and having argued that the thermostat's flow is in this class, we now
+show how to proceed with our overestimation of existence of points in
+|r_src| which the flow function map to points in |r_dst|. 
+%As discussed
+%in section \ref{abs.space.constr}, the abstract space for our
+%thermostat was constructed as the product of two abstract spaces based
+%on temperature and clock intervals, respectively. 
+Regions in the abstract space for our thermostat are basically pairs
+of regions in the composite spaces, so |r_src| and |r_dst| can be
+written as |(r_src_temp, r_src_clock)| and |(r_dst_temp,
+r_dst_clock)|, respectively, where each of these four components are
+intervals.
 
@@
 \end{code}
-(defined in terms of things like |overestimate_CRle| shown in section \ref{estimation}) to overestimate whether the following three ranges overlap:
+(defined in terms of things like |overestimate_CRle| shown in section \ref{estimation}) to overestimate whether the following three intervals overlap:
 \begin{enumerate}
@@
 \end{enumerate}
-Overlap of 2 and 3 is equivalent to existence of a point in |r_src| from which one can flow to a point in |r_dst|.
-After all, if these two range inverses overlap, then there is a duration |d| that takes a certain temperature value in |r_src_temp| to a value in |r_dst_temp| and also takes a certain clock value in |r_src_clock| to a value in |r_dst_clock|.
-If 2 and 3 do \emph{not} overlap, then either it takes so long for the temperature to flow from |r_src_temp| to |r_dst_temp| that any clock value in |r_src_clock| would ``overshoot'' |r_dst_clock|, or vice versa.
-Finally, if 1 does not overlap with 2 and 3, then apparently one could only flow backward in time, which is not permitted.
-Hence, overlap of these three ranges is a necessary condition for existence of concrete flow from points in |r_src| to points in |r_dst|, and so our |abstract.cont_trans| overestimator may justifiably return ``false'' when the overlap overestimator manages to prove absence of overlap.
+For a visual explanation, one may consult the left drawing in Figure
+\ref{fig:abstraction} and view |r_src_clock| as $[a_1,b_1]$,
+|r_dst_clock| as $[c_1,d_1]$ etc.  Overlap of 2 and 3 is equivalent to
+existence of a point in |r_src| from which one can flow to a point in
+|r_dst|. After all, if these two range inverses overlap, then there is
+a duration |d| that takes a certain temperature value in |r_src_temp|
+to a value in |r_dst_temp| and also takes a certain clock value in
+|r_src_clock| to a value in |r_dst_clock|. If 2 and 3 do \emph{not}
+overlap, then either it takes so long for the temperature to flow from
+|r_src_temp| to |r_dst_temp| that any clock value in |r_src_clock|
+would ``overshoot'' |r_dst_clock|, or vice versa. Finally, if 1 does
+not overlap with 2 and 3, then apparently one could only flow backward
+in time, which is not permitted. Hence, overlap of these three ranges
+is a necessary condition for existence of concrete flow from points in
+|r_src| to points in |r_dst|, and so our |abstract.cont_trans|
+overestimator may justifiably return ``false'' when the overlap
+overestimator manages to prove absence of overlap.
 
@@
 
+\weg{
 Having shown at last the ``real'' flow test condition used in the overestimation, we can show more precisely how drift would arise if we avoided region overlap (thereby eliminating drift problems at the specification level) but did not bother to employ a redundancy underestimator.
@@
 
-\section{Overestimating Discrete Abstract Transitions}
+\subsection{Overestimating Discrete Abstract Transitions}
 \label{over_disc_trans}
@@
 Basically, we simply take as an argument an underestimation of whether a particular reset function is the identity function, and if the underestimator says that it is, then we only make abstract discrete transition to abstract states with the same region as the source state.
-If on the other hand the underestimator does \emph{not} indicate that the reset function in question is the identity function, then we simply take the bounds of the source region, map them with the reset function, and overestimate whether the resulting region overlaps with the candidate destination (for this to be valid, the reset function must be nondecreasing\footnote{When we say that a \emph{flow} function is ``nondecreasing'', we mean that it does not decrease over time, and so this is really a statement about flow direction.
+If on the other hand the underestimator does \emph{not} indicate that the reset function in question is the identity function, then we simply take the bounds of the source region, map them with the reset function, and overestimate whether the resulting region overlaps with the candidate destination (for this to be valid, the reset function must be nondecreasing\fotonote{When we say that a \emph{flow} function is ``nondecreasing'', we mean that it does not decrease over time, and so this is really a statement about flow direction.
 But when we say that a \emph{reset} function (which does not take a time argument) is ``nondecreasing'', we simply mean that it does not flip ranges around, which is a rather mundane property.}
@@
 % todo: mention that we don't discuss boring things like over_initial and the abstract safety list, even though they're all treated uniformly: the user must specify these as lists of opensquares when setting up the abstraction.
-
-\section{Conclusions}
+}
+\section{Related work}
+
+\emph{Verification} of hybrid systems is an active field of research and
+there is a number of tools developed with this goal in mind;
+see \cite{HST} for a comprehensive list.
+Most of them are based on abstract refinement methods, either
+using box representations \cite{Kow96,Kow98}
+or with polyhedra approximations \cite{alur,Chu99,Cla03}.
+
+Many of those tools are implemented in MATLAB \cite{MATLAB}
+and those using some general programming language of choice
+most often rely on standard floating point arithmetic, that
+comes with its rounding errors. Few tools that address 
+this problem include 
+PHAVer \cite{PhaVER}, that relies on the Parma Polyhedra Library \cite{Parma} 
+for exact computations with non-convex polyhedra 
+and 
+HSolver \cite{HSolver}, that is based on the constraint solver 
+RSolver \cite{RSolver}.
+
+\emph{Formal verification} becomes more and more important, especially in the
+field of hybrid systems, which are used to model safety critical systems 
+of growing complexity. There has been previous work on using general purpose 
+theorem provers for verification of hybrid systems: see \cite{Mum01,Hen98} 
+and \cite{Man98,Man01} for works using, respectively, PVS and STeP. 
+KeYmaera \cite{KeYmaera} is a dedicated interactive theorem prover for 
+specification and verification logic for hybrid systems. It combines
+deductive, real algebraic, and computer algebraic prover technologies
+and allows users to model hybrid systems, specify their properties
+and prove them in a semi-automated way.
+
+However, to the best of our knowledge, none of the previous work nor tools 
+relies on a precise model of real number computations completely verified 
+in a theorem prover; such as the model of CoRN used in this work.
+
+\section{Conclusions and further research}
 \label{conclusions}
-
-\begin{itemize}
-\item nice showcase of proof-by-computation-on-computable-reals
-\item systematic use of estimators to make tactic-like optional-deciders, at each level in the stack
-\item computable reals do complicate things
-\item heuristic for bound selection doesn't work out of the box. manual tweaking obviously not ideal. more experimentation required
-\item region overlap
-\item drift and redundance
-\item code stats(?)
-\item requirements on system: separability of flow and reset, invariant stability
-\item lots of room for more clever heuristics with less restrictive preconditions (e.g. separability)
-\end{itemize}
+The presented verification of hybrid systems in Coq gives a nice
+showcase of proof-by-computation-on-computable-reals. The computable
+reals in CoRN do really complicated things for us, by approximating
+values for various real number exprseeions at great precision. The
+development also contains some nice layers of abstraction, involving
+the sophisticated use of type classes, e.g.\ the systematic use of
+estimators to make tactic-like optional-deciders, at each level in the
+stack and the use of the double negation monad.
+
+It is to be seen how far this automated verification approach can be
+taken, given the fact that we have limited ourselves to hybrid systems
+with separable reset and flow functions that are monotone, and with a
+stable invariants. There is still a lot of room for more clever
+heuristics with less restrictive preconditions. The heuristic in
+\cite{alur} for bound selection doesn't work out of the box, but
+manual tweaking is obviously not ideal, so some more experimentation
+is required here. Finally, in case safety cannot be proved, one would
+like the system to also generate an ``offending trace'' automatically,
+which can the be inspected by the user.
 

### 2010-03-22 ITP10/hybrid.tex
Add ITP10 paper sources.
Ignore-this: 49de24983449dd9b9daef53ef1da6000

darcs-hash:20100322174102-bab43-dcc3cc0af50eb50c30ef52b815fff99223eb4669.gz

--- 2010-03-22 ITP10/hybrid-12oclock.tex
@@
 
-\newcommand{\DN}{{\sf DN}}
+\newcommand{\DN}{{\sf DN}\,}
 \newcommand{\IR}{{\mathbf R}}
@@
 \title{\thetitle}
-\subtitle{An Implementation of the Abstraction Method In Coq}
-
-\author{Herman Geuvers\inst{1,2}, Adam Koprowski\inst{3}, Dan Synek\inst{1}, \\ Eelis van der Weegen\inst{1}}
+%\subtitle{An Implementation of the Abstraction Method In Coq}
+
+\author{Herman Geuvers\inst{1,2}, Adam Koprowski\inst{3}, Dan Synek\inst{1}, Eelis van der Weegen\inst{1}}
 \authorrunning{H. Geuvers, A. Koprowski, D. Synek, E. van der Weegen}
 
+%Institute for Computing and Information Sciences\\ 
+%Faculty of Mathematics and Computer Science\\ 
 \institute{
-Institute for Computing and Information Sciences\\ 
 Radboud University Nijmegen 
 \and
-Faculty of Mathematics and Computer Science\\ 
 Technical University Eindhoven
@@
 using the abstraction method introduced by Alur, Dang and Ivan\v ci\'c
-(2006).  The development includes: a formalization of the structure of
+(2006).  The development includes: a formalisation of the structure of
 hybrid systems; a set of utilities for the construction of an abstract
@@
 relies on the computable real number implementation part of the CoRN
-library of formalized constructive mathematics in Coq.  The obtained
+library of formalised constructive mathematics in Coq.  The obtained
 safety proofs are fully certified and do not use floating point real
@@
 
-\weg{ This technical report documents our development of a hybrid
-  system safety prover, implemented in Coq using the abstraction
-  method introduced by Alur in \cite{alur}. The development includes:
-  a formalization of the structure of hybrid systems; a systematic
-  approach and generic set of support utilities for the construction
-  of an abstract system (consisting of decidable ``overestimators'' of
-  abstract transitions and initiality) faithfully representing a
-  (concrete) hybrid system; a translation of abstract systems to
-  graphs enabling decision of abstract state reachability using a
-  certified graph reachability algorithm; a proof of an example hybrid
-  system (taken from \cite{alur}) generated using this tool stack. The
-  development critically relies on the computable real number
-  implementation part of the CoRN library of formalized constructive
-  mathematics. \todo{Hm, anything else worth mentioning?}  }
-
 \end{abstract}
@@
 
-\nocite{*}
+%\nocite{*}
 
@@
 artifacts inherent in this approach limits the confidence one can
-justifiably have in safety judgments made by such an
-implementation. In this paper, we describe a reimplementation of the
+justifiably have in safety judgements made by such an
+implementation. In this paper, we describe a re-implementation of the
 basic technique of Alur et al.\ in Coq, a proof assistant based on a
@@
 developed by Russell O'Connor \cite{oconnor}, and part of the CoRN
-library of formalized constructive mathematics \cite{corn}. Indeed,
+library of formalised constructive mathematics \cite{corn}. Indeed,
 this development showcases its use in a concrete and practical
@@
 The motivation for this work is threefold. 
-
 First, we want to avoid the use of floating point number computation,
@@
 that we need a symbolic proof.) In Nijmegen we have developed a
-verified library of exact real arithmetic, formalized in the proof
-assistant Coq, as part of CoRN \cite{corn,oconnor}, our Constructice
+verified library of exact real arithmetic, formalised in the proof
+assistant Coq, as part of CoRN \cite{corn,oconnor}, our Constructive
 Coq Repository at Nijmegen. The algorithms (for $e^x$, $\cos$, $\sin$
@@
 system, which is then verified. However, this abstraction itself is
-not verified. In this formalization we want to also verify the
+not verified. In this formalisation we want to also verify the
 abstraction itself: show that the abstract system models the concrete
 one.
-
-% todo: document that fallback-to-making transitions
-% todo: document separability, and why we don't use alur's technique
-% read my mail about DN to see if all its contents is now part of section bla
-
- % In our case, the specification of the program we write is this:
- % given a description of a concrete system, the program . The whole
- % implementation, including the computational part, serves only to
- % prove a single theorem. Hence, one essentially only needs to trust
- % Coq to be convinced of the safety proofs created by our
- % development.
-
-%A key ingredient in our development
-
-%Wherever we say ``|CR|'', we are really referring to the type
-%|CRplain| of computable reals, part of C-CoRN.
-
-%The computational steps become part of the complete proof of the
-%safety theorem.
-
-%We formally prove safety of the thermostat example used by Alur.
-
-\weg{
-\paragraph{Organization}
-
-Throughout this report, we present in parallel our general formalized
-framework (with particular emphasis on differences with Alur's
-presentation), and the way it is used to obtain a safety proof for an
-example hybrid system, taken from \cite{alur}, modeling the operation
-of a thermostat. In section \ref{concrete}, we begin by defining the
-structure of normal (concrete) hybrid systems, their semantics,
-reachability, and safety. We also define the concrete thermostat and
-the safety condition we wish to prove. In section \ref{dn}, we take a
-moment to discuss some of the limits on decidability of properties of
-computable real numbers, as these complicate matters in the remainder
-of the development. Next, in section \ref{estimation} we describe
-notions of underestimation and overestimation that are applied
-throughout the development. Then, in section \ref{abstraction}, we
-begin our presentation of the abstraction method by discussing
-abstract regions, states, and spaces. We build on these in section
-\ref{abs.reach}, where we first develop the ``obvious'' notions of
-abstract system and reachability, discuss why they fail, and then
-rework them to embrace reachability sharing. In section
-\ref{underestimatingsafety} we show how, assuming abstract
-reachability is decidable, we may prove hybrid system safety by
-computation. Then, in section \ref{graphs}, we show how an abstract
-system consisting of transition and initiality ``overestimators'' can
-be used to decide abstract reachability by construction of a directed
-graph corresponding to the system, and the use of an ordinary but
-verified graph reachability algorithm. Finally, in sections
-\ref{over_cont_trans} and \ref{over_disc_trans} we describe the
-implementation of the aforementioned overestimators. We end with
-conclusions in section \ref{conclusions}.
-}
-
-% proceed to define abstraction bla. Following that, we show that a
-% simple (certified) graph reachability can be employed to compute
-% reachability in the graph corresponding to the abstract system, and
-% that the resulting reachability judgement can be translated back
-% through the abstract system all the way to the concrete system. We
-% end with conclusions in section \ref{conclusions}.
 
@@
 permitted values for the continuous variables. 
-\weg{
-We will use this in the
-definitions of transition relations in a moment. The morphism field
-expresses that this invariant respects |Point|'s setoid equality. We
-further require that it is \emph{stable}, where |stable P| is defined
-as |~ ~P -> P|. We discuss the reason for this stability requirement
-in section \ref{dn}.
-}
 For the thermostat, the invariant is as follows: $$\Inv_{\Heat}(c,T) := T\leq 10\wedge c\leq 3,\quad \Inv_{\Cool}(C,t) := 5 \leq T,\quad \Inv_{\Chec} := c\leq 1.$$
@@
 $\Init(l,c,T) := l= \Heat \wedge c= 0 \wedge 5 \leq T \leq 10.$
-\weg{
-\begin{code}
-  {s: State pipe location s = Heat /\ 5 <= temp s <= 10 /\ clock s == 0}
-\end{code}
-where |location|, |temp|, and |clock| are the obvious projections.
-} 
 There is the obvious requirement that the invariant holds at each
@@
 with the set of initial states, determine the set of \emph{reachable}
-states, representing the possible behaviors exhibited by a
+states, representing the possible behaviours exhibited by a
 hypothetical real-world implementation of the hybrid system (as
 software running on a device with sensors and actuators).
-
-%\subsection{Flow and Continuous Transitions}
-%\label{concreteflow}
 
@@
 function corresponding to the $\Heat$ location will have the
-temperature increase with time, modeling the effect of the heater
+temperature increase with time, modelling the effect of the heater
 component in the imagined thermostat device.
-
-%format Duration = CR"_{\ge 0}"
-%format CRge0 = CR"_{\ge 0}"
 
@@
 differential equations, which are flow functions $\Phi$ which satisfy:
-\begin{eqnarray*}
-  \Phi( p, 0) &=& p \\
-  \Phi (p, t + t') &=& \Phi (\Phi( p, t), t')
-\end{eqnarray*}
-\weg{
-We further require that $\Phi$ is a morphism respecting |Point|'s
-setoid equality. We bundle functions with these properties as a record
-type called |Flow| (not shown), which is parametrized over the
-continuous state space type, and is equipped with an implicit coercion
-to the |phi| function it contains.
-\todo{Unlike Alur, we don't have ``input'' (why?).}
-}
-
+$$\Phi( p, 0) = p \quad \mbox{and} \quad \Phi (p, t + t') = \Phi (\Phi( p, t), t')$$
 We now say that there is a (concrete) \emph{continuous transition}
@@
 seen in the following definition of the discrete transition relation:
-
 $$ (l,p) \distrans (l',p') := \guard_{l,l'}(p) \wedge \reset_{l,l'}(p)= p' \wedge \Inv_l(p) \wedge \Inv_{l'}(p')$$
-
-\weg{
-\begin{code}
-  Definition disc_trans: relation State := fun (l, p) (l', p') =>
-    guard (l, p) l' /\ reset l l' p = p' /\
-    invariant (l, p) /\ invariant (l', p').
-\end{code}
-}
-
 It will be obvious from Figure \ref{fig:thermostat} what the guards
@@
 
-\weg{
-\begin{code}
-Definition thermo_reset (l l': Location) (p: Point): Point :=
-  ( match l, l' with
-    | Cool, Heat | Heat, Check | Check, Heat => 0
-    | _, _ => fst p
-    end
-  , snd p).
-\end{code}
-}
-
-%\subsection{Traces and Safety}
-
 A transition is either continuous or discrete $\trans:= \distrans \cup \contrans$.
-We now say that a state $s$ is \emph{reachable} if there is an initial state $i$ from which one can, by a finite number of transitions, end up in $s$:
-%format trans_refl_closure (x) = x "^{*}"
+We now say that a state $s$ is \emph{reachable} if there is an initial state $i$ from which one can, by a finite number of transitions, end up in $s$ ($\ttrans$ is the transitive reflexive closure of $\trans$):
 $$\Reach(s):=
     \Exists{i\in \State}{\Init(i) \wedge i \ttrans s}.$$
-Here, $\ttrans$ is the transitive reflexive closure of $\trans$.
-
-As mentioned before, the set of reachable states represents the
-possible behaviors exhibited by a hypothetical real-world
+\weg{As mentioned before, the set of reachable states represents the
+possible behaviours exhibited by a hypothetical real-world
 implementation of the hybrid system (as software running on a device
 with sensors and actuators).
+}
 
@@
 keep the temperature above $4.5$ degrees at all times, and so the safe
-states are defined to be those in which we have $T > 4.5$: $$\Safe(l,c,T) := T >4.5 \quad and \quad \Unsafe(l,c,T) := T\leq 4.5.$$
+states are defined to be those in which we have $T > 4.5$: $\Safe(l,c,T) := T >4.5$ and $\Unsafe(l,c,T) := T\leq 4.5.$
 The goal, then, is to verify that the reachable states are all safe.
@@
 
-\weg{
-\section{Motivation}
-
-1. Verify the correctness of the abstraction itself
-
-2. Avoid Floating Point arithmetic
-
-3. Avoid correctness proofs of systems that are ``almost wrong''.
-   (Add one arrow to Thermostat!)
-
-- One cannot decide x>0 of a measured value
-- One can't define a function on the basis of distinguishing x<=y \/ x >y
-}
-
-\section{Formalization}
-We now describe the formalization in Coq and the design choices
+\section{Formalisation}
+We now describe the formalisation in Coq and the design choices
 made. Some of the design choices were due to specifics in Coq or its
@@
 latter design choices, because we feel they are most relevant.
+Technical details, including a tar file with the formalization can be found at {\tt http://www.xs4all.nl/~weegen/eelis/research/hybrid/} or \cite{hybrid-techreport}.
 
@@
   ; guard: State -> Location -> Prop
-  ; reset: Location -> Location -> Point -> Point
-  }.
+  ; reset: Location -> Location -> Point -> Point  }.
 \end{code}
@@
 %algorithms to complete.
-
+\weg{ 
 A place where we definitely need the stability of the invariant is in
@@
 is a trivial tautology.
-
+}
 A question that immediately arises is: How does one actually use this
@@
 
-%format return_DN
 %format bind_DN
@@
 \item 
-different {\em representations\/} of the same point $(c,T)$ may be
+different {\em representations\/} of a point $(c,T)$ may be
 sent to different rectangles,
 \item
-a point that is $<\epsilon$ outside the rectangle $R$ may still be sent to $R$.
+a point that is less than $\epsilon$ outside the rectangle $R$ may still be sent to $R$.
 \end{itemize}
@@
 suppose we add a fourth location $\Off$ to the thermostat of Figure
-\ref{fig:thermostat}, with $\cdot{T} = -1$, $\cdot{c} =1$ and an arrow
+\ref{fig:thermostat}, with $\dot{T} = -1$, $\dot{c} =1$ and an arrow
 from $\Heat$ to $\Off$ with guard $c\geq 2 \wedge T<9$. Clearly, if
-the system can end up in location $\Off$, its is unsafe. Now, using
+the system can end up in location $\Off$, it is unsafe. Now, using
 the classical abstraction method, there is no transition to any state
@@
 
-So, to summarize, if $s\contrans s'$, then we don't require $s'$ to
-$A(s')$ to be in the list $\overcontrans(A(s))$, but we just require
-$s'$ to be in the $\bigcup \overcontrans(A(s))$. (Basically it should
-be in one of the regions in the list $\overcontrans(A(s))$.) This
-simple change relieves us from having to determine the exact regions
-that points are in: they just should be covered. The functions
-$\overcontrans$ and $\overdistrans$ yield a notion of {\em trace\/} in the abstract
-hybrid system in the obvious way: starting from $R_0$, take an $R_1$ in $\overcontrans( R_0)$, then an $R_2$ in $\overdistrans(R_1)$, and so forth.
-
-\weg{
-The abstraction method for verification of (concrete) hybrid system safety as we implement it can be summarized as follows:
-\begin{enumerate}
-\item build an abstract hybrid system corresponding to the concrete hybrid system;
-\item 
-\label{steptwo} 
-show that each trace in the concrete system corresponds to some trace
-in the abstract system (and as a corollary, that reachability in the
-concrete system implies reachability in the abstract system, and most
-importantly, that unreachability in the abstract system implies
-unreachability in the concrete system);
-\item 
-\label{stepthree} 
-run a certified graph reachability algorithm on the (finite, discrete)
-abstract system to verify that no unsafe \emph{abstract} states are
-reachable;
-\item 
-conclude from \ref{steptwo} and \ref{stepthree} that no unsafe
-\emph{concrete} states are reachable either.
-\end{enumerate}
-}
+So, to summarize, if $s\contrans s'$, then we don't require
+$A(s')$ to be in the list $\overcontrans(A(s))$, but we only require
+$s'$ to be in the $\bigcup \overcontrans(A(s))$. 
+%(Basically it should
+%be in one of the regions in the list $\overcontrans(A(s))$.) 
+This simple change relieves us from having to determine the exact
+regions that points are in: they just should be covered. The functions
+$\overcontrans$ and $\overdistrans$ yield a notion of {\em trace\/} in
+the abstract hybrid system in the obvious way: starting from $R_0$,
+take an $R_1$ in $\overcontrans( R_0)$, then an $R_2$ in
+$\overdistrans(R_1)$, and so forth.
 
@@
 a region corresponding to a subset of the continuous state space.
-
 From now on we will use a ``|concrete.|'' prefix for names like
@@
     ; regions_cover: forall (l: Location) (p: Point),
-        invariant (l, p) -> DN { r: Region | p `elem` r }
-    }.
+        invariant (l, p) -> DN { r: Region | p `elem` r }  }.
 \end{code}
@@
 the notation ``|x `elem` y|'', prettier than ``|in_region x
-y|''. 
-%|in_region_mor| states that |in_region| respects |Point|'s
-%setoid equality.
-
-|regions_cover| expresses that each concrete point belonging to a
-valid state must be represented by a region---a crucial ingredient
+y|''. |regions_cover| expresses that each concrete point belonging to
+a valid state must be represented by a region---a crucial ingredient
 when arguing that unreachability in the abstract system implies
@@
 
-%\todo{|CRle_unapplied| can also be underestimated, right?}
-
-%format A_0
-%format A_n
-%format a_0
-%format a_n
-
-\weg{
-The types of underestimators and overerestimators often merely repeat
-a certain predicate's parameters, as in the |overestimate_CRle|
-example above. Using a bit of type class magic to achieve variadicity,
-we can rid us of this repetition. Given a predicate |P: A_0 -> ... ->
-A_n -> Prop|, we define |underestimator P| as the
-type 
-\begin{code}forall (a_0: A_0) ... (a_n: A_n), underestimation (P a_0 ... a_n).\end{code} 
-(And similar for |overestimator|.) With
-these, the type of |overestimate_CRle| may be written as |Qpos ->
-overestimator (CRle_unapplied)|.
-We use the same techniques for real decision procedures (in the few places we use those): we define |decision P| as |{P}+{~P}|, and given a predicate |P: A_0 -> ...
--> A_n -> Prop|, we define |decider P| as the type |forall (a_0: A_0) ...
-(a_n: A_n), decision (P a_0 ...
-a_n)|.
-
-%\subsection{Automating estimator/decider composition}
-
-Underestimators, overestimators and deciders can be combined to form
-underestimators (resp. overestimators, deciders) for things like
-conjunctions and quantifications over finite domains. For instance,
-here is a combinator that forms conjunction overestimators:
-
-\begin{code}
-Program Definition overestimate_conj {P Q: Prop}
-  (x: overestimation P) (y: overestimation Q):
-    overestimation (P /\ Q) := x && y.
-Next Obligation.
-  intros [A B].
-  destruct x. destruct y.
-  simpl in H.
-  destruct (andb_false_elim _ _ H); intuition.
-Qed.
-\end{code}
-%\todo{clean up proof}
-Applying these combinators by hand to form estimators/deciders for
-composite propositions is somewhat tedious, but fortunately this
-process can be automated if we make |overestimation|,
-|underestimation|, and |decision| type classes:
-
-\begin{code}
-Class overestimation (P: Prop): Set
-  := overestimate: { b: bool | b = false -> ~ P }.
-\end{code}
-
-If we now say |overestimate P|, the type class instance resolution
-mechanism will try to find (or build) a term of type |overestimation
-P| by recursively applying declared type class instances. For
-instance, if we declare |overestimate_conj| above as a type class
-instance (instead of a plain definition) and then say |overestimate (A
-/\ B)| where |A| and |B| are already known to be overestimatable
-(meaning the resolution mechanism is able to build terms of types
-|overestimation A| and |overestimation B|), then the resolution
-mechanism will be able to build a term of type |overestimation (A
-/\ B)| by applying |overestimate_conj|.
-
-We show one nontrivial example to illustrate the convenience of this mechanism: at one point in the development we have Coq automatically construct an overestimator of the property
-\begin{code}
-  exists u: abstract.State, In u astates /\ reachable u,
-\end{code}
-simply by saying
-\begin{code}
-  overestimate (exists u: abstract.State, In u astates /\ reachable u).
-\end{code}
-To construct this overestimator, Coq recursively applies a variety of
-instances that show that (1) decidable properties are overestimatable;
-(2) existentially quantified decidable properties over finite domains
-are decidable; (3) |abstract.State| is a finite domain; (4)
-conjunctions of decidable properties are decidable; (5) list
-membership is decidable if equality is decidable for the element type;
-(6) equality is decidable for products (such as |abstract.State|) if
-equality is decidable for the components; (7) equality is decidable
-for |Location|s; (8) equality is decidable for abstract (symbolic)
-regions; (9) reachability in the abstract system is decidable.
-
-In \cite{hintsinunification}, Asperti et al. use a similar technique
-(but implemented using unification hints instead of type classes) to
-automate reflection of algebraic expressions in concrete syntax into a
-an inductively defined abstract syntax.
-}
-
-\weg{
-\section{Abstraction: Regions, States, and Spaces}
-
-We can lift the containment relation between points and regions to a containment relation between concrete states and abstract states:
-\begin{code}
-  Instance abs: Container concrete.State abstract.State
-    := fun (l, p) (l', r) => l = l' /\ p `elem` r.
-\end{code}
-Again, the |Container| application reduces to |concrete.State -> abstract.State -> Prop|, but the use of |Container| lets us say |s `elem` s'| instead of |abs s s'|.
-\todo{Coq trunk has nameless instances, I believe, which would let us dispense with the pointless |abs| name altogether.}
-
-%format inf = "\infty "
-%format closesquare = "\hspace{-1mm}] "
-
-% Suppose one has a very simple concrete system with the trivial invariant |True| and only a single continuous variable in |CR|.
-And suppose one wishes to abstract this system using two regions, corresponding to the ranges |(inf, 2 closesquare| and |(2, inf)|, respectively.
-Then implementing |regions_cover| amounts to deciding, for any |x|, whether |x <= 2| or |2 < x|.
-As discussed in section \ref{bla}, bla.
-And as before, the solution involved the double negation monad |DN|.
-
-% Unfortunately, the constructive real number representation we use does not feature decidability of this particular query, because they fundamentally only let one observe arbitrarily close approximations, and if |x| happens to be 2, no approximation of it can decisely determine whether |x <= 2| or |2 < x| (because 2 will always lie within the error range of the approximation).
-
-% One solution to this problem is to make sure one's abstract regions have nontrivial overlap.
-For example, one could instead have regions corresponding to |(inf, 2.01)| and |(2, inf)|.
-An approximation of |x| with error margin 0.05 would then let one find at least one range that definitely contains |x|.
-Unfortunately, as we will see later, region overlap can result in undesireable transitions (and therefore traces), which can adversely affect the abstract system's utility.
-We therefore use a different solution, which is where |DN| comes in.
-
-% As it turns out, there are propositions that are not provable themselves, but whose double negation \emph{is} provable.
-The proposition |x <= 2 \/ 2 < x| is an example of such a proposition.
-Defining |DN P| as |~ ~P|, |DN (x <= 2 \/ 2 < x)| is provable (for any |x|).
-This |DN| is a monad, which means that one can reason ``inside'' of it, deriving ever more doubly negated propositions, possibly making use of lemmas/functions with result types also in |DN|.
-Such lemmas/functions are ``weak'' in the sense that their result can only be used in proofs inside the |DN| monad.
-
-% Now, there exist propositions |P| for which |DN P| is equivalent to |P| (these are called ``stable'' propositions).
-One important example are propositions of the form |~ P|.
-To prove such a proposition, one may do so inside the |DN| monad, making use of propositions which are only true inside |DN|, until one has derived |DN P|, at which point |P| may be extracted.
-
-% In our hybrid system development, we willl ultimately only use |regions_cover| in a proof of |...
--> ~ concrete_reachable s| (for some universally quantified variable s), which, due to its head type being a negation, is stable.
-Hence, this proof can be performed inside the DN monad, which means it is sufficient for |regions_cover| to have the weaker |DN| result, which in turn means it may make use of DN-only lemmas such as the trichotomy property for |CR|, which in turn means we need not resort to non-trivial region overlap.
-
-% The result type of |regions_cover|, |DN (sig (in_region p))|
-% On the other hand, |regions_cover| 
-}
-
 \subsection{Abstract Space Construction}
@@
 
-When building an abstract system, one is in principle free to divide the continuous state space up whichever way one likes. However:
-\begin{itemize}
-\item if the regions are too fine-grained, there will have to be very many of them to cover the continuous state space of the concrete system, resulting in poor performance;
-\item if the regions are too coarse, they will fail to capture the subtleties of the hybrid system that actually make it safe (if indeed it is safe at all);
-\item careless use of region overlap can result in undesireable abstract transitions (and therefore traces), adversely affecting the abstract system's utility 
+When building an abstract system, one is in principle free to divide
+the continuous state space up whichever way one likes. However, if the
+regions are too fine-grained, there will have to be very many of them
+to cover the continuous state space of the concrete system, resulting
+in poor performance. On the other hand, if the regions are too coarse,
+they will fail to capture the subtleties of the hybrid system that
+allow to prove it safe (if indeed it is safe at all).  Furthermore,
+careless use of region overlap can result in undesireable abstract
+transitions (and therefore traces), adversely affecting the abstract
+system's utility
 %(as we will discuss in detail in section \ref{trouble}).
-\end{itemize}
-
-\weg{ Like |Alur|, we use regions formed by multiplying intervals on
-  individual continuous variables. That is, for the thermostat, we
-  first define a |Space| partitioning the continous state space into
-  regions corresponding to clock intervals (which in the development
-  amounts to enumerating the interval bounds, thanks to some generic
-  space definition utilities we wrote), then define a |Space|
-  partitioning the continuous state space into regions corresponding
-  to temperature intervals, and then take the product of these to
-  obtain a |Space| where regions correspond to ``squares'' in the
-  continuous state space.
-
-\todo{Mention optional unbounded-ness? Of no theoretical interest, but makes implementation a bit more work.}
-}
 
@@
 proof. 
-\weg{
-We give one example of why such tweaking was required.
-
-Following the heuristic, Alur derives the following two regions:
-%format r_0
-%format r_1
-%format p_0
-%format p_1
-\begin{code}
-  r_0 := { (c, t): Point | c <= 0 /\ 5 <= t <=6 }
-  r_1 := { (c, t): Point | 2 <= c <= 3 /\ 6 < t < 9 }
-\end{code}
-
-In the |Heat| location, where both the clock and the temperature increase linearly, the latter twice as fast as the former, there is a concrete continuous transition from |p_0 := (0, 5) `elem` r_0| to |p_1 := (2, 9)|.
-|p_1| is not an element of |r_1|, and since the floating point representation Alur uses presumably easily lets one conclude that |5+2*2 </ 9|, Alur's procedure manages to determine that there is no continuous transition from any point in |r_0| to a point in |r_1|, justifying suppression of any abstract continuous transition from |(Heat, r_0)| to |(Heat, r_1)| (we will discuss abstract transitions in detail in the next section).
-
-Unfortunately, our computable reals only let us compute arbitrarily close approximations of |5+2*2|. 
-Consequently, in the process of determining whether there ought to be an abstract continuous transition from |(Heat, r_0)| to |(Heat, r_1)|, our procedure (which we will describe in \ref{over_cont_trans}) will fail to conclude with certainty that this value does not lie below 9, will thus be unable to definitively establish the absence of concrete continuous transitions from points in |r_0| to points in |r_1|, and will be forced to include the transition, to ensure that the abstract transition relation respects its concrete counterpart (in a way discussed in detail in the next section).
-
-As it happens, the thermostat's safety indirectly depends on unreachability of |(Heat, r_1)|, and since |(Heat, r_0)| is a reachable state, the addition of the transition mentioned above makes the abstract system unsafe, obviously preventing us from using it to conclude safety of the concrete system.
-
-% Hence, by the fallback mechanism we will describe in \ref{bla}, the procedure \emph{will} default to emitting an abstract continuous transition from |(Heat, r_0)| to |(Heat, r_1)|.
-
-To correct this kind of problem, we tweaked many of the bounds, nudging them slightly toward one side or the other.
-For this particular example, we changed the 9 bound to 8.9.
-That way, the comparison reduces to |5+2*2 </ 8.9|, which a sufficiently close approximation \emph{can} automatically establish.
-
-\todo{Mention that the epsilon we use in computation to determine whether there's a transition from some point to another (and in the side conditions and box comparisons and such) is a parameter of the abstraction procecudure.
-for the actual safety proof, we take this epsilon to be 0.001, which makes the 0.1 nudge in things like 8.9 large enough.}
-
-\todo{Note that since the abstract system is only an overestimation of the concrete system, its unsafety does not imply unsafety of the concrete system.}
-
-\todo{Emphasize that changing the abstraction interval bounds does not mean our result isn't as "pure" as Alur's, because the whole abstraction business only serves to prove concrete system safety, and we do prove safety of the same concrete system as Alur.}
-}
 Having to do this "tweaking" manually is obviously not idea. One may want to develop heuristics for this.
-\
-
-% The main idea of abstraction is to come up with a finite set of "regions", each corresponding to a subset of the continuous state space, and to analyze traces through these regions instead of traces through points.
 
@@
 
- %(* todo: hide the nodup part? *)
-% todo: embed AState := ?
-
-\weg{ \todo{mention that the reason the space definition could be
-    mostly automated is that |regions_cover| is essentially a whole
-    bunch of if-then-else's (which are only constructive inside |DN|)}
-}
-
 \subsection{Abstract Transitions and Reachability}
 \label{abs.reach}
-
-% hm, where shoudl we introduce alternating stuff?
-
 Once we have a satisfactory abstract |Space|, our goal is to construct
@@
 simply by executing the abstract reachability overestimator. We first overestimate the continuous transitions; we need the following definition for that.
-
-\weg{
-It seems
-reasonable, then, to look for an overestimatable predicate
-|abstract.reachable| such that |reachable_respect| holds:
-\begin{code}
-Definition reachable_respect: Prop :=
-  forall (s: concrete.State), concrete.reachable s ->
-    forall (s': abstract.State), s `elem` s' -> abstract.reachable s'.
-\end{code}
-After all, this would imply
-\begin{code}
-forall (s: concrete.State) (s': abstract.State),
-   s `elem` s' -> ~ abstract.reachable s' -> ~ concrete.reachable s,
-\end{code}
-expressing that to conclude unreachability of a concrete state, one
-need only establish unreachability of an abstract state that contains
-it. However, as abstract states may overlap, the above definition of
-|reachable_respect| leads to problems, so this is not
-actually the definition we use in the development. We first introduce |shared_cover|, expressing that a collection of abstract states together covers a collection of concrete states.
-}
-
-\weg{
-However, in order
-to see the motivation for the actual definition we use, let us proceed
-as if without foresight for a few more moments.
-
-\subsection{The Straightforward (but Inadequate) Approach}
-\label{straightforward}
-
-The definition of |reachable_respect| above suggests that the property we would wish to overestimate is simply
-\begin{code}
-Definition abstract.reachable (s: abstract.State): Prop
-  := exists s': concrete.State, s' `elem` s /\ concrete.reachable s'.
-\end{code}
-In order to overestimate it, we could introduce the following obvious abstract versions of initiality and continuous and discrete transitions:
-\begin{code}
-  Definition abstract.initial (s: abstract.State): Prop
-    := exists c `elem` s, concrete.initial c.
-  Definition abstract.cont_trans: relation abstract.State
-    := fun s s' => exists (c `elem` s) (c' `elem` s'), concrete.cont_trans c c'.
-  Definition abstract.disc_trans: relation abstract.State
-    := fun s s' => exists (c `elem` s) (c' `elem` s'), concrete.disc_trans c c'.
-\end{code}
-We could then prove (by induction over traces) that |abstract.reachable| is included in the transitive closure of |abstract.cont_trans `union` abstract.disc_trans| starting at |abstract.initial|:\footnote{One would actually take the \emph{alternating} transitive closure, but we will get to that later.}
-
-\begin{code}
-  Definition abstract.trans: relation abstract.State :=
-    abstract.cont_trans `union` abstract.disc_trans.
-
-  Definition reachable_by_abstract_trace (s: abstract.State): Prop
-    := exists i, abstract.initial i /\ trans_refl_closure (abstract.trans) i s.
-
-  Lemma: abstract.reachable `subsetof` reachable_by_abstract_trace.
-\end{code}
-This would show that |abstract.reachable| could be overestimated by overestimating |reachable_by_abstract_trace|.
-For the latter, we would first define an \emph{abstract system} as a triple containing overestimators for |abstract.initial|, |abstract.cont_trans|, and |abstract.disc_trans|:
-\begin{code}
-Record abstract.System: Type :=
-  { over_initial: overestimator abstract.initial
-  ; over_cont_trans: overestimator abstract.cont_trans
-  ; over_disc_trans: overestimator abstract.disc_trans
-  }.
-\end{code}
-Given an instance of this record, we would then overestimate |reachable_by_abstract_trace| by constructing a graph with vertices representing abstract states and edges representing overestimated abstract transitions, and running an ordinary (but verified) graph reachability algorithm on this graph.
-
-Thus, the task would be reduced to construction of the three overestimators.
-
-\subsection{Clouds on the Horizon: Drift}
-\label{trouble}
-
-All of the above seems perfectly reasonable, but having gotten to the level of abstract transition overestimators, we can now see where things would (and did, in early versions of our development) go wrong.
-Unfolding the |overestimator| type specified for the |over_cont_trans| member in |abstract.System|, we get
-\begin{code}
-  forall (s s': abstract.State), { b: bool | b = false -> ~ abstract.cont_trans s s' }.
-\end{code}
-Hence, |over_cont_trans| would need to satisfy:
-\begin{code}
-  forall (s s': abstract.State), abstract.cont_trans s s' ->
-    abstract.cont_trans_over s s' = true
-\end{code}
-Which rewrites to:
-\begin{code}
-  forall (c c': concrete.State), concrete.cont_trans c c' ->
-  forall (s s': abstract.State), c `elem` s) -> c' `elem` s' ->
-    over_cont_trans s s' = true
-\end{code}
-In other words, the existence of a (continuous) transition from one concrete state to another would force inclusion of transitions between any two abstract states that contain the respective concrete states.
-When regions do not overlap, this is perfectly appropriate.
-However, consider the implications in the following example involving overlapping regions.
-
-Suppose our continuous state space is |CRge0|, and our abstract regions are intervals of the form |[n, n+1]| with |n `elem` nat|.
-Further suppose that in some location |l|, the flow function monotonically increases.
-Then by the above, |over_cont_trans| must yield |true| given |(l, [1, 2])| and |(l, [0, 1])| (in that order), because in |l|, there actually is a concrete continuous transition from a point in the former to a point in the latter, namely from 1 to 1 (since the continuous transition relation is reflexive).
-Hence, we would get an abstract transition from |(l, [1, 2])| to |(l, [0, 1])|, essentially introducing abstract flow in the opposite direction of the concrete flow.
-This can clearly have disastrous consequences for unreachability of abstract states containing unsafe states, which can render the abstraction useless.
-
-Intuitively, the transition from |(l, [1, 2])| and |(l, [0, 1])| seems \emph{redundant}, because the only point in |[0, 1]| that can be flowed to from |[1, 2]| is a point that is actually still in |[1, 2]|.
-Making |[0, 1]| reachable on account of this redundant transition seems wasteful.
-Let us formalize this notion of redundance:
-\begin{code}
-  Definition redundant_cont_trans (l: Location): relation Region
-    := fun r r' => forall p, p `elem` r -> forall t, 0 <= t -> flow l p t `elem` r' -> flow l p t `elem` r.
-\end{code}
-This relation can be lifted to abstract states. An analogous definition can be given for abstract discrete transitions.
-
-\todo{Discuss strong redundancy which holds over space composition.}
-
-We call the phenomenon of redundant transitions being generated by abstract transition overestimators: ``drift''.
-With the specifications for abstract transition overestimators given in the previous section, drift is an inescapable consequence when regions overlap, occurring not only for continuous abstract transitions (as shown above), but also for discrete abstract transitions (as will be discussed in section \ref{over_disc_trans}).
-
-It is important to note that drift is not just a consequence of ill-chosen specifications or regions.
-Suppose we use strictly non-overlapping regions.
-While this would solve the problem at the specification level (in that the specification would no longer force the overestimators to produce redundant transitions), the overestimator \emph{implementations} would still be bound to yield redundant transitions when no specific countermeasures are employed.
-As a trivial example, consider non-overlapping regions [0,1) and [1,2), again with monotonically increasing flow.
-To justify omission of an abstract continous transition from [1,2) to [0,1), an overestimator implementation would need to prove that |1 </ 1|, which is not decidable or meaningfully underestimatable given only arbitrarily close approximations of 1.
-
-Hence, our drift countermeasures will be twofold.
-First, we will first ensure that at the \emph{specification} level, transition overestimators are no longer forced to emit redundant transitions.
-Then, in sections \ref{over_cont_trans} and \ref{over_disc_trans}, we show how our overestimator \emph{implementations} exploit this freedom and use a redundancy hint mechanism to detect and omit redundant transitions.
-
-\todo{Describe somewhere how things go \emph{completely} wrong when regions have surface overlap.}
-
-\subsection{Specification Drift Avoidance: Sharing Overestimators}
-
-To avoid drift at the specification level, we must either alter the specifications of abstract transition overestimators to account for redundancy (which will necessitate analogous changes in the specifications of abstract reachability and respect, as we will see in a moment), or avoid region overlap.
-
-Practically speaking, using non-overlapping regions implies the use of non-strict equalities, as in the example above.
-As it happens, non-strict and strict constructive inequalities on real numbers live in different universes: the former live in |Prop|, while the latter live in |Type| \todo{explain why}.
-Since derived propositions (such as |p `elem` r| for a point |p| and a region |r|) all inherit this trait, we would essentially have to give up on |Prop| for the entire development.
-Unfortunately, several components in the Coq system, including the highly useful subset coercion functionality of the |Program| family of commands, currently only support |Prop|.
-While these are strictly engineering concerns that may or may not have theoretical substance, they do matter when developing a working system.
-
-On the other hand, changing the abstract transition overestimator specification, abstract reachability, and respect specifications, turns out to be a relatively local change, affecting only a few hundred lines of code, most of them in the one module that introduces these concepts and proves some key lemmas about them.
-
-From an engineering perspective, then, the choice is easy. We will now describe the changes in abstract transition overestimator specification in detail.
-
-% todo: The continuous hint mechanism only works with border-overlap.
-% todo: how about making abstraction bounds rationals?
-% Drift at the specification level (as discussed in the previous section) \emph{can} be avoided by restricting oneself to abstract spaces whose regions do not overlap, but drift at the implementation level cannot.
-% blabla evil---even though we're using alternating transitions.
-% Discrete abstract transitions suffer the same fate, but in a slightly
-
-\noindent In the discussion of the first example in the previous section, we judged the redundant abstract transition from |(l, [1, 2]|) to |(l, [0, 1])| to be ``wasteful'' because the destination of the sole concrete continuous transition from 1 to 1 that spawned it was already covered by |[1, 2]|, and so if |[1, 2]| is reachable, there should be no reason to make |[0, 1]| reachable as well.
-Implicit in this intuition is the idea that regions ought to \emph{share} the burden of being reachable on behalf of the points they represent: if 1 is a reachable concrete point, then reachability of |[1, 2]| should remove the need for reachability of |[0, 1]| (to represent reachability of 1), and vice versa.
-
-Unfortunately, as shown in the previous section, the overestimator specifications used in the definition of |abstract.System| given in section \ref{straightforward} do not permit this sharing.
-Let us consider how they would need to be changed to allow for it.
-We can easily make the unfolded specification of |over_cont_trans| shown in the previous section permit sharing by substituting one universal quantifier with an existential one:
-\begin{code}
-  forall c c', concrete.cont_trans c c' ->
-  forall (s: abstract.State), c `elem` s ->
-  exists (s': abstract.State), c' `elem` s' /\
-    over_cont_trans s s' = true
-\end{code}
-Applied to the example above, the existence of a concrete continuous transition from 1 to 1 now only results in the requirement that
-\begin{code}
-  forall (s: abstract.State), (l, 1) `elem` s ->
-  exists (s': abstract.State), (l, 1) `elem` s' /\
-    over_cont_trans s s' = true
-\end{code}
-This means |over_cont_trans| now has the freedom to return |false| given |(l, [1, 2])| and |(l, [0, 1])| (thereby suppressing creation of that abstract transition) if it returns |true| given |(l, [0, 1])| and |(l, [0, 1])| (which it will have to anyway).
-
-To give |over_cont_trans| and |over_disc_trans| new types that embody this new specification, to replace the old |overestimator| types, we first introduce |shared_cover|:
-}
 
@@
 contains it.
-%\footnote{The double negation is here for much the same
-%  reasons as were given in section \ref{regionsandstates} for the
-%  double negation in the type of |regions_cover|.} 
 We state this in
@@
 
-\weg{
-Stated in terms of |shared_cover|, a reasonably straightforward type for |over_cont_trans| would be:
-\begin{code}
-over_cont_trans: forall s: abstract.State,
-  { p: abstract.State -> bool |
-    shared_cover (fun c' => exists c `elem` s, concrete.cont_trans c c') p }
-\end{code}
-That is, |over_cont_trans| would be a function that, given an abstract state, returns a boolean predicate on abstract states that, when interpreted as a container in the obvious way, covers all concrete states reachable by a single concrete continuous transition from a concrete state contained in the original abstract state.
-We make a few last adjustments to arrive at the actual type used in the development.
-
-First, the abstract transition only actually needs to cover points at which the invariant holds.
-Integrating that fact in the type of |over_cont_trans| allows us to use |regions_cover| in its implementation:
-\begin{code}
-over_cont_trans: forall s: abstract.State,
-  { p: abstract.State -> bool | shared_cover
-    (fun c' => concrete.invariant c' /\ exists c `elem` s, concrete.cont_trans c c') p }
-\end{code}
-Using some |Container| utilities, we may write this in a more point-free fashion:
-%format `inter` = "\cap "
-%format `compose` = "\circ "
-%format utilflip (a) = a "^{-1}"
-%format prettysig = "\Sigma"
-\begin{code}
-over_cont_trans: forall s: abstract.State,
-  { p: abstract.State -> bool | shared_cover
-    (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p }
-\end{code}
-The use of a boolean predicate is less than ideal because it means that to enumerate all abstract states directly reachable from some given abstract state (which we will have to do eventually when computing reachability in a graph whose edges are built from these overestiamtors), one basically needs to filter an exhaustive list of all abstract regions using the predicate, while one could easily imagine that clever overestimator implementations could exploit locality to exclude certain classes of potential transition destinations a priori from consideration.
-To facilitate such overestimators (which we do not develop in this report), we replace the boolean predicate with a \emph{list} of abstract states.
-Since lists are also |Container|s, the rest of the definition remains unchanged:
-\begin{code}
-over_cont_trans: forall s: abstract.State,
-  { p: list abstract.State | shared_cover
-    (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p }
-\end{code}
-
-Lists may contain duplicates.
-The graph reachability algorithm which we will use in section \ref{graphs} relies on absence of duplicates in the edge relation (expressed as a function returning lists) for its termination proof.
-While we could filter out duplicates at a much later stage to satisfy that requirement, we choose to let the no-duplicates requirement propagate through to our overestimator specification, avoiding premature pessimization:
-}
-
 We now specify what the type of $\overcontrans$ should be.
@@
 So, |over_cont_trans s| should produce a list of abstract states |p|
-without duplicates (to enhance the graph reachability algorithm that
-we don't detail here, but that will do the actual computation of the
-reachability set), such that |p| is a shared cover of the collection
+without duplicates,
+%(to enhance the graph reachability algorithm that
+%we don't detail here, but that will do the actual computation of the
+%reachability set), 
+such that |p| is a shared cover of the collection
 of concrete states |c| that satisfy the invariant and whose set of
@@
 \mid \Inv(c) \wedge s\cap \{c' \mid c' \contrans c\} \neq \emptyset\}$.
-
-\weg{
-The |vm_compute| tactic we will ultimately use to run our program
-tends to reduce terms in |Prop| a bit too eagerly, adversely affecting
-performance. So to work around this, we have introduced a bit of laziness in
-the form of a little ``unit abstraction'' that we will not detail here.
-}
-
 We similarly specify |over_disc_trans| as an overestimator for
@@
 |concrete.initial|.
-
- \weg{
-\begin{code}
-over_cont_trans: forall s: abstract.State,
-  { p: list abstract.State | () -> (NoDup l /\ shared_cover
-    (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p) }
-\end{code}
-This will still let us use the property in proofs, but will stop |vm_compute| from unnecessarily evaluating its proof term.
-
-Finally, we actually need the same definition for abstract discrete transitions, so we factor out the common part:
-\begin{code}
-Definition sharing_transition_overestimator
-  (R: relation concrete.State): Set := forall s: abstract.State,
-    { p: list abstract.State | () -> (NoDup l /\ shared_cover
-      (concrete.invariant `inter` (overlap s `compose` utilflip R)) p) }.
-\end{code}
-We can now show the definition of |abstract.System| as it appears in the development:
-\begin{code}
-  Record abstract.System: Type :=
-    { over_initial: overestimator (overlap concrete.initial)
-    ; over_disc_trans: sharing_transition_overestimator concrete.disc_trans
-    ; over_cont_trans: sharing_transition_overestimator concrete.cont_trans
-    }.
-\end{code}
-
-\subsection{Propagating Sharing Upwards}
-
-Consider again the definition of |abstract.reachable| given in section \ref{straightforward}:
-\begin{code}
-Definition abstract.reachable (s: abstract.State): Prop
-  := exists s': concrete.State, s' `elem` s /\ concrete.reachable s'.
-\end{code}
-Clearly, this definition does not let abstract states share the burden
-of reachability, and can therefore not be overestimated using our
-retyped abstract transition overestimators. But recall that it
-followed directly from
-}
 
@@
 \end{code}
+\weg{
 This property follows from the following new definition of |reachable_respect| we will use:
@@
 \end{code}
-% This is only natural--in the example, to conclude unreachability of the concrete state |(l, 1)|, it is no longer sufficient to establish unreachability of the abstract state |(l, [0, 1])|.
-\weg{
-This new definition does not give rise to a nice simple definition of
-|abstract.reachable| distinct from the simple transitive closure of
-initiality and transition judgements made by the overestimators
-(analogous to |reachable_by_abstract_trace|), so we just take the
-latter and prove |reachable_respect| by induction over traces. There
-is, however, one last subtlety involved.
-
-%  Let us see what happens when we redefine it directly in terms of our abstract system overestimators, much like how |reachable_by_abstract_trace| was defined earlier:
-% \begin{code}
-% Definition trans (b: bool): relation abstract.State :=
-%   fun s1 s2 => s2 `elem` if b
-%     then cont_trans_over ahs s1
-%     else disc_trans_over ahs s1.
-% 
-% Definition abstract.reachable (s: abstrat.State): Prop :=
-%   exists is: State ap,
-%     overestimation_bool (initial_dec ahs is) = true /\ alternate trans is s.
-% \end{code}
-% By induction over traces, we can prove
-% \begin{code}
-% shared_cover concrete.reachable abstract.reachable,
-% \end{code}
-% which rewrites to
-% Having changed the abstract transition overestimator specifications to permit sharing, the definition of |abstract.reachable| given in \ref{inadequate} is no longer appropriate, 
-% we can no longer use these to overestimate |abstract.reachable| as defined in \ref{inadequate}.
-After all, when the burden of reachability is shared among abstract states, some abstract states may actually be unreachable even though they contain reachable concrete states.
-% 
-% In section \ref{inadequate}, we defined |reachable_by_abstract_trace| in terms of |abstract.cont_trans| and |abstract.desc_trans|, which were the simplistic ``is there a concrete transition between the regions?'' relations that our abstract transition overestimators overestimated, but which had no notion of sharing.
-Having changed the abstract transition overestimator specifications to permit sharing, we must mimic this change in |abstract.cont_trans| and |abstract.desc_trans| for |reachable_by_abstract_trace| to remain representative.
-Hence, we redefine these relations to simply reflect overestimator judgements:
-% \begin{code}
-% Definition abstract.cont_trans: relation abstract.State
-%   := fun s1 s2 => s2 ∈ cont_trans_over ahs s1.
-% 
-% Definition abstract.disc_trans: relation abstract.State
-%   := fun s1 s2 => s2 ∈ disc_trans_over ahs s1.
-% \end{code}
-% Unfortunately, after these changes, |reachable_by_abstract_trace `subsetof` abstract.reachable| is no longer provable, because after all, when the burden of reachability is shared among abstract states, some abstract states may actually be unreachable even though they contain reachable concrete states.
-However, we \emph{can} prove (again, by induction over traces) a variation of that bypasses |abstract.reachable| altogether, and embraces sharing:
-% \begin{code}
-% (* SharedCover concrete.reachable reachable *)
-% 
-% Definition reachable_respect: Prop :=
-%   forall (s: concrete.State),
-%     (forall s': abstract.State, s `elem` s' -> ~ reachable_by_abstract_trace s') ->
-%     ~ concrete.reachable s.
-% \end{code}
-% Where the previous definition of |unreachable_respect| stated that to conclude unreachability of a concrete state one needed only to prove unreachability of \emph{any} abstract state that contained it, this revised definition requires one to prove unreachability of \emph{all} abstract states containing the concrete state.
-We can see this in the example: to conclude unreachability of the concrete state |(l, 1)|, it is no longer sufficient to establish unreachability of the abstract state |(l, [0, 1])|.
-% 
-% At this point, what's left is to implement the three overestimators that make up an abstract system, and to use them to overestimate |reachable_by_abstract_trace|.
-However, before we do that, we make one more change.
 }
@@
 hybrid system safety, and in particular, lets us obtain a proof of
-thermostat safety. So suppose we have
-\begin{code}
-  reachable_dec: decider abstract.reachable
-\end{code}
+thermostat safety. So suppose we have |reachable_dec: decider abstract.reachable|
+\weg{
 And suppose we are given the following specification of unsafe concrete states, covered by a finite list of abstract states:
@@
 \end{code}
+}
 Recall that |ThermoSafe| was defined as |thermo_unsafe `subsetof` concrete.unreachable| in section \ref{concrete}.
@@
 Proof.
-  apply (underestimation_true under_unsafe_unreachable).
-  vm_compute. sp1mm reflexivity.
+  apply (underestimation_true under_thermo_unsafe_unreachable).
+  vm_compute.  reflexivity.
 Qed.
@@
 |underestimation_true| is a tiny utility of type |forall P (o: underestimation P), o = true -> P|, whose application in the proof reduces the goal to
-\begin{code}under_thermo_unsafe_unreachable = true.\end{code} The |vm_compute| tactic invocation then forces evaluation of the left hand side, which will in turn evaluate |over_thermo_unsafe_reachable|, which will evaluate |reachable_dec|, which will (as we will see in the next section) evaluate the |abstract.System| overestimators (which we will build in sections \ref{over_cont_trans} and \ref{over_disc_trans}).
+\begin{code}under_thermo_unsafe_unreachable = true.\end{code} The |vm_compute| tactic invocation then forces evaluation of the left hand side, which will in turn evaluate |over_thermo_unsafe_reachable|, which will evaluate |reachable_dec|, which will evaluate the overestimators of the continuous and discrete transitions.
 This process, which takes about 35 seconds on a modern desktop machine, eventually reduces |under_thermo_unsafe_unreachable| to |true|, leaving |true = true|, proved by |reflexivity|.
 
-We can now also clearly see what happens when the abstraction method ``fails'' due to poor region selection, overly simplistic transition/initiality overestimators, or plain old unsafety of the system.
+We can now also clearly see what happens when the abstraction method ``fails'' due to poor region selection, overly simplistic transition/initiality overestimators, or plain unsafety of the system.
 In all these cases, |vm_compute| reduces |under_thermo_unsafe_unreachable| to |false|, and the subsequent |reflexivity| invocation will fail.
@@
 for the thermostat case.
-%latter
-%in sections \ref{over_cont_trans} and \ref{over_disc_trans}.
-
- % using  the particular heuristics and techniques we used to implement the decidable overestimators for abstract initiality and transitions, the specifications (types) of which were given in section \ref{abs.trans}.
-
-% next up:
-% - defining cont_trans
-% - defining disc_trans
-% - definign initial
-% - square_abstraction
-
-\weg{
-\section{Graphs for Reachability Decision}
-\label{graphs}
-
-Given an |abstract.System| containing overestimators for abstract transitions and initiality, we can decide |abstract.reachable| (as defined in the last section) by constructing a directed graph in which vertices and edges correspond to abstract states and transitions, respectively, and then using an ordinary graph reachability algorithm.
-Because the hybrid system safety proofs we ultimately wish to produce obviously depend on the correctness of the reachability determinations obtained this way, we have implemented a certified graph reachability algorithm inside Coq.
-\todo{say something about algorithm?
-does it have a name?} For efficiency reasons, it computes a list of all vertices reachable from a given list of initial vertices in one go:
-\begin{code}
-graph_reachables: forall (g : DiGraph) (start: list Vertex), NoDup start ->
-  {l : list Vertex | forall w : Vertex, In w l <-> digraph.reachable start w}
-\end{code}
-Here, |DiGraph| and |digraph.reachable| are defined as
-\begin{code}
-Record DiGraph: Type := Build_DiGraph
-  { Vertex: Set
-  ; Vertex_eq_dec: EqDec Vertex eq
-  ; vertices: ExhaustiveList Vertex
-  ; edges: Vertex -> list Vertex
-  ; edges_NoDup: forall v, NoDup (edges v)
-  }.
-
-Definition edge: relation Vertex := fun v w => In w (edges v).
-
-Definition digraph.reachable (start: list Vertex) (v: Vertex): Prop :=
-  exists s `elem` start, trans_refl_closure edge s v.
-\end{code}
-
-Naively equating vertices with abstract states and edges with abstract transitions (either continuous or discrete) would make vertex reachability in the graph equivalent to potentially-non-alternating abstract state reachability, which as we saw in section \ref{alter} would result in many needlessly reachable abstract states, potentially rendering the abstraction useless.
-
-To make vertex reachability in the graph equivalent to \emph{alternating} abstract state reachability, we first add a |bool| component to |Vertex|:
-\begin{code}
-Definition Vertex: Set := prod bool abstract.State.
-\end{code}
-Next, we only add edges corresponding to abstract \emph{continuous} transitions from a vertex |v| with |fst v = true| to a vertex |w| with |fst w = false|, and only add edges corresponding to abstract \emph{discrete} transitions from a vertex |v| with |fst v = false| to a vertex |w| with |fst w = true|:
-
-\begin{code}
-  Definition nexts (v: Vertex): list abstract.State :=
-    let (k, s) := v in
-      if k then over_cont_trans s
-      else over_disc_trans s.
-
-  Definition edges (v: Vertex): list Vertex := map (pair (negb (fst v))) (nexts v).
-
-  Definition graph: DiGraph := Build_DiGraph Vertex edges.
-    (*{-"\text{ Remaining record fields omitted for the sake of exposition. }"-}*)
-\end{code}
-This ensures that paths through the graph alternate between vertices |v| with |fst v = true| and vertices with |fst v = false|, and that the corresponding abstract traces alternate between continuous and discrete transitions.
-A few simple inductions now show that if we take as |start| those vertices selected by our initiality overestimator, |abstract.reachable| is equivalent to |digraph.reachable|:
-\begin{code}
-  Definition start: list Vertex := { v: Vertex | abstract.over_initial (snd v) }.
-
-  Theorem graph_respect: forall s: abstract.State ap,
-    abstract.reachable s <-> exists b: bool, digraph.reachable start (b, s).
-\end{code}
-
-Combining |graph_respect|, |abstract.unreachable_respect|, and |reachables|, it is now straightforward to construct a |decider abstract.reachable|.
-\todo{Show some more detail?}
-
-Through the definitions of |edges|, |nexts|, and |start|, this decider builds on the three overestimators in |abstract.System|.
-In the next two sections, we discuss how to define an |abstract.System| instance for the discretization (described in section \ref{abs.space.constr}) of the concrete thermostat's continuous state space.
-}
 
@@
 We now discuss the implementation of  |over_cont_trans|:
-\weg{
-\begin{code}
-  over_cont_trans: sharing_transition_overestimator concrete.cont_trans
-\end{code}
-where |sharing_transition_overestimator| is defined as
-\begin{code}
-Definition sharing_transition_overestimator
-  (R: relation concrete.State): Set := forall s: abstract.State,
-    { l: list abstract.State | () -> (NoDup l /\ shared_cover
-      (concrete.invariant `inter` (overlap s `compose` utilflip R)) l) }.
-\end{code}
-where |shared_cover| is defined as
-\begin{code}
-  Definition shared_cover
-    `{Container concrete.State C} `{Container abstract.State D}
-      (cs: C) (ss: D): Prop :=
-        forall s: concrete.State, In s cs -> DN (exists r: abstract.State, In s r /\ In r ss).
-\end{code}
-
-\subsection{Avoiding Drift}
-
-A first observation is that since the concrete continuous transition relation is reflexive, |over_cont_trans| must be reflexive as well (meaning its result list must include the abstract state it received as an argument).
-By the sharing principle, this means that redundantly reachable regions (whose only reachable points lie within the source region) are accounted for, which in turn means the problem can be reduced to the overestimation of non-redundant transitions, characterized by the following relation:
-\begin{code}
-Definition substantial_cont_trans (l: Location): relation Region
-  := fun r r' => exists p p', p `elem` r /\ p' `elem` r' /\ ~ p' `elem` r /\ can_flow l p q.
-\end{code}
-Hence, we seek to build an |overestimator substantial_cont_trans|.
-Note that even though redundantly reachable regions are ``accounted for'' specification-wise, we still need a mechanism enabling us to notice their redundancy, so that our overestimator may return |false|.
-Unfortunately, detecting whether a particular abstract continuous transition would be redundant is undecidable in general.
-For now, let us simply assume existence of a redundance underestimator, and worry about its implementation later:
-
-% 
-% 
-% not notice their redundancy.
-% 
-% In section \ref{trouble}, we illustrated how drift would cause problems if we gave the |over_cont_trans| member in |abstract.System| the type
-% \begin{code}
-% overestimator abstract.cont_trans.
-% \end{code}
-% While appealing in its simplicity---such overestimators only ever consider two regions at a time---this specification would force inclusion of redundant transitions like the one from |[1,2]| to |[0,1]| in the example.
-To counteract this outcome, we reformulated the specification for abstract continuous transitions as a predicate |cont_respect| on functions that, given an abstract state, yield a list of regions representing abstract continuous transitions.
-This permitted such functions to exploit the sharing principle and omit redundant transitions.
-
-\begin{code}
-Variable under_redundant: underestimator redundant_cont_trans.
-\end{code}
-The idea now is to build a drift-oblivious |overestimator| for the simplistic |abstract.cont_trans| relation, which \emph{will} emit redundant transitions, but to only run it if the redundancy underestimator does not indicate redundance.
-That is, supposing we have
-\begin{code}
-Variable simple_over_cont_trans: overestimator abstract.cont_trans,
-\end{code}
-we can define
-
-%format r_src
-%format r_dst
-%format p_src
-%format p_dst
-%format negb (x) = ~ x
-\begin{code}
-  Program Definition over_subst_cont_trans (l: Location) (r_src r_dst: Region)
-    : overestimation (substantial_cont_trans l r_src r_dst)
-    := negb (under_redundant l r r') && simple_over_cont_trans l r_src r_dst.
-\end{code}
-
-Thus, the problem is reduced to construction of |under_redundant| and |simple_over_cont_trans|.
-We discuss the latter in the next section.
-For the former, we observe that the redundant transitions we worry about appear as transitions from a region to a region adjacent to it in the opposite direction of the flow corresponding to the current location.
-This means that if we know which regions are adjacent, and further know when (e.g.
-in which locations) individual flow components are monotonic (and in which direction), we can underestimate redundancy quite well.
-Since in our development abstract regions are formed by explicitly enumerating interval bounds, we have a-priori knowledge of which regions will be adjacent.
-Hence, the only additional work we have to do is to provide as many proofs of flow monotonicity (in the different locations and for the different flow components) as possible, with which some generic utilities can construct a proper redundancy underestimator.
-
-Having taken care of drift, all we need now (to complete the definition of |over_cont_trans|) is an implementation of |simple_over_cont_trans|, which we present in the next section.
-
-\subsection{Simple Transition Overestimation}
-
-Recall that |abstract.cont_trans| is defined as
-\begin{code}
-Definition abstract.cont_trans: relation abstract.State
-  := fun s s' => exists (c `elem` s) (c' `elem` s'), concrete.cont_trans c c'.
-\end{code}
-Since location does not change in continuous transitions, and location equality is decidable (since locations are just names), we can easily reduce the problem down to overestimation of
-\begin{code}
-Definition region_cont_trans (l: Location): relation Region
-  := fun r_src r_dst => exists (p_src `elem` r_src) (p_dst `elem` r_dst'), can_flow l p_src p_dst,
-\end{code}
-where |can_flow| is still
-\begin{code}
-Definition can_flow (l: Location): relation Point
-  := fun p_src p_dst => exists d: Duration, flow l p_src d = p_dst /\
-    forall t, 0 <= t <= d -> invariant (l, flow l p_src t).
-\end{code}
-So given a location |l| and regions |r_src| and |r_dst|, we are looking to establish absence of concrete transitions from points in |r_src| to points in |r_dst|.
-
-\todo{story about how Alur does it, and why we don't just do the same}
-
-%\subsection{Ruling out Transitions by Invariant Absence}
-
-\noindent Let us first consider whether we can rule out transitions by looking at the invariant, which |can_flow| demands must hold at every point along the way.
-Clearly, if we are able to determine that there do not exists paths from points in |r_src| to points in |r_dst| with the invariant holding at each point along the way, then there need be no continuous transition between |r_src| and |r_dst|.
-Unfortunately, \emph{deciding} this property is too hard.
-However, suppose we can at least overestimate whether the invariant holds somewhere in an abstract region:
-\begin{code}
-  Variable invariant_overestimator: overestimator abstract.invariant.
-\end{code}
-With |invariant_overestimator|, we can start defining |simple_over_cont_trans| as follows:
-\begin{code}
-  Definition simple_over_cont_trans: overestimator abstract.cont_trans
-    := fun r_src r_dst =>
-      invariant_overestimator r_src &&
-      invariant_overestimator r_dst && (*{-"\text{ .. further tests, to be discussed. }"-}*).
-\end{code}
-That is, if either |~ invariant_overestimator r_src| or |~ invariant_overestimator r_dst|, then we conclude that there can be no concrete transition from a point in |r_src| to a point in |r_dst|.
-To keep matters simple, we will not bother to check the invariant anywhere else.
-This does mean that we might generate a spurious abstract transition from |r_src| to |r_dst| if the only reason for lack of concrete transitions between |r_src| and |r_dst| is that the invariant be broken somewhere in the middle of the flow path.
-Fortunately, this turns out not to cause problems for the thermostat hybrid system.
-\todo{Can anything be said about to what extent this simplification is justifiable or even necessary for hybrid systems in general?} In our development, the definition of |invariant_overestimator| is almost automatic, requiring only that the thermostat's invariant be reformulated in terms of (possibly unbounded) squares in the continuous state space, whose intersection with abstract regions can be automatically overestimated.
-
-Next, let us consider }
-
-Given two regions |r_src| and |r_dst|, we have to think how we can
-rule out the existence of concrete continuous transitions from points
-in |r_src| to points |r_dst|. Clearly, if we can determine that there
+Given two regions |r_src| and |r_dst|, 
+%we want to
+%rule out the existence of concrete continuous transitions from points
+%in |r_src| to points |r_dst|. Clearly, 
+if we can determine that there
 are no points in |r_src| which the flow function maps to points in
-|r_dst|, then there need be no continuous transition between |r_src|
-and |r_dst|. Equally clearly, this is utterly impossible to
+|r_dst|, then there need be no abstract continuous transition between |r_src|
+and |r_dst|. Clearly, this is impossible to
 meaningfully overestimate for a general flow function and general
@@
 
-% Finally, we get to the real ``meat'' of our overestimation: overestimating whether there is a point in |r_src| that the flow function maps to a point in |r_dst| in the first place.
-
 The notion of {\em separability\/} has already been discussed in
 Section \ref{sec:hybsys}.
-%: a flow function on a product space is
-%\emph{separable} if it can be written as the product of flow function
-%on the respective component spaces. In $\IR^2$, a flow function is
-%separable if future values of the first component only depend on past
-%values of that same component, and the similarly for the second
-%component. Being defined as the product of two flow functions on |CR|
-%(|temp_flow| and |temp_flow|), 
-The thermostat's flow function (in any
-location) is separable by construction.
+%The thermostat's flow function (in any
+%location) is separable by construction.
 
@@
 are intervals in $\IR$, then $F(a,b)$ is an interval that contains all
-$t$ for which $\phi(x,t)\in b$ for some $x\in a$. So $F(a,b)$ should
-contain all durations $t$ that take some elements of $a$ to some
-element of $b$ via flow $\phi$.
-
+$t$ for which $\phi(x,t)\in b$ for some $x\in a$. 
+%So $F(a,b)$ should
+%contain all durations $t$ that take some elements of $a$ to some
+%element of $b$ via flow $\phi$.
 Range invertibility is a less demanding alternative to point invertibility: $\phi^{-1}$ is the {\em point inverse\/} of $\phi$ if $\forall x,y\in \IR (\phi (x, \phi^-1(x,y)) =y)$.
 So a point inverse $\phi^{-1}(x,y)$ computes the exact time it takes to go from $x$ to $y$ via flow $\phi$.
-\weg{
-\begin{code}
-  exists (point_inverse: CR -> CR -> CR), forall p q: CR, f p (point_inverse p q) = q
-\end{code}
-}
-We need range invertibility because the exponential flow functions are
-not point invertible.
+%We need range invertibility because the exponential flow functions are
+%not point invertible.
 
@@
 show that the thermostat's flow functions are range-invertible.
-
-%format r_src_temp = r_src _temp
-%format r_dst_temp = r_dst _temp
-%format r_src_clock = r_src _clock
-%format r_dst_clock = r_dst _clock
-
 Having defined the class of separable range-invertible flow functions,
@@
 |r_src| which the flow function map to points in |r_dst|. 
-%As discussed
-%in section \ref{abs.space.constr}, the abstract space for our
-%thermostat was constructed as the product of two abstract spaces based
-%on temperature and clock intervals, respectively. 
 Regions in the abstract space for our thermostat are basically pairs
@@
 
-% todo: these paragraphs make it seem as though this algorithm was hard coded for temperature/clock, while it just works because the thermostat has product flow/space.
-%format halfrange (a) (b) = [a, b)
-
-\weg{
-Having shown at last the ``real'' flow test condition used in the overestimation, we can show more precisely how drift would arise if we avoided region overlap (thereby eliminating drift problems at the specification level) but did not bother to employ a redundancy underestimator.
-Suppose there were clock intervals [0, 1) and [1, 2), and a temperature interval [5, 6).
-Let us consider what our simple |abstract.cont_trans| overestimator would return given the abstract states |(Heat, (half_range 5 6, halfrange 1 2)| and |(Heat, (halfrange 5 6, halfrange 0 1)|.
-After determining that the locations are equal as required, and that the invariant holds in both regions, our overestimator would proceed to overestimate overlap of the ranges mentioned above.
-In this particular case, with |clock_flow| simply being linear, it would compute that
-\begin{code}range_inverse clock_flow r_src_clock r_dst_clock = halfrange (-2) 0\end{code} Unfortunately, the overestimator would not be able to determine that [0, inf) does not overlap with [-2, 0), because given only two approximations of 0, it cannot determine that |0 <= 0|.
-Thus, it would have no choice but to bless the transition.
-
-\subsection{Overestimating Discrete Abstract Transitions}
-\label{over_disc_trans}
-
-We now turn our attention to the implementation of the |over_disc_trans| component in |abstract.System|. Recall (from section \ref{conc.disc_trans}) the definition of concrete discrete transitions:
-\begin{code}
-  Definition concrete.disc_trans: relation concrete.State := fun s s' =>
-    guard s (location s') /\
-     reset (location s) (location s') (point s) = point s' /\
-      invariant s /\ invariant s'.
-\end{code}
-
-To overestimate whether there exists a concrete discrete transition from a concrete state in one abstract state to a concrete state in another abstract state, we first have |over_disc_trans| simply check whether the invariant holds at both ends (exactly as in the continuous case), and (in the same way) whether the guard holds at the origin.
-The interesting part is the overestimation of whether there is a point in the source region that the reset function maps to a point in the destination region.
-This problem sounds rather similar to that of flow overestimation discussed in section \ref{over_cont_trans}, but the similarity is superficial for two reasons: reset functions are much simpler as they don't have a time parameter, and discrete transitions suffer from a substantially different variety of drift.
-
-Suppose one's continuous state space is |CRge0|, and one's abstract regions are intervals of the form |[n, n+1]| with |n `elem` nat|.
-Further suppose that there is a discrete transition from location |l| to location |l'| with guard |g = const True| and reset function |r = id|.
-And suppose we are overestimating whether there needs to be an abstract discrete transition from |(l, [0, 1])| to |(l', [1, 2])|.
-Including this transition seems wasteful, because the only point in |[1, 2]| that the reset function maps a point in |[0, 1]| to is 1, but that point is still in |[0, 1]|.
-
-For continuous transitions, we avoided drift by explicitly filtering out transitions found to be redundant by a redundance underestimator that relied on reflexivity of the continuous transition relation, knowledge of adjacent regions, and knowledge of where the flow function was monotonic.
-However, the discrete transition relation is not necessarily reflexive, so we use a slightly different mechanism.
-Basically, we simply take as an argument an underestimation of whether a particular reset function is the identity function, and if the underestimator says that it is, then we only make abstract discrete transition to abstract states with the same region as the source state.
-If on the other hand the underestimator does \emph{not} indicate that the reset function in question is the identity function, then we simply take the bounds of the source region, map them with the reset function, and overestimate whether the resulting region overlaps with the candidate destination (for this to be valid, the reset function must be nondecreasing\fotonote{When we say that a \emph{flow} function is ``nondecreasing'', we mean that it does not decrease over time, and so this is really a statement about flow direction.
-But when we say that a \emph{reset} function (which does not take a time argument) is ``nondecreasing'', we simply mean that it does not flip ranges around, which is a rather mundane property.}
-
-\todo{In the code, this is actually done in a more roundabout way, including an extra case for constant reset functions, but I really think this should not be necessary.
-Preliminary experiments with removing the third case mysteriously made the safety proof fail though, so deeper investigation is needed.}
-
-For all this to work for a system with a continuous state space in |pow CR n|, the reset function must additionally be separable. Our thermostat's reset functions are trivially separable.
-
-% todo: mention that we don't discuss boring things like over_initial and the abstract safety list, even though they're all treated uniformly: the user must specify these as lists of opensquares when setting up the abstraction.
-}
 \section{Related work}
@@
 \section{Conclusions and further research}
-\label{conclusions}
 The presented verification of hybrid systems in Coq gives a nice
@@
 is required here. Finally, in case safety cannot be proved, one would
-like the system to also generate an ``offending trace'' automatically,
+like the system to generate an ``offending trace'' automatically,
 which can the be inspected by the user.
 
-\bibliographystyle{splncs}
-\addcontentsline{toc}{chapter}{Bibliography}
+%\bibliographystyle{splncs}
+\bibliographystyle{plain}
+%\addcontentsline{toc}{chapter}{Bibliography}
 \bibliography{paper}

### 2010-03-22 ITP10/paper.tex
Add ITP10 paper sources.
Ignore-this: 49de24983449dd9b9daef53ef1da6000

darcs-hash:20100322174102-bab43-dcc3cc0af50eb50c30ef52b815fff99223eb4669.gz

--- 2010-03-22 ITP10/hybrid.tex
@@
 \usepackage{amsmath}
-%include format.tex
+%% ODER: format ==         = "\mathrel{==}"
+%% ODER: format /=         = "\neq "
+%
+%
+\makeatletter
+\@ifundefined{lhs2tex.lhs2tex.sty.read}%
+  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
+   \newcommand\SkipToFmtEnd{}%
+   \newcommand\EndFmtInput{}%
+   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
+  }\SkipToFmtEnd
+
+\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
+\usepackage{amstext}
+\usepackage{amssymb}
+\usepackage{stmaryrd}
+\DeclareFontFamily{OT1}{cmtex}{}
+\DeclareFontShape{OT1}{cmtex}{m}{n}
+  {<5><6><7><8>cmtex8
+   <9>cmtex9
+   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
+\DeclareFontShape{OT1}{cmtex}{m}{it}
+  {<-> ssub * cmtt/m/it}{}
+\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
+\DeclareFontShape{OT1}{cmtt}{bx}{n}
+  {<5><6><7><8>cmtt8
+   <9>cmbtt9
+   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
+\DeclareFontShape{OT1}{cmtex}{bx}{n}
+  {<-> ssub * cmtt/bx/n}{}
+\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
+
+\newcommand{\Sp}{\hskip.33334em\relax}
+
+\newcommand{\Conid}[1]{\mathit{#1}}
+\newcommand{\Varid}[1]{\mathit{#1}}
+\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
+\newcommand{\plus}{\mathbin{+\!\!\!+}}
+\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
+\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
+\newcommand{\sequ}{\mathbin{>\!\!\!>}}
+\renewcommand{\leq}{\leqslant}
+\renewcommand{\geq}{\geqslant}
+\usepackage{polytable}
+
+%mathindent has to be defined
+\@ifundefined{mathindent}%
+  {\newdimen\mathindent\mathindent\leftmargini}%
+  {}%
+
+\def\resethooks{%
+  \global\let\SaveRestoreHook\empty
+  \global\let\ColumnHook\empty}
+\newcommand*{\savecolumns}[1][default]%
+  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
+\newcommand*{\restorecolumns}[1][default]%
+  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
+\newcommand*{\aligncolumn}[2]%
+  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
+
+\resethooks
+
+\newcommand{\onelinecommentchars}{\quad-{}- }
+\newcommand{\commentbeginchars}{\enskip\{-}
+\newcommand{\commentendchars}{-\}\enskip}
+
+\newcommand{\visiblecomments}{%
+  \let\onelinecomment=\onelinecommentchars
+  \let\commentbegin=\commentbeginchars
+  \let\commentend=\commentendchars}
+
+\newcommand{\invisiblecomments}{%
+  \let\onelinecomment=\empty
+  \let\commentbegin=\empty
+  \let\commentend=\empty}
+
+\visiblecomments
+
+\newlength{\blanklineskip}
+\setlength{\blanklineskip}{0.66084ex}
+
+\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
+\let\hspre\empty
+\let\hspost\empty
+\newcommand{\NB}{\textbf{NB}}
+\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
+
+\EndFmtInput
+\makeatother
+%
+%
+%
+%
+%
+%
+% This package provides two environments suitable to take the place
+% of hscode, called "plainhscode" and "arrayhscode". 
+%
+% The plain environment surrounds each code block by vertical space,
+% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
+% similar to formulas. Note that if these dimensions are changed,
+% the spacing around displayed math formulas changes as well.
+% All code is indented using \leftskip.
+%
+% Changed 19.08.2004 to reflect changes in colorcode. Should work with
+% CodeGroup.sty.
+%
+\ReadOnlyOnce{polycode.fmt}%
+\makeatletter
+
+\newcommand{\hsnewpar}[1]%
+  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
+
+% can be used, for instance, to redefine the code size, by setting the
+% command to \small or something alike
+\newcommand{\hscodestyle}{}
+
+% The command \sethscode can be used to switch the code formatting
+% behaviour by mapping the hscode environment in the subst directive
+% to a new LaTeX environment.
+
+\newcommand{\sethscode}[1]%
+  {\expandafter\let\expandafter\hscode\csname #1\endcsname
+   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
+
+% "compatibility" mode restores the non-polycode.fmt layout.
+
+\newenvironment{compathscode}%
+  {\par\noindent
+   \advance\leftskip\mathindent
+   \hscodestyle
+   \let\\=\@normalcr
+   \(\pboxed}%
+  {\endpboxed\)%
+   \par\noindent
+   \ignorespacesafterend}
+
+\newcommand{\compaths}{\sethscode{compathscode}}
+
+% "plain" mode is the proposed default.
+% It should now work with \centering.
+% This required some changes. The old version
+% is still available for reference as oldplainhscode.
+
+\newenvironment{plainhscode}%
+  {\hsnewpar\abovedisplayskip
+   \advance\leftskip\mathindent
+   \hscodestyle
+   \let\hspre\(\let\hspost\)%
+   \pboxed}%
+  {\endpboxed%
+   \hsnewpar\belowdisplayskip
+   \ignorespacesafterend}
+
+\newenvironment{oldplainhscode}%
+  {\hsnewpar\abovedisplayskip
+   \advance\leftskip\mathindent
+   \hscodestyle
+   \let\\=\@normalcr
+   \(\pboxed}%
+  {\endpboxed\)%
+   \hsnewpar\belowdisplayskip
+   \ignorespacesafterend}
+
+% Here, we make plainhscode the default environment.
+
+\newcommand{\plainhs}{\sethscode{plainhscode}}
+\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
+\plainhs
+
+% The arrayhscode is like plain, but makes use of polytable's
+% parray environment which disallows page breaks in code blocks.
+
+\newenvironment{arrayhscode}%
+  {\hsnewpar\abovedisplayskip
+   \advance\leftskip\mathindent
+   \hscodestyle
+   \let\\=\@normalcr
+   \(\parray}%
+  {\endparray\)%
+   \hsnewpar\belowdisplayskip
+   \ignorespacesafterend}
+
+\newcommand{\arrayhs}{\sethscode{arrayhscode}}
+
+% The mathhscode environment also makes use of polytable's parray 
+% environment. It is supposed to be used only inside math mode 
+% (I used it to typeset the type rules in my thesis).
+
+\newenvironment{mathhscode}%
+  {\parray}{\endparray}
+
+\newcommand{\mathhs}{\sethscode{mathhscode}}
+
+% texths is similar to mathhs, but works in text mode.
+
+\newenvironment{texthscode}%
+  {\(\parray}{\endparray\)}
+
+\newcommand{\texths}{\sethscode{texthscode}}
+
+% The framed environment places code in a framed box.
+
+\def\codeframewidth{\arrayrulewidth}
+\RequirePackage{calc}
+
+\newenvironment{framedhscode}%
+  {\parskip=\abovedisplayskip\par\noindent
+   \hscodestyle
+   \arrayrulewidth=\codeframewidth
+   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
+   \hline\framedhslinecorrect\\{-1.5ex}%
+   \let\endoflinesave=\\
+   \let\\=\@normalcr
+   \(\pboxed}%
+  {\endpboxed\)%
+   \framedhslinecorrect\endoflinesave{.5ex}\hline
+   \endtabular
+   \parskip=\belowdisplayskip\par\noindent
+   \ignorespacesafterend}
+
+\newcommand{\framedhslinecorrect}[2]%
+  {#1[#2]}
+
+\newcommand{\framedhs}{\sethscode{framedhscode}}
+
+% The inlinehscode environment is an experimental environment
+% that can be used to typeset displayed code inline.
+
+\newenvironment{inlinehscode}%
+  {\(\def\column##1##2{}%
+   \let\>\undefined\let\<\undefined\let\\\undefined
+   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
+   \def\fromto##1##2##3{##3}%
+   \def\nextline{}}{\) }%
+
+\newcommand{\inlinehs}{\sethscode{inlinehscode}}
+
+% The joincode environment is a separate environment that
+% can be used to surround and thereby connect multiple code
+% blocks.
+
+\newenvironment{joincode}%
+  {\let\orighscode=\hscode
+   \let\origendhscode=\endhscode
+   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
+   %\let\SaveRestoreHook=\empty
+   %\let\ColumnHook=\empty
+   %\let\resethooks=\empty
+   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
+  {\origendhscode
+   \global\let\hscode=\orighscode
+   \global\let\endhscode=\origendhscode}%
+
+\makeatother
+\EndFmtInput
+%
+
+ % Without these, lhs2tex uses three-line equal signs.
+
+  % so that parameters declared with {...} are spaced like those declared with (...)
+
 %\usepackage{rotating}
@@
 
-\begin{code}
-Record System: Type :=
-  { Point: CSetoid
-  ; Location: Set
-  ; Location_eq_dec: EqDec Location eq
-  ; locations: ExhaustiveList Location
-  ; State :=  Location x Point
-  ; initial: State -> Prop
-  ; invariant: State -> Prop
-  ; invariant_initial: initial subsetof invariant
-  ; invariant_stable: forall s, Stable (invariant s)
-  ; flow: Location -> Flow Point
-  ; guard: State -> Location -> Prop
-  ; reset: Location -> Location -> Point -> Point  }.
-\end{code}
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{52}{@{}>{\hspre}c<{\hspost}@{}}%
+\column{52E}{@{}l@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\ensuremath{\mathbf{Record}}\;\Conid{System}\mathbin{:}\Conid{Type}\mathbin{:=}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\hspace{0.9mm}\{\Conid{Point}\mathbin{:}\Conid{CSetoid}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Conid{Location}\mathbin{:}\Conid{Set}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Conid{Location\char95 eq\char95 dec}\mathbin{:}\Conid{EqDec}\;\Conid{Location}\;\Varid{eq}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Varid{locations}\mathbin{:}\Conid{ExhaustiveList}\;\Conid{Location}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Conid{State}\mathbin{:=}{}\<[15]%
+\>[15]{}\Conid{Location}\;\Varid{x}\;\Conid{Point}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Varid{initial}\mathbin{:}\Conid{State}\to \Conid{Prop}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Varid{invariant}\mathbin{:}\Conid{State}\to \Conid{Prop}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Varid{invariant\char95 initial}\mathbin{:}\Varid{initial}\;\Varid{subsetof}\;\Varid{invariant}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Varid{invariant\char95 stable}\mathbin{:}\ensuremath{\forall}\;\Varid{s},\Conid{Stable}\;(\Varid{invariant}\;\Varid{s}){}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Varid{flow}\mathbin{:}\Conid{Location}\to \Conid{Flow}\;\Conid{Point}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Varid{guard}\mathbin{:}\Conid{State}\to \Conid{Location}\to \Conid{Prop}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{};\Varid{reset}\mathbin{:}\Conid{Location}\to \Conid{Location}\to \Conid{Point}\to \Conid{Point}{}\<[52]%
+\>[52]{}\mskip1.5mu\}.{}\<[52E]%
+\ColumnHook
+\end{hscode}\resethooks
 
@@
 
-%format bind_DN
-\begin{code}
-  return_DN: forall A, A -> DN A
-  bind_DN: forall A B, DN A -> (A -> DN B) -> DN B
-\end{code}
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\Varid{return\char95 DN}\mathbin{:}\ensuremath{\forall}\;\Conid{A},\Conid{A}\to \Conid{DN}\;\Conid{A}{}\<[E]%
+\\
+\>[3]{}\Varid{bind}_{\Conid{DN}}\mathbin{:}\ensuremath{\forall}\;\Conid{A}\;\Conid{B},\Conid{DN}\;\Conid{A}\to (\Conid{A}\to \Conid{DN}\;\Conid{B})\to \Conid{DN}\;\Conid{B}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 
@@
 a region corresponding to a subset of the continuous state space.
-From now on we will use a ``|concrete.|'' prefix for names like
-|State| defined in section \ref{concrete}, which now have abstract
-counterparts. |Region| is a field from a record type |Space| bundling
+From now on we will use a ``\ensuremath{\Varid{concrete}.}'' prefix for names like
+\ensuremath{\Conid{State}} defined in section \ref{concrete}, which now have abstract
+counterparts. \ensuremath{\Conid{Region}} is a field from a record type \ensuremath{\Conid{Space}} bundling
 region sets with related requirements:
 
-\begin{code}
-  Record Space: Type :=
-    { Region: Set
-    ; Region_eq_dec: EqDec Region eq
-    ; regions: ExhaustiveList Region
-    ; NoDup_regions: NoDup regions
-    ; in_region: Container Point Region
-    ; regions_cover: forall (l: Location) (p: Point),
-        invariant (l, p) -> DN { r: Region | p `elem` r }  }.
-\end{code}
-
-The |Container Point Region| type specified for |in_region| reduces to
-|Point -> Region -> Prop|. |Container| is a type class that provides
-the notation ``|x `elem` y|'', prettier than ``|in_region x
-y|''. |regions_cover| expresses that each concrete point belonging to
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{60}{@{}>{\hspre}c<{\hspost}@{}}%
+\column{60E}{@{}l@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\ensuremath{\mathbf{Record}}\;\Conid{Space}\mathbin{:}\Conid{Type}\mathbin{:=}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\hspace{0.9mm}\{\Conid{Region}\mathbin{:}\Conid{Set}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{};\Conid{Region\char95 eq\char95 dec}\mathbin{:}\Conid{EqDec}\;\Conid{Region}\;\Varid{eq}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{};\Varid{regions}\mathbin{:}\Conid{ExhaustiveList}\;\Conid{Region}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{};\Conid{NoDup\char95 regions}\mathbin{:}\Conid{NoDup}\;\Varid{regions}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{};\Varid{in\char95 region}\mathbin{:}\Conid{Container}\;\Conid{Point}\;\Conid{Region}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{};\Varid{regions\char95 cover}\mathbin{:}\ensuremath{\forall}\;(\Varid{l}\mathbin{:}\Conid{Location})\;(\Varid{p}\mathbin{:}\Conid{Point}),{}\<[E]%
+\\
+\>[5]{}\hsindent{4}{}\<[9]%
+\>[9]{}\Varid{invariant}\;(\Varid{l},\Varid{p})\to \Conid{DN}\;\hspace{0.9mm}\{\Varid{r}\mathbin{:}\Conid{Region}\mid \Varid{p}\in \Varid{r}\mskip1.5mu\}{}\<[60]%
+\>[60]{}\mskip1.5mu\}.{}\<[60E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+The \ensuremath{\Conid{Container}\;\Conid{Point}\;\Conid{Region}} type specified for \ensuremath{\Varid{in\char95 region}} reduces to
+\ensuremath{\Conid{Point}\to \Conid{Region}\to \Conid{Prop}}. \ensuremath{\Conid{Container}} is a type class that provides
+the notation ``\ensuremath{\Varid{x}\in \Varid{y}}'', prettier than ``\ensuremath{\Varid{in\char95 region}\;\Varid{x}\;\Varid{y}}''.
+\ensuremath{\Varid{regions\char95 cover}} expresses that each concrete point belonging to
 a valid state must be represented by a region---a crucial ingredient
@@
 
-It is \emph{necessary} because |regions_cover| boils down to a
+It is \emph{necessary} because \ensuremath{\Varid{regions\char95 cover}} boils down to a
 (partial) function that, given a concrete point, must select an
@@
 point lies. As we saw in section \ref{dn}, that kind of decidability
-is only available inside |DN| unless all region borders have
+is only available inside \ensuremath{\Conid{DN}} unless all region borders have
 nontrivial overlap, which as we will see later is undesireable.
@@
 Fortunately, the double negation is also \emph{sufficient}, because we
-will ultimately only use |regions_cover| in a proof of of |... -> ~
-concrete.reachable s| (for some universally quantified variable s),
+will ultimately only use \ensuremath{\Varid{regions\char95 cover}} in a proof of of \ensuremath{\mathbin{...}\to \mathord{\sim}\Varid{concrete}.\Varid{reachable}\;\Varid{s}} (for some universally quantified variable s),
 which, due to its head type being a negation, is stable, and can
-therefore be proved in and then extracted from |DN|. Hence, we only
-need |regions_cover|'s result in |DN|.
+therefore be proved in and then extracted from \ensuremath{\Conid{DN}}. Hence, we only
+need \ensuremath{\Varid{regions\char95 cover}}'s result in \ensuremath{\Conid{DN}}.
 
@@
 
-We define |underestimation P| to be either a proof of
-|P|, or not. In Coq, either of the following will do:
-\begin{code}
-Definition underestimation (P: Prop): Set := option P.
-Definition underestimation (P: Prop): Set := { b: bool | b = true -> P }.
-\end{code}
-The latter tends to work better with the |Program| family of commands
+We define \ensuremath{\Varid{underestimation}\;\Conid{P}} to be either a proof of
+\ensuremath{\Conid{P}}, or not. In Coq, either of the following will do:
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\ensuremath{\mathbf{Definition}}\;\Varid{underestimation}\;(\Conid{P}\mathbin{:}\Conid{Prop})\mathbin{:}\Conid{Set}\mathbin{:=}\Varid{option}\;\Conid{P}.{}\<[E]%
+\\
+\>[B]{}\ensuremath{\mathbf{Definition}}\;\Varid{underestimation}\;(\Conid{P}\mathbin{:}\Conid{Prop})\mathbin{:}\Conid{Set}\mathbin{:=}\hspace{0.9mm}\{\Varid{b}\mathbin{:}\Varid{bool}\mid \Varid{b}\mathrel{=}\Varid{true}\to \Conid{P}\mskip1.5mu\}.{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+The latter tends to work better with the \ensuremath{\ensuremath{\mathbf{Program}}} family of commands
 \cite{sozeau} which have special support for dependent pairs. Using
-|Program|, an underestimation of the second variety may be
+\ensuremath{\ensuremath{\mathbf{Program}}}, an underestimation of the second variety may be
 provisionally defined strictly as a bool, and then separately proved
 to be a valid underestimation in a proof obligation generated by
-|Program|. The second form also nicely illustrates why we call this an
-underestimation: it may be |false| even when |P| holds. We can now
+\ensuremath{\ensuremath{\mathbf{Program}}}. The second form also nicely illustrates why we call this an
+underestimation: it may be \ensuremath{\Varid{false}} even when \ensuremath{\Conid{P}} holds. We can now
 describe the functionality of our program by saying that it
 underestimates hybrid system safety, yielding a term of type
-|underestimation Safe|, where |Safe| is a proposition expressing
+\ensuremath{\Varid{underestimation}\;\Conid{Safe}}, where \ensuremath{\Conid{Safe}} is a proposition expressing
 safety of a hybrid system.
@@
 because they can be trivially ``proved'' by taking
-|false|/|None|. Hence, the value of our program is not witnessed by
+\ensuremath{\Varid{false}}/\ensuremath{\Conid{None}}. Hence, the value of our program is not witnessed by
 the mere fact that it manages to produce terms of type
-|underestimation Safe|, but rather by the fact that when run, it
-actually manages to returns |true|/|Just P| for the hybrid system we
+\ensuremath{\Varid{underestimation}\;\Conid{Safe}}, but rather by the fact that when run, it
+actually manages to returns \ensuremath{\Varid{true}}/\ensuremath{\Conid{Just}\;\Conid{P}} for the hybrid system we
 are interested in (e.g.\ the thermostat). It is for this reason that
@@
 The opposite of an underestimation is an overestimation:
-\begin{code}
-Definition overestimation (P: Prop): Set := { b: bool | b = false -> ~ P }.
-\end{code}
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\ensuremath{\mathbf{Definition}}\;\Varid{overestimation}\;(\Conid{P}\mathbin{:}\Conid{Prop})\mathbin{:}\Conid{Set}\mathbin{:=}\hspace{0.9mm}\{\Varid{b}\mathbin{:}\Varid{bool}\mid \Varid{b}\mathrel{=}\Varid{false}\to \mathord{\sim}\Conid{P}\mskip1.5mu\}.{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 
@@
 
-%format CRle_unapplied = <= "_{" CR "}"
-%format CRle a b = a <= "_{" CR "}" b
-%format Qpos = "\mathbb{Q}^{+}"
-%format overestimate_CRle = overestimate "_{" CRle_unapplied "}"
-
-%format epsilon = "\epsilon "
-
-\begin{code}
-overestimate_CRle (epsilon: Qpos): forall x y: CR, overestimation (CRle x y)
-\end{code}
-As discussed earlier, |CRle_unapplied| is not
-decidable. |overestimate_CRle| merely makes a ``best effort'' to prove
-|~ (CRle x y)| using |epsilon|-approximations. A smaller |epsilon|
-will result in fewer spurious |true| results.
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\Varid{overestimate}_{\leq _{\Conid{CR}}}\;(\epsilon \mathbin{:}\mathbb{Q}^{+})\mathbin{:}\ensuremath{\forall}\;\Varid{x}\;\Varid{y}\mathbin{:}\Conid{CR},\Varid{overestimation}\;(\Varid{x}\leq _{\Conid{CR}}\Varid{y}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+As discussed earlier, \ensuremath{\leq _{\Conid{CR}}} is not
+decidable. \ensuremath{\Varid{overestimate}_{\leq _{\Conid{CR}}}} merely makes a ``best effort'' to prove
+\ensuremath{\mathord{\sim}(\Varid{x}\leq _{\Conid{CR}}\Varid{y})} using \ensuremath{\epsilon }-approximations. A smaller \ensuremath{\epsilon }
+will result in fewer spurious \ensuremath{\Varid{true}} results.
 
@@
 \label{abs.reach}
-Once we have a satisfactory abstract |Space|, our goal is to construct
+Once we have a satisfactory abstract \ensuremath{\Conid{Space}}, our goal is to construct
 an overestimatable notion of abstract reachability implied by concrete
@@
 
-\begin{code}
-  Definition shared_cover
-    `{Container concrete.State C} `{Container abstract.State D}
-      (cs: C) (ss: D): Prop :=
-        forall s: concrete.State, In s cs -> DN (exists r: abstract.State, In s r /\ In r ss).
-\end{code}
-
-The details of the |Container| type class are of no interest to us
-right now; one may think of |C| and |D| as the types of lists over
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\ensuremath{\mathbf{Definition}}\;\Varid{shared\char95 cover}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}`\hspace{0.9mm}\{\Conid{Container}\;\Varid{concrete}.\Conid{State}\;\Conid{C}\mskip1.5mu\}`\hspace{0.9mm}\{\Conid{Container}\;\Varid{abstract}.\Conid{State}\;\Conid{D}\mskip1.5mu\}{}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{}(\Varid{cs}\mathbin{:}\Conid{C})\;(\Varid{ss}\mathbin{:}\Conid{D})\mathbin{:}\Conid{Prop}\mathbin{:=}{}\<[E]%
+\\
+\>[7]{}\hsindent{2}{}\<[9]%
+\>[9]{}\ensuremath{\forall}\;\Varid{s}\mathbin{:}\Varid{concrete}.\Conid{State},\Conid{In}\;\Varid{s}\;\Varid{cs}\to \Conid{DN}\;(\ensuremath{\exists }\;\Varid{r}\mathbin{:}\Varid{abstract}.\Conid{State},\Conid{In}\;\Varid{s}\;\Varid{r}\ensuremath{\wedge }\Conid{In}\;\Varid{r}\;\Varid{ss}).{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+The details of the \ensuremath{\Conid{Container}} type class are of no interest to us
+right now; one may think of \ensuremath{\Conid{C}} and \ensuremath{\Conid{D}} as the types of lists over
 concrete states, respectively over abstract states.  What matters is
@@
 We state this in
-terms of |Container|s so that the definition applies to any types
+terms of \ensuremath{\Conid{Container}}s so that the definition applies to any types
 which have a reasonable notion of containment.
@@
 We now specify what the type of $\overcontrans$ should be.
-\begin{code}
-over_cont_trans: forall s: abstract.State,
-  { p: list abstract.State | NoDup p /\ shared_cover
-    (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p }
-\end{code}
-So, |over_cont_trans s| should produce a list of abstract states |p|
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\Varid{over\char95 cont\char95 trans}\mathbin{:}\ensuremath{\forall}\;\Varid{s}\mathbin{:}\Varid{abstract}.\Conid{State},{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\hspace{0.9mm}\{\Varid{p}\mathbin{:}\Varid{list}\;\Varid{abstract}.\Conid{State}\mid \Conid{NoDup}\;\Varid{p}\ensuremath{\wedge }\Varid{shared\char95 cover}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}(\Varid{concrete}.\Varid{invariant}\mathbin{`\Varid{inter}`}(\Varid{overlap}\;\Varid{s}\mathbin{`\Varid{compose}`}\Varid{utilflip}\;(\Varid{concrete}.\Varid{cont\char95 trans})))\;\Varid{p}\mskip1.5mu\}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+So, \ensuremath{\Varid{over\char95 cont\char95 trans}\;\Varid{s}} should produce a list of abstract states \ensuremath{\Varid{p}}
 without duplicates,
@@
 %reachability set), 
-such that |p| is a shared cover of the collection
-of concrete states |c| that satisfy the invariant and whose set of
-origins under |concrete.cont_trans| have an overlap with |s|. In more
-mathematical terms: |p| should form a shared cover of $\{ c \in \State
+such that \ensuremath{\Varid{p}} is a shared cover of the collection
+of concrete states \ensuremath{\Varid{c}} that satisfy the invariant and whose set of
+origins under \ensuremath{\Varid{concrete}.\Varid{cont\char95 trans}} have an overlap with \ensuremath{\Varid{s}}. In more
+mathematical terms: \ensuremath{\Varid{p}} should form a shared cover of $\{ c \in \State
 \mid \Inv(c) \wedge s\cap \{c' \mid c' \contrans c\} \neq \emptyset\}$.
-We similarly specify |over_disc_trans| as an overestimator for
-|concrete_disc_trans| and |over_initial| as an overestimator of
-|concrete.initial|.
-
-We now consider the properties we require for |abstract.reachable|. The following basic specification we had in mind:
-
-\begin{code}
-Definition reachable_respect: Prop :=
-  forall (s: concrete.State), concrete.reachable s ->
-    forall (s': abstract.State), s `elem` s' -> abstract.reachable s'.
-\end{code}
+We similarly specify \ensuremath{\Varid{over\char95 disc\char95 trans}} as an overestimator for
+\ensuremath{\Varid{concrete\char95 disc\char95 trans}} and \ensuremath{\Varid{over\char95 initial}} as an overestimator of
+\ensuremath{\Varid{concrete}.\Varid{initial}}.
+
+We now consider the properties we require for \ensuremath{\Varid{abstract}.\Varid{reachable}}. The following basic specification we had in mind:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\ensuremath{\mathbf{Definition}}\;\Varid{reachable\char95 respect}\mathbin{:}\Conid{Prop}\mathbin{:=}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\ensuremath{\forall}\;(\Varid{s}\mathbin{:}\Varid{concrete}.\Conid{State}),\Varid{concrete}.\Varid{reachable}\;\Varid{s}\to {}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\ensuremath{\forall}\;(\Varid{s'}\mathbin{:}\Varid{abstract}.\Conid{State}),\Varid{s}\in \Varid{s'}\to \Varid{abstract}.\Varid{reachable}\;\Varid{s'}.{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 Which we chose because it implies
-\begin{code}
-forall (s: concrete.State)
-  (exists s': abstract.State, s `elem` s' /\ ~ abstract.reachable s') ->
-    ~ concrete.reachable s,
-\end{code}
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\ensuremath{\forall}\;(\Varid{s}\mathbin{:}\Varid{concrete}.\Conid{State})\;{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}(\ensuremath{\exists }\;\Varid{s'}\mathbin{:}\Varid{abstract}.\Conid{State},\Varid{s}\in \Varid{s'}\ensuremath{\wedge }\mathord{\sim}\Varid{abstract}.\Varid{reachable}\;\Varid{s'})\to {}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\mathord{\sim}\Varid{concrete}.\Varid{reachable}\;\Varid{s},{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 This expresses that to conclude unreachability of a concrete state,
@@
 abstract states containing the concrete state. Hence, what we really
-want is an |abstract.reachable| satisfying:
-\begin{code}
-forall s: concrete.State,
-  (forall s': abstract.State, s `elem` s' -> ~ abstract.reachable s') ->
-  ~ concrete.reachable s.
-\end{code}
+want is an \ensuremath{\Varid{abstract}.\Varid{reachable}} satisfying:
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\ensuremath{\forall}\;\Varid{s}\mathbin{:}\Varid{concrete}.\Conid{State},{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}(\ensuremath{\forall}\;\Varid{s'}\mathbin{:}\Varid{abstract}.\Conid{State},\Varid{s}\in \Varid{s'}\to \mathord{\sim}\Varid{abstract}.\Varid{reachable}\;\Varid{s'})\to {}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\mathord{\sim}\Varid{concrete}.\Varid{reachable}\;\Varid{s}.{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 \weg{
-This property follows from the following new definition of |reachable_respect| we will use:
-\begin{code}
-Definition reachable_respect: Prop :=
-  shared_cover concrete.reachable abstract.reachable.
-\end{code}
+This property follows from the following new definition of \ensuremath{\Varid{reachable\char95 respect}} we will use:
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\ensuremath{\mathbf{Definition}}\;\Varid{reachable\char95 respect}\mathbin{:}\Conid{Prop}\mathbin{:=}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{shared\char95 cover}\;\Varid{concrete}.\Varid{reachable}\;\Varid{abstract}.\Varid{reachable}.{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 }
@@
 In the next section we show that thanks to decidability of our
-transition and initiality overstimators, |abstract.reachable| is
+transition and initiality overstimators, \ensuremath{\Varid{abstract}.\Varid{reachable}} is
 decidable. But first we show how a
-decision procedure for |abstract.reachable| lets us underestimate
+decision procedure for \ensuremath{\Varid{abstract}.\Varid{reachable}} lets us underestimate
 hybrid system safety, and in particular, lets us obtain a proof of
-thermostat safety. So suppose we have |reachable_dec: decider abstract.reachable|
+thermostat safety. So suppose we have \ensuremath{\Varid{reachable\char95 dec}\mathbin{:}\Varid{decider}\;\Varid{abstract}.\Varid{reachable}}
 \weg{
 And suppose we are given the following specification of unsafe concrete states, covered by a finite list of abstract states:
-\begin{code}
-  Variables
-    (unsafe: concrete.State -> Prop)
-    (astates: list abstract.State)
-    (astates_cover_unsafe: forall s, unsafe s -> forall r, s `elem` r -> r `elem` astates).
-\end{code}
-
-Then, using |reachable_dec| and |unreachable_respect|, we can easily define
-\begin{code}
-  Definition over_unsafe_reachable: overestimation (overlap unsafe concrete.reachable).
-\end{code}
-Taking |unsafe := thermo_unsafe| and a suitable abstract cover, we obtain
-\begin{code}
-  Definition over_thermo_unsafe_reachable:
-    overestimation (overlap thermo_unsafe concrete.reachable).
-\end{code}
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\ensuremath{\mathbf{Variables}}\;{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}(\Varid{unsafe}\mathbin{:}\Varid{concrete}.\Conid{State}\to \Conid{Prop})\;{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}(\Varid{astates}\mathbin{:}\Varid{list}\;\Varid{abstract}.\Conid{State})\;{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}(\Varid{astates\char95 cover\char95 unsafe}\mathbin{:}\ensuremath{\forall}\;\Varid{s},\Varid{unsafe}\;\Varid{s}\to \ensuremath{\forall}\;\Varid{r},\Varid{s}\in \Varid{r}\to \Varid{r}\in \Varid{astates}).{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Then, using \ensuremath{\Varid{reachable\char95 dec}} and \ensuremath{\Varid{unreachable\char95 respect}}, we can easily define
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\ensuremath{\mathbf{Definition}}\;\Varid{over\char95 unsafe\char95 reachable}\mathbin{:}\Varid{overestimation}\;(\Varid{overlap}\;\Varid{unsafe}\;\Varid{concrete}.\Varid{reachable}).{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+Taking \ensuremath{\Varid{unsafe}\mathbin{:=}\Varid{thermo\char95 unsafe}} and a suitable abstract cover, we obtain
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\ensuremath{\mathbf{Definition}}\;\Varid{over\char95 thermo\char95 unsafe\char95 reachable}\mathbin{:}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\Varid{overestimation}\;(\Varid{overlap}\;\Varid{thermo\char95 unsafe}\;\Varid{concrete}.\Varid{reachable}).{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 }
-Recall that |ThermoSafe| was defined as |thermo_unsafe `subsetof` concrete.unreachable| in section \ref{concrete}.
-Since we trivially have |~ overlap unsafe concrete.reachable -> ThermoSafe|, we also have:
-\begin{code}
-Definition under_thermo_unsafe_unreachable: underestimation ThermoSafe.
-\end{code}
+Recall that \ensuremath{\Conid{ThermoSafe}} was defined as \ensuremath{\Varid{thermo\char95 unsafe}\mathbin{`\Varid{subsetof}`}\Varid{concrete}.\Varid{unreachable}} in section \ref{concrete}.
+Since we trivially have \ensuremath{\mathord{\sim}\Varid{overlap}\;\Varid{unsafe}\;\Varid{concrete}.\Varid{reachable}\to \Conid{ThermoSafe}}, we also have:
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\ensuremath{\mathbf{Definition}}\;\Varid{under\char95 thermo\char95 unsafe\char95 unreachable}\mathbin{:}\Varid{underestimation}\;\Conid{ThermoSafe}.{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 Finally, we \emph{run} the underestimation:
-\begin{code}
-Theorem: ThermoSafe.
-Proof.
-  apply (underestimation_true under_thermo_unsafe_unreachable).
-  vm_compute.  reflexivity.
-Qed.
-\end{code}
-|underestimation_true| is a tiny utility of type |forall P (o: underestimation P), o = true -> P|, whose application in the proof reduces the goal to
-\begin{code}under_thermo_unsafe_unreachable = true.\end{code} The |vm_compute| tactic invocation then forces evaluation of the left hand side, which will in turn evaluate |over_thermo_unsafe_reachable|, which will evaluate |reachable_dec|, which will evaluate the overestimators of the continuous and discrete transitions.
-This process, which takes about 35 seconds on a modern desktop machine, eventually reduces |under_thermo_unsafe_unreachable| to |true|, leaving |true = true|, proved by |reflexivity|.
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\ensuremath{\mathbf{Theorem}}\mathbin{:}\Conid{ThermoSafe}.{}\<[E]%
+\\
+\>[B]{}\ensuremath{\mathbf{Proof}}.{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{apply}\;(\Varid{underestimation\char95 true}\;\Varid{under\char95 thermo\char95 unsafe\char95 unreachable}).{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{vm\char95 compute}.{}\<[16]%
+\>[16]{}\Varid{reflexivity}.{}\<[E]%
+\\
+\>[B]{}\ensuremath{\mathbf{Qed}}.{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+\ensuremath{\Varid{underestimation\char95 true}} is a tiny utility of type \ensuremath{\ensuremath{\forall}\;\Conid{P}\;(\Varid{o}\mathbin{:}\Varid{underestimation}\;\Conid{P}),\Varid{o}\mathrel{=}\Varid{true}\to \Conid{P}}, whose application in the proof reduces the goal to
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\Varid{under\char95 thermo\char95 unsafe\char95 unreachable}\mathrel{=}\Varid{true}.{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+ The \ensuremath{\Varid{vm\char95 compute}} tactic invocation then forces evaluation of the left hand side, which will in turn evaluate \ensuremath{\Varid{over\char95 thermo\char95 unsafe\char95 reachable}}, which will evaluate \ensuremath{\Varid{reachable\char95 dec}}, which will evaluate the overestimators of the continuous and discrete transitions.
+This process, which takes about 35 seconds on a modern desktop machine, eventually reduces \ensuremath{\Varid{under\char95 thermo\char95 unsafe\char95 unreachable}} to \ensuremath{\Varid{true}}, leaving \ensuremath{\Varid{true}\mathrel{=}\Varid{true}}, proved by \ensuremath{\Varid{reflexivity}}.
 
 We can now also clearly see what happens when the abstraction method ``fails'' due to poor region selection, overly simplistic transition/initiality overestimators, or plain unsafety of the system.
-In all these cases, |vm_compute| reduces |under_thermo_unsafe_unreachable| to |false|, and the subsequent |reflexivity| invocation will fail.
+In all these cases, \ensuremath{\Varid{vm\char95 compute}} reduces \ensuremath{\Varid{under\char95 thermo\char95 unsafe\char95 unreachable}} to \ensuremath{\Varid{false}}, and the subsequent \ensuremath{\Varid{reflexivity}} invocation will fail.
 
 This concludes the high level story of our development. What remains
-are the implementation of |reachable_dec| in terms of the decidable
+are the implementation of \ensuremath{\Varid{reachable\char95 dec}} in terms of the decidable
 overestimators for abstract initiality and continuous and discrete
@@
 algorithm, that we don't detail here. The overestimator for continuous
-transitions, |over_cont_trans| will be detailed in the next section
+transitions, \ensuremath{\Varid{over\char95 cont\char95 trans}} will be detailed in the next section
 for the thermostat case.
@@
 
-We now discuss the implementation of  |over_cont_trans|:
-Given two regions |r_src| and |r_dst|, 
+We now discuss the implementation of  \ensuremath{\Varid{over\char95 cont\char95 trans}}:
+Given two regions \ensuremath{\Varid{r\char95 src}} and \ensuremath{\Varid{r\char95 dst}}, 
 %we want to
@@
 if we can determine that there
-are no points in |r_src| which the flow function maps to points in
-|r_dst|, then there need be no abstract continuous transition between |r_src|
-and |r_dst|. Clearly, this is impossible to
+are no points in \ensuremath{\Varid{r\char95 src}} which the flow function maps to points in
+\ensuremath{\Varid{r\char95 dst}}, then there need be no abstract continuous transition between \ensuremath{\Varid{r\char95 src}}
+and \ensuremath{\Varid{r\char95 dst}}. Clearly, this is impossible to
 meaningfully overestimate for a general flow function and general
@@
 
-A flow function |f| on |CR| is {\em range invertible\/} if
-\begin{code}
-  exists (range_inverse: OpenRange -> OpenRange -> OpenRange),
-    forall (a: OpenRange) (p: CR), p `elem` a ->
-    forall (b: OpenRange) (d: Duration), f p d `elem` b -> d `elem` range_inverse a b
-\end{code}
-Here, |OpenRange| represents potentially unbounded intervals in $\IR$
+A flow function \ensuremath{\Varid{f}} on \ensuremath{\Conid{CR}} is {\em range invertible\/} if
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\ensuremath{\exists }\;(\Varid{range\char95 inverse}\mathbin{:}\Conid{OpenRange}\to \Conid{OpenRange}\to \Conid{OpenRange}),{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\ensuremath{\forall}\;(\Varid{a}\mathbin{:}\Conid{OpenRange})\;(\Varid{p}\mathbin{:}\Conid{CR}),\Varid{p}\in \Varid{a}\to {}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\ensuremath{\forall}\;(\Varid{b}\mathbin{:}\Conid{OpenRange})\;(\Varid{d}\mathbin{:}\Conid{Duration}),\Varid{f}\;\Varid{p}\;\Varid{d}\in \Varid{b}\to \Varid{d}\in \Varid{range\char95 inverse}\;\Varid{a}\;\Varid{b}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+Here, \ensuremath{\Conid{OpenRange}} represents potentially unbounded intervals in $\IR$
 (with bounds closed if present. In other words, if $\phi : \IR^2
@@
 show how to proceed with our overestimation of existence of points in
-|r_src| which the flow function map to points in |r_dst|. 
+\ensuremath{\Varid{r\char95 src}} which the flow function map to points in \ensuremath{\Varid{r\char95 dst}}. 
 Regions in the abstract space for our thermostat are basically pairs
-of regions in the composite spaces, so |r_src| and |r_dst| can be
-written as |(r_src_temp, r_src_clock)| and |(r_dst_temp,
-r_dst_clock)|, respectively, where each of these four components are
+of regions in the composite spaces, so \ensuremath{\Varid{r\char95 src}} and \ensuremath{\Varid{r\char95 dst}} can be
+written as \ensuremath{(\Varid{r\char95 src\char95 temp},\Varid{r\char95 src\char95 clock})} and \ensuremath{(\Varid{r\char95 dst\char95 temp},\Varid{r\char95 dst\char95 clock})}, respectively, where each of these four components are
 intervals.
 
-We now simply use an |OpenRange| overlap overestimator of type
-\begin{code}
-Qpos -> forall a b: OpenRange, overestimation (overlap a b)
-\end{code}
-(defined in terms of things like |overestimate_CRle| shown in section \ref{estimation}) to overestimate whether the following three intervals overlap:
+We now simply use an \ensuremath{\Conid{OpenRange}} overlap overestimator of type
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbb{Q}^{+}\to \ensuremath{\forall}\;\Varid{a}\;\Varid{b}\mathbin{:}\Conid{OpenRange},\Varid{overestimation}\;(\Varid{overlap}\;\Varid{a}\;\Varid{b}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+(defined in terms of things like \ensuremath{\Varid{overestimate}_{\leq _{\Conid{CR}}}} shown in section \ref{estimation}) to overestimate whether the following three intervals overlap:
 \begin{enumerate}
-\item |[0, inf]|
-\item |range_inverse temp_flow r_src_temp r_dst_temp|
-\item |range_inverse clock_flow r_src_clock r_dst_clock|
+\item \ensuremath{[\mskip1.5mu \mathrm{0},\Varid{inf}\mskip1.5mu]}
+\item \ensuremath{\Varid{range\char95 inverse}\;\Varid{temp\char95 flow}\;\Varid{r\char95 src\char95 temp}\;\Varid{r\char95 dst\char95 temp}}
+\item \ensuremath{\Varid{range\char95 inverse}\;\Varid{clock\char95 flow}\;\Varid{r\char95 src\char95 clock}\;\Varid{r\char95 dst\char95 clock}}
 \end{enumerate}
 For a visual explanation, one may consult the left drawing in Figure
-\ref{fig:abstraction} and view |r_src_clock| as $[a_1,b_1]$,
-|r_dst_clock| as $[c_1,d_1]$ etc.  Overlap of 2 and 3 is equivalent to
-existence of a point in |r_src| from which one can flow to a point in
-|r_dst|. After all, if these two range inverses overlap, then there is
-a duration |d| that takes a certain temperature value in |r_src_temp|
-to a value in |r_dst_temp| and also takes a certain clock value in
-|r_src_clock| to a value in |r_dst_clock|. If 2 and 3 do \emph{not}
+\ref{fig:abstraction} and view \ensuremath{\Varid{r\char95 src\char95 clock}} as $[a_1,b_1]$,
+\ensuremath{\Varid{r\char95 dst\char95 clock}} as $[c_1,d_1]$ etc.  Overlap of 2 and 3 is equivalent to
+existence of a point in \ensuremath{\Varid{r\char95 src}} from which one can flow to a point in
+\ensuremath{\Varid{r\char95 dst}}. After all, if these two range inverses overlap, then there is
+a duration \ensuremath{\Varid{d}} that takes a certain temperature value in \ensuremath{\Varid{r\char95 src\char95 temp}}
+to a value in \ensuremath{\Varid{r\char95 dst\char95 temp}} and also takes a certain clock value in
+\ensuremath{\Varid{r\char95 src\char95 clock}} to a value in \ensuremath{\Varid{r\char95 dst\char95 clock}}. If 2 and 3 do \emph{not}
 overlap, then either it takes so long for the temperature to flow from
-|r_src_temp| to |r_dst_temp| that any clock value in |r_src_clock|
-would ``overshoot'' |r_dst_clock|, or vice versa. Finally, if 1 does
+\ensuremath{\Varid{r\char95 src\char95 temp}} to \ensuremath{\Varid{r\char95 dst\char95 temp}} that any clock value in \ensuremath{\Varid{r\char95 src\char95 clock}}
+would ``overshoot'' \ensuremath{\Varid{r\char95 dst\char95 clock}}, or vice versa. Finally, if 1 does
 not overlap with 2 and 3, then apparently one could only flow backward
@@
 is a necessary condition for existence of concrete flow from points in
-|r_src| to points in |r_dst|, and so our |abstract.cont_trans|
+\ensuremath{\Varid{r\char95 src}} to points in \ensuremath{\Varid{r\char95 dst}}, and so our \ensuremath{\Varid{abstract}.\Varid{cont\char95 trans}}
 overestimator may justifiably return ``false'' when the overlap

### 2010-03-24 ITP10/hybrid.tex
Merge introduction and motivation sections in ITP paper. Saves one page.
Ignore-this: 6b8537a38e3e5651475431178e8d9265

darcs-hash:20100324175748-bab43-835811384164ce8ce3f4253713f8f25f24ab492c.gz

@@
 (2006).  The development includes: a formalisation of the structure of
-hybrid systems; a set of utilities for the construction of an abstract
+hybrid systems; a framework for the construction of an abstract
 system (consisting of decidable ``overestimators'' of abstract
 transitions and initiality) faithfully representing a (concrete)
-hybrid system; a translation of abstract systems to graphs enabling
+hybrid system; a translation of abstract systems to graphs, enabling
 the decision of abstract state reachability using a certified graph
 reachability algorithm; a proof of the safety of an example hybrid
-system generated using this tool stack.  The development critically
-relies on the computable real number implementation part of the CoRN
-library of formalised constructive mathematics in Coq.  The obtained
-safety proofs are fully certified and do not use floating point real
-numbers, but arbitrary precise approximations of computable real
-numbers. It also uses a nice interplay between constructive and
+system generated using this tool stack. 
+To produce fully certified safety proofs without relying on floating point computations, the development critically relies on the computable real number implementation part of the CoRN library of constructive mathematics formalised in Coq.
+The development also features a nice interplay between constructive and 
 classical logic via the double negation monad.
@@
 
-%\setcounter{tocdepth}{2}
-
-%\tableofcontents
-%\todo{How do I get rid of that silly author index at the top of the TOC?!}
-
 %\nocite{*}
@@
 In \cite{alur}, Alur et al.\ present an automated method for hybrid system
-safety verification based on the construction of an \emph{abstract}
-hybrid system (essentially a finite automaton) corresponding to the
-hybrid system of interest. The abstract system is constructed such
-that traces in the original system are represented in the abstract
-system. Consequently, one can draw conclusions about reachability of
-states in the concrete system from analysis of state reachability in
-the abstract system. Since the abstract system is an entirely finite
-discrete object (unlike the concrete system), reachability can simply
-be computed using any graph reachability algorithm. Thus, the
-abstraction method brings the safety verification problem from a
-continuous and infinite domain into a discrete and finite domain,
-where the problem can be dealt with using standard graph algorithms.
-%where it is amenable to what amounts to brute force.
+safety verification in which one derives from the hybrid system of interest an \emph{abstract}
+hybrid system, which is essentially a finite automaton whose traces are sufficiently representative of traces in the original system that unreachability in the abstract system (which can be decided using a standard graph algorithm) implies unreachability in the concrete system (which, being governed by continuous behaviors, cannot be decided so readily).
+Thus, the abstraction method brings the safety verification problem from a continuous and infinite domain into a discrete and finite domain, where it can be dealt with using standard graph algorithms.
 
 The prototype implementation described in \cite{alur} was developed in
-a conventional programming language using libraries that use ordinary
-floating point arithmetic. The potential for bugs and floating point
-artifacts inherent in this approach limits the confidence one can
-justifiably have in safety judgements made by such an
-implementation. In this paper, we describe a re-implementation of the
-basic technique of Alur et al.\ in Coq, a proof assistant based on a
-rich type theory that also functions as a programming language,
-letting one develop ``certified'' programs: programs whose correctness
-is proved formally within the system. In our case, the program will be
-one that produces machine-checked proofs of hybrid system safety,
-obtained using the abstraction method.
-
-Our development relies crucially on computations involving real
-numbers, for which we use the computable real number implementation
-developed by Russell O'Connor \cite{oconnor}, and part of the CoRN
-library of formalised constructive mathematics \cite{corn}. Indeed,
-this development showcases its use in a concrete and practical
-application. Subtleties regarding the nature of these computable reals
-are reflected in this development in several ways, which we will
-discuss in some detail.
-
-\section{Motivation}
-The motivation for this work is threefold. 
-First, we want to avoid the use of floating point number computation,
-because rounding errors may affect the correctness of the safety
-verification that has been performed. Instead, we want to use exact
-real arithmetic, which deals with infinitely precise real numbers and
-more importantly, infinitely precise algorithms for computing with
-real numbers. This means that, e.g.\ an algorithm for the function
-$f(x) = e^x$ can produce arbitrary many output precision (``digits'')
-by asking for more input precision. So, using exact real number
-algorithms, we can decide $\cos(\sin(e^2))<\cos(\sin(\frac{1}{e^2}))$
-by computing better and better approximations of the value of both
-expressions and observing at a certain point that the inequality
-holds. (Note that with this method we will never be able to determine
-in finite time that two such expressions denote the same number; for
-that we need a symbolic proof.) In Nijmegen we have developed a
-verified library of exact real arithmetic, formalised in the proof
-assistant Coq, as part of CoRN \cite{corn,oconnor}, our Constructive
-Coq Repository at Nijmegen. The algorithms (for $e^x$, $\cos$, $\sin$
-etc.) have been written in the programming language of Coq and have
-been proved correct using its logic. Coq is sufficiently fast that the
-algorithms can also be executed in reasonable time within the system,
-as can be witnessed by our online certified calculator
-\url{http://prover.cs.ru.nl/calc.html}. So, from this perspective,
-this work can also be seen as a case study in the use of this formal
-repository of exact real arithmetic: Can we use its computational
-capacities to verify safety of hybrid systems? Can we use exact real
-arithmetic to model the environment and verify the correctness of the
-controller?
-
-A second related point is that hybrid systems include continuous
-values. Some of these are controlled by the system itself, like a
-clock that can be reset, a valve that can be opened or a gas handle
-that can be opened. Others represent characteristics of the
-environment and can only be observed, like speed, pressure and
-temperature. Due to the fact that we deal with physical systems,
-measurements are always imprecise, and setting a valve in a certain
-position %may on a continuous scale 
-is also imprecise. 
-%Of course, we don't know exactly how imprecise, and we may assume
-%that better measuring instruments may arise
-This implies that when a controller makes a case distinction on the
-basis of $T>20$ or $T\leq 20$ -- e.g.\ if the temperature is larger
-that $20$, turn the heater off, if it is $20$ or below, keep it on --
-is always only ``approximately'' effective: The actual heater may
-remain on until just above $20$ degrees, because of a measuring
-imprecision. In a verification, one should actually take these things
-into account and we believe that constructive analysis provides the
-right way to model and verify hybrid systems. In constructive
-analysis, we do not have (for $x,y\in\IR$) $x>y \vee x\leq y$, so we
-cannot define an algorithm on the reals by making such a case
-distinction. On the other hand, we {\em do\/} have, for all
-$\epsilon>0$, $x+\epsilon>y \vee x-\epsilon< y$, which means that we
-can get an $\epsilon$-close enough approximation to a decision between
-$x>y$ and $x\leq y$.  So, the present work can also be seen as a
-confirmation of our belief that constructive analysis is the
-proper setting to model and verify hybrid systems.
-
-As a third point, the method to verify safety of hybrid systems of
-\cite{alur} uses so called ``predicate abstraction''. We will detail
-this approach in the next section; for now it suffices to say that it
-involves mapping the hybrid system to another, finite transition
-system, which is then verified. However, this abstraction itself is
-not verified. In this formalisation we want to also verify the
-abstraction itself: show that the abstract system models the concrete
-one.
+a conventional programming language, only has an informal correctness argument,
+and uses ordinary floating point numbers to approximate the real numbers that are used in said argument.
+These factors limit the confidence one can justifiably have in safety judgements computed by this implementation, because (1) it is easy for bugs to creep into uncertified programs; (2) it is easy to make mistakes in informal correctness arguments; and (3) floating point computations are subject to rounding errors and representation artifacts.
+
+The primary goal of our work is to increase this degree of confidence by developing a \emph{certified} reimplementation of the abstraction technique in Coq, a proof assistant based on a rich type theory that also functions as a (purely functional) programming language.
+The Coq system lets us develop the algorithms and their formal correctness proofs in tandem in a unified environment, addressing (1) and (2) above.
+
+To address (3), we replace the floating point numbers with exact computable reals instead of floating point computation, using the certified exact real arithmetic library developed by O'Connor \cite{oconnor} for CoRN, our Coq repository of formalised constructive mathematics.
+\cite{corn} This change is much more than a simple change of representation, however; because computable reals only permit observation of arbitrarily close approximations, certain key operations on them (namely naive comparisons) are not decidable.
+The consequences of this manifest themselves in our development in several ways, which we discuss in some detail.
+Hence, our development also serves to showcase O'Connor's certified exact real arithmetic library applied to a concrete and practical problem.
+
+On a separate note, we argue that the use of computable reals is not just a pragmatic choice necessitated by the need to actually compute, but is actually fundamentally appropriate considering their role in hybrid systems, where they represent physical quantities acted upon by a device with sensors and actuators.
+And in the real world, too, measurements are approximate.
+
+The end result of our work is a framework with which one can specify (inside Coq) a concrete hybrid system, set some abstraction parameters, derive an abstract system, and use it to compute (either inside Coq itself or via extraction to OCaml) a safety proof for the concrete system.
 

### 2010-03-24 ITP10/hybrid.tex
ITP: Various improvements and shortenings.
Ignore-this: 3c36131e57a7dfc9e293d76ab4fccfc4

darcs-hash:20100324204311-bab43-3fc9a163dd5bbada82503f4e9ba520f079466d63.gz

@@
 
-The primary goal of our work is to increase this degree of confidence by developing a \emph{certified} reimplementation of the abstraction technique in Coq, a proof assistant based on a rich type theory that also functions as a (purely functional) programming language.
+Our goal is to increase this degree of confidence by developing a \emph{certified} reimplementation of the abstraction technique in Coq, a proof assistant based on a rich type theory that also functions as a (purely functional) programming language.
 The Coq system lets us develop the algorithms and their formal correctness proofs in tandem in a unified environment, addressing (1) and (2) above.
@@
 
-On a separate note, we argue that the use of computable reals is not just a pragmatic choice necessitated by the need to actually compute, but is actually fundamentally appropriate considering their role in hybrid systems, where they represent physical quantities acted upon by a device with sensors and actuators.
+On a separate note, we argue that the use of computable reals is not just a pragmatic choice necessitated by the need to compute, but is actually fundamentally appropriate considering their role in hybrid systems, where they represent physical quantities acted upon by a device with sensors and actuators.
 And in the real world, too, measurements are approximate.
@@
 \label{sec:hybsys}
-We briefly introduce the notion of a {\em linear hybrid system}, as
-used by \cite{alur}, and we show how the abstraction method is used in
-that paper in order to verify safety. There are various notions
-(formal definitions) of hybrid systems
-\cite{henziger,lynchvaandrager}. To illustrate the definition, we give
-the example of a thermostat, which is the same as \cite{alur} uses as
-a running example to illustrate the abstraction method. We will use
-the same example to illustrate our formal verification in Coq.
+There are many varieties of hybrid systems \cite{henziger,lynchvaandrager}.
+We follow \cite{alur} in considering so-called \emph{linear hybrid systems} (to be defined momentarily).
+To illustrate the definition as well as the abstraction process, we use as an example a hybrid system expressing the operation of a thermostat (this is the same example used in \cite{alur}), shown in Figure \ref{fig:thermostat}.
+
+In general, a hybrid system is a model of how a software system (running on a device with sensors and actuators), described as a finite set of \emph{locations}
+%\footnote{We reserve the term ``state'' for something else.}
+with (discrete) transitions between them, acts
+on and responds to a set of continuous variables (called the
+\emph{continuous state space}), typically representing physical
+properties of some environment (such as temperature and pressure).
+
+%format IR = "\mathbb{R}"
+%format pow x y = x "^{" y "}"
+
+%In \cite{alur}, Alur requires that the continuous state space be a
+%convex polyhedron in |pow IR n|.
+%In our definition of a hybrid system,
+%we initially let the continuous state space be an arbitrary
+%(constructive) setoid, called |Point|. We also explicitly require
+%decidability of location equality.
+
+The thermostat has three locations: $\Heat$, $\Cool$, and $\Chec$.
+The first two represent states in which the thermostat heats and cools the environment it operates in, respectively.
+$\Chec$ represents a self-diagnostic state in which the thermostat does not heat or cool.
+The continuous state space of the thermostat consists of two continuous variables denoting an internally resettable clock ($c$) and the temperature ($T$) in the environment in which the thermostat operates.
+
+With each location is associated an \emph{invariant} predicate defining the set of
+permitted values for the continuous variables while in that location.
 
@@
 
-A hybrid system is a model of how a software system, described as a
-finite set of \emph{locations}, $\Loc$ 
-%\footnote{We reserve the term ``state''
-%  for something else.} 
-with (discrete) transitions between them, acts
-on and responds to a set of continuous variables (called the
-\emph{continuous state space}), typically representing physical
-properties of some environment (such as temperature and pressure).
-
-In \cite{alur}, Alur requires that the continuous state space be a
-convex polyhedron in $\IR^ n$. 
-%In our definition of a hybrid system,
-%we initially let the continuous state space be an arbitrary
-%(constructive) setoid, called |Point|. We also explicitly require
-%decidability of location equality.
-In the running example of Figure \ref{fig:thermostat},
-the discrete component consists of three locations (called $\Heat$, $\Cool$,
-and $\Chec$), whose continuous state space is $\IR^2$,
-where $\IR$ are the \emph{computable} reals as defined in CoRN. 
-The $\Heat$ and $\Cool$ locations represent states in which the
-thermostat heats and cools the environment it operates in,
-respectively. The $\Chec$ state is a self-diagnostic state in which
-the thermostat does not heat or cool. The first $\IR$ component in the
-continuous state space represents an internally resettable clock,
-while the second represents the temperature in the environment in
-which the thermostat operates.
-
-A \emph{State} in a hybrid system is a pair $(l, p)$ consisting of a
-location $l$ and a point $p$ in the continuous state space
-(representing the value of the continuous variables). We use $\State$
-for the set of states of a hybrid system.
-
-The {\em invariant\/} predicate defines, for each location, a set of
-permitted values for the continuous variables. 
-For the thermostat, the invariant is as follows: $$\Inv_{\Heat}(c,T) := T\leq 10\wedge c\leq 3,\quad \Inv_{\Cool}(C,t) := 5 \leq T,\quad \Inv_{\Chec} := c\leq 1.$$
-
-A hybrid system also has a set of \emph{initial states}, for which we
-will use the predicate $\Init(s)$.  For the thermostat example,
+As shown in Figure \ref{fig:thermostat}, the invariants for the thermostat are as follows: $$\Inv_{\Heat} := c\leq 3 \wedge T\leq 10,\quad \Inv_{\Cool} := 5 \leq T,\quad \Inv_{\Chec} := c\leq 1.$$
+
+A hybrid system has a set of \emph{initial states}, determined by a predicate $\Init$. For the thermostat example,
 $\Init(l,c,T) := l= \Heat \wedge c= 0 \wedge 5 \leq T \leq 10.$
-There is the obvious requirement that the invariant holds at each
-initial state.
+
+%There is the obvious requirement that the invariant holds at each initial state.
 
@@
 
-Each location in a concrete system has an accompanying \emph{flow
-  function} which describes how the continuous variables change over
-time while the system is in that location. The idea is that then
-different locations correspond to different uses of actuators
+Each location in a hybrid system has an accompanying \emph{flow function} which describes how the continuous variables change over
+time while the system is in that location. The idea is that different locations correspond to different uses of actuators
 available to the software system, the effects of which are described
@@
 differential equations, which are flow functions $\Phi$ which satisfy:
-$$\Phi( p, 0) = p \quad \mbox{and} \quad \Phi (p, t + t') = \Phi (\Phi( p, t), t')$$
+$$\Phi(p, 0) = p \quad \mbox{and} \quad \Phi (p, d + d') = \Phi (\Phi(p, d), d')$$
 We now say that there is a (concrete) \emph{continuous transition}
@@
 
-In case we are in $\IR ^2$, the flow function can be expressed as the
-product of two flow functions: $\Phi_l(c,T,t) = (\phi_{l,1}(c,T,t),
-\phi_{l,2}(c,T,t))$. In the thermostat example, as in many other
-examples of hybrid systems, $\phi_{l,1}(c,T,t)$ does not depend on $T$
-and $\phi_{l,2}(c,T)$ does not depend on $c$. We call this feature
-{\em separability\/} of the flow function; our development currently
+A flow function on |pow IR 2| can be expressed as the
+product of two flow functions: $\Phi_l((c,T),d) = (\phi_{l,c}((c,T),d),
+\phi_{l,T}((c,T),d))$. In the thermostat example, as in many other
+examples of hybrid systems, $\phi_{l,c}((c,T),d)$ does not actually depend on $T$
+and $\phi_{l,T}((c,T),d)$ does not actually depend on $c$. We call this feature
+{\em separability\/} of the flow function. Our development currently
 relies heavily on this property. Separability makes the form of the
 flow functions simpler:
-$$\Phi_l(c,T,t) = (\phi_{l,1}(c,t), \phi_{l,2}(T,t))$$
-
-In the thermostat, for all locations, $\phi_{l,1}(c,d) = c + d$ (for
-the clock) and $\phi_{\Heat,2}(T,d)= T + 2 * d$,
-$\phi_{\\Cool,2}(T,d)= T * e^{-d}$, $\phi_{\Chec,2}(T,d)= T * e^{-
-  \frac{1}{2}d}$.
+$$\Phi_l((c,T),t) = (\phi_{l,1}(c,t), \phi_{l,2}(T,t))$$
+
+In the thermostat, $\phi_{l,c}(c,d) = c + d$ for all locations $l$, $\phi_{\Heat,T}(T,d)= T + 2 * d$,
+$\phi_{\\Cool,T}(T,d)= T * e^{-d}$, $\phi_{\Chec,T}(T,d)= T * e^{- \frac{1}{2}d}$.
 
@@
 logic of the software system. Each such transition is comprised of two
-components: a \emph{guard} predicate, and a \emph{reset} function. The
-former defines a subset of the continuous state space in which the
-transition is enabled (permitted), while the latter describes an
+components: a \emph{guard} predicate defining a subset of the continuous state
+space in which the transition is enabled (permitted), and a \emph{reset} function
+describing an
 instantaneous change applied as a side effect of the transition, as
@@
 complete state space. For the thermostat example, the intent is to
-keep the temperature above $4.5$ degrees at all times, and so the safe
-states are defined to be those in which we have $T > 4.5$: $\Safe(l,c,T) := T >4.5$ and $\Unsafe(l,c,T) := T\leq 4.5.$
+keep the temperature above $4.5$ degrees at all times, and so we define $\Safe(c,T) := T > 4.5$ (and $\Unsafe(c,T)$ as its complement).
 The goal, then, is to verify that the reachable states are all safe.
@@
 \section{Formalisation}
-We now describe the formalisation in Coq and the design choices
-made. Some of the design choices were due to specifics in Coq or its
-type theory CiC, while other are due to our choice to work with exact
-real arithmetic and a constructive logic. We will mostly discuss the
-latter design choices, because we feel they are most relevant.
-Technical details, including a tar file with the formalization can be found at {\tt http://www.xs4all.nl/~weegen/eelis/research/hybrid/} or \cite{hybrid-techreport}.
+We now describe the Coq formalisation and the design choices made, which fall into two categories: those concerning the specifics of Coq and its type theory CiC (which we will not pay much attention to), and those addressing concerns relating to the use of computable reals and  constructive logic.
+
+The complete sources of the development, as well as a technical report describing the formalization in more detail \cite{hybrid-techreport}, can be found at \newline {\tt http://www.eelis.net/research/hybrid/}.
 
@@
 
-A property $P$ ({\tt P: Prop}) is {\em stable\/} if $\neg \neg P \rightarrow P$ holds. 
+A property |P| is {\em stable\/} if $\neg \neg P$ implies |P|.
 Classically, this is trivially true, but constructively
@@
 
-In constructive logic, a proof of a
-disjunction $A\vee B$ (where $A$ and $B$ are arbitrary propositions) is
-either a proof of $A$ or a proof of $B$. Similarly, to have a
-proof of $X \rightarrow A \vee B$ is to have a function that, given an $X$,
-either returns a proof of $A$, or a proof of $B$. With this in mind,
-suppose we try to implement $\leltdec: \forall
-%x, y \in \IR (x\leq y \vee y<x)$. Then given $x$ and $y$ in
+Constructively, a proof of a disjunction |A \/ B| is
+either a proof of $A$ or a proof of $B$. Similarly, to have a constructive
+proof of |X -> A \/ B| is to have a function that, given an $X$,
+returns either a proof of $A$, or a proof of $B$. With this in mind,
+suppose we try to implement |le_lt_dec: forall (x, y: CR), (x <= y \/ y < x)|, where |CR| are the constructive reals. Then given $x$ and $y$ in
 $\IR$, we are to produce a proof either of $x \leq y$ or of $y <
@@
 \begin{enumerate}
-\item $\leltdecoverlap: \forall x y: \IR. ( x < y \rightarrow \forall z. (z \leq y \vee x \leq z))$
-\item $\leltdecDN: \forall x y: \IR. \neg  \neg  (x \leq y \vee y < x)$
+\item |leltdec_overlap: forall (x y: CR), ( x < y -> forall z, (z <= y \/ x <= z))|
+\item |leltdecDN: forall (x y: CR), ~ ~ (x <= y \/ y < x)|
 \end{enumerate}
@@
 the monad may be used freely in proofs of additional properties in the
-monad. For instance, one may $bind_{\DN}$ a proof of $\DN (x \leq y \vee y < x)$ 
+monad. For instance, one may |bind_DN| a proof of $\DN (x \leq y \vee y < x)$
 (obtained from $\leltdecDN$ above) with a proof of $(x \leq y \vee y < x) \rightarrow \DN P$, yielding a proof of $\DN P$.
@@
 will ultimately only use |regions_cover| in a proof of of |... -> ~
-concrete.reachable s| (for some universally quantified variable s),
+concrete.reachable s| (for some universally quantified variable $s$),
 which, due to its head type being a negation, is stable, and can
@@
 proof, even though the program's purpose is to produce proofs.
+
 The opposite of an underestimation is an overestimation:
@@
 transitions (and therefore traces), adversely affecting the abstract
-system's utility
+system's utility.
 %(as we will discuss in detail in section \ref{trouble}).
@@
 
+%format PowerSet = "\mathcal{P}"
+%format `In` = "\in "
 \begin{code}
   Definition shared_cover
-    `{Container concrete.State C} `{Container abstract.State D}
-      (cs: C) (ss: D): Prop :=
-        forall s: concrete.State, In s cs -> DN (exists r: abstract.State, In s r /\ In r ss).
+    (cs: PowerSet(concrete.State)) (ss: PowerSet(abstract.State)): Prop :=
+        forall s: concrete.State, s `In` cs -> DN (exists r: abstract.State, s `In` r /\ r `In` ss).
 \end{code}
@@
 We now specify what the type of $\overcontrans$ should be.
+%format `compose` = "\circ "
+%format `inter` = "\cap "
 \begin{code}
@@
   { p: list abstract.State | NoDup p /\ shared_cover
-    (concrete.invariant `inter` (overlap s `compose` utilflip (concrete.cont_trans))) p }
+    (concrete.invariant `inter` (overlap s `compose` flip concrete.cont_trans)) p }
 \end{code}
@@
 
+%format `subsetof` = "\subseteq "
+
 In the next section we show that thanks to decidability of our
@@
 hybrid system safety, and in particular, lets us obtain a proof of
-thermostat safety. So suppose we have |reachable_dec: decider abstract.reachable|
+thermostat safety. So suppose we have |reachable_dec: decider abstract.reachable|.
 \weg{

### 2010-04-01 ITP10/hybrid.tex
ITP: Space reduction and lots of small tweaks.

@@
 On a separate note, we argue that the use of computable reals is not just a pragmatic choice necessitated by the need to compute, but is actually fundamentally appropriate considering their role in hybrid systems, where they represent physical quantities acted upon by a device with sensors and actuators.
-And in the real world, too, measurements are approximate.
+In the real world, too, measurements are approximate.
 
@@
 \label{sec:hybsys}
-There are many varieties of hybrid systems \cite{henziger,lynchvaandrager}.
-We follow \cite{alur} in considering so-called \emph{linear hybrid systems} (to be defined momentarily).
-To illustrate the definition as well as the abstraction process, we use as an example a hybrid system expressing the operation of a thermostat (this is the same example used in \cite{alur}), shown in Figure \ref{fig:thermostat}.
-
 In general, a hybrid system is a model of how a software system (running on a device with sensors and actuators), described as a finite set of \emph{locations}
@@
 
+There are many varieties of hybrid systems \cite{henziger,lynchvaandrager}.
+We follow \cite{alur} in considering so-called \emph{linear hybrid systems}\weg{ (to be defined momentarily)}.
+To illustrate the definition and the abstraction process, we use as an example a hybrid system expressing the operation of a thermostat (this is the same example used in \cite{alur}), shown in Figure \ref{fig:thermostat}.
+
 %format IR = "\mathbb{R}"
 %format pow x y = x "^{" y "}"
+%format cross = "\times "
 
@@
 
-The thermostat has three locations: $\Heat$, $\Cool$, and $\Chec$.
-The first two represent states in which the thermostat heats and cools the environment it operates in, respectively.
-$\Chec$ represents a self-diagnostic state in which the thermostat does not heat or cool.
-The continuous state space of the thermostat consists of two continuous variables denoting an internally resettable clock ($c$) and the temperature ($T$) in the environment in which the thermostat operates.
+The thermostat has three locations.
+The first two, $\Heat$ and $\Cool$, represent states in which the thermostat heats and cools the environment it operates in, respectively.
+The third, $\Chec$, represents a self-diagnostic state in which the thermostat does not heat or cool.
+The continuous state space of the thermostat consists of two continuous variables denoting an internally resettable clock $c$ and the temperature $T$ in the environment in which the thermostat operates.
 
@@
 permitted values for the continuous variables while in that location.
+The invariants for the thermostat are: $$\Inv_{\Heat} := c\leq 3 \wedge T\leq 10,\quad \Inv_{\Cool} := 5 \leq T,\quad \Inv_{\Chec} := c\leq 1.$$
 
@@
 
-As shown in Figure \ref{fig:thermostat}, the invariants for the thermostat are as follows: $$\Inv_{\Heat} := c\leq 3 \wedge T\leq 10,\quad \Inv_{\Cool} := 5 \leq T,\quad \Inv_{\Chec} := c\leq 1.$$
-
-A hybrid system has a set of \emph{initial states}, determined by a predicate $\Init$. For the thermostat example,
+The \emph{initial states} of a hybrid system are determined by a predicate $\Init$. For the thermostat,
 $\Init(l,c,T) := l= \Heat \wedge c= 0 \wedge 5 \leq T \leq 10.$
@@
 
-The remaining parts of a hybrid system (i.e.\ {\em flow}, {\em guard}
-and {\em reset}) describe transitions between states, which, together
+The remaining parts of a hybrid system (described below) describe transitions between states, which, together
 with the set of initial states, determine the set of \emph{reachable}
@@
 time while the system is in that location. The idea is that different locations correspond to different uses of actuators
-available to the software system, the effects of which are described
-by the flow function. For instance, in the thermostat, the flow
+available to the software system, the effects of which are reflected in the flow function. For instance, in the thermostat, the flow
 function corresponding to the $\Heat$ location will have the
-temperature increase with time, modelling the effect of the heater
+temperature increase over time, modelling the effect of the heater
 component in the imagined thermostat device.
@@
 We now say that there is a (concrete) \emph{continuous transition}
-from a state $(l, p)$ to a state $(l', p')$ if $l = l'$ and there is a
-non-negative duration $d$ such that $p' = \Phi_l( p,d)$ with the
+from a state $(l, p)$ to a state $(l, p')$ if there is a
+non-negative duration $d$ such that $p' = \Phi_l(p,d)$ with the
 invariant for $l$ holding at every point along the way:
-$$ (l,p) \contrans (l',p') := l=l' \wedge \Exists{d\in
-\Dur}{\Phi_l(p, d) = p' \wedge \Forall{t}{0 \leq t\leq d \rightarrow
-\Inv_l(\Phi_l (p, t))}}.$$
+$$ (l,p) \contrans (l',p') := l = l' \wedge \Exists{d\in
+\Dur}{\Phi_l(p, d) = p' \wedge \Forall{0 \leq t\leq d}{\Inv_l(\Phi_l (p, t))}}.$$
 
@@
 flow functions simpler:
-$$\Phi_l((c,T),t) = (\phi_{l,1}(c,t), \phi_{l,2}(T,t))$$
-
-In the thermostat, $\phi_{l,c}(c,d) = c + d$ for all locations $l$, $\phi_{\Heat,T}(T,d)= T + 2 * d$,
+$$\Phi_l((c,T),t) = (\phi_{l,c}(c,t), \phi_{l,T}(T,t))$$
+
+In the thermostat, $\phi_{l,c}(c,d) = c + d$ for all locations $l$, $\phi_{\Heat,T}(T,d)= T + 2 d$,
 $\phi_{\\Cool,T}(T,d)= T * e^{-d}$, $\phi_{\Chec,T}(T,d)= T * e^{- \frac{1}{2}d}$.
@@
 It will be obvious from Figure \ref{fig:thermostat} what the guards
-and reset functions are. The reset function resets the clock for all
+and reset functions are.
+
+\weg{
+The reset function resets the clock for all
 but one of these transitions, and leaves the temperature variable as
@@
 value. However, this would simply make the system unimplementable.
-
-A transition is either continuous or discrete $\trans:= \distrans \cup \contrans$.
+}
+
+A transition is either continuous or discrete: $\trans:= \distrans \cup \contrans$.
 We now say that a state $s$ is \emph{reachable} if there is an initial state $i$ from which one can, by a finite number of transitions, end up in $s$ ($\ttrans$ is the transitive reflexive closure of $\trans$):
@@
 
-The idea now is that the purpose of a hybrid system is typically to
-keep the continuous variables within certain limits. In other words,
-to limit the set of reachable states to some ``safe'' subset of the
-complete state space. For the thermostat example, the intent is to
+The object of hybrid system safety verification is to show that the set of reachable states is a subset of a predefined set of ``safe'' states. For the thermostat, the intent is to
 keep the temperature above $4.5$ degrees at all times, and so we define $\Safe(c,T) := T > 4.5$ (and $\Unsafe(c,T)$ as its complement).
-The goal, then, is to verify that the reachable states are all safe.
 
@@
 There are uncountably many traces in a hybrid system, so safety is
-undecidable in general. In concrete situations, safety may be (easily)
-provable, however, for example, if one finds the proper {\em proof
-  invariant}. As proof-invariants are often hard to find, one would
-like to have more automated heuristic methods for establishing
-safety. This is what the {\em predicate abstraction method\/} of
-\cite{alur} does. In general one divides the continuous state space
+undecidable in general. In concrete cases, however, safety may be (easily)
+provable if one finds the proper {\em proof invariant}. Unfortunately these
+are often hard to find, so we prefer methods that are more easily automated. The \emph{predicate abstraction} method of \cite{alur} is one such method.
+
+The idea is to divide the continuous state space
 into a finite number of convex subsets (polygons), $A_1, \ldots, A_n$,
@@
 abstract state space, one immediately defines {\em abstract continuous
-  transitions\/} and {\em abstract discrete transitions\/} as follows.
+  transitions\/} and {\em abstract discrete transitions\/} (both potentially undecidable) as follows.
 \begin{eqnarray*}
@@
 
-This yields a finite transition system (where it may be undecidable if
-there is a transition between two states!) which has the
-obvious property indicated in Figure \ref{diag:abstraction}.
+Traces in the finite transition system constructed this way are sufficiently representative (see Figure \ref{diag:abstraction}.) of those in the original (concrete) system that one can conclude safety of the latter from safety of the abstract system:
+$$\text{if}\ \Forall{a\in \AState}{\AReach(a) \rightarrow \ASafe(a)}, \text{then}\ \Forall{s\in \State}{
+\Reach(s) \rightarrow \Safe(s)}.$$
+
 \begin{figure}[htb!]
@@
 \end{figure}
-With this property, one can conclude safety of the concrete system
-from safety of the abstract system: $$\text{if}\ \Forall{a\in \AState}{\AReach(a) \rightarrow \ASafe(a)}, \text{then}\ \Forall{s\in \State}{
-\Reach(s) \rightarrow \Safe(s)}.$$
 
@@
 of diagram \ref{diag:abstraction}, so we can {\em overestimate\/}
-$\acontrans$ and $\adistrans$ (that is: replace it with a transition
-relation that allows more transitions). Second, there is a fairly good
-heuristic for making the abstraction (how to divide the continuous
-state space into regions) and how to decide whether there should be an
+$\acontrans$ and $\adistrans$ (i.e. replace it with a transition
+relation that allows more transitions). Second, there are good
+heuristics for how to divide the continuous
+state space into regions, and how to decide whether there should be an
 abstract transition from one abstract state to another.
-This is indicated in figure \ref{fig:abstraction}.
+
+This is indicated in figure \ref{fig:abstraction}. The left side illustrates
+the challenge: given abstract regions $A$ and $B$, we are to determine
+whether nonnegative flow durations might permit flow from points in $A$ to points in $B$.
+Following the overestimation property just mentioned, we introduce an abstract transition from $A$ to $B$ whenever we cannot positively rule out such paths.
+
+On the right side we see the abstract state space indicated for the
+location $\Heat$. The abstract state space consists of rectangles,
+possibly degenerate (extending to $-\infty$ or $+\infty$). For the
+bounds of these rectangles one takes the values that occur in the
+specification of the hybrid system (see Figure
+\ref{fig:thermostat}). According to \cite{alur}, this is a good
+choice. (In case one cannot prove safety in this, there is of course
+the opportunity for the user to refine it.)  The gray area indicates
+that from these states also abstract discrete transitions are
+possible. The dashed area is unreachable, because of the invariant for
+the $\Heat$ location. %prevents these states from being reachable.  
+All the abstract
+transitions from the rectangle $[0.5,1) \times[5,6)$ are shown: as the
+    temperature flow function for $\Heat$ is $\phi_{\Heat,2}(T,d)= T +
+    2 * d$, and the clock flow function is $\phi_{\Heat,1}(c,d) = c +
+    d$, these are all the possible abstract transitions.
+
+Using the abstraction method, \cite{alur} proves the correctness of the thermostat.
 
@@
 \end{center}
-
-On the left side we schematically indicate how one can try to
-``follow'' the path that region $A$ takes when time passes and verify
-whether this path intersects with region $B$. The idea of
-overestimation is that one should only {\em not\/} put an arrow $A
-\acontrans B$ if we positively know that the path of $A$ will avoid
-$B$, which is the case for example for paths that follow the dashed lines.
-
-On the right side we see the abstract state space indicated for the
-location $\Heat$. The abstract state space consists of rectangles,
-possibly degenerated (extending to $-\infty$ or $+\infty$). For the
-bounds of these rectangles one takes the values that occur in the
-specification of the hybrid system (see Figure
-\ref{fig:thermostat}). According to \cite{alur}, this is a good
-choice. (In case one cannot prove safety in this, there is of course
-the opportunity for the user to refine it.)  The gray area indicates
-that from these states also abstract discrete transitions are
-possible. The dashed area is unreachable, because of the invariant for
-the $\Heat$ location %prevents these states from being reachable.  
-All the abstract
-transitions from the rectangle $[0.5,1) \times[5,6)$ are shown: as the
-    temperature flow function for $\Heat$ is $\phi_{\Heat,2}(T,d)= T +
-    2 * d$, and the clock flow function is $\phi_{\Heat,1}(c,d) = c +
-    d$, these are all the possible abstract transitions.
-
+\weg{ % because this is explained by Alur already and it is not of particular relevance in our development; we don't refer to it ever again.
 There is a slight subtlety with the abstract transitions that we want
@@
         of $\adistrans$ gives us such a transition anyway.
+}
 % \ADAM{I will try to add some more explanation}
-Using this method, \cite{alur} proves the correctness of the thermostat.
-
 \section{Formalisation}
-We now describe the Coq formalisation and the design choices made, which fall into two categories: those concerning the specifics of Coq and its type theory CiC (which we will not pay much attention to), and those addressing concerns relating to the use of computable reals and  constructive logic.
+We now describe the Coq formalisation and the design choices made.
+We will not pay much attention to the specifics of Coq and its type theory CiC, and will instead focus on concerns relating to the use of computable reals and constructive logic.
 
@@
 parts of which we discuss in the remainder of this section.
-
 \begin{code}
@@
   ; locations: ExhaustiveList Location
-  ; State :=  Location x Point
+  ; State :=  Location cross Point
   ; initial: State -> Prop
@@
 \end{code}
-
 This Coq definition defines the record type of ``Systems'', which
@@
 and a field ``Location'', representing the set of locations. Here, we
-take Point to be an arbitrary Constructive Setoid, which is basically
+take Point to be an arbitrary constructive setoid, which is basically
 just a type with an equality on it. For Location, we assume a
@@
 
-\subsection{Stability and Double Negation}
+\subsection{Stability, Double Negation, and Computable Reals}
 \label{dn}
 
+\weg{ % wordt later ook genoemd
 A property |P| is {\em stable\/} if $\neg \neg P$ implies |P|.
@@
 not.
+}
 %Especially when we work with real numbers, we don't have $\leltdec: \forall
@@
 
-Constructively, a proof of a disjunction |A \/ B| is
-either a proof of $A$ or a proof of $B$. Similarly, to have a constructive
-proof of |X -> A \/ B| is to have a function that, given an $X$,
+Constructively, a proof of |X -> A \/ B| is a function that, given an $X$,
 returns either a proof of $A$, or a proof of $B$. With this in mind,
 suppose we try to implement |le_lt_dec: forall (x, y: CR), (x <= y \/ y < x)|, where |CR| are the constructive reals. Then given $x$ and $y$ in
-$\IR$, we are to produce a proof either of $x \leq y$ or of $y <
+|CR|, we are to produce a proof either of $x \leq y$ or of $y <
 x$. Unfortunately, the nature of computable reals only lets us observe
-arbitrarily close approximations of $x$ and $y$. Now suppose $x =
-y$. Then no matter how closely we approximate $x$ and $y$, the error
+arbitrarily close approximations of $x$ and $y$. If it happens to be the case that $x =
+y$, then no matter how closely we approximate $x$ and $y$, the error
 margins (however small) will always leave open the possibility that
@@
 Computable reals do admit two variations of the proposition:
+%format leltdec_overlap
+%format leltdec_DN
+%format ~ = "\neg "
 \begin{enumerate}
 \item |leltdec_overlap: forall (x y: CR), ( x < y -> forall z, (z <= y \/ x <= z))|
-\item |leltdecDN: forall (x y: CR), ~ ~ (x <= y \/ y < x)|
+\item |leltdec_DN: forall (x y: CR), ~ ~ (x <= y \/ y < x)|
 \end{enumerate}
@@
 use. Nevertheless, this is the path we will take in our development,
-because just taking $\leltdec$ as an axiom amounts to cheating.
+because just taking |le_lt_dec| as an axiom amounts to cheating.
 %and partly because in some parts of the development, we may need to
@@
 that double negation, as a function on propositions, is a monad
-\cite{wadler95monads}. Writing $\DN P$ for $\neg  \neg  P$, we have the following two key
-operations that make $\DN$ a monad:
-
+\cite{wadler95monads}. Writing |DN P| for |~ ~ P|, we have the following two key
+operations that make |DN| a monad:
+
+%format return_DN = "\Varid{return}_{DN} "
 %format bind_DN
@@
 the monad may be used freely in proofs of additional properties in the
-monad. For instance, one may |bind_DN| a proof of $\DN (x \leq y \vee y < x)$
-(obtained from $\leltdecDN$ above) with a proof of $(x \leq y \vee y < x) \rightarrow \DN P$, yielding a proof of $\DN P$.
-
-Thus, $\DN$ establishes a ``proving context'' in which one may make
-use of lemmas yielding results inside $\DN$ that may not hold outside
-of it (such as $\leltdecDN$), as well as lemmas yielding results
-not in $\DN$, which can always be injected into $\DN$ using
-$return_{\DN}$. The catch is that such proofs always end up with results
-in $\DN$, which begs the question: what good is any of this? In
-particular, can $\leltdecDN$ be used to prove anything not doubly
+monad. For instance, one may |bind_DN| a proof of $\Varid{DN} (x \leq y \vee y < x)$
+(obtained from |leltdec_DN| above) with a proof of $(x \leq y \vee y < x) \rightarrow \Varid{DN} P$, yielding a proof of |DN P|.
+
+Thus, |DN| establishes a ``proving context'' in which one may make
+use of lemmas yielding results inside |DN| that may not hold outside
+of it (such as |leltdec_DN|), as well as lemmas yielding results
+not in |DN|, which can always be injected into |DN| using
+|return_DN|. The catch is that such proofs always end up with results
+in |DN|, which begs the question: what good is any of this? In
+particular, can |leltdec_DN| be used to prove anything not doubly
 negated?
 
-As it happens, there is the class of \emph{stable} propositions (that
-are equivalent to their own double negation). Examples include
+As it happens, some propositions are \emph{stable} in the sense that they are constructively equivalent to their own double negation. Examples include
 negations, non-strict inequalities on real numbers, and any decidable
@@
 
+Requiring that hybrid system invariants are stable effectively lets us use classical reasoning when showing that invariants holds in certain states.
+One instance where we need this is in the proof of transitivity of the concrete continuous transition.
+
+\weg{ %referred to something that was already cut
 We have seen why we required invariant stability:
@@
 then pull the latter out of $\DN$ on account of its stability.
+}
 
@@
 \end{itemize}
-The second is very problematic, because the property for the
-abstraction function $A$, as depicted in diagram
-\ref{diag:abstraction} is now just false.
-
-These may seem just anoying problems,
-caused by the fact that our logic is constructive, but we claim it is
-not. It is is a fact of physical reality that one can only get a
-finite approximation of a measured value, so case distinctions like
-$x\leq 0 \vee 0< x$ are unrealistic. Moreover, we claim that the
+The second is very problematic, because it means the property for the
+abstraction function $A$ depicted in diagram
+\ref{diag:abstraction} no longer holds.
+
+We argue that these problems are not merely annoying byproducts
+of our use of constructive logic and computable reals, but actually
+reflect the profound limitation of physical reality that one can only
+ever measure quantities approximately, making case distinctions like
+$x\leq 0 \vee 0< x$ simply unrealistic.
+
+Moreover, we claim that the
 classical abstraction method allows one to prove the safety of systems
@@
 \item
-We replace the {\em abstract continuous transition\/} relation
-$\acontrans$ by a function $\overcontrans$ that takes a region $R_0$
-as input and outputs a {\em list of regions\/} including $R_0$: $(R_o,
+We replace the abstract transition relations $\acontrans$ and $\distrans$ by functions $\overcontrans$ and $\overdistrans$ that take a region $R_0$
+as input and output a {\em list of regions\/} including $R_0$: $(R_o,
 R_1, \ldots, R_n)$ in such a way that $\cup_{0\leq i\leq n}R_i$ is an
-overapproximation of the set of states reachable by a continuous step
+overapproximation of the set of states reachable by a continuous (resp. discrete) step
 from a state in $R_0$.\\
-We do something similar for the {\em abstract discrete transition}, and obtain $\overdistrans :\Region \rightarrow List \Region$.
 \item
-We loosen the requirement on the abstraction function $A$; for $s \in \State$, we only require $\DN {\exists r\in \Region}{(s\in r)}$.
+We loosen the requirement on the abstraction function $A$; for $s \in \State$, we only require $\Varid{DN}(\Exists {r\in \Region}{s\in r})$.
 \end{itemize}
 
-So, to summarize, if $s\contrans s'$, then we don't require
+To summarize, if $s\contrans s'$, then we don't require
 $A(s')$ to be in the list $\overcontrans(A(s))$, but we only require
@@
 
-We define |underestimation P| to be either a proof of
-|P|, or not. In Coq, either of the following will do:
-\begin{code}
-Definition underestimation (P: Prop): Set := option P.
+% Eelis: discussion of |option P| variant removed in what follows
+We define |underestimation P| to be either a proof of |P|, or not:
+\begin{code}
 Definition underestimation (P: Prop): Set := { b: bool | b = true -> P }.
 \end{code}
-The latter tends to work better with the |Program| family of commands
-\cite{sozeau} which have special support for dependent pairs. Using
-|Program|, an underestimation of the second variety may be
+Using the |Program| family of commands
+\cite{sozeau}, an underestimation may be
 provisionally defined strictly as a bool, and then separately proved
 to be a valid underestimation in a proof obligation generated by
-|Program|. The second form also nicely illustrates why we call this an
-underestimation: it may be |false| even when |P| holds. We can now
+|Program|. The |bool| in the definition nicely illustrates why we call this an
+``underestimation'': it may be |false| even when |P| holds. We can now
 describe the functionality of our program by saying that it
@@
 because they can be trivially ``proved'' by taking
-|false|/|None|. Hence, the value of our program is not witnessed by
+|false|. Hence, the value of our program is not witnessed by
 the mere fact that it manages to produce terms of type
 |underestimation Safe|, but rather by the fact that when run, it
-actually manages to returns |true|/|Just P| for the hybrid system we
+actually manages to returns |true| for the hybrid system we
 are interested in (e.g.\ the thermostat). It is for this reason that
@@
 proof. 
-Having to do this "tweaking" manually is obviously not idea. One may want to develop heuristics for this.
+Having to do this ``tweaking'' manually is obviously not ideal. One may want to develop heuristics for this.
 
@@
   Definition shared_cover
-    (cs: PowerSet(concrete.State)) (ss: PowerSet(abstract.State)): Prop :=
+    (cs: concrete.State -> Prop) (ss: abstract.State -> Prop): Prop :=
         forall s: concrete.State, s `In` cs -> DN (exists r: abstract.State, s `In` r /\ r `In` ss).
@@
 
-The details of the |Container| type class are of no interest to us
-right now; one may think of |C| and |D| as the types of lists over
-concrete states, respectively over abstract states.  What matters is
-that a ``container'' of concrete states is said to be sharedly-covered
-by a ``container'' of abstract states if for each of the concrete
+A set of concrete states is said to be sharedly-covered
+by a set of abstract states if for each of the concrete
 states in the former there is an abstract state in the latter that
 contains it.
-We state this in
-terms of |Container|s so that the definition applies to any types
-which have a reasonable notion of containment.
-
-We now specify what the type of $\overcontrans$ should be.
+
+We now specify what the type of |over_cont_trans| should be.
 %format `compose` = "\circ "
@@
 
-We now consider the properties we require for |abstract.reachable|. The following basic specification we had in mind:
-
-\begin{code}
-Definition reachable_respect: Prop :=
+We now consider the properties we require for |abstract.reachable|. An obvious candidate is:
+\begin{code}
   forall (s: concrete.State), concrete.reachable s ->
@@
 \end{code}
-Which we chose because it implies
+Which would nicely imply:
 \begin{code}
@@
 \end{code}
-This expresses that to conclude unreachability of a concrete state,
+This would expresses that to conclude unreachability of a concrete state,
 one need only establish unreachability of \emph{any} abstract state
@@
 \end{code}
-Finally, we \emph{run} the underestimation:
+Using a tiny utility |underestimation_true| of type |forall P (o: underestimation P), o = true -> P|, we can now \emph{run} this underestimator to obtain a proof of the thermostat system's safety:
 \begin{code}
@@
 \end{code}
-|underestimation_true| is a tiny utility of type |forall P (o: underestimation P), o = true -> P|, whose application in the proof reduces the goal to
+The first |apply| reduces the goal to
 \begin{code}under_thermo_unsafe_unreachable = true.\end{code} The |vm_compute| tactic invocation then forces evaluation of the left hand side, which will in turn evaluate |over_thermo_unsafe_reachable|, which will evaluate |reachable_dec|, which will evaluate the overestimators of the continuous and discrete transitions.
@@
 reals in CoRN do really complicated things for us, by approximating
-values for various real number exprseeions at great precision. The
+values for various real number expressions at great precision. The
 development also contains some nice layers of abstraction, involving
@@
 
-It is to be seen how far this automated verification approach can be
+It remains to be seen how far this automated verification approach can be
 taken, given the fact that we have limited ourselves to hybrid systems

### 2010-04-08 ITP10/hybrid.tex
final ? updates

@@
 \input{diagrams}
-\newcommand{\thetitle}{Automated Machine-Checked Hybrid System Safety Proofs}
+\newcommand{\thetitle}{Automated Machine-Checked Hybrid System Safety Proofs\footnote{This research was supported by the BRICKS/FOCUS project 642.000.501, Advancing the Real use of Proof Assistants}}
 \newcommand{\eelis}{Eelis van der Weegen$^{\dagger}$}
@@
 \newcommand{\weg}[1]{}
-
 \newcommand{\DN}{{\sf DN}\,}
 \newcommand{\IR}{{\mathbf R}}
+\newcommand{\ST}{{\sf T}}
+\newcommand{\Sc}{{\sf c}}
 \newcommand{\Heat}{{\sf Heat}}
@@
 \newcommand{\Chec}{{\sf Check}}
-\newcommand{\Off}{{\rm Off}} 
+\newcommand{\Off}{{\sf Off}} 
 \newcommand{\Loc}{{\rm Loc}}
@@
 hybrid systems; a framework for the construction of an abstract
-system (consisting of decidable ``overestimators'' of abstract
+system (consisting of decidable ``over-estimators'' of abstract
 transitions and initiality) faithfully representing a (concrete)
@@
 safety verification in which one derives from the hybrid system of interest an \emph{abstract}
-hybrid system, which is essentially a finite automaton whose traces are sufficiently representative of traces in the original system that unreachability in the abstract system (which can be decided using a standard graph algorithm) implies unreachability in the concrete system (which, being governed by continuous behaviors, cannot be decided so readily).
+hybrid system, which is essentially a finite automaton whose traces are sufficiently representative of traces in the original system that unreachability in the abstract system (which can be decided using a standard graph algorithm) implies unreachability in the concrete system (which, being governed by continuous behaviours, cannot be decided so readily).
 Thus, the abstraction method brings the safety verification problem from a continuous and infinite domain into a discrete and finite domain, where it can be dealt with using standard graph algorithms.
@@
 
-To address (3), we replace the floating point numbers with exact computable reals instead of floating point computation, using the certified exact real arithmetic library developed by O'Connor \cite{oconnor} for CoRN, our Coq repository of formalised constructive mathematics.
-\cite{corn} This change is much more than a simple change of representation, however; because computable reals only permit observation of arbitrarily close approximations, certain key operations on them (namely naive comparisons) are not decidable.
-The consequences of this manifest themselves in our development in several ways, which we discuss in some detail.
-Hence, our development also serves to showcase O'Connor's certified exact real arithmetic library applied to a concrete and practical problem.
+To address (3), we replace the floating point numbers with exact
+computable reals instead of floating point computation, using the
+certified exact real arithmetic library developed by O'Connor
+\cite{oconnor} for CoRN, our Coq repository of formalised constructive
+mathematics. \cite{corn} This change is much more than a simple change
+of representation, however; because computable reals only permit
+observation of arbitrarily close approximations, certain key
+operations on them (namely naive comparisons) are not decidable. The
+consequences of this manifest themselves in our development in several
+ways, which we discuss in some detail. Hence, our development also
+serves to showcase O'Connor's certified exact real arithmetic library
+applied to a concrete and practical problem.
 
@@
 \label{sec:hybsys}
-In general, a hybrid system is a model of how a software system (running on a device with sensors and actuators), described as a finite set of \emph{locations}
+A hybrid system is a model of how a software system (running on a device with sensors and actuators), described as a finite set of \emph{locations}
 %\footnote{We reserve the term ``state'' for something else.}
@@
 
-There are many varieties of hybrid systems \cite{henziger,lynchvaandrager}.
-We follow \cite{alur} in considering so-called \emph{linear hybrid systems}\weg{ (to be defined momentarily)}.
-To illustrate the definition and the abstraction process, we use as an example a hybrid system expressing the operation of a thermostat (this is the same example used in \cite{alur}), shown in Figure \ref{fig:thermostat}.
+There are many varieties of hybrid systems
+\cite{henziger,lynchvaandrager}. We follow \cite{alur} in considering
+so-called \emph{linear hybrid systems}\weg{ (to be defined
+  momentarily)}. To illustrate the definition and the abstraction
+process, we use the example of a hybrid system describing
+a thermostat (this is the same example used in \cite{alur}), shown
+in Figure \ref{fig:thermostat}. (The linearity in \cite{alur} refers
+to invariants, guards etc.; our development doesn't use that.)
 
@@
 permitted values for the continuous variables while in that location.
-The invariants for the thermostat are: $$\Inv_{\Heat} := c\leq 3 \wedge T\leq 10,\quad \Inv_{\Cool} := 5 \leq T,\quad \Inv_{\Chec} := c\leq 1.$$
+The invariants for the thermostat are: $$\Inv_{\Heat}(c,T) :=  T\leq 10 \wedge c\leq 3,\quad \Inv_{\Cool}(c,T) := T \geq 5,\quad \Inv_{\Chec}(c,T) := c\leq 1.$$
 
@@
 The \emph{initial states} of a hybrid system are determined by a predicate $\Init$. For the thermostat,
-$\Init(l,c,T) := l= \Heat \wedge c= 0 \wedge 5 \leq T \leq 10.$
+$\Init(l,c,T)$ is defined as $l= \Heat \wedge c= 0 \wedge 5 \leq T \leq 10.$
 
@@
 
-The remaining parts of a hybrid system (described below) describe transitions between states, which, together
-with the set of initial states, determine the set of \emph{reachable}
-states, representing the possible behaviours exhibited by a
-hypothetical real-world implementation of the hybrid system (as
-software running on a device with sensors and actuators).
-
-Each location in a hybrid system has an accompanying \emph{flow function} which describes how the continuous variables change over
-time while the system is in that location. The idea is that different locations correspond to different uses of actuators
-available to the software system, the effects of which are reflected in the flow function. For instance, in the thermostat, the flow
-function corresponding to the $\Heat$ location will have the
-temperature increase over time, modelling the effect of the heater
-component in the imagined thermostat device.
-
-In the canonical definition of hybrid systems, flow functions are
-specified as solutions to differential equations describing the
-dynamics of the continuous variables. We follow \cite{alur} in
-abstracting from these, taking instead the solutions of these
-differential equations, which are flow functions $\Phi$ which satisfy:
-$$\Phi(p, 0) = p \quad \mbox{and} \quad \Phi (p, d + d') = \Phi (\Phi(p, d), d')$$
-We now say that there is a (concrete) \emph{continuous transition}
-from a state $(l, p)$ to a state $(l, p')$ if there is a
-non-negative duration $d$ such that $p' = \Phi_l(p,d)$ with the
-invariant for $l$ holding at every point along the way:
-$$ (l,p) \contrans (l',p') := l = l' \wedge \Exists{d\in
-\Dur}{\Phi_l(p, d) = p' \wedge \Forall{0 \leq t\leq d}{\Inv_l(\Phi_l (p, t))}}.$$
-
-A flow function on |pow IR 2| can be expressed as the
-product of two flow functions: $\Phi_l((c,T),d) = (\phi_{l,c}((c,T),d),
-\phi_{l,T}((c,T),d))$. In the thermostat example, as in many other
-examples of hybrid systems, $\phi_{l,c}((c,T),d)$ does not actually depend on $T$
-and $\phi_{l,T}((c,T),d)$ does not actually depend on $c$. We call this feature
-{\em separability\/} of the flow function. Our development currently
-relies heavily on this property. Separability makes the form of the
-flow functions simpler:
-$$\Phi_l((c,T),t) = (\phi_{l,c}(c,t), \phi_{l,T}(T,t))$$
-
-In the thermostat, $\phi_{l,c}(c,d) = c + d$ for all locations $l$, $\phi_{\Heat,T}(T,d)= T + 2 d$,
-$\phi_{\\Cool,T}(T,d)= T * e^{-d}$, $\phi_{\Chec,T}(T,d)= T * e^{- \frac{1}{2}d}$.
-
-Where continuous transitions describe the flow of continuous
-variables, \emph{discrete} transitions between locations describe the
+%The remaining parts of a hybrid system (described below) describe
+%transitions between states, which, together with the set of initial
+%states, determine the set of \emph{reachable} states, representing the
+%possible behaviours exhibited by a hypothetical real-world
+%implementation of the hybrid system (as software running on a device
+%with sensors and actuators).
+The \emph{discrete} transitions between locations describe the
 logic of the software system. Each such transition is comprised of two
@@
 seen in the following definition of the discrete transition relation:
-$$ (l,p) \distrans (l',p') := \guard_{l,l'}(p) \wedge \reset_{l,l'}(p)= p' \wedge \Inv_l(p) \wedge \Inv_{l'}(p')$$
-It will be obvious from Figure \ref{fig:thermostat} what the guards
-and reset functions are.
+$$ (l,p) \distrans (l',p') := \guard_{l,l'}(p) \wedge
+\reset_{l,l'}(p)= p' \wedge \Inv_l(p) \wedge \Inv_{l'}(p')$$ It will
+be clear from Figure \ref{fig:thermostat} what the guards and reset
+functions are. Note the inherent non-determinism in a Hybrid Systems
+specification: when in $\Cool$, the system can jump to $\Heat$
+whenever the temperature $T$ is in the interval $[5,6]$.
+
+Each location in a hybrid system has an accompanying \emph{flow function} which describes how the continuous variables change over
+time while the system is in that location. The idea is that different locations correspond to different uses of actuators
+available to the software system, the effects of which are reflected in the flow function. In the thermostat example, the flow
+function corresponding to the $\Cool$ location has the
+temperature decrease over time. 
+%modelling the effect of the heater
+%component in the imagined thermostat device.
+This is expressed via the differential equation $\dot{T} = - T$,
+which is the usual short hand for $T'(t) = - T(t)$, where $T'(t)$
+denotes the derivative of the temperature function over time $t$.
+
+In the canonical definition of hybrid systems, flow functions are
+specified as solutions to differential equations (or differential
+inclusions) describing the dynamics of the continuous variables. We
+follow \cite{alur} in abstracting from these, taking instead the
+solutions of these differential equations, which are flow functions
+$\Phi$ which satisfy:
+$$\Phi(p, 0) = p \quad \mbox{and} \quad \Phi (p, d + d') = \Phi
+(\Phi(p, d), d')$$ 
+The idea is that $\Phi(p,d)$ denotes the value of the
+continuous variable after duration $d$, starting from the value
+$p$. We say that there is a (concrete) \emph{continuous
+  transition} from a state $(l, p)$ to a state $(l, p')$ if there is a
+non-negative duration $d$ such that $p' = \Phi_l(p,d)$ with the
+invariant for $l$ holding at every point along the way:
+$$ (l,p) \contrans (l,p') := \Exists{d\in
+\Dur}{\Phi_l(p, d) = p' \wedge \Forall{0 \leq t\leq d}{\Inv_l(\Phi_l (p, t))}}.$$
+
+A flow function on |pow IR 2| can be expressed as the
+product of two flow functions: $\Phi_l((c_0,T_0),t) = (\phi_{l,\Sc}((c_0,T_0),t),
+\phi_{l,\ST}((c_0,T_0),t))$. In the thermostat example, as in many other
+examples of hybrid systems, $\phi_{l,\Sc}((c_0,T_0),t)$ does not actually depend on $T_0$
+and $\phi_{l,\ST}((c_0,T_0),t)$ does not actually depend on $c_0$. We call this feature
+{\em separability\/} of the flow function. Our development currently
+relies heavily on this property. Separability makes the form of the
+flow functions simpler:
+$$\Phi_l((c_0,T_0),t) = (\phi_{l,\Sc}(c_0,t), \phi_{l,\ST}(T_0,t))$$
+
+In the thermostat, $\phi_{l,\Sc}(c_0,t) = c_0 + t$ for all locations
+$l$, $\phi_{\Heat,\ST}(T_0,t)= T_0 + 2 t$, $\phi_{\Chec,\ST}(T_0,t)=
+T_0 * e^{- \frac{1}{2}t}$ and $\phi_{\Cool,\ST}(T_0,t)= T_0 *
+e^{-t}$. So $\phi'_{\Cool,\ST}(T_0,t)= - \phi_{\Cool,\ST}(T_0,t)$,
+solving the differential equation $\dot{T} = - T$ for the $\Cool$
+location.
 
@@
 
-A transition is either continuous or discrete: $\trans:= \distrans \cup \contrans$.
-We now say that a state $s$ is \emph{reachable} if there is an initial state $i$ from which one can, by a finite number of transitions, end up in $s$ ($\ttrans$ is the transitive reflexive closure of $\trans$):
+A transition is either continuous or discrete: $\trans:= \distrans
+\cup \contrans$. A finite sequence of transitions constitutes a {\em trace}. We now say that a state $s$ is \emph{reachable} if
+there is an initial state $i$ from which there is a trace to $s$.
+% one can, by a finite number
+%of transitions, end up in $s$ (
+If we denote by $\ttrans$ the transitive reflexive
+closure of $\trans$, that is:
 $$\Reach(s):=
     \Exists{i\in \State}{\Init(i) \wedge i \ttrans s}.$$
+
 \weg{As mentioned before, the set of reachable states represents the
@@
 
-The object of hybrid system safety verification is to show that the set of reachable states is a subset of a predefined set of ``safe'' states. For the thermostat, the intent is to
+The objective of hybrid system safety verification is to show that the set of reachable states is a subset of a predefined set of ``safe'' states. For the thermostat, the intent is to
 keep the temperature above $4.5$ degrees at all times, and so we define $\Safe(c,T) := T > 4.5$ (and $\Unsafe(c,T)$ as its complement).
@@
 \begin{eqnarray*}
-(l,P) \acontrans (l',Q) &:=& l=l' \wedge \Exists{p\in P, q\in Q}{(l,p) \contrans (l,q)}\\
+(l,P) \acontrans (l,Q) &:=& \Exists{p\in P, q\in Q}{(l,p) \contrans (l,q)}\\
 (l,P) \adistrans (l',Q) &:=& \Exists{p\in P, q\in Q}{(l,p) \distrans (l,q)}.
 \end{eqnarray*}
-Now {\em abstract reachability\/} can be defined in the obvious way as
+Now {\em abstract reachability\/} can be defined as
 follows $\AReach(a) := \Exists{s_0\in\State}{\Init(s_0) \wedge A(s_0)
@@
 system from safety of the abstract system. We only need the property
-of diagram \ref{diag:abstraction}, so we can {\em overestimate\/}
+of Figure \ref{diag:abstraction}, so we can {\em over-estimate\/}
 $\acontrans$ and $\adistrans$ (i.e. replace it with a transition
@@
 
-This is indicated in figure \ref{fig:abstraction}. The left side illustrates
-the challenge: given abstract regions $A$ and $B$, we are to determine
-whether nonnegative flow durations might permit flow from points in $A$ to points in $B$.
-Following the overestimation property just mentioned, we introduce an abstract transition from $A$ to $B$ whenever we cannot positively rule out such paths.
-
-On the right side we see the abstract state space indicated for the
+This is indicated in figure \ref{fig:abstraction}. The left hand side
+illustrates the challenge: given abstract regions $A$ and $B$, we are
+to determine whether some flow duration permits flow from points in
+$A$ to points in $B$.  Following the over-estimation property just
+mentioned, we introduce an abstract transition from $A$ to $B$
+whenever we cannot positively rule this out.
+
+On the right hand side we see the abstract state space indicated for the
 location $\Heat$. The abstract state space consists of rectangles,
@@
 \ref{fig:thermostat}). According to \cite{alur}, this is a good
-choice. (In case one cannot prove safety in this, there is of course
-the opportunity for the user to refine it.)  The gray area indicates
+choice. (In case one cannot prove safety, there is of course
+the opportunity for the user to refine the bounds.)  The grey area indicates
 that from these states also abstract discrete transitions are
@@
 transitions from the rectangle $[0.5,1) \times[5,6)$ are shown: as the
-    temperature flow function for $\Heat$ is $\phi_{\Heat,2}(T,d)= T +
-    2 * d$, and the clock flow function is $\phi_{\Heat,1}(c,d) = c +
-    d$, these are all the possible abstract transitions.
+    temperature flow function for $\Heat$ is $\phi_{\Heat,\ST}(T_0,t)= T_0 +
+    2 * t$, and the clock flow function is $\phi_{\Heat,\Sc}(c_0,t) = c_0 +
+    t$, these are all the possible abstract transitions.
 
@@
 \section{Formalisation}
-We now describe the Coq formalisation and the design choices made.
-We will not pay much attention to the specifics of Coq and its type theory CiC, and will instead focus on concerns relating to the use of computable reals and constructive logic.
-
-The complete sources of the development, as well as a technical report describing the formalization in more detail \cite{hybrid-techreport}, can be found at \newline {\tt http://www.eelis.net/research/hybrid/}.
+We now describe the Coq formalisation and the design choices made.  We
+will not pay much attention to the specifics of Coq and its type
+theory CiC, and will instead focus on concerns relating to the use of
+computable reals and constructive logic.  The complete sources of the
+development are available on the web, as well as a technical report
+describing the formalisation in more detail \cite{hybrid-techreport}.
+%, can be found
+%at \newline {\tt http://www.eelis.net/research/hybrid/}.
 
@@
 We now want to define an abstract system and an abstraction function
-satisfying the properties indicated in diagram
+satisfying the properties indicated in Figure
 \ref{diag:abstraction}. However, this is not possible, because we
@@
 The second is very problematic, because it means the property for the
-abstraction function $A$ depicted in diagram
+abstraction function $A$ depicted in Figure
 \ref{diag:abstraction} no longer holds.
@@
 
-The good thing is that we don't really need the property in Diagram
+The good thing is that we don't really need the property in Figure
 \ref{diag:abstraction}, because we can do something better. 
@@
 \item
-We replace the abstract transition relations $\acontrans$ and $\distrans$ by functions $\overcontrans$ and $\overdistrans$ that take a region $R_0$
+We replace the abstract relations $\acontrans$ and $\adistrans$ by functions $\overcontrans$ and $\overdistrans$ that take a region $R_0$
 as input and output a {\em list of regions\/} including $R_0$: $(R_o,
 R_1, \ldots, R_n)$ in such a way that $\cup_{0\leq i\leq n}R_i$ is an
-overapproximation of the set of states reachable by a continuous (resp. discrete) step
+over-approximation of the set of states reachable by a continuous (resp. discrete) step
 from a state in $R_0$.\\
@@
 
-To summarize, if $s\contrans s'$, then we don't require
+To summarise, if $s\contrans s'$, then we don't require
 $A(s')$ to be in the list $\overcontrans(A(s))$, but we only require
@@
 $\overcontrans$ and $\overdistrans$ yield a notion of {\em trace\/} in
-the abstract hybrid system in the obvious way: starting from $R_0$,
+the abstract hybrid system in the straightforward way: starting from $R_0$,
 take an $R_1$ in $\overcontrans( R_0)$, then an $R_2$ in
@@
 is only available inside |DN| unless all region borders have
-nontrivial overlap, which as we will see later is undesireable.
+nontrivial overlap, which 
+%as we will see later 
+is undesirable.
 
@@
 
-\subsection{Underestimation and Overestimation}
+\subsection{Under-Estimation and Over-Estimation}
 \label{estimation}
 
-% It \emph{will} always terminate, so we are not writing a semidecision procedure either.
+% It \emph{will} always terminate, so we are not writing a semi-decision procedure either.
 Ultimately, in our development we are writing a program that
@@
 \begin{code}
-Definition underestimation (P: Prop): Set := { b: bool | b = true -> P }.
-\end{code}
-Using the |Program| family of commands
-\cite{sozeau}, an underestimation may be
-provisionally defined strictly as a bool, and then separately proved
-to be a valid underestimation in a proof obligation generated by
-|Program|. The |bool| in the definition nicely illustrates why we call this an
-``underestimation'': it may be |false| even when |P| holds. We can now
+Definition under-estimation (P: Prop): Set := { b: bool | b = true -> P }.
+\end{code}
+%Using the |Program| family of commands
+%\cite{sozeau}, an underestimation may be
+%provisionally defined strictly as a bool, and then separately proved
+%to be a valid underestimation in a proof obligation generated by
+%|Program|. 
+The |bool| in the definition nicely illustrates why we call this an
+``under-estimation'': it may be |false| even when |P| holds. We can now
 describe the functionality of our program by saying that it
-underestimates hybrid system safety, yielding a term of type
+under-estimates hybrid system safety, yielding a term of type
 |underestimation Safe|, where |Safe| is a proposition expressing
@@
 
-Considered as theorems, underestimations are not very interesting,
+Considered as theorems, under-estimations are not very interesting,
 because they can be trivially ``proved'' by taking
@@
 |underestimation Safe|, but rather by the fact that when run, it
-actually manages to returns |true| for the hybrid system we
+actually manages to return |true| for the hybrid system we
 are interested in (e.g.\ the thermostat). It is for this reason that
@@
 
-The opposite of an underestimation is an overestimation:
+The opposite of an under-estimation is an over-estimation:
 \begin{code}
@@
 states, we may equivalently express the functionality of our
-development by saying that it overestimates unsafe state
+development by saying that it over-estimates unsafe state
 reachability. Indeed, most subroutines in our programs will be
-overestimators rather than underestimators. Notions of overestimation
-and underestimation trickle down through all layers of our
+over-estimators rather than under-estimators. Notions of over-estimation
+and under-estimation trickle down through all layers of our
 development, down to basic arithmetic. For instance, we employ
-functions such as:
+functions such as (recall that |CR| denotes the type of constructive
+reals):
 
@@
 allow to prove it safe (if indeed it is safe at all).  Furthermore,
-careless use of region overlap can result in undesireable abstract
+careless use of region overlap can result in undesirable abstract
 transitions (and therefore traces), adversely affecting the abstract
@@
 proof. 
-Having to do this ``tweaking'' manually is obviously not ideal. One may want to develop heuristics for this.
+Having to do this ``tweaking'' manually is clearly not ideal. One may want to develop heuristics for this.
 
@@
 lies in the fact that our bounds are always inclusive, which means
-adjacent regions overlap in lines. We will discuss this in more detail
-later.
+adjacent regions overlap in lines. 
+%We will discuss this in more detail
+%later.
 
@@
 Once we have a satisfactory abstract |Space|, our goal is to construct
-an overestimatable notion of abstract reachability implied by concrete
+an over-estimatable notion of abstract reachability implied by concrete
 reachability, so that concrete unreachability results may be obtained
-simply by executing the abstract reachability overestimator. We first overestimate the continuous transitions; we need the following definition for that.
+simply by executing the abstract reachability over-estimator. We first over-estimate the continuous transitions; we need the following definition for that.
 
@@
 \mid \Inv(c) \wedge s\cap \{c' \mid c' \contrans c\} \neq \emptyset\}$.
-We similarly specify |over_disc_trans| as an overestimator for
-|concrete_disc_trans| and |over_initial| as an overestimator of
+We similarly specify |over_disc_trans| as an over-estimator for
+|concrete_disc_trans| and |over_initial| as an over-estimator of
 |concrete.initial|.
@@
 \end{code}
-Which would nicely imply:
+because it implies
+%Which would nicely imply:
 \begin{code}
@@
 \end{code}
-This would expresses that to conclude unreachability of a concrete state,
-one need only establish unreachability of \emph{any} abstract state
+This expresses that to conclude unreachability of a concrete state,
+one only needs to establish unreachability of \emph{one} abstract state
 that contains it. However, this definition neglects to facilitate {\em
-  sharing}: remember that a concrete state may be in more then one
-abstract state. When abstract states ``share'' the burden of
-reachability, one should establish unreachability of \emph{all}
+  sharing}: a concrete state may be in more than one
+abstract state. So, if a concrete state is in one abstract state which is unreachable, it may still be in another abstract state which {\em is\/} reachable.
+% When abstract states ``share'' the burden of
+%reachability, 
+One should establish unreachability of \emph{all}
 abstract states containing the concrete state. Hence, what we really
@@
 
-\subsection{Underestimating Safety}
+\subsection{Under-Estimating Safety}
 \label{underestimatingsafety}
@@
 
-In the next section we show that thanks to decidability of our
-transition and initiality overstimators, |abstract.reachable| is
-decidable. But first we show how a
-decision procedure for |abstract.reachable| lets us underestimate
-hybrid system safety, and in particular, lets us obtain a proof of
-thermostat safety. So suppose we have |reachable_dec: decider abstract.reachable|.
-\weg{
-And suppose we are given the following specification of unsafe concrete states, covered by a finite list of abstract states:
+%In the next section we show that thanks to decidability of our
+%transition and initiality overestimators, |abstract.reachable| is
+%decidable. But first 
+We now show how a decision procedure for |abstract.reachable| lets us
+under-estimate hybrid system safety, and in particular, lets us obtain
+a proof of thermostat safety. (The construction of the decision
+procedure itself is detailed in the next section.) So suppose we have
+|reachable_dec: decider abstract.reachable|.
+\weg{ And suppose we are
+  given the following specification of unsafe concrete states, covered
+  by a finite list of abstract states:
 \begin{code}
@@
 }
-Recall that |ThermoSafe| was defined as |thermo_unsafe `subsetof` concrete.unreachable| in section \ref{concrete}.
+%Recall that 
+|ThermoSafe| is defined as |thermo_unsafe `subsetof` concrete.unreachable|.
+% in section \ref{concrete}. 
 Since we trivially have |~ overlap unsafe concrete.reachable -> ThermoSafe|, we also have:
@@
 \end{code}
-Using a tiny utility |underestimation_true| of type |forall P (o: underestimation P), o = true -> P|, we can now \emph{run} this underestimator to obtain a proof of the thermostat system's safety:
+Using a tiny utility |underestimation_true| of type |forall P (o: underestimation P), o = true -> P|, we can now \emph{run} this under-estimator to obtain a proof of the thermostat system's safety:
 \begin{code}
@@
 The first |apply| reduces the goal to
-\begin{code}under_thermo_unsafe_unreachable = true.\end{code} The |vm_compute| tactic invocation then forces evaluation of the left hand side, which will in turn evaluate |over_thermo_unsafe_reachable|, which will evaluate |reachable_dec|, which will evaluate the overestimators of the continuous and discrete transitions.
+\begin{code}under_thermo_unsafe_unreachable = true.\end{code} The |vm_compute| tactic invocation then forces evaluation of the left hand side, which will in turn evaluate |over_thermo_unsafe_reachable|, which will evaluate |reachable_dec|, which will evaluate the over-estimators of the continuous and discrete transitions.
 This process, which takes about 35 seconds on a modern desktop machine, eventually reduces |under_thermo_unsafe_unreachable| to |true|, leaving |true = true|, proved by |reflexivity|.
 
-We can now also clearly see what happens when the abstraction method ``fails'' due to poor region selection, overly simplistic transition/initiality overestimators, or plain unsafety of the system.
+We can now also clearly see what happens when the abstraction method ``fails'' due to poor region selection, overly simplistic transition/initiality over-estimators, or plain unsafety of the system.
 In all these cases, |vm_compute| reduces |under_thermo_unsafe_unreachable| to |false|, and the subsequent |reflexivity| invocation will fail.
@@
 are the implementation of |reachable_dec| in terms of the decidable
-overestimators for abstract initiality and continuous and discrete
-transitions, and the implementation of those overestimators
+over-estimators for abstract initiality and continuous and discrete
+transitions, and the implementation of those over-estimators
 themselves. The former is a formally verified graph reachability
-algorithm, that we don't detail here. The overestimator for continuous
+algorithm, that we don't detail here. The over-estimator for continuous
 transitions, |over_cont_trans| will be detailed in the next section
@@
 
-\subsection{Overestimating Continuous Abstract Transitions}
+\subsection{Over-Estimating Continuous Abstract Transitions}
 \label{over_cont_trans}
 
-We now discuss the implementation of  |over_cont_trans|:
+We now discuss the implementation of  |over_cont_trans|.
 Given two regions |r_src| and |r_dst|, 
@@
 are no points in |r_src| which the flow function maps to points in
-|r_dst|, then there need be no abstract continuous transition between |r_src|
+|r_dst|, then we don't put an abstract continuous transition between |r_src|
 and |r_dst|. Clearly, this is impossible to
-meaningfully overestimate for a general flow function and general
+meaningfully over-estimate for a general flow function and general
 regions. However, the thermostat posesses three key properties that
@@
 %element of $b$ via flow $\phi$.
-Range invertibility is a less demanding alternative to point invertibility: $\phi^{-1}$ is the {\em point inverse\/} of $\phi$ if $\forall x,y\in \IR (\phi (x, \phi^-1(x,y)) =y)$.
-So a point inverse $\phi^{-1}(x,y)$ computes the exact time it takes to go from $x$ to $y$ via flow $\phi$.
+Range invertibility is a less demanding alternative to point
+invertibility: $\phi^{-1}$ is the {\em point inverse\/} of $\phi$ if
+$\forall x,y\in \IR (\phi (x, \phi^-1(x,y)) =y)$. So a point inverse
+$\phi^{-1}(x,y)$ computes the exact time $t$ it takes to go from $x$
+to $y$ via flow $\phi$. A range inverse computes an interval that
+contains this $t$.
 %We need range invertibility because the exponential flow functions are
@@
 
-In the formalization we use a modest library
+In the formalisation we use a modest library
 of flow functions when defining the thermostat's flow. Included in
@@
 and having argued that the thermostat's flow is in this class, we now
-show how to proceed with our overestimation of existence of points in
+show how to proceed with our over-estimation of existence of points in
 |r_src| which the flow function map to points in |r_dst|. 
@@
 
-We now simply use an |OpenRange| overlap overestimator of type
+We now simply use an |OpenRange| overlap over-estimator of type
 \begin{code}
@@
 \end{code}
-(defined in terms of things like |overestimate_CRle| shown in section \ref{estimation}) to overestimate whether the following three intervals overlap:
+(defined in terms of things like |overestimate_CRle| shown in section \ref{estimation}) to over-estimate whether the following three intervals overlap:
 \begin{enumerate}
@@
 |r_src| to points in |r_dst|, and so our |abstract.cont_trans|
-overestimator may justifiably return ``false'' when the overlap
-overestimator manages to prove absence of overlap.
+over-estimator may justifiably return ``false'' when the overlap
+over-estimator manages to prove absence of overlap.
 
@@
 most often rely on standard floating point arithmetic, that
-comes with its rounding errors. Few tools that address 
+comes with its rounding errors. Some tools that address 
 this problem include 
@@
 
-However, to the best of our knowledge, none of the previous work nor tools 
-relies on a precise model of real number computations completely verified 
-in a theorem prover; such as the model of CoRN used in this work.
+However, to the best of our knowledge, none of the work or tools
+discussed above rely on a precise model of real number computations
+completely verified in a theorem prover; such as the model of CoRN
+used in this work.
 
@@
 
-It remains to be seen how far this automated verification approach can be
-taken, given the fact that we have limited ourselves to hybrid systems
-with separable reset and flow functions that are monotone, and with a
-stable invariants. There is still a lot of room for more clever
-heuristics with less restrictive preconditions. The heuristic in
+It remains to be seen how far this automated verification approach can
+be taken, given the fact that we have limited ourselves to hybrid
+systems where we have a {\em solution\/} to the differential equation
+as a flow function, this flow function is {\em separable} (Section
+\ref{sec:hybsys}) and {\em range invertible} (Section
+\ref{over_cont_trans}) and the invariants are {\em stable} (Section
+\ref{concrete}). Finally, the reset functions should not be too
+strange, see \cite{hybrid-techreport} for details. If the flow
+functions are given, the largest part of the work is in producing {\em
+  useful\/} range inverse functions. Note that we always have the
+trivial range inverse function, that just returns $\IR$, but that is
+not useful.  We want a function that actually helps to exclude certain
+abstract continuous transitions.
+
+There is still a lot of room for more clever
+heuristics, possibly with less restrictive preconditions. The heuristic in
 \cite{alur} for bound selection doesn't work out of the box, but
@@
 like the system to generate an ``offending trace'' automatically,
-which can the be inspected by the user.
+which can then be inspected by the user.
+
+\paragraph{Acknowledgements} 
+We thank the referees for their useful suggestions; due
+to space limitations we have not been able to incorporate all of them.
 
